### 2017-02-14 paper/oopsla17-submission/evaluation.tex
init paper

\section{Evaluation}
\label{sec:evaluation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2017-02-20 paper/oopsla17-submission/related.tex
init paper
throw in a few paragraphs on related work for fault localization to get us started

--- 2017-02-14 paper/oopsla17-submission/evaluation.tex
@@
-\section{Evaluation}
-\label{sec:evaluation}
+\section{Related Work}
+\label{sec:related-work}
+
+In this section we describe two relevant aspects of related work:
+software engineering approaches to fault localization, and 
+programming languages approaches to pinpointing type errors.
+
+Given a defect, fault localization is the task of identifying
+``suspicious'' program elements (e.g., lines, statements) that are likely
+implicated in the defect (i.e., that should be changed to fix the defect). 
+Perhaps the best-known is Tarantula, which uses a simple mathematical
+formula based on measured information from dynamic normal and buggy
+runs~\cite{FIXME}. Other similar approaches, including those of
+Jacquard~\cite{FIXME} and Ochiai~\cite{FIXME} consider alternate features
+of information or refined formulae and generally obtain more precise
+results; see Wong for a survey~\cite{FIXME}. While some researchers have
+approached such fault localization with an eye toward optimality (e.g.,
+Harman et al. determine optimal coefficients~\cite{FIXME}), in general
+such fault localization approaches are limited by their reliance on either
+running tests or including relevant features. For example, Tarantula-based 
+techniques require a normal and a buggy run of the program. By contrast,
+we consider incomplete programs with type errors that may not be executed
+in any standard sense. Similarly, the features available influence the
+classes of defects that can be localized. A fault localization scheme based
+purely on control flow features lines will have difficulty with cross-site
+scripting or SQL code injection attacks, which follow the same control flow
+path on normal and buggy runs (differing only in the user-supplied data).
+By contrast, we FIXME-ERIC.
+
+Languages with static type systems and type inference produce type errors
+that novices often perceive as difficult to interpret~\cite{FIXME}.  For
+example, approaches based on Hindley-Milner or other constraint systems 
+typically have the issue that when the constraint $A=B$ is violated, 
+the system cannot know whether the user should intend to fix $A$ or $B$ and
+must thus report the discrepancy in a generic manner. As a result, a number
+of approaches have been proposed to either give clearer error messages
+(e.g.,~\cite{FIXME}) or more precisely localize the type error
+(e.g.,~\cite{FIXME}). The technique most related to our work is FIXME-ERIC. 
+It has been shown to be quite good at FIXME-ERIC. However, we choose to 
+focus on FIXME-ERIC, and our approach contains FIXME-ERIC, which is not
+present in previous work.
 

### 2017-03-03 paper/oopsla17-submission/related.tex
sketch outline of paper

@@
 formula based on measured information from dynamic normal and buggy
-runs~\cite{FIXME}. Other similar approaches, including those of
-Jacquard~\cite{FIXME} and Ochiai~\cite{FIXME} consider alternate features
+runs~\cite{Jones2002-us}. Other similar approaches, including those of
+Jaccard~\cite{Chen2002-qz} and
+Ochiai~\cite{Abreu2006-fn,Abreu2007-mu} consider alternate features
 of information or refined formulae and generally obtain more precise
-results; see Wong for a survey~\cite{FIXME}. While some researchers have
+results; see Wong for a survey~\cite{Wong2009-pd}. While some researchers have
 approached such fault localization with an eye toward optimality (e.g.,
-Harman et al. determine optimal coefficients~\cite{FIXME}), in general
+Harman et al. determine optimal coefficients~\cite{Yoo2013-rw}), in general
 such fault localization approaches are limited by their reliance on either

### 2017-03-03 paper/oopsla17-submission/learning.tex
start drafting technical bits

\section{Learning to Blame}
\label{sec:learning}

\subsection{Syntax}
\label{sec:syntax}
\input{syntax}
%
\autoref{fig:syntax} describes the syntax of \lang, a simple lambda
calculus with integers, booleans, pairs, and lists.

\subsection{Features}
\label{sec:features}
The first issue we must tackle is formulating our learning task in
machine learning terms.
%
We are given expressions $e$, but the learning algorithms expect to work
with \emph{feature vectors} --- vectors of real numbers, where each
column describes a particular feature of the input.
%
Thus, our first task is to convert expressions to feature vectors.

We choose to model a program as a \emph{set} of feature vectors, where
each element corresponds a sub-expression in the program. We group the
features into four categories.

\paragraph{Local syntactic features}
These features describe the syntactic category of each sub-expression
$e$.
%
In other words, for each production of $e$ in \autoref{fig:syntax} we
introduce a feature that is enabled (set to $1$) \emph{iff} the
sub-expression was built with that production.

\paragraph{Contextual syntactic features}
These are like local syntactic features, but lifted to describe the
parent and children of the current sub-expression.
%
If a particular $e$ does not have children (\eg a variable $x$) or a
parent (\ie the root expression), we leave the corresponding features
disabled (set to $0$).
%
This gives us a notion of the \emph{context} in which an expression
occurs, similar to the \emph{n-grams} commonly found in linguistic
models.

Instead of just describing the immediate context, we could describe
whether a particular syntax element occurs in the neighboring
sub-expressions (or even a count of how many times it occurs).
%
Such fuzzier notions of context may enable increased precision in the
model, but they also introduce opportunities for \emph{overfitting} ---
where the model memorizes particular inputs rather than learning general
patterns.
%
We will investigate (\ES{maybe..}) the impact of these alternatives
in \autoref{sec:evaluation}.

\paragraph{Expression size}
We also add a feature representing the \emph{size} of each expression,
\ie how many sub-expressions does it contain?
%
This allows the model to learn that, \eg, expressions closer to the
leaves are more likely to be blamed than expressions closer to the root.

\paragraph{Typing features}

\subsection{Labels}
\label{sec:labels}

\subsection{Models}
\label{sec:models}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2017-03-06 paper/oopsla17-submission/learning.tex
more features

@@
 each element corresponds a sub-expression in the program. We group the
-features into four categories.
+features into the following categories.
 
@@
 \paragraph{Typing features}
+A natural way of summarizing the context in which an expression occurs
+is with \emph{types}.
+%
+A difficulty that arises here is that, due to the parametric type
+constructors $\tfun{\cdot}{\cdot}$, $\tprod{\cdot}{\cdot}$, and
+$\tlist{\cdot}$, there is an \emph{infinite} set of possible types ---
+but we must have a \emph{finite} set of features.
+%
+Thus, we add features for each type \emph{constructor} that describe
+whether a given type \emph{mentions} the type constructor.
+%
+For example, the type $\tint$ would only enable the $\tint$ feature,
+while the type $\tfun{\tint}{\tbool}$ would enable the
+$\tfun{\cdot}{\cdot}$, $\tint$, and $\tbool$ features.
+%
+We add these features for parent and child expressions to summarize the
+context, but also for the current expression, as the type of an
+expression is not always clear \emph{syntactically} (\eg for variables
+and applications).
+\ES{note the possibility of traversal bias from typing features}
+
+\paragraph{Type error slice}
+Finally, we would like the model to be able to distinguish between
+changes that could fix the error, and changes that
+\emph{cannot possibly} fix the error.
+%
+Thus, we compute a minimal type error \emph{slice} for the program
+(\ie the set of expressions that contribute to the error), and add a
+feature that is enabled \emph{iff} an expression is part of the slice.
+%
+If the program contains multiple type errors, we will compute a minimal
+slice for each error.
 

### 2017-03-06 paper/oopsla17-submission/learning.tex
talk about labels

@@
 \label{sec:learning}
-
-\subsection{Syntax}
-\label{sec:syntax}
+Next, we describe our approach to localizing type errors, in the context
+of \lang (\autoref{fig:syntax}), a simple lambda calculus with integers,
+booleans, pairs, and lists.
+%
 \input{syntax}
 %
-\autoref{fig:syntax} describes the syntax of \lang, a simple lambda
-calculus with integers, booleans, pairs, and lists.
+First, we define the inputs to the model, a set of \emph{features}
+(\autoref{sec:features}) that we will use to describe programs.
+%
+Second, we define the expected outputs of the model, a set of
+\emph{labels} (\autoref{sec:labels}) that we will use to assign blame.
+%
+Third, we define the actual \emph{models} (\autoref{sec:models}) that we
+will evaluate.
 
@@
 %
-We are given expressions $e$, but the learning algorithms expect to work
+We are given programs over $e$, but the learning algorithms expect to work
 with \emph{feature vectors} --- vectors of real numbers, where each
-column describes a particular feature of the input.
+column describes a particular aspect of the input.
 %
-Thus, our first task is to convert expressions to feature vectors.
+Thus, our first task is to convert programs to feature vectors.
 
 We choose to model a program as a \emph{set} of feature vectors, where
-each element corresponds a sub-expression in the program. We group the
+each element corresponds an expression in the program. We group the
 features into the following categories.
@@
 \paragraph{Local syntactic features}
-These features describe the syntactic category of each sub-expression
-$e$.
+These features describe the syntactic category of each expression $e$.
 %
@@
 introduce a feature that is enabled (set to $1$) \emph{iff} the
-sub-expression was built with that production.
+expression was built with that production.
 
@@
 These are like local syntactic features, but lifted to describe the
-parent and children of the current sub-expression.
+parent and children of the current expression.
 %
@@
 whether a particular syntax element occurs in the neighboring
-sub-expressions (or even a count of how many times it occurs).
+expressions (or even a count of how many times it occurs).
 %
@@
 %
-Thus, we add features for each type \emph{constructor} that describe
-whether a given type \emph{mentions} the type constructor.
+Thus, we abstract the type of an expression to the set of type
+constructors it \emph{mentions}, and add features for each type
+constructor that describe whether a given type mentions the type
+constructor.
 %
@@
 and applications).
+%
 \ES{note the possibility of traversal bias from typing features}
@@
 \label{sec:labels}
+We define the outputs of the model to be a simple ``good'' or ``bad''
+label, where ``good'' means the expression \emph{should not} change and
+``bad'' means the expression \emph{should} change.
+%
+This allows us to predict whether any individual expression should
+change, but we would actually like to predict the \emph{most likely}
+expressions to change.
+%
+Many learning algorithms produce not only a prediction, but also a
+metric that can be interpreted as a \emph{confidence} in the prediction.
+%
+Thus, we add a post-processing step where we \emph{rank} the expressions
+by the model's confidence that they will change, and select only the top
+$k$ (in practice $k=3$) to present to the user.
 

### 2017-03-06 paper/oopsla17-submission/related.tex
misc

--- 2017-02-20 paper/oopsla17-submission/related.tex
@@
 formula based on measured information from dynamic normal and buggy
-runs~\cite{FIXME}. Other similar approaches, including those of
-Jacquard~\cite{FIXME} and Ochiai~\cite{FIXME} consider alternate features
+runs~\citep{Jones2002-us}. Other similar approaches, including those of
+\citet{Chen2002-qz} and \citet{Abreu2006-fn,Abreu2007-mu} consider alternate features
 of information or refined formulae and generally obtain more precise
-results; see Wong for a survey~\cite{FIXME}. While some researchers have
+results; see \citet{Wong2009-pd} for a survey. While some researchers have
 approached such fault localization with an eye toward optimality (e.g.,
-Harman et al. determine optimal coefficients~\cite{FIXME}), in general
+\citet{Yoo2013-rw} determine optimal coefficients), in general
 such fault localization approaches are limited by their reliance on either
@@
 path on normal and buggy runs (differing only in the user-supplied data).
-By contrast, we FIXME-ERIC.
+By contrast, we \ES{TODO}.
 
 Languages with static type systems and type inference produce type errors
-that novices often perceive as difficult to interpret~\cite{FIXME}.  For
-example, approaches based on Hindley-Milner or other constraint systems 
+that novices often perceive as difficult to
+interpret~\citep{Wand1986-nw,Beaven1993-hb,Duggan1996-by,Yang2000-yr,McAdam1998-ub}.
+For example, approaches based on Hindley-Milner or other constraint systems
 typically have the issue that when the constraint $A=B$ is violated, 
@@
 (e.g.,~\cite{FIXME}) or more precisely localize the type error
-(e.g.,~\cite{FIXME}). The technique most related to our work is FIXME-ERIC. 
-It has been shown to be quite good at FIXME-ERIC. However, we choose to 
-focus on FIXME-ERIC, and our approach contains FIXME-ERIC, which is not
+(e.g.,~\cite{FIXME}). The technique most related to our work is \ES{TODO}.
+It has been shown to be quite good at \ES{TODO}. However, we choose to
+focus on \ES{TODO}, and our approach contains \ES{TODO}, which is not
 present in previous work.

### 2017-03-07 paper/oopsla17-submission/learning.tex
misc
examples of features

@@
 %
-Third, we define the actual \emph{models} (\autoref{sec:models}) that we
+Third, we describe the actual \emph{models} (\autoref{sec:models}) that we
 will evaluate.
@@
 \label{sec:features}
+
+\begin{figure}[ht]
+\begin{minipage}{0.6\linewidth}
+\begin{lstlisting}
+  let rec sumList xs =
+    match xs with
+    | []     -> []
+    | hd::tl -> hd + sumList tl
+\end{lstlisting}
+\end{minipage}
+\begin{minipage}{0.3\linewidth}
+\begin{lstlisting}[numbers=left]
+[]
+hd + sumList tl
+sumList tl
+tl
+\end{lstlisting}
+\end{minipage}
+\caption{An ill-typed program (left) and a selection of its
+  sub-expressions (right).}
+\label{fig:sumList}
+\end{figure}
+\begin{table}[ht]
+\begin{tabular}{lrrrrrrr}
+\toprule
+Expression
+  & \IsNil & \IsCaseListP & \CountVarP & \ExprSize
+  & \HasTypeIntCOne & \HasTypeList & \InSlice \\
+\midrule
+\lstinline![]!
+  & 1 & 1 & 3 & 1 & 0 & 1 & 1 \\
+\lstinline!hd + sumList tl!
+  & 0 & 1 & 3 & 5 & 1 & 0 & 1 \\
+\lstinline!sumList tl!
+  & 0 & 0 & 3 & 3 & 0 & 1 & 1 \\
+\lstinline!tl!
+  & 0 & 0 & 2 & 1 & 0 & 1 & 0 \\
+\bottomrule
+\end{tabular}
+\caption{A selection of the features we would extract from the
+  \lstinline!sumList! program in \autoref{fig:sumList}. A feature is
+  considered \emph{enabled} if it has a non-zero value, and
+  \emph{disabled} otherwise. A ``-P'' suffix indicates that the feature
+  describes the parent of the current expression, a ``-C$n$'' suffix
+  indicates that the feature describes the $n$-th (left-to-right) child
+  of the current expression.  Note that, since we rely on a partial
+  typing derivation, we are subject to the well-known traversal bias and
+  label the expression \lstinline!sumList tl! as having type
+  $\tlist{\cdot}$. The model will have to learn to correct for this
+  bias.}
+\label{tab:sumList}
+\end{table}
+
 The first issue we must tackle is formulating our learning task in
@@
 We choose to model a program as a \emph{set} of feature vectors, where
-each element corresponds an expression in the program. We group the
-features into the following categories.
+each element corresponds an expression in the program.
+%
+Thus, given the \lstinline!sumList! program in \autoref{fig:sumList} we
+would first split it into its constituent sub-expressions and then
+transform each sub-expression into a single feature vector.
+%
+We group the features into five categories, using \autoref{tab:sumList}
+as a running example of the feature extraction process.
 
@@
 In other words, for each production of $e$ in \autoref{fig:syntax} we
-introduce a feature that is enabled (set to $1$) \emph{iff} the
-expression was built with that production.
+introduce a feature that is enabled (set to $1$) if the expression was
+built with that production, and disabled (set to $0$) otherwise.
+%
+For example, the \IsNil feature in \autoref{tab:sumList} describes
+whether the current expression is the empty list $\enil$.
+
+We distinguish between matching on a list vs.\ on a pair, as this
+affects the typing derivation.
+%
+We also assume that all pattern matches are well-formed --- \ie all
+patterns must match on the same type.
+%
+Ill-formed match expressions would lead to a type error; however, they
+are already effectively localized to the match expression itself.
+%
+This is not a \emph{fundamental} limitation though, one could easily add
+features that specify whether a match \emph{contains} a particular
+pattern, and thus have a match expression that enables multiple
+features. \ES{This is actually sort of what we do, we just ignore
+  programs that have patterns of different types..}
 
@@
 %
+For example, the \IsCaseListP feature in \autoref{tab:sumList} describes
+whether current expression's \emph{parent} matches on a list.
+%
 If a particular $e$ does not have children (\eg a variable $x$) or a
 parent (\ie the root expression), we leave the corresponding features
-disabled (set to $0$).
+disabled.
 %
@@
 expressions (or even a count of how many times it occurs).
+%
+For example, the \CountVarP feature in \autoref{tab:sumList} describes
+how many variables are contained in the expression \emph{rooted} at the
+current expression's parent.
 %
@@
 %
+For example, the \ExprSize feature in \autoref{tab:sumList} is set to 3
+for the expression \lstinline!sumList tl! as it contains 3 expressions:
+the two variables and the application itself.
+%
 This allows the model to learn that, \eg, expressions closer to the
@@
 %
+Of course, the programs we are given are \emph{untypeable}, but we can
+still extract a \emph{partial} typing derivation from the type checker
+and use it to provide more information to the model.
+
 A difficulty that arises here is that, due to the parametric type
@@
 $\tfun{\cdot}{\cdot}$, $\tint$, and $\tbool$ features.
-%
+
 We add these features for parent and child expressions to summarize the
 context, but also for the current expression, as the type of an
-expression is not always clear \emph{syntactically} (\eg for variables
-and applications).
+expression is not always clear \emph{syntactically}.
 %
-\ES{note the possibility of traversal bias from typing features}
+For example, the expressions \lstinline!tl! and \lstinline!sumList tl!
+in \autoref{tab:sumList} both enable the feature \HasTypeList, as they
+are both inferred to have a type that mentions the $\tlist{\cdot}$
+constructor.
+
+Note that our use of typing features in an \emph{ill-typed} program
+subjects us to the infamous traversal
+bias~\cite{McAdam1998-ub,Yang2000-yr} --- in particular with the
+\lstinline!sumList tl! expression, which might alternatively be assigned
+the type $\tint$.
+%
+We rely on the model to counteract this bias.
 
@@
 %
-If the program contains multiple type errors, we will compute a minimal
-slice for each error.
+The \InSlice feature in \autoref{tab:sumList} indicates whether an
+expression is part of such a minimal slice, and is enabled for all of
+the sampled expressions except for \lstinline!tl!, which does not affect
+the type error.
+%
+If the program contains multiple type errors, we will compute
+a minimal slice for each error.
 
@@
 \label{sec:models}
+\ES{TODO: trade-off between ease of training/interpreting and
+  complexity, \ie actually fitting the data..}
+
+\paragraph{Linear models}
+The simplest model we investigate is a linear model, which learns a
+linear function of the features.
+%
+The goal is to learn a set of weights $W$ and biases $b$ such that the
+function
+%
+$$
+y = Wx + b
+$$
+%
+effectively maps feature vectors $x$ to labels $y$.
+
+\ES{TODO: talk about logistic regression and the loss function?}
+
+Linear models are popular for their ease of training and of interpreting
+the resulting model; however, they can be quite limited in their
+applicability.
+%
+As the name implies, linear models will perform poorly when asked to
+learn an inherently \emph{nonlinear} function.
+%
+This limitation can be mitigated to some extent by adding more features
+to the model (\eg combinations of the existing features \ES{CITE?}), but
+the fundamental issue remains.
+
+\paragraph{Decision trees}
+\ES{TODO: Maybe Huma?}
+
+\paragraph{Random forests}
+\ES{TODO: Maybe Huma?}
+
+\paragraph{Neural networks}
+The last (and most complex) model we discuss is Artificial Neural
+Networks (ANN).
+
+The fundamental building block of an ANN is the linear model, the
+nonlinearity arises from the \emph{activation functions} that govern
+whether an individual linear model propagates its signal to the next layer.
 

### 2017-03-08 paper/oopsla17-submission/introduction.tex
init paper
blurb about contributions

--- 2017-02-14 paper/oopsla17-submission/evaluation.tex
@@
-\section{Evaluation}
-\label{sec:evaluation}
+\section{Introduction}
+\label{sec:introduction}
+
+\subsection{Contributions}
+\label{sec:contributions}
+Thus, we identify the following novel contributions of our work.
+%
+\begin{enumerate}
+\item An experimental evaluation of state-of-the-art type-error
+  localization techniques based not on expert \emph{opinion}, but
+  on the eventual \emph{fix} implemented by a novice user.
+  %
+  \ES{need to explain \emph{why} this is superior..}
+\item A machine learning approach to \emph{modeling} (novice) type
+  errors that can outperform the state-of-the-art by 10--30\% in
+  localizing errors.
+  %
+  Our model can be trained on a relatively small amount of data --- we
+  used a single class of around 50 students --- and appears to
+  generalize to other instances of the same class.
+  %
+  \ES{and hopefully completely separate classes!}
+  %
+  In contrast to many approaches to \emph{fault localization} from the
+  software engineering community, our model does not use any linguistic
+  modeling techniques, \eg \emph{n-grams} over the token stream; rather,
+  it relies entirely on features of the abstract syntax tree and a
+  partial typing derivation.
+  %
+  \ES{if there's time, investigate adding Wes' n-gram model}
+\end{enumerate}
 

### 2017-03-08 paper/oopsla17-submission/learning.tex
blurb about contributions

@@
 %
-We are given programs over $e$, but the learning algorithms expect to work
+We are given programs over \lang, but the learning algorithms expect to work
 with \emph{feature vectors} --- vectors of real numbers, where each
@@
 
-\subsection{Models}
+\subsection{Learning Algorithms}
 \label{sec:models}
-\ES{TODO: trade-off between ease of training/interpreting and
-  complexity, \ie actually fitting the data..}
+\ES{TODO: mention that this is a SUPERVISED, CLASSIFICATION problem}
+%
+There are a large number of learning algorithms to choose from, existing
+on a spectrum that balances expressiveness with ease of training (and of
+interpreting the learned model).
+%
+In this section we present the learning algorithms we will investigate,
+beginning with the simplest.
+%
+For each algorithm, we describe \ES{TODO...}
 
-\paragraph{Linear models}
-The simplest model we investigate is a linear model, which learns a
+\paragraph{Logistic regression}
+The simplest model we investigate is logistic regression, which learns a
 linear function of the features.
@@
 
-Linear models are popular for their ease of training and of interpreting
-the resulting model; however, they can be quite limited in their
-applicability.
+Linear models like logistic regression are popular for their ease of
+training and of interpreting the resulting model; however, they can be
+limited in their applicability.
 %
-As the name implies, linear models will perform poorly when asked to
+As the name implies, linear models may perform poorly when asked to
 learn an inherently \emph{nonlinear} function.

### 2017-03-10 paper/oopsla17-submission/evaluation.tex
hmm, graphs

@@
 \label{sec:evaluation}
+
+\pgfplotstableset{col sep=comma}
+\pgfplotstableread{../../data/sp14/baseline.csv}{\spring}
+\pgfplotstablevertcat{\spring}{../../data/sp14/ocaml/results.csv}
+\pgfplotstablevertcat{\spring}{../../data/sp14/mycroft/results.csv}
+\pgfplotstablevertcat{\spring}{../../data/sp14/sherrloc/results.csv}
+\pgfplotstablevertcat{\spring}{../../data/sp14/op+type+size/hidden-500/results.csv}
+
+\definecolor{blue1}{HTML}{DEEBF7}
+\definecolor{blue2}{HTML}{9ECAE1}
+\definecolor{blue3}{HTML}{3182BD}
+
+\begin{figure}[ht]
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+  ybar stacked,
+  title={Spring 2014},
+  ylabel={Accuracy},
+  bar width=20pt,
+  %ymin=0,
+  ymax=1,
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\%},
+  ymajorgrids = true,
+  symbolic x coords={baseline, ocaml, mycroft, sherrloc, op+type+size/hidden-500},
+  enlarge x limits=0.25,
+  xtick=data,
+  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc, \hiddenFH},
+  x tick label style={rotate=45},
+  reverse legend,
+  legend style={legend pos = north west},
+]
+\addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\spring};
+\addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\spring};
+\addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\spring};
+\legend{Top 1, Top 2, Top 3}
+% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/ocaml/results.csv};
+% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/mycroft/results.csv};
+% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/op+type+size/hidden-500/results.csv};
+% \addplot[draw=black, fill=blue!40] table[col sep=comma, x=tool, y=top-2] {../../data/sp14/op+type+size/hidden-500/results.csv};
+% \addplot[draw=black, fill=blue!60] table[col sep=comma, x=tool, y=top-3] {../../data/sp14/op+type+size/hidden-500/results.csv};
+\end{axis}
+\end{tikzpicture}
+
+\caption{reuslts!}
+\label{fig:results}
+\end{figure}
 

### 2017-03-10 paper/oopsla17-submission/evaluation.tex
graphs

@@
 \pgfplotstablevertcat{\spring}{../../data/sp14/sherrloc/results.csv}
+\pgfplotstablevertcat{\spring}{../../data/sp14/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\spring}{../../data/sp14/op+type+size/hidden-10/results.csv}
 \pgfplotstablevertcat{\spring}{../../data/sp14/op+type+size/hidden-500/results.csv}
@@
   ymajorgrids = true,
-  symbolic x coords={baseline, ocaml, mycroft, sherrloc, op+type+size/hidden-500},
-  enlarge x limits=0.25,
+  symbolic x coords={baseline, ocaml, mycroft, sherrloc,
+                     op+type+size/linear, op+type+size/hidden-10,
+                     op+type+size/hidden-500},
+  % enlarge x limits=0.25,
   xtick=data,
-  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc, \hiddenFH},
+  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,
+               \linear, \hiddenT, \hiddenFH},
   x tick label style={rotate=45},

### 2017-03-10 paper/oopsla17-submission/evaluation.tex
pretty graph

@@
   ybar stacked,
+  width=12cm,
+  height=8cm,
   title={Spring 2014},
@@
   bar width=20pt,
-  %ymin=0,
+  ymin=0.2,
   ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\%},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
   ymajorgrids = true,
-  symbolic x coords={baseline, ocaml, mycroft, sherrloc,
+  symbolic x coords={baseline, spacer1, ocaml, mycroft, sherrloc, spacer2,
                      op+type+size/linear, op+type+size/hidden-10,
@@
   xtick=data,
+  xtick style={draw=none},
   xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,

### 2017-03-13 paper/oopsla17-submission/evaluation.tex
add FA15 to comparison

--- 2017-03-10 paper/oopsla17-submission/evaluation.tex
@@
 \pgfplotstableset{col sep=comma}
-\pgfplotstableread{../../data/sp14/baseline.csv}{\spring}
-\pgfplotstablevertcat{\spring}{../../data/sp14/ocaml/results.csv}
-\pgfplotstablevertcat{\spring}{../../data/sp14/mycroft/results.csv}
-\pgfplotstablevertcat{\spring}{../../data/sp14/sherrloc/results.csv}
-\pgfplotstablevertcat{\spring}{../../data/sp14/op+type+size/hidden-500/results.csv}
+
+\pgfplotstableread{../../data/sp14/baseline.csv}{\FeatureBench}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+
+\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
+\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
+\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
+\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
+\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
+\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
+% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+
+\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-10/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
+
+\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-10/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-500/results.csv}
 
@@
 \begin{axis}[
-  ybar stacked,
+  %ybar stacked,
+  width=12cm,
+  height=8cm,
   title={Spring 2014},
@@
   bar width=20pt,
-  %ymin=0,
+  %ymin=0.2,
   ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\%},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
   ymajorgrids = true,
-  symbolic x coords={baseline, ocaml, mycroft, sherrloc, op+type+size/hidden-500},
-  enlarge x limits=0.25,
+  symbolic x coords={op+type+size/hidden-10, op+type+size/hidden-25, op+type+size/hidden-50,
+                     op+type+size/hidden-100, op+type+size/hidden-250, op+type+size/hidden-500},
+  % enlarge x limits=0.25,
   xtick=data,
-  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc, \hiddenFH},
+  xtick style={draw=none},
+  xticklabels={\hiddenT, \hiddenTF, \hiddenF, \hiddenH, \hiddenTHF, \hiddenFH},
   x tick label style={rotate=45},
@@
 ]
-\addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\spring};
-\addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\spring};
-\addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\spring};
+% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
+% \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
+% \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
+\addplot table[x=tool, y=top-1] {\HiddenBench};
+\addplot table[x=tool, y=top-2] {\HiddenBench};
+\addplot table[x=tool, y=top-3] {\HiddenBench};
 \legend{Top 1, Top 2, Top 3}
-% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/ocaml/results.csv};
-% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/mycroft/results.csv};
-% \addplot[draw=black, fill=blue!20] table[col sep=comma, x=tool, y=top-1] {../../data/sp14/op+type+size/hidden-500/results.csv};
-% \addplot[draw=black, fill=blue!40] table[col sep=comma, x=tool, y=top-2] {../../data/sp14/op+type+size/hidden-500/results.csv};
-% \addplot[draw=black, fill=blue!60] table[col sep=comma, x=tool, y=top-3] {../../data/sp14/op+type+size/hidden-500/results.csv};
 \end{axis}
 \end{tikzpicture}
+\caption{reuslts!}
+\label{fig:results}
+\end{figure}
 
+\makeatletter
+\newcommand\resetstackedplots{
+\makeatletter
+\pgfplots@stacked@isfirstplottrue
+\makeatother
+\addplot [forget plot,draw=none] coordinates{
+  (baseline,0) (spacer1,0)
+  (ocaml,0) (mycroft,0) (sherrloc,0) (spacer2,0)
+  (op+type+size/linear,0) (op+type+size/hidden-10,0) (op+type+size/hidden-500,0)
+};
+}
+\makeatother
+
+\begin{figure}[ht]
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+  ybar stacked,
+  width=14cm,
+  height=8cm,
+  title={Comparison of Type-Error Localization Techniques},
+  ylabel={Accuracy},
+  bar width=0.5cm,
+  ymin=0.2,
+  ymax=1,
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = true,
+  symbolic x coords={baseline, spacer1, ocaml, mycroft, sherrloc, spacer2,
+                     op+type+size/linear, op+type+size/hidden-10,
+                     op+type+size/hidden-500},
+  % enlarge x limits=0.5,
+  xtick=data,
+  xtick style={draw=none},
+  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,
+               \linear, \hiddenT, \hiddenFH},
+  x tick label style={rotate=45},
+  reverse legend,
+  transpose legend,
+  legend style={legend pos = north west, legend columns=4},
+]
+\addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=tool, y=top-1] {\SpringBench};
+\addplot[draw=black, fill=blue2, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\SpringBench};
+\addplot[draw=black, fill=blue3, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\SpringBench};
+\addlegendentry{Top 1}
+\addlegendentry{Top 2}
+\addlegendentry{Top 3}
+\addlegendimage{empty legend}
+\addlegendentry{FA15}
+
+\resetstackedplots
+
+\addplot[draw=black, fill=blue1, bar shift=.25cm] table[x=tool, y=top-1] {\FallBench};
+\addplot[draw=black, fill=blue2, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\FallBench};
+\addplot[draw=black, fill=blue3, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\FallBench};
+\addlegendentry{Top 1}
+\addlegendentry{Top 2}
+\addlegendentry{Top 3}
+\addlegendentry{SP14}
+\addlegendimage{empty legend}
+%\legend{Top 1, Top 2, Top 3}
+\end{axis}
+\end{tikzpicture}
 \caption{reuslts!}

### 2017-03-13 paper/oopsla17-submission/evaluation.tex
another graph

@@
 
-\pgfplotstableread{../../data/sp14/baseline.csv}{\FeatureBench}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+\pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
+\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
+\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
+\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
+\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
+\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
+\pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
+\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
 
@@
 \begin{axis}[
-  %ybar stacked,
+  % ybar stacked,
   width=12cm,
   height=8cm,
-  title={Spring 2014},
+  title={Impact of Feature Set on Accuracy},
+  ylabel={Accuracy},
+  %ymin=0.2,
+  ymax=1,
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = true,
+  symbolic x coords={op+type+size, op+context+type+size, op+context-has+type+size, op+context-count+type+size},
+  % enlarge x limits=0.25,
+  xtick=data,
+  xtick style={draw=none},
+  xticklabels={Type, Context, Context-Has, Context-Count},
+  x tick label style={rotate=45},
+  reverse legend,
+  transpose legend,
+  legend style={legend pos = outer north east, legend columns=4},
+]
+% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
+% \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
+% \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
+
+\addplot[mark options={fill=blue1, scale=1.5}, mark=square*]
+  table[x=features, y=top-1] {\FeatureHiddenFHBench};
+\addplot[mark options={fill=blue2, scale=1.5}, mark=square*]
+  table[x=features, y=top-2] {\FeatureHiddenFHBench};
+\addplot[mark options={fill=blue3, scale=1.5}, mark=square*]
+  table[x=features, y=top-3] {\FeatureHiddenFHBench};
+\addlegendentry{Top 1}
+\addlegendentry{Top 2}
+\addlegendentry{Top 3}
+\addlegendimage{empty legend}
+\addlegendentry{\hiddenFH}
+
+\addplot[mark options={fill=blue1, scale=1.5}, mark=*]
+  table[x=features, y=top-1] {\FeatureLinearBench};
+\addplot[mark options={fill=blue2, scale=1.5}, mark=*]
+  table[x=features, y=top-2] {\FeatureLinearBench};
+\addplot[mark options={fill=blue3, scale=1.5}, mark=*]
+  table[x=features, y=top-3] {\FeatureLinearBench};
+\addlegendentry{Top 1}
+\addlegendentry{Top 2}
+\addlegendentry{Top 3}
+\addlegendimage{empty legend}
+\addlegendentry{\linear}
+
+\end{axis}
+\end{tikzpicture}
+\caption{reuslts!}
+\label{fig:results}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+  ybar stacked,
+  width=12cm,
+  height=8cm,
+  title={Impact of Hidden Layer Size on Accuracy},
   ylabel={Accuracy},
@@
 ]
+\addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
+\addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
+\addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
 % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-% \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-\addplot table[x=tool, y=top-1] {\HiddenBench};
-\addplot table[x=tool, y=top-2] {\HiddenBench};
-\addplot table[x=tool, y=top-3] {\HiddenBench};
+% \addplot[draw=black, fill=blue2] table[x=tool, y=top-2] {\HiddenBench};
+% \addplot[draw=black, fill=blue3] table[x=tool, y=top-3] {\HiddenBench};
 \legend{Top 1, Top 2, Top 3}

### 2017-03-13 paper/oopsla17-submission/evaluation.tex
another dimension of shrinking

@@
   xtick style={draw=none},
-  xticklabels={Type, Context, Context-Has, Context-Count},
+  xticklabels={Type, Context-Is, Context-Has, Context-Count},
   x tick label style={rotate=45},

### 2017-03-20 paper/oopsla17-submission/evaluation.tex
some examples for the gallery

@@
 \pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/decision-tree/results.csv}
 \pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
 
@@
 \pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/linear/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/decision-tree/results.csv}
 \pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-500/results.csv}
+\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-250/results.csv}
 
@@
   (ocaml,0) (mycroft,0) (sherrloc,0) (spacer2,0)
-  (op+type+size/linear,0) (op+type+size/hidden-10,0) (op+type+size/hidden-500,0)
+  (op+type+size/linear,0)
+  (op+type+size/decision-tree,0)
+  (op+type+size/hidden-10,0)
+  (op+type+size/hidden-250,0)
 };
@@
   symbolic x coords={baseline, spacer1, ocaml, mycroft, sherrloc, spacer2,
-                     op+type+size/linear, op+type+size/hidden-10,
-                     op+type+size/hidden-500},
+                     op+type+size/linear,
+                     op+type+size/decision-tree,
+                     op+type+size/hidden-10,
+                     op+type+size/hidden-250},
   % enlarge x limits=0.5,
@@
   xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,
-               \linear, \hiddenT, \hiddenFH},
+               \linear, \dectree, \hiddenT, \hiddenTHF},
   x tick label style={rotate=45},
@@
 
+\input{gallery}
+
 %%% Local Variables:

### 2017-03-20 paper/oopsla17-submission/gallery.tex
some examples for the gallery

\subsection{Qualitative Evaluation}
\label{sec:qualitative}

\ES{\hiddenT locations are underlined, \sherrloc locations are bold}
\paragraph{Padding a list}

\begin{ecode}
  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)

  let padZero l1 l2 =
    let l = List.length l1 - List.length l2 in
    if l < 0 then
      (clone 0 ((-1) * l) @ l2, l2)
    else
      (l1, _=clone 0 l=___ :: l2__)
\end{ecode}

\begin{ecode}
  let rec sepConcat sep sl =
    match sl with
    | [] -> ""
    | h::t ->
        let f a x = a ^ (sep ^ x) in
        let base = h in
        let l = t in
        List.fold_left f base l

  let stringOfList f l = sepConcat "; " __[ "["; ___=List.map f l=___; "]" ]__
\end{ecode}

\begin{ecode}
  let rec wwhile (f,b) =
    match f b with
    | (x,false ) -> x
    | (x,true ) -> wwhile (f, x)

  let fixpoint (f,b) =
    wwhile (let g = __let bb = f b in ___=(bb, (bb = b))=_ in (g, b))
\end{ecode}

\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ __[ ___=digitsOfInt (n / 10)=___ ]__
\end{ecode}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2017-03-22 paper/oopsla17-submission/evaluation.tex
ugh, start typign up the eval

@@
 
-\pgfplotstableset{col sep=comma}
+\input{data}
 
-\pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+\subsection{Experimental Setup}
+\label{sec:experimental-setup}
 
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+\subsection{Type Error Slice}
+\label{sec:type-error-slice}
+The \InSlice feature should be highly predictive --- a fix must change
+at least one expression in the type-error slice.
+%
+Thus, our first experiment seeks to quantify the impact of \InSlice by
+comparing the accuracy of a linear model on three sets of features.
+%
+\begin{enumerate}
+\item A baseline with only local syntactic features.
+\item extends (1) with \InSlice.
+\item discards samples from (1) where \InSlice is \emph{disabled}.
+\end{enumerate}
+%
+The key difference between (2) and (3) is that a model for (2) must
+\emph{learn} that \InSlice is a strong predictor.
+%
+In contrast, a model for (3) must only learn about the syntactic
+features, the decision to discard samples where \InSlice is disabled has
+already been made by a human.
+%
+This has a few additional advantages: it reduces the search space by a
+factor of 7 on average, and it guarantees that any prediction made by
+the classifier can fix the type error.
+%
+We expect that (2) will perform better than (1) as it has more
+information, and that (3) will perform better than (2) as it does not
+have to learn the importance of \InSlice.
 
-\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/linear/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
+We tested our hypothesis with a linear model cross-validated ($k=10$)
+over the combined SP14/FA15 dataset. We used a learning rate
+$\alpha=0.001$, L2 regularization rate $\lambda=0.001$, and mini-batch
+size of 200. We trained for a single epoch on feature sets (1) and
+(2), and for 8 epochs on (3), so that the total number of training samples
+would be roughly equal for each feature set.
 
-\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-250/results.csv}
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Feature Set  & Top-1  & Top-2  & Top-3  & Recall \\
+    \midrule
+    Local Syntax & 23.6\% & 42.6\% & 56.3\% & 19.6\% \\
+    With Slice   & 46.4\% & 65.0\% & 75.4\% & 30.4\% \\
+    Only Slice   & 54.9\% & 71.4\% & 82.5\% & 57.6\% \\
+    \bottomrule
+  \end{tabular}
+  \caption{
+    Impact of type-error slice on accuracy.
+    \ES{TODO: expand caption}
+    \ES{TODO: load these numbers from CSV}
+  }\label{tab:type-error-slice}
+\end{table}
+
+\autoref{tab:type-error-slice} shows the results of our experiment.
+%
+As expected the baseline performs the worst, with a mere 23.6\% Top-1
+accuracy.
+%
+Adding \InSlice improves the results substantially with a 46.4\% Top-1
+accuracy, demonstrating the importance of a minimal error slice.
+%
+However, filtering out expressions that are not part of the slice
+\emph{further} improves the results to 54.9\% Top-1 accuracy.
+%
+Clearly, some decisions are too important to be left to a machine.
+
+Note also the jump in Recall when we filter out expressions that are not
+part of the error slice.
+%
+Reducing the search space not only improves our chances of making a
+single correct prediction, it also allows us to make \emph{multiple}
+correct predictions per program.
+
+Given the decisive benefits of filtering out expressions that do not
+belong to the type-error slice, we will assume going forward that
+\emph{all} data sets have been filtered.
+
+\subsection{Contextual Features}
+\label{sec:contextual-features}
+
+Next, we will investigate the relative impact of the other three classes
+of features discussed in \autoref{sec:features}.
+%
+For this we consider again a baseline of only local syntactic features,
+extended by each combination of
+%
+(1) expression size,
+(2) contextual syntactic features, and
+(3) typing features.
+%
+As before we perform a 10-fold cross-validation with a learning rate and
+L2 regularization rate of 0.001 and a mini-batch size of 200, but we
+train for a full 10 epochs.
+
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{lrrrr}
+    \toprule
+    Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+    \midrule
+    Local Syntax                & 55.0\% & 71.6\% & 82.6\% & 57.7\% \\
+    \midrule
+    + Size                      & 55.8\% & 72.8\% & 82.5\% & 57.3\% \\
+    + Contextual Syntax         & 59.9\% & 77.7\% & 86.4\% & 63.0\% \\
+    + Types                     & 62.2\% & 77.7\% & 85.7\% & 62.2\% \\
+    \midrule
+    + Size + Contextual Syntax  & 60.2\% & 77.9\% & 86.0\% & 62.5\% \\
+    + Size + Types              & 62.0\% & 78.2\% & 85.6\% & 62.3\% \\
+    + Contextual Syntax + Types & 63.2\% & 80.3\% & 87.9\% & 65.4\% \\
+    \midrule
+    + All                       & 62.3\% & 80.0\% & 88.0\% & 65.4\% \\
+    \bottomrule
+  \end{tabular}
+  \caption{
+    Impact of contextual features on accuracy.
+    \ES{TODO: expand caption}
+    \ES{TODO: load these numbers from CSV}
+  }\label{tab:type-error-slice}
+\end{table}
 
@@
 
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
+% \begin{figure}[ht]
+% \centering
+% \begin{tikzpicture}
+% \begin{axis}[
+%   % ybar stacked,
+%   width=12cm,
+%   height=8cm,
+%   title={Impact of Feature Set on Accuracy},
+%   ylabel={Accuracy},
+%   %ymin=0.2,
+%   ymax=1,
+%   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+%   ytick style={draw=none},
+%   ymajorgrids = true,
+%   symbolic x coords={op+type+size, op+context+type+size, op+context-has+type+size, op+context-count+type+size},
+%   % enlarge x limits=0.25,
+%   xtick=data,
+%   xtick style={draw=none},
+%   xticklabels={Type, Context-Is, Context-Has, Context-Count},
+%   x tick label style={rotate=45},
+%   reverse legend,
+%   transpose legend,
+%   legend style={legend pos = outer north east, legend columns=4},
+% ]
+% % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
+% % \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
+% % \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
+
+% \addplot[mark options={fill=blue1, scale=1.5}, mark=square*]
+%   table[x=features, y=top-1] {\FeatureHiddenFHBench};
+% \addplot[mark options={fill=blue2, scale=1.5}, mark=square*]
+%   table[x=features, y=top-2] {\FeatureHiddenFHBench};
+% \addplot[mark options={fill=blue3, scale=1.5}, mark=square*]
+%   table[x=features, y=top-3] {\FeatureHiddenFHBench};
+% \addlegendentry{Top 1}
+% \addlegendentry{Top 2}
+% \addlegendentry{Top 3}
+% \addlegendimage{empty legend}
+% \addlegendentry{\hiddenFH}
+
+% \addplot[mark options={fill=blue1, scale=1.5}, mark=*]
+%   table[x=features, y=top-1] {\FeatureLinearBench};
+% \addplot[mark options={fill=blue2, scale=1.5}, mark=*]
+%   table[x=features, y=top-2] {\FeatureLinearBench};
+% \addplot[mark options={fill=blue3, scale=1.5}, mark=*]
+%   table[x=features, y=top-3] {\FeatureLinearBench};
+% \addlegendentry{Top 1}
+% \addlegendentry{Top 2}
+% \addlegendentry{Top 3}
+% \addlegendimage{empty legend}
+% \addlegendentry{\linear}
+
+% \end{axis}
+% \end{tikzpicture}
+% \caption{reuslts!}
+% \label{fig:results}
+% \end{figure}
+
+% \begin{figure}[ht]
+% \centering
+% \begin{tikzpicture}
+% \begin{axis}[
 %   ybar stacked,
-  width=12cm,
-  height=8cm,
-  title={Impact of Feature Set on Accuracy},
-  ylabel={Accuracy},
-  %ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={op+type+size, op+context+type+size, op+context-has+type+size, op+context-count+type+size},
-  % enlarge x limits=0.25,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={Type, Context-Is, Context-Has, Context-Count},
-  x tick label style={rotate=45},
-  reverse legend,
-  transpose legend,
-  legend style={legend pos = outer north east, legend columns=4},
-]
+%   width=12cm,
+%   height=8cm,
+%   title={Impact of Hidden Layer Size on Accuracy},
+%   ylabel={Accuracy},
+%   bar width=20pt,
+%   %ymin=0.2,
+%   ymax=1,
+%   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+%   ytick style={draw=none},
+%   ymajorgrids = true,
+%   symbolic x coords={op+type+size/hidden-10, op+type+size/hidden-25, op+type+size/hidden-50,
+%                      op+type+size/hidden-100, op+type+size/hidden-250, op+type+size/hidden-500},
+%   % enlarge x limits=0.25,
+%   xtick=data,
+%   xtick style={draw=none},
+%   xticklabels={\hiddenT, \hiddenTF, \hiddenF, \hiddenH, \hiddenTHF, \hiddenFH},
+%   x tick label style={rotate=45},
+%   reverse legend,
+%   legend style={legend pos = north west},
+% ]
 % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
@@
 % \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-
-\addplot[mark options={fill=blue1, scale=1.5}, mark=square*]
-  table[x=features, y=top-1] {\FeatureHiddenFHBench};
-\addplot[mark options={fill=blue2, scale=1.5}, mark=square*]
-  table[x=features, y=top-2] {\FeatureHiddenFHBench};
-\addplot[mark options={fill=blue3, scale=1.5}, mark=square*]
-  table[x=features, y=top-3] {\FeatureHiddenFHBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{\hiddenFH}
-
-\addplot[mark options={fill=blue1, scale=1.5}, mark=*]
-  table[x=features, y=top-1] {\FeatureLinearBench};
-\addplot[mark options={fill=blue2, scale=1.5}, mark=*]
-  table[x=features, y=top-2] {\FeatureLinearBench};
-\addplot[mark options={fill=blue3, scale=1.5}, mark=*]
-  table[x=features, y=top-3] {\FeatureLinearBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{\linear}
-
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
-  ybar stacked,
-  width=12cm,
-  height=8cm,
-  title={Impact of Hidden Layer Size on Accuracy},
-  ylabel={Accuracy},
-  bar width=20pt,
-  %ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={op+type+size/hidden-10, op+type+size/hidden-25, op+type+size/hidden-50,
-                     op+type+size/hidden-100, op+type+size/hidden-250, op+type+size/hidden-500},
-  % enlarge x limits=0.25,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={\hiddenT, \hiddenTF, \hiddenF, \hiddenH, \hiddenTHF, \hiddenFH},
-  x tick label style={rotate=45},
-  reverse legend,
-  legend style={legend pos = north west},
-]
-\addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-\addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-\addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% \addplot[draw=black, fill=blue2] table[x=tool, y=top-2] {\HiddenBench};
-% \addplot[draw=black, fill=blue3] table[x=tool, y=top-3] {\HiddenBench};
-\legend{Top 1, Top 2, Top 3}
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
+% % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
+% % \addplot[draw=black, fill=blue2] table[x=tool, y=top-2] {\HiddenBench};
+% % \addplot[draw=black, fill=blue3] table[x=tool, y=top-3] {\HiddenBench};
+% \legend{Top 1, Top 2, Top 3}
+% \end{axis}
+% \end{tikzpicture}
+% \caption{reuslts!}
+% \label{fig:results}
+% \end{figure}
 

### 2017-03-23 paper/oopsla17-submission/evaluation.tex
say something about experimental setup

@@
 \label{sec:evaluation}
+\input{data}
 
-\input{data}
+We have implemented our technique for localizing type errors for a
+purely functional subset of \ocaml with polymorphic types and functions.
+%
+We seek to answer three questions in our evaluation:
+%
+\begin{description}
+\item[1. Feature Utility (\autoref{sec:feature-utility})]
+  %
+  How much do the features described in \autoref{sec:features}
+  contribute to our predictions?
+\item[2. Blame Accuracy (\autoref{sec:quantitative-evaluation})]
+  %
+  How often are our predictions \emph{correct}?
+  % For how many ill-typed programs can we accurately predict the source
+  % of the error?
+\item[3. Explaining Predictions (\autoref{sec:qualitative})]
+  %
+  Can we explain individual predictions (correct or not)?
+\end{description}
+\ES{may want to swap (1) and (2)..}
 
@@
 
-\subsection{Type Error Slice}
+We answer our questions on two sets of data gathered from the
+undergraduate Programming Languages course at
+\if@anonymous
+AUTHOR INSTITUTION.
+\else
+UC San Diego (IRB \#140608).
+\fi
+%
+We recorded each interaction with the \ocaml top-level system over the
+course of the first three assignments, which allowed us to capture
+ill-typed programs and, crucially, their subsequent fixes.
+%
+The first dataset comes from the Spring 2014 class (\SPRING), from which
+we extracted 2,333 pairs of programs, and the second comes from the Fall
+2015 class (\FALL), from which we extracted 2,273 pairs of programs.
+%
+The extracted programs are relatively small, but they demonstrate a
+range of functional programming idioms, \eg higher-order functions and
+(polymorphic) algebraic data types.
+
+We identified the fixes for each ill-typed program with an
+expression-level diff~\citep{XXXX}.
+%
+We consider two sources of changes:
+%
+\begin{enumerate}
+\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
+  is rewritten to $\eapp{g}{x}$, we will consider the expression $f$
+  changed, as it has been replaced by $g$.
+\item If a new expression has been inserted \emph{around} an existing
+  expression, \eg if $\eapp{f}{x}$ is rewritten to
+  $\eplus{\eapp{f}{x}}{1}$, we will consider the application expression
+  $\eapp{f}{x}$ (but not $f$ or $x$) changed, as the $+$ operator now
+  occupies the original location of the application.
+\end{enumerate}
+%
+As the students may have made many, potentially unrelated, changes
+between compilations, we discard any program pairs where more than 40\%
+of the sub-expressions have changed.
+%
+We picked 40\% as an estimate of the inflection point where we could
+still retain the large majority of program pairs.
+
+\subsection{Feature Utility}
+\label{sec:feature-utility}
+
+\subsubsection{Type Error Slice}
 \label{sec:type-error-slice}
@@
 \begin{table}[ht]
+  \begin{minipage}{0.49\linewidth}
   \centering
+  \linear
   \begin{tabular}{lrrrr}
@@
   \end{tabular}
+  \end{minipage}
+  \begin{minipage}{0.49\linewidth}
+  \centering
+  \hiddenF
+  \begin{tabular}{lrrrr}
+    \toprule
+    Feature Set  & Top-1  & Top-2  & Top-3  & Recall \\
+    \midrule
+    Local Syntax & 30.9\% & 47.9\% & 58.6\% & 20.6\% \\
+    With Slice   & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
+    Only Slice   & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
+    \bottomrule
+  \end{tabular}
+  \end{minipage}
   \caption{
@@
 
-\subsection{Contextual Features}
+\subsubsection{Contextual Features}
 \label{sec:contextual-features}
@@
     + Size                      & 55.8\% & 72.8\% & 82.5\% & 57.3\% \\
-    + Contextual Syntax         & 59.9\% & 77.7\% & 86.4\% & 63.0\% \\
+    + Context                   & 59.9\% & 77.7\% & 86.4\% & 63.0\% \\
     + Types                     & 62.2\% & 77.7\% & 85.7\% & 62.2\% \\
     \midrule
-    + Size + Contextual Syntax  & 60.2\% & 77.9\% & 86.0\% & 62.5\% \\
+    + Size + Context            & 60.2\% & 77.9\% & 86.0\% & 62.5\% \\
     + Size + Types              & 62.0\% & 78.2\% & 85.6\% & 62.3\% \\
-    + Contextual Syntax + Types & 63.2\% & 80.3\% & 87.9\% & 65.4\% \\
+    + Context + Types           & 63.2\% & 80.3\% & 87.9\% & 65.4\% \\
     \midrule
@@
   \end{tabular}
+  \end{minipage}
+  % \begin{minipage}{0.49\linewidth}
+  % \centering
+  % \hiddenF
+  % \begin{tabular}{lrrrr}
+  %   \toprule
+  %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+  %   \midrule
+  %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
+  %   \midrule
+  %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
+  %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
+  %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
+  %   \midrule
+  %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
+  %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
+  %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
+  %   \midrule
+  %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
+  %   \bottomrule
+  % \end{tabular}
+  % \end{minipage}
+  % \\
+  \begin{minipage}{0.49\linewidth}
+  \centering
+  \hiddenFH
+  \begin{tabular}{lrrrr}
+    \toprule
+    Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+    \midrule
+    Local Syntax                & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
+    \midrule
+    + Size                      & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
+    + Context                   & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
+    + Types                     & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
+    \midrule
+    + Size + Context            & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
+    + Size + Types              & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
+    + Context + Types           & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
+    \midrule
+    + All                       & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
+    \bottomrule
+  \end{tabular}
+  \end{minipage}
   \caption{

### 2017-03-23 paper/oopsla17-submission/learning.tex
say something about experimental setup

@@
 \begin{minipage}{0.6\linewidth}
-\begin{lstlisting}
+\begin{code}
   let rec sumList xs =
@@
     | hd::tl -> hd + sumList tl
-\end{lstlisting}
+\end{code}
 \end{minipage}
 \begin{minipage}{0.3\linewidth}
-\begin{lstlisting}[numbers=left]
+\begin{code}[numbers=left]
 []
@@
 tl
-\end{lstlisting}
+\end{code}
 \end{minipage}
@@
 \begin{table}[ht]
-\begin{tabular}{lrrrrrrr}
+\begin{tabular}{lrrrrrr}
 \toprule
 Expression
-  & \IsNil & \IsCaseListP & \CountVarP & \ExprSize
+  & \IsNil & \IsCaseListP & \ExprSize
   & \HasTypeIntCOne & \HasTypeList & \InSlice \\
@@
 \lstinline![]!
-  & 1 & 1 & 3 & 1 & 0 & 1 & 1 \\
+  & 1 & 1 & 1 & 0 & 1 & 1 \\
 \lstinline!hd + sumList tl!
-  & 0 & 1 & 3 & 5 & 1 & 0 & 1 \\
+  & 0 & 1 & 5 & 1 & 0 & 1 \\
 \lstinline!sumList tl!
-  & 0 & 0 & 3 & 3 & 0 & 1 & 1 \\
+  & 0 & 0 & 3 & 0 & 1 & 1 \\
 \lstinline!tl!
-  & 0 & 0 & 2 & 1 & 0 & 1 & 0 \\
+  & 0 & 0 & 1 & 0 & 1 & 0 \\
 \bottomrule
@@
 pattern, and thus have a match expression that enables multiple
-features. \ES{This is actually sort of what we do, we just ignore
-  programs that have patterns of different types..}
+features.
 
@@
 
-Instead of just describing the immediate context, we could describe
-whether a particular syntax element occurs in the neighboring
-expressions (or even a count of how many times it occurs).
-%
-For example, the \CountVarP feature in \autoref{tab:sumList} describes
-how many variables are contained in the expression \emph{rooted} at the
-current expression's parent.
-%
-Such fuzzier notions of context may enable increased precision in the
-model, but they also introduce opportunities for \emph{overfitting} ---
-where the model memorizes particular inputs rather than learning general
-patterns.
-%
-We will investigate (\ES{maybe..}) the impact of these alternatives
-in \autoref{sec:evaluation}.
+% Instead of just describing the immediate context, we could describe
+% whether a particular syntax element occurs in the neighboring
+% expressions (or even a count of how many times it occurs).
+% %
+% For example, the \CountVarP feature in \autoref{tab:sumList} describes
+% how many variables are contained in the expression \emph{rooted} at the
+% current expression's parent.
+% %
+% Such fuzzier notions of context may enable increased precision in the
+% model, but they also introduce opportunities for \emph{overfitting} ---
+% where the model memorizes particular inputs rather than learning general
+% patterns.
+% %
+% We will investigate (\ES{maybe..}) the impact of these alternatives
+% in \autoref{sec:evaluation}.
 

### 2017-03-23 paper/oopsla17-submission/evaluation-accuracy.tex
evaluation

\subsection{Blame Accuracy}
\label{sec:quantitative}

In this experiment we compare the accuracy of our predictions to the
state of the art in type-error localization.

\paragraph{Baseline}
We provide two baselines for the comparison: a random choice of location
from a minimized type-error slice, and the \ocaml compiler that
programmers typically interact with.

\paragraph{State of the Art}
\mycroft~\citep{Loncaric2016-uk} localizes type errors by searching for
a minimal subset of typing constraints that can be removed, such that
the resulting system is satisfiable.
%
When multiple such subsets exist it can enumerate them, though it has no
notion of which subsets are \emph{more likely} to be correct, and thus
the order is arbitrary.
%
\sherrloc~\citep{Zhang2014-lv} localizes errors by searching the typing
constraint graph for constraints that participate in many unsatisfiable
paths and few satisfiable paths.
%
It can also enumerate multiple predictions, in descending order of
likelihood.

\paragraph{Classifiers}
We compare five classifiers, each trained on a full set of 276 features:
44 local syntactic features, 176 contextual syntactic features, 55
typing features, and a single expression size feature.
%
\ES{should explain the make-up of these groups}
%
We preemptively discard expressions that are not part of the minimal
type error slice, we will explain the rationale for this in
\autoref{sec:feature-utility}.
%
Our classifiers are:
%
\begin{enumerate}
\item A logistic regression trained with a learning rate
  $\alpha = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
  mini-batch size of 200.
\item A decision tree \ES{HUMA: hyperparams?}.
\item A random forest with 30 estimators \ES{HUMA: hyperparams?}.
\item Two multi-layer perceptrons, both trained with $\alpha = 0.001$,
  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
  contains a single hidden layer of 10 neurons, and the second contains
  a hidden layer of 500 neurons. This allows us to investigate how well
  the MLP can \emph{compress} its model.
\end{enumerate}
%
All classifiers were trained for 20 iterations one dataset before being
evaluated on the other.

\input{evaluation-accuracy-graph}

\autoref{fig:accuracy-results} shows the results of our experiment.

### 2017-03-23 paper/oopsla17-submission/evaluation-utility.tex
evaluation

--- 2017-03-22 paper/oopsla17-submission/evaluation.tex
@@
-\section{Evaluation}
-\label{sec:evaluation}
+\subsection{Feature Utility}
+\label{sec:feature-utility}
 
-\input{data}
-
-\subsection{Experimental Setup}
-\label{sec:experimental-setup}
-
-\subsection{Type Error Slice}
+\subsubsection{Type Error Slice}
 \label{sec:type-error-slice}
@@
 \end{table}
-
-\definecolor{blue1}{HTML}{DEEBF7}
-\definecolor{blue2}{HTML}{9ECAE1}
-\definecolor{blue3}{HTML}{3182BD}
-
-% \begin{figure}[ht]
-% \centering
-% \begin{tikzpicture}
-% \begin{axis}[
-%   % ybar stacked,
-%   width=12cm,
-%   height=8cm,
-%   title={Impact of Feature Set on Accuracy},
-%   ylabel={Accuracy},
-%   %ymin=0.2,
-%   ymax=1,
-%   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-%   ytick style={draw=none},
-%   ymajorgrids = true,
-%   symbolic x coords={op+type+size, op+context+type+size, op+context-has+type+size, op+context-count+type+size},
-%   % enlarge x limits=0.25,
-%   xtick=data,
-%   xtick style={draw=none},
-%   xticklabels={Type, Context-Is, Context-Has, Context-Count},
-%   x tick label style={rotate=45},
-%   reverse legend,
-%   transpose legend,
-%   legend style={legend pos = outer north east, legend columns=4},
-% ]
-% % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% % \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-% % \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-
-% \addplot[mark options={fill=blue1, scale=1.5}, mark=square*]
-%   table[x=features, y=top-1] {\FeatureHiddenFHBench};
-% \addplot[mark options={fill=blue2, scale=1.5}, mark=square*]
-%   table[x=features, y=top-2] {\FeatureHiddenFHBench};
-% \addplot[mark options={fill=blue3, scale=1.5}, mark=square*]
-%   table[x=features, y=top-3] {\FeatureHiddenFHBench};
-% \addlegendentry{Top 1}
-% \addlegendentry{Top 2}
-% \addlegendentry{Top 3}
-% \addlegendimage{empty legend}
-% \addlegendentry{\hiddenFH}
-
-% \addplot[mark options={fill=blue1, scale=1.5}, mark=*]
-%   table[x=features, y=top-1] {\FeatureLinearBench};
-% \addplot[mark options={fill=blue2, scale=1.5}, mark=*]
-%   table[x=features, y=top-2] {\FeatureLinearBench};
-% \addplot[mark options={fill=blue3, scale=1.5}, mark=*]
-%   table[x=features, y=top-3] {\FeatureLinearBench};
-% \addlegendentry{Top 1}
-% \addlegendentry{Top 2}
-% \addlegendentry{Top 3}
-% \addlegendimage{empty legend}
-% \addlegendentry{\linear}
-
-% \end{axis}
-% \end{tikzpicture}
-% \caption{reuslts!}
-% \label{fig:results}
-% \end{figure}
-
-% \begin{figure}[ht]
-% \centering
-% \begin{tikzpicture}
-% \begin{axis}[
-%   ybar stacked,
-%   width=12cm,
-%   height=8cm,
-%   title={Impact of Hidden Layer Size on Accuracy},
-%   ylabel={Accuracy},
-%   bar width=20pt,
-%   %ymin=0.2,
-%   ymax=1,
-%   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-%   ytick style={draw=none},
-%   ymajorgrids = true,
-%   symbolic x coords={op+type+size/hidden-10, op+type+size/hidden-25, op+type+size/hidden-50,
-%                      op+type+size/hidden-100, op+type+size/hidden-250, op+type+size/hidden-500},
-%   % enlarge x limits=0.25,
-%   xtick=data,
-%   xtick style={draw=none},
-%   xticklabels={\hiddenT, \hiddenTF, \hiddenF, \hiddenH, \hiddenTHF, \hiddenFH},
-%   x tick label style={rotate=45},
-%   reverse legend,
-%   legend style={legend pos = north west},
-% ]
-% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-% \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-% % \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% % \addplot[draw=black, fill=blue2] table[x=tool, y=top-2] {\HiddenBench};
-% % \addplot[draw=black, fill=blue3] table[x=tool, y=top-3] {\HiddenBench};
-% \legend{Top 1, Top 2, Top 3}
-% \end{axis}
-% \end{tikzpicture}
-% \caption{reuslts!}
-% \label{fig:results}
-% \end{figure}
-
-\makeatletter
-\newcommand\resetstackedplots{
-\makeatletter
-\pgfplots@stacked@isfirstplottrue
-\makeatother
-\addplot [forget plot,draw=none] coordinates{
-  (baseline,0) (spacer1,0)
-  (ocaml,0) (mycroft,0) (sherrloc,0) (spacer2,0)
-  (op+type+size/linear,0)
-  (op+type+size/decision-tree,0)
-  (op+type+size/hidden-10,0)
-  (op+type+size/hidden-250,0)
-};
-}
-\makeatother
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
-  ybar stacked,
-  width=14cm,
-  height=8cm,
-  title={Comparison of Type-Error Localization Techniques},
-  ylabel={Accuracy},
-  bar width=0.5cm,
-  ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={baseline, spacer1, ocaml, mycroft, sherrloc, spacer2,
-                     op+type+size/linear,
-                     op+type+size/decision-tree,
-                     op+type+size/hidden-10,
-                     op+type+size/hidden-250},
-  % enlarge x limits=0.5,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,
-               \linear, \dectree, \hiddenT, \hiddenTHF},
-  x tick label style={rotate=45},
-  reverse legend,
-  transpose legend,
-  legend style={legend pos = north west, legend columns=4},
-]
-\addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=tool, y=top-1] {\SpringBench};
-\addplot[draw=black, fill=blue2, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\SpringBench};
-\addplot[draw=black, fill=blue3, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\SpringBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{FA15}
-
-\resetstackedplots
-
-\addplot[draw=black, fill=blue1, bar shift=.25cm] table[x=tool, y=top-1] {\FallBench};
-\addplot[draw=black, fill=blue2, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\FallBench};
-\addplot[draw=black, fill=blue3, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\FallBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendentry{SP14}
-\addlegendimage{empty legend}
-%\legend{Top 1, Top 2, Top 3}
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
-
-\input{gallery}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "main"
-%%% End:

### 2017-03-23 paper/oopsla17-submission/evaluation.tex
evaluation

--- 2017-03-20 paper/oopsla17-submission/evaluation.tex
@@
 \label{sec:evaluation}
+\input{data}
 
-\pgfplotstableset{col sep=comma}
+We have implemented our technique for localizing type errors for a
+purely functional subset of \ocaml with polymorphic types and functions.
+%
+We seek to answer three questions in our evaluation:
+%
+\begin{description}
+\item[1. Blame Accuracy (\autoref{sec:quantitative})]
+  %
+  How often are our predictions \emph{correct}?
+  % For how many ill-typed programs can we accurately predict the source
+  % of the error?
+\item[2. Feature Utility (\autoref{sec:feature-utility})]
+  %
+  Which feature sets are most important in general?
+  % How much do the features described in \autoref{sec:features}
+  % contribute to our predictions?
+\item[3. Explaining Predictions (\autoref{sec:qualitative})]
+  %
+  Can we explain \emph{individual} predictions (correct or not)?
+\end{description}
+\ES{may want to swap (1) and (2)..}
 
-\pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
-\pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+\subsection{Experimental Setup}
+\label{sec:experimental-setup}
 
-\pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
-\pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
-% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}
+We answer our questions on two sets of data gathered from the
+undergraduate Programming Languages course at
+\if@anonymous
+AUTHOR INSTITUTION.
+\else
+UC San Diego (IRB \#140608).
+\fi
+%
+We recorded each interaction with the \ocaml top-level system over the
+course of the first three assignments, which allowed us to capture
+ill-typed programs and, crucially, their subsequent fixes.
+%
+The first dataset comes from the Spring 2014 class (\SPRING), from which
+we extracted 2,333 pairs of programs, and the second comes from the Fall
+2015 class (\FALL), from which we extracted 2,273 pairs of programs.
+%
+The extracted programs are relatively small, but they demonstrate a
+range of functional programming idioms, \eg higher-order functions and
+(polymorphic) algebraic data types.
 
-\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/linear/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
+We identified the fixes for each ill-typed program with an
+expression-level diff~\citep{Lempsink2009-xf}.
+%
+We consider two sources of changes:
+%
+\begin{enumerate}
+\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
+  is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
+  changed, as it has been replaced by $g$.
+\item If a new expression has been inserted \emph{around} an existing
+  expression, \eg if $\eapp{f}{x}$ is rewritten to
+  $\eplus{\eapp{f}{x}}{1}$, we will mark the application expression
+  $\eapp{f}{x}$ (but not $f$ or $x$) as changed, as the $+$ operator now
+  occupies the original location of the application.
+\end{enumerate}
+%
+As the students may have made many, potentially unrelated, changes
+between compilations, we discard any program pairs where more than 40\%
+of the sub-expressions have changed.
+%
+We picked 40\% as an estimate of the inflection point where we could
+still retain the large majority of program pairs.
 
-\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+type+size/hidden-250/results.csv}
-
-\definecolor{blue1}{HTML}{DEEBF7}
-\definecolor{blue2}{HTML}{9ECAE1}
-\definecolor{blue3}{HTML}{3182BD}
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
-  % ybar stacked,
-  width=12cm,
-  height=8cm,
-  title={Impact of Feature Set on Accuracy},
-  ylabel={Accuracy},
-  %ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={op+type+size, op+context+type+size, op+context-has+type+size, op+context-count+type+size},
-  % enlarge x limits=0.25,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={Type, Context-Is, Context-Has, Context-Count},
-  x tick label style={rotate=45},
-  reverse legend,
-  transpose legend,
-  legend style={legend pos = outer north east, legend columns=4},
-]
-% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% \addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-% \addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-
-\addplot[mark options={fill=blue1, scale=1.5}, mark=square*]
-  table[x=features, y=top-1] {\FeatureHiddenFHBench};
-\addplot[mark options={fill=blue2, scale=1.5}, mark=square*]
-  table[x=features, y=top-2] {\FeatureHiddenFHBench};
-\addplot[mark options={fill=blue3, scale=1.5}, mark=square*]
-  table[x=features, y=top-3] {\FeatureHiddenFHBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{\hiddenFH}
-
-\addplot[mark options={fill=blue1, scale=1.5}, mark=*]
-  table[x=features, y=top-1] {\FeatureLinearBench};
-\addplot[mark options={fill=blue2, scale=1.5}, mark=*]
-  table[x=features, y=top-2] {\FeatureLinearBench};
-\addplot[mark options={fill=blue3, scale=1.5}, mark=*]
-  table[x=features, y=top-3] {\FeatureLinearBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{\linear}
-
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
-  ybar stacked,
-  width=12cm,
-  height=8cm,
-  title={Impact of Hidden Layer Size on Accuracy},
-  ylabel={Accuracy},
-  bar width=20pt,
-  %ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={op+type+size/hidden-10, op+type+size/hidden-25, op+type+size/hidden-50,
-                     op+type+size/hidden-100, op+type+size/hidden-250, op+type+size/hidden-500},
-  % enlarge x limits=0.25,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={\hiddenT, \hiddenTF, \hiddenF, \hiddenH, \hiddenTHF, \hiddenFH},
-  x tick label style={rotate=45},
-  reverse legend,
-  legend style={legend pos = north west},
-]
-\addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-\addplot[draw=black, fill=blue2] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\HiddenBench};
-\addplot[draw=black, fill=blue3] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\HiddenBench};
-% \addplot[draw=black, fill=blue1] table[x=tool, y=top-1] {\HiddenBench};
-% \addplot[draw=black, fill=blue2] table[x=tool, y=top-2] {\HiddenBench};
-% \addplot[draw=black, fill=blue3] table[x=tool, y=top-3] {\HiddenBench};
-\legend{Top 1, Top 2, Top 3}
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
-
-\makeatletter
-\newcommand\resetstackedplots{
-\makeatletter
-\pgfplots@stacked@isfirstplottrue
-\makeatother
-\addplot [forget plot,draw=none] coordinates{
-  (baseline,0) (spacer1,0)
-  (ocaml,0) (mycroft,0) (sherrloc,0) (spacer2,0)
-  (op+type+size/linear,0)
-  (op+type+size/decision-tree,0)
-  (op+type+size/hidden-10,0)
-  (op+type+size/hidden-250,0)
-};
-}
-\makeatother
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-\begin{axis}[
-  ybar stacked,
-  width=14cm,
-  height=8cm,
-  title={Comparison of Type-Error Localization Techniques},
-  ylabel={Accuracy},
-  bar width=0.5cm,
-  ymin=0.2,
-  ymax=1,
-  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
-  ytick style={draw=none},
-  ymajorgrids = true,
-  symbolic x coords={baseline, spacer1, ocaml, mycroft, sherrloc, spacer2,
-                     op+type+size/linear,
-                     op+type+size/decision-tree,
-                     op+type+size/hidden-10,
-                     op+type+size/hidden-250},
-  % enlarge x limits=0.5,
-  xtick=data,
-  xtick style={draw=none},
-  xticklabels={\baseline, \ocaml, \mycroft, \sherrloc,
-               \linear, \dectree, \hiddenT, \hiddenTHF},
-  x tick label style={rotate=45},
-  reverse legend,
-  transpose legend,
-  legend style={legend pos = north west, legend columns=4},
-]
-\addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=tool, y=top-1] {\SpringBench};
-\addplot[draw=black, fill=blue2, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\SpringBench};
-\addplot[draw=black, fill=blue3, bar shift=-.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\SpringBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendimage{empty legend}
-\addlegendentry{FA15}
-
-\resetstackedplots
-
-\addplot[draw=black, fill=blue1, bar shift=.25cm] table[x=tool, y=top-1] {\FallBench};
-\addplot[draw=black, fill=blue2, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-2} - \thisrow{top-1}] {\FallBench};
-\addplot[draw=black, fill=blue3, bar shift=.25cm] table[x=tool, y expr=\thisrow{top-3} - \thisrow{top-2}] {\FallBench};
-\addlegendentry{Top 1}
-\addlegendentry{Top 2}
-\addlegendentry{Top 3}
-\addlegendentry{SP14}
-\addlegendimage{empty legend}
-%\legend{Top 1, Top 2, Top 3}
-\end{axis}
-\end{tikzpicture}
-\caption{reuslts!}
-\label{fig:results}
-\end{figure}
+\input{evaluation-accuracy}
+\input{evaluation-utility}
 

### 2017-03-24 paper/oopsla17-submission/evaluation-accuracy.tex
more eval

@@
 
-\paragraph{Classifiers}
-We compare five classifiers, each trained on a full set of 276 features:
-44 local syntactic features, 176 contextual syntactic features, 55
-typing features, and a single expression size feature.
+Comparing source locations from multiple tools with their own parsers is
+not trivial.
+%
+In order to ensure a fair comparison when evaluating \mycroft and
+\sherrloc, we ignored program where they predicted locations that our
+oracle could not match with a program expression --- 6--8\% of programs
+for \mycroft and 3--4\% for \sherrloc.
+%
+We also ignored programs where \mycroft or \sherrloc timed out (after
+one minute) or where they encountered an unsupported language feature ---
+another 5\% for \mycroft and 12--13\% for \sherrloc.
+
+\paragraph{Our Classifiers}
+We evaluate five classifiers, each trained on the full set of 276
+features: 44 local syntactic features, 176 contextual syntactic
+features, 55 typing features, and a single expression size feature.
 %
@@
 %
-All classifiers were trained for 20 iterations one dataset before being
-evaluated on the other.
+All classifiers were trained for 20 iterations on one dataset before
+being evaluated on the other.
 
@@
 
+\paragraph{Results}
 \autoref{fig:accuracy-results} shows the results of our experiment.
+%
+Localizing the type errors in our benchmarks amounted on average to
+selecting one of 2--3 correct locations out of a slice of 10--11.
+%
+Our baseline of selecting at random is able to achieve 30\% Top 1
+accuracy (58\% Top 3), while \ocaml achieves a Top 1 accuracy of 45\%.
+%
+Interestingly, one only needs two \emph{random} guesses to outperform
+\ocaml, with 47\% accuracy.
+%
+\sherrloc outperforms both baselines with 56\% Top 1 accuracy (84\% Top
+3), while \mycroft actually underperforms \ocaml with 38--41\% Top 1
+accuracy.
+%
+Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
+ranging from 60--62\% Top 1 accuracy (85--88\% Top 3) for the \linear
+classifier to 71--73\% Top 1 accuracy (91\% Top 3) for the \hiddenFH.
+
+Interestingly there is little variation in accuracy between classifiers
+with the exception of the \linear model, they all achieve around 70\%
+Top 1 accuracy and around 90\% Top 3 accuracy.
+%
+In particular, the \hiddenT only loses around 2\% accuracy compared to
+the \hiddenFH, which suggests that it can substantially compress the
+model.
+%
+We also note that our classifiers consistently perform better when
+trained on the \FALL programs and tested on the \SPRING programs than
+vice versa, they appear to generalize better from the \FALL data.

### 2017-03-24 paper/oopsla17-submission/evaluation-utility.tex
more eval

@@
 \label{sec:feature-utility}
+We have shown that we can train a classifier to effectively localize
+type errors, but which of the 276 features that we use are contributing
+the most?
+%
+In order to answer this question we investigate the performance of
+classifiers trained on various subsets of the features.
 
@@
 Thus, our first experiment seeks to quantify the impact of \InSlice by
-comparing the accuracy of a linear model on three sets of features.
+comparing the accuracy of a linear model on three sets of features:
 %
@@
 \item A baseline with only local syntactic features.
-\item extends (1) with \InSlice.
-\item discards samples from (1) where \InSlice is \emph{disabled}.
+\item The local syntactic features of (1) extended with \InSlice.
+\item The same features as (1), but we preemptively discard samples
+  where \InSlice is \emph{disabled}.
 \end{enumerate}
 %
-The key difference between (2) and (3) is that a model for (2) must
+The key difference between (2) and (3) is that a classifier for (2) must
 \emph{learn} that \InSlice is a strong predictor.
 %
-In contrast, a model for (3) must only learn about the syntactic
+In contrast, a classifier for (3) must only learn about the syntactic
 features, the decision to discard samples where \InSlice is disabled has
@@
 %
-This has a few additional advantages: it reduces the search space by a
-factor of 7 on average, and it guarantees that any prediction made by
-the classifier can fix the type error.
+This has a few additional advantages: it reduces the set of candidate
+locations by a factor of 7 on average, and it guarantees that any
+prediction made by the classifier can fix the type error.
 %
-We expect that (2) will perform better than (1) as it has more
-information, and that (3) will perform better than (2) as it does not
-have to learn the importance of \InSlice.
+We expect that (2) will perform better than (1) as it contains more
+information, and that (3) will perform better than (2) as the classifier
+does not have to learn the importance of \InSlice.
 
@@
 \begin{table}[ht]
-  \begin{minipage}{0.49\linewidth}
   \centering
-  \linear
-  \begin{tabular}{lrrrr}
+  \begin{tabular}{lrcrrrrcrrrr}
     \toprule
-    Feature Set  & Top-1  & Top-2  & Top-3  & Recall \\
+                 &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenF       \\
+                                   \cmidrule{4-7}                        \cmidrule{9-12}
+    Feature Set  & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
     \midrule
-    Local Syntax & 23.6\% & 42.6\% & 56.3\% & 19.6\% \\
-    With Slice   & 46.4\% & 65.0\% & 75.4\% & 30.4\% \\
-    Only Slice   & 54.9\% & 71.4\% & 82.5\% & 57.6\% \\
+    Local Syntax & 47          & & 23.6\% & 42.6\% & 56.3\% & 19.6\% & & 30.9\% & 47.9\% & 58.6\% & 20.6\% \\
+    + Slice      & 48          & & 46.4\% & 65.0\% & 75.4\% & 30.4\% & & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
+    Only Slice   & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
     \bottomrule
   \end{tabular}
-  \end{minipage}
-  \begin{minipage}{0.49\linewidth}
-  \centering
-  \hiddenF
-  \begin{tabular}{lrrrr}
-    \toprule
-    Feature Set  & Top-1  & Top-2  & Top-3  & Recall \\
-    \midrule
-    Local Syntax & 30.9\% & 47.9\% & 58.6\% & 20.6\% \\
-    With Slice   & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
-    Only Slice   & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
-    \bottomrule
-  \end{tabular}
-  \end{minipage}
   \caption{
@@
 \begin{table}[ht]
-  \begin{minipage}{0.49\linewidth}
   \centering
-  \linear
-  \begin{tabular}{lrrrr}
+  \begin{tabular}{lrcrrrrcrrrr}
     \toprule
-    Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+                             &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
+                                               \cmidrule{4-7}                        \cmidrule{9-12}
+    Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
     \midrule
-    Local Syntax                & 55.0\% & 71.6\% & 82.6\% & 57.7\% \\
+    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
     \midrule
-    + Size                      & 55.8\% & 72.8\% & 82.5\% & 57.3\% \\
-    + Context                   & 59.9\% & 77.7\% & 86.4\% & 63.0\% \\
-    + Types                     & 62.2\% & 77.7\% & 85.7\% & 62.2\% \\
+    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
+    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
+    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
     \midrule
-    + Size + Context            & 60.2\% & 77.9\% & 86.0\% & 62.5\% \\
-    + Size + Types              & 62.0\% & 78.2\% & 85.6\% & 62.3\% \\
-    + Context + Types           & 63.2\% & 80.3\% & 87.9\% & 65.4\% \\
+    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
+    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
+    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
     \midrule
-    + All                       & 62.3\% & 80.0\% & 88.0\% & 65.4\% \\
+    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
     \bottomrule
   \end{tabular}
-  \end{minipage}
   % \begin{minipage}{0.49\linewidth}
@@
   % \\
-  \begin{minipage}{0.49\linewidth}
-  \centering
-  \hiddenFH
-  \begin{tabular}{lrrrr}
-    \toprule
-    Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
-    \midrule
-    Local Syntax                & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
-    \midrule
-    + Size                      & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
-    + Context                   & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
-    + Types                     & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
-    \midrule
-    + Size + Context            & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
-    + Size + Types              & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
-    + Context + Types           & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
-    \midrule
-    + All                       & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
-    \bottomrule
-  \end{tabular}
-  \end{minipage}
+  % \begin{minipage}{0.49\linewidth}
+  % \centering
+  % \hiddenFH
+  % \begin{tabular}{lrrrr}
+  %   \toprule
+  %   Feature Set                 \\
+  %   \midrule
+  %   Local Syntax                \\
+  %   \midrule
+  %   + Size                      \\
+  %   + Context                   \\
+  %   + Types                     \\
+  %   \midrule
+  %   + Size + Context            \\
+  %   + Size + Types              \\
+  %   + Context + Types           \\
+  %   \midrule
+  %   + All                       \\
+  %   \bottomrule
+  % \end{tabular}
+  % \end{minipage}
   \caption{

### 2017-03-24 paper/oopsla17-submission/evaluation.tex
more eval

@@
 
-\subsection{Experimental Setup}
-\label{sec:experimental-setup}
+\subsection{Methodology}
+\label{sec:methodology}
 
@@
 
+\paragraph{Accuracy Metric}
+All of the tools we compare (with the exception of \ocaml) can produce a
+list of potential error locations, which might in the limit include the
+entire type error slice.
+%
+However, in a study of fault localization techniques,
+\citet{Kochhar2016-oc} show that most developers will not consider more
+than around five potential error locations before falling back to manual
+debugging.
+%
+Type errors are relatively simple in comparison to general fault
+localization, thus we limit our evaluation to the top three predictions
+of each tool.
+%
+We evaluate each tool on whether a changed expression occurred in its
+top 1, top 2, or top 3 predictions.
+
 \input{evaluation-accuracy}

### 2017-03-24 paper/oopsla17-submission/evaluation-accuracy.tex
more eval

@@
 %
-All classifiers were trained for 20 iterations on one dataset before
-being evaluated on the other.
+All classifiers were trained for 20 epochs on one dataset
+--- \ie they were shown each program 20 times ---
+before being evaluated on the other.
 
@@
 %
-Our baseline of selecting at random is able to achieve 30\% Top 1
-accuracy (58\% Top 3), while \ocaml achieves a Top 1 accuracy of 45\%.
+Our baseline of selecting at random is able to achieve 30\% Top-1
+accuracy (58\% Top-3), while \ocaml achieves a Top-1 accuracy of 45\%.
 %
@@
 %
-\sherrloc outperforms both baselines with 56\% Top 1 accuracy (84\% Top
-3), while \mycroft actually underperforms \ocaml with 38--41\% Top 1
+\sherrloc outperforms both baselines with 56\% Top-1 accuracy (84\% Top
+3), while \mycroft actually underperforms \ocaml with 38--41\% Top-1
 accuracy.
@@
 Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
-ranging from 60--62\% Top 1 accuracy (85--88\% Top 3) for the \linear
-classifier to 71--73\% Top 1 accuracy (91\% Top 3) for the \hiddenFH.
+ranging from 60--62\% Top-1 accuracy (85--88\% Top-3) for the \linear
+classifier to 71--73\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
 
@@
 with the exception of the \linear model, they all achieve around 70\%
-Top 1 accuracy and around 90\% Top 3 accuracy.
+Top-1 accuracy and around 90\% Top-3 accuracy.
 %

### 2017-03-24 paper/oopsla17-submission/evaluation-utility.tex
more eval

@@
 \item A baseline with only local syntactic features.
-\item The local syntactic features of (1) extended with \InSlice.
+\item The features of (1) extended with \InSlice.
 \item The same features as (1), but we preemptively discard samples
@@
 would be roughly equal for each feature set.
-
+\ES{introduce the MLP too}
+\ES{use MLP-500 instead?}
 \begin{table}[ht]
@@
     Local Syntax       & 47          & & 23.6\% & 42.6\% & 56.3\% & 19.6\% & & 30.9\% & 47.9\% & 58.6\% & 20.6\% \\
-    + Slice      & 48          & & 46.4\% & 65.0\% & 75.4\% & 30.4\% & & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
-    Only Slice   & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
+    + \InSlice         & 48          & & 46.4\% & 65.0\% & 75.4\% & 30.4\% & & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
+    Filter by \InSlice & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
     \bottomrule
@@
 Given the decisive benefits of filtering out expressions that do not
-belong to the type-error slice, we will assume going forward that
-\emph{all} data sets have been filtered.
+belong to the type-error slice, we choose to filter all programs by
+\InSlice.
 
@@
 %
-As before we perform a 10-fold cross-validation with a learning rate and
-L2 regularization rate of 0.001 and a mini-batch size of 200, but we
-train for a full 10 epochs.
+As before we perform a 10-fold cross-validation with $\alpha = 0.001$,
+$\lambda = 0.001$, and a mini-batch size of 200, but we
+train for a full 20 epochs.
 
@@
   % \end{minipage}
-  % \\
-  % \begin{minipage}{0.49\linewidth}
-  % \centering
-  % \hiddenFH
-  % \begin{tabular}{lrrrr}
-  %   \toprule
-  %   Feature Set                 \\
-  %   \midrule
-  %   Local Syntax                \\
-  %   \midrule
-  %   + Size                      \\
-  %   + Context                   \\
-  %   + Types                     \\
-  %   \midrule
-  %   + Size + Context            \\
-  %   + Size + Types              \\
-  %   + Context + Types           \\
-  %   \midrule
-  %   + All                       \\
-  %   \bottomrule
-  % \end{tabular}
-  % \end{minipage}
   \caption{
@@
     \ES{TODO: load these numbers from CSV}
-  }\label{tab:type-error-slice}
+  }\label{tab:contextual-features}
 \end{table}
+
+\autoref{tab:contextual-features} summarizes the results of our experiment.
+%
+As we can see, the \linear classifier and the \hiddenFH start off
+competitive when given only local syntactic features, but the \hiddenFH
+quickly begins to outperform as more features are added.
+%
+\ES{HEREHEREHERHE}

### 2017-03-24 paper/oopsla17-submission/introduction.tex
more eval

@@
 \begin{enumerate}
-\item An experimental evaluation of state-of-the-art type-error
-  localization techniques based not on expert \emph{opinion}, but
-  on the eventual \emph{fix} implemented by a novice user.
+\item An large-scale evaluation of state-of-the-art type error
+  localization techniques on over 4,500 student programs.
   %
-  \ES{need to explain \emph{why} this is superior..}
+  We collected a time-series of \ocaml programs from our students, and
+  use their subsequent \emph{fixes} to ill-typed programs as
+  \emph{oracles} for the location of the error.
+  %
+  This is the most extensive evaluation of the accuracy of type error
+  localization techniques that we are aware of.
+  % based not on expert \emph{opinion}, but
+  % on the eventual \emph{fix} implemented by a novice user.
+  %
+  % \ES{need to explain \emph{why} this is superior..}
 \item A machine learning approach to \emph{modeling} (novice) type
-  errors that can outperform the state-of-the-art by 10--30\% in
+  errors that can outperform the state-of-the-art by 15--30\% in
   localizing errors.
   %
-  Our model can be trained on a relatively small amount of data --- we
-  used a single class of around 50 students --- and appears to
+  Our classifiers can be trained on a relatively small amount of data --- we
+  used a single class of around 50 students --- and appear to
   generalize to other instances of the same class.

### 2017-03-24 paper/oopsla17-submission/evaluation-accuracy.tex
minor

@@
   mini-batch size of 200.
-\item A decision tree \ES{HUMA: hyperparams?}.
-\item A random forest with 30 estimators \ES{HUMA: hyperparams?}.
+\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
+\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
 \item Two multi-layer perceptrons, both trained with $\alpha = 0.001$,
@@
 
-Interestingly there is little variation in accuracy between classifiers
+Surprisingly, there is little variation in accuracy between classifiers
 with the exception of the \linear model, they all achieve around 70\%
@@
 %
-In particular, the \hiddenT only loses around 2\% accuracy compared to
-the \hiddenFH, which suggests that it can substantially compress the
-model.
+This would appear to suggest that the model they learn is relatively
+simple.
+%
+In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
+hidden features than the \hiddenFH, it only loses around 2\% accuracy.
+% In particular, notice that the \hiddenT only loses around 2\% accuracy
+% compared to the \hiddenFH,
 %

### 2017-03-24 paper/oopsla17-submission/evaluation-utility.tex
more eval

@@
 
+\paragraph{Results}
 \autoref{tab:contextual-features} summarizes the results of our experiment.
@@
 quickly begins to outperform as more features are added.
+
+\ExprSize appears to be the weakest feature, improving \linear Top-1
+accuracy by only 1\% and \hiddenFH by only 4\%.
 %
-\ES{HEREHEREHERHE}
+In contrast, the contextual syntactic features improve \linear Top-1
+accuracy by 5\% (\resp 15\%), and the typing features improve
+Top-1 accuracy by 7\% (\resp 17\%).
+%
+Furthermore, while \ExprSize does provide some benefit when it is the
+only additional feature, it does not appear to provide any real increase
+in accuracy when added alongside the contextual or typing features.
+
+As one might expect, the typing features are more beneficial than the
+contextual syntactic features.
+%
+They improve Top-1 accuracy by an additional 2\%, and are much more
+compact --- we only have 55 typing features compared to 176 contextual
+syntactic features.
+%
+This aligns with our intuition that types should be a good summary of
+the context of an expression.
+%
+However, typing features do not \emph{subsume} contextual syntactic
+features, we can gain an additional 1\% Top-1 accuracy (\resp 4\%) by
+adding \emph{both}.

### 2017-03-24 paper/oopsla17-submission/gallery.tex
qualitative

@@
 
-\ES{\hiddenT locations are underlined, \sherrloc locations are bold}
-\paragraph{Padding a list}
+Next, we present a \emph{qualitative} evaluation that compares the
+predictions made by our classifiers with those of \sherrloc.
+%
+In particular, we demonstrate, with a series of example programs from
+our student dataset, how our classifiers are able to use past student
+mistakes to make more accurate predictions of future fixes.
+%
+For each example, we provide
+%
+(1) the code,
+%
+(2) \sherrloc's prediction (in \textbf{bold}), and
+%
+(3) the prediction of our Decision Tree (\underline{underlined}).
+%
+We choose the Decision Tree classifier for this section as its model
+is more easily interpreted than the MLP.
+%
+We will also attempt, for each program, to explain \emph{why} the
+Decision Tree made its prediction, by analyzing the paths induced
+by the programs.
 
-\begin{ecode}
-  let rec clone x n =
-    if n <= 0 then
-      []
-    else
-      x :: clone x (n - 1)
-
-  let padZero l1 l2 =
-    let l = List.length l1 - List.length l2 in
-    if l < 0 then
-      (clone 0 ((-1) * l) @ l2, l2)
-    else
-      (l1, _=clone 0 l=___ :: l2__)
-\end{ecode}
-
-\begin{ecode}
-  let rec sepConcat sep sl =
-    match sl with
-    | [] -> ""
-    | h::t ->
-        let f a x = a ^ (sep ^ x) in
-        let base = h in
-        let l = t in
-        List.fold_left f base l
-
-  let stringOfList f l = sepConcat "; " __[ "["; ___=List.map f l=___; "]" ]__
-\end{ecode}
-
-\begin{ecode}
-  let rec wwhile (f,b) =
-    match f b with
-    | (x,false ) -> x
-    | (x,true ) -> wwhile (f, x)
-
-  let fixpoint (f,b) =
-    wwhile (let g = __let bb = f b in ___=(bb, (bb = b))=_ in (g, b))
-\end{ecode}
-
+\paragraph{Extracting the Digits of an Integer}
+Our first program is a simple recursive function |digitsOfInt| that
+extracts the digits of an |int|.
+%
 \begin{ecode}
@@
 \end{ecode}
+%
+Unfortunately, the student has decided to wrap the recursive call to
+|digitsOfInt| with a list literal, even though |digitsOfInt| already
+returns an |int list|.
+%
+Thus, the list literal is inferred to have type |int list list|, which
+is incompatible with the |int list| on the left of the |@| (list append)
+operator.
+%
+Both \sherrloc and the \ocaml compiler blame the recursive call for
+returning a |int list| rather than |int|, but the recursive call is
+actually correct!
+%
+As our Decision Tree correctly points out, the fault lies with the list
+literal \emph{surrounding} the recursive call, remove it and the type
+error disappears.
+%
+\ES{TODO: WHY?}
+
+\paragraph{Padding a list}
+Our next program, |padZero|, is given two |int list|s as input, and must
+left-pad the shorter one with enough zeros that the two output lists
+have equal length.
+
+The student first defines a helper function |clone|, which takes an
+input element |x| and an |int| |n|, and returns a list containing |n|
+|x|s.
+%
+\begin{ecode}
+  let rec clone x n =
+    if n <= 0 then
+      []
+    else
+      x :: clone x (n - 1)
+\end{ecode}
+%
+Then she defines |padZero| with a simple branch that determines which
+list is shorter, followed by a |clone| with the appropriate number of
+zeros.
+%
+\lstset{firstnumber=last}
+\begin{ecode}
+  let padZero l1 l2 =
+    let n = List.length l1 - List.length l2 in
+    if n < 0 then
+      (clone 0 ((-1) * n) @ l2, l2)
+    else
+      (l1, _=clone 0 n=___ :: l2__)
+\end{ecode}
+\lstset{firstnumber=1}
+%
+Alas, our student has accidentally used the |::| operator rather than
+the |@| operator in the |else| branch.
+%
+\sherrloc and \ocaml correctly determine that she cannot cons the
+|int list| returned by |clone| onto |l2|, which is another |int list|,
+but they decide to \emph{blame} the call to |clone|, while our Decision
+Tree correctly blames the |::| operator.
+%
+\ES{TODO: WHY? (might be too similar to previous?)}
+
+% ES: decision tree gets this one wrong, may want to find something else
+% \begin{ecode}
+%   let rec sepConcat sep sl =
+%     match sl with
+%     | [] -> ""
+%     | h::t ->
+%         let f a x = a ^ (sep ^ x) in
+%         List.fold_left f h t
+
+%   let stringOfList f l = sepConcat "; " __[ "["; ___=List.map f l=___; "]" ]__
+% \end{ecode}
+
+\paragraph{Computing the Fixed Point of a Function}
+Finally, our students must write a |fixpoint| function that computes the
+fixed point of a given function |f|, starting from a base case |b|.
+%
+As a hint we first have them write a |wwhile| function that performs the
+functional equivalent of a while-loop, repeatedly passing a function's
+output back in until it receives a (boolean) signal to stop.
+%
+\begin{ecode}
+  let rec wwhile (f, b) =
+    match f b with
+    | (x, false) -> x
+    | (x, true)  -> wwhile (f, x)
+\end{ecode}
+\lstset{firstnumber=last}
+%
+The |fixpoint| function can then be written as a clever instantiation of
+the arguments to |wwhile|.
+%
+\begin{ecode}
+  let fixpoint (f, b) =
+    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
+    wwhile (g, b)
+\end{ecode}
+\lstset{firstnumber=1}
+%
+Sadly, our student has forgotten that |g| should itself be a function.
+%
+As a result, she passes a pair of a \emph{pair} and a starting value to
+|wwhile|, rather than a pair of a \emph{function} and a starting value.
+%
+\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
+actually blames the use of |g|), but identifies the construction of the
+pair inside |g| as the most likely culprit, while our Decision Tree
+correctly identifies the \emph{definition} of |g| as the source of the
+error.
+\ES{TODO: WHY?}
 

### 2017-03-26 paper/oopsla17-submission/evaluation-accuracy.tex
feature selection

@@
 \paragraph{Our Classifiers}
-We evaluate five classifiers, each trained on the full set of 276
-features: 44 local syntactic features, 176 contextual syntactic
-features, 55 typing features, and a single expression size feature.
-%
-\ES{should explain the make-up of these groups}
+We evaluate five classifiers, each trained on the full set of features.
+% features: 44 local syntactic features, 176 contextual syntactic
+% features, 55 typing features, and a single expression size feature.
+% %
+% \ES{should explain the make-up of these groups}
 %
 We preemptively discard expressions that are not part of the minimal
-type error slice, we will explain the rationale for this in
-\autoref{sec:feature-utility}.
+type error slice --- we will explain the rationale for this in
+\autoref{sec:feature-utility} --- and thus the final feature count is
+276.
 %
@@
 In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
-hidden features than the \hiddenFH, it only loses around 2\% accuracy.
+hidden neurons than the \hiddenFH, it only loses around 2\% accuracy.
 % In particular, notice that the \hiddenT only loses around 2\% accuracy

### 2017-03-26 paper/oopsla17-submission/evaluation-utility.tex
feature selection

@@
 
-Next, we will investigate the relative impact of the other three classes
-of features discussed in \autoref{sec:features}.
+Continuing on, we will investigate the relative impact of the other
+three classes of features discussed in \autoref{sec:features}, assuming
+we have discarded expressions not in the type error slice.
 %
-For this we consider again a baseline of only local syntactic features,
-extended by each combination of
+For this experiment we consider again a baseline of only local syntactic
+features, extended by each combination of
 %

### 2017-03-26 paper/oopsla17-submission/evaluation.tex
feature selection

@@
 
+\subsubsection{Feature Selection}
+We extract a set of 277 features from each sub-expression in a student
+program, including:
+%
+\begin{enumerate}
+\item 44 local syntactic features. In addition to the syntax of \lang,
+  we support the full range of arithmetic operators (integer and
+  floating point), equality and comparison operators, character and
+  string literals, and a user-defined |expr| type of simple arithmetic
+  expressions. We will discuss the challenge of supporting other
+  user-defined types in \autoref{sec:discussion}.
+\item 176 contextual syntactic features. For each sub-expression we
+  additionally extract the local syntactic features of its parent and
+  first, second, and third (left-to-right) children. If an expression
+  does not have a parent or children, these features will simply be
+  disabled. If an expression has more than three children, the
+  classifiers will be unable to learn anything about the additional
+  children.
+\item 55 typing features. In addition to the types of \lang, we support
+  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
+  mentioned above. These features are extracted for each sub-expression
+  and for the contextual sub-expressions.
+\item 1 feature denoting the size of each sub-expression.
+\item 1 feature denoting whether each sub-expression is part of the
+  minimal type error slice.
+\end{enumerate}
+
+\subsubsection{Blame Oracle}
 We identified the fixes for each ill-typed program with an
@@
 
-\paragraph{Accuracy Metric}
+\subsubsection{Accuracy Metric}
 All of the tools we compare (with the exception of \ocaml) can produce a

### 2017-03-26 paper/oopsla17-submission/gallery.tex
feature selection

@@
 Finally, our students must write a |fixpoint| function that computes the
-fixed point of a given function |f|, starting from a base case |b|.
+fixed point of a given function |f|, starting from an initial value |b|.
 %

### 2017-03-26 paper/oopsla17-submission/evaluation-utility.tex
discuss recall

@@
 
-We tested our hypothesis with a linear model cross-validated ($k=10$)
-over the combined SP14/FA15 dataset. We used a learning rate
-$\alpha=0.001$, L2 regularization rate $\lambda=0.001$, and mini-batch
-size of 200. We trained for a single epoch on feature sets (1) and
-(2), and for 8 epochs on (3), so that the total number of training samples
-would be roughly equal for each feature set.
-\ES{introduce the MLP too}
-\ES{use MLP-500 instead?}
+We tested our hypothesis with the \linear and
+%
+\hiddenFH\footnote{A layer of 500 neurons is excessive when have so few
+  input features --- we use \hiddenFH for continuity with the
+  surrounding sections.}
+%
+classifiers, cross-validated ($k=10$) over the combined SP14/FA15
+dataset. We used a learning rate $\alpha=0.001$, L2 regularization rate
+$\lambda=0.001$, and mini-batch size of 200. We trained for a single
+epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
+total number of training samples would be roughly equal for each feature
+set.
+%
+In addition to the accuracy metric from before, we report each
+classifier's \emph{recall} --- \ie ``How many true changes can we
+remember?'' --- defined as
+\lstDeleteShortInline{|} % sigh...
+$$
+\frac{|\mathsf{predicted} \cap \mathsf{oracle}|}
+     {|\mathsf{oracle}|}
+$$
+\lstMakeShortInline{|}
+where $\mathsf{predicted}$ is limited to the top 3 predictions of the
+classier, and $\mathsf{oracle}$ is limited to the expressions that
+changed \emph{and} were part of the type error slice. We make the
+latter distinction as (1) changes that are not part of the type error
+slice are noise in the data set, and (2) it makes the comparison between
+feature sets easier to interpret since $\mathsf{oracle}$ never changes.
+%
 \begin{table}[ht]
@@
     \toprule
-                       &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenF       \\
+                       &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
                                          \cmidrule{4-7}                        \cmidrule{9-12}
@@
     \midrule
-    Local Syntax       & 47          & & 23.6\% & 42.6\% & 56.3\% & 19.6\% & & 30.9\% & 47.9\% & 58.6\% & 20.6\% \\
-    + \InSlice         & 48          & & 46.4\% & 65.0\% & 75.4\% & 30.4\% & & 54.5\% & 70.5\% & 81.5\% & 34.7\% \\
-    Filter by \InSlice & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.7\% & 72.3\% & 82.9\% & 58.0\% \\
+    Local Syntax       & 47          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
+    + \InSlice         & 48          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
+    Filter by \InSlice & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
     \bottomrule

### 2017-03-26 paper/oopsla17-submission/evaluation-utility.tex
misc

@@
 \begin{enumerate}
-\item A baseline with only local syntactic features.
+\item A baseline with \emph{only} local syntactic features.
 \item The features of (1) extended with \InSlice.
@@
 %
+\lstDeleteShortInline{|} % sigh...
 In addition to the accuracy metric from before, we report each
@@
 remember?'' --- defined as
-\lstDeleteShortInline{|} % sigh...
 $$
@@
 $$
-\lstMakeShortInline{|}
 where $\mathsf{predicted}$ is limited to the top 3 predictions of the
@@
 feature sets easier to interpret since $\mathsf{oracle}$ never changes.
+% NOTE: keep this at the end of the para or it screws up spacing...
+\lstMakeShortInline{|}
 %
 \begin{table}[ht]
+  \caption{
+    Impact of type-error slice on accuracy.
+    \ES{TODO: expand caption}
+    \ES{TODO: load these numbers from CSV}
+  }\label{tab:type-error-slice}
   \centering
@@
   \end{tabular}
-  \caption{
-    Impact of type-error slice on accuracy.
-    \ES{TODO: expand caption}
-    \ES{TODO: load these numbers from CSV}
-  }\label{tab:type-error-slice}
 \end{table}
@@
 
-Note also the jump in Recall when we filter out expressions that are not
-part of the error slice.
-%
-Reducing the search space not only improves our chances of making a
-single correct prediction, it also allows us to make \emph{multiple}
-correct predictions per program.
+% Note also the jump in Recall when we filter out expressions that are not
+% part of the error slice.
+% %
+% Reducing the search space not only improves our chances of making a
+% single correct prediction, it also allows us to make \emph{multiple}
+% correct predictions per program.
 
@@
 \begin{table}[ht]
+  \caption{
+    Impact of contextual features on accuracy.
+    \ES{TODO: expand caption}
+    \ES{TODO: load these numbers from CSV}
+  }\label{tab:contextual-features}
   \centering
@@
   % \end{minipage}
-  \caption{
-    Impact of contextual features on accuracy.
-    \ES{TODO: expand caption}
-    \ES{TODO: load these numbers from CSV}
-  }\label{tab:contextual-features}
 \end{table}

### 2017-03-26 paper/oopsla17-submission/learning.tex
misc

@@
 \midrule
-\lstinline![]!
+|[]|
   & 1 & 1 & 1 & 0 & 1 & 1 \\
-\lstinline!hd + sumList tl!
+|hd + sumList tl|
   & 0 & 1 & 5 & 1 & 0 & 1 \\
-\lstinline!sumList tl!
+|sumList tl|
   & 0 & 0 & 3 & 0 & 1 & 1 \\
-\lstinline!tl!
+|tl|
   & 0 & 0 & 1 & 0 & 1 & 0 \\
@@
 %
-Thus, given the \lstinline!sumList! program in \autoref{fig:sumList} we
+Thus, given the |sumList| program in \autoref{fig:sumList} we
 would first split it into its constituent sub-expressions and then
@@
 
-\paragraph{Local syntactic features}
+\subsubsection{Local syntactic features}
 These features describe the syntactic category of each expression $e$.
@@
 
-\paragraph{Contextual syntactic features}
+\subsubsection{Contextual syntactic features}
 These are like local syntactic features, but lifted to describe the
@@
 
-\paragraph{Expression size}
+\subsubsection{Expression size}
 We also add a feature representing the \emph{size} of each expression,
@@
 For example, the \ExprSize feature in \autoref{tab:sumList} is set to 3
-for the expression \lstinline!sumList tl! as it contains 3 expressions:
+for the expression |sumList tl| as it contains 3 expressions:
 the two variables and the application itself.
@@
 
-\paragraph{Typing features}
+\subsubsection{Typing features}
 A natural way of summarizing the context in which an expression occurs
@@
 %
-For example, the expressions \lstinline!tl! and \lstinline!sumList tl!
+For example, the expressions |tl| and |sumList tl|
 in \autoref{tab:sumList} both enable the feature \HasTypeList, as they
@@
 bias~\cite{McAdam1998-ub,Yang2000-yr} --- in particular with the
-\lstinline!sumList tl! expression, which might alternatively be assigned
+|sumList tl| expression, which might alternatively be assigned
 the type $\tint$.
@@
 
-\paragraph{Type error slice}
+\subsubsection{Type error slice}
 Finally, we would like the model to be able to distinguish between
@@
 expression is part of such a minimal slice, and is enabled for all of
-the sampled expressions except for \lstinline!tl!, which does not affect
+the sampled expressions except for |tl|, which does not affect
 the type error.

### 2017-03-26 paper/oopsla17-submission/gallery.tex
failed predictions

--- 2017-03-24 paper/oopsla17-submission/gallery.tex
@@
 \paragraph{Extracting the Digits of an Integer}
+% data/sp14/1655.ml
 Our first program is a simple recursive function |digitsOfInt| that
@@
 \paragraph{Padding a list}
+% data/sp14/0306.ml
 Our next program, |padZero|, is given two |int list|s as input, and must
@@
 \paragraph{Computing the Fixed Point of a Function}
+% data/sp14/0941.ml
 Finally, our students must write a |fixpoint| function that computes the
-fixed point of a given function |f|, starting from a base case |b|.
+fixed point of a given function |f|, starting from an initial value |b|.
 %
@@
 
+\subsubsection{Failed Predictions}
+\label{sec:failed-predictions}
+Of course, our classifiers are sometimes wrong, we focus next on the
+10\% of programs where \emph{none} of our Top-3 predictions are correct.
+
+% data/sp14/0148.ml
+\begin{ecode}
+  let rec clone x n =
+    let loop acc n =
+      if n <= 0 then
+        acc
+      else
+        clone ==([==_=x=_==] @ acc)== (n - 1) in
+    loop [] n
+\end{ecode}
+\ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
+
+% data/sp14/0234.ml
+\begin{ecode}
+  let pipe fs =
+    let f a x z = _=z=___ a__ in
+    let base y = y in
+    List.fold_left f base fs
+\end{ecode}
+\ES{TODO: fix is to swap a and z, \sherrloc is correct, \ocaml blames use of f in fold-left..}
+
+% data/sp14/0255.ml
+\begin{ecode}
+  let rec wwhile (f, b) =
+    let check = f b in
+    let (x, y) = (f, b) in
+    if __y__ = false then
+      x
+    else
+      wwhile (f, ==x==)
+\end{ecode}
+\ES{marking ``y'' seems totally bogus, why is it even part of the error
+  slice? (other locations are ``f b'' and ``false'' though, whihc is not
+  any better)}
+
 %%% Local Variables:

### 2017-03-27 paper/oopsla17-submission/gallery.tex
more gallery

@@
 As a hint we first have them write a |wwhile| function that performs the
-functional equivalent of a while-loop, repeatedly passing a function's
+functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
 output back in until it receives a (boolean) signal to stop.
@@
 Of course, our classifiers are sometimes wrong, we focus next on the
-10\% of programs where \emph{none} of our Top-3 predictions are correct.
+$\sim$10\% of programs where \emph{none} of our Top-3 predictions are correct.
 
+\paragraph{Cloning an Element}
 % data/sp14/0148.ml
+Consider the following implementation of the |clone| function from
+before.
+%
 \begin{ecode}
@@
 \end{ecode}
-\ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
+% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
+%
+The student has defined a helper function |loop| with an accumulator
+|acc|, likely meant to call itself tail-recursively.
+%
+Unfortunately, she has called the top-level function |clone| rather than
+|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
+for the |x| argument to |clone|.
 
+Our classifier incorrectly predicts that the use of |x| in the recursive
+call is the most likely source of the error.
+\ES{TOOD: WHY??}
+%
+Our second prediction coincides with \sherrloc (and \ocaml), blaming the
+the first argument to |clone|.
+%
+This is also incorrect, but may be more helpful than our first
+prediction --- if our student decides that she has certainly provided
+the correct \emph{argument}, an alternative explanation is that
+perhaps she has called the wrong \emph{function}.
+%
+Our final prediction is the entire body of |clone|, which is hardly
+helpful.
+
+\paragraph{Composing Functions}
 % data/sp14/0234.ml
+Next, let us consider the |pipe| function that composes a list of
+functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
+function |fun x -> f (g (h x))|.
+%
 \begin{ecode}
@@
 \end{ecode}
-\ES{TODO: fix is to swap a and z, \sherrloc is correct, \ocaml blames use of f in fold-left..}
+%
+Leaving aside issues of \emph{functional} correctness, the error in the
+student's code is that she has swapped function and argument in the
+application |z a|.
+%
+\sherrloc correctly blames the use of |z|, while our classifier
+(incorrectly) blames the entire application (\ocaml blames the use of
+|f| on line 4).
+\ES{TODO: WHY??}
+%
 
+\paragraph{A Functional \texttt{while}-loop}
 % data/sp14/0255.ml
+Finally, consider the following attempt at the functional
+\texttt{while}-loop from before.
+%
 \begin{ecode}
@@
 \end{ecode}
+% ES: the let (x, y) = (f, b) is actually a match, but this seems easier to present..
+Our student
+
 \ES{marking ``y'' seems totally bogus, why is it even part of the error
-  slice? (other locations are ``f b'' and ``false'' though, whihc is not
+  slice? (other locations are ``f b'' and ``false'' though, which is not
   any better)}

### 2017-03-27 paper/oopsla17-submission/evaluation-utility.tex
more eval

@@
 The \InSlice feature should be highly predictive --- a fix must change
-at least one expression in the type-error slice.
+at least one expression in the type error slice.
 %
@@
 %
-\hiddenFH\footnote{A layer of 500 neurons is excessive when have so few
+\hiddenFH\footnote{A layer of 500 neurons is excessive when we have so few
   input features --- we use \hiddenFH for continuity with the
@@
   \caption{
-    Impact of type-error slice on accuracy.
-    \ES{TODO: expand caption}
+    Impact of Type Error Slice on Accuracy.
     \ES{TODO: load these numbers from CSV}
@@
 %
-As expected the baseline performs the worst, with a mere 23.6\% Top-1
-accuracy.
+As expected the baseline performs the worst, with a mere 23.6\% \linear
+Top-1 accuracy.
 %
-Adding \InSlice improves the results substantially with a 46.4\% Top-1
+Adding \InSlice improves the results substantially with a 46.4\% \linear Top-1
 accuracy, demonstrating the importance of a minimal error slice.
@@
 However, filtering out expressions that are not part of the slice
-\emph{further} improves the results to 54.9\% Top-1 accuracy.
+\emph{further} improves the results to 54.9\% \linear Top-1 accuracy.
 %
-Clearly, some decisions are too important to be left to a machine.
+Interestingly, while the \hiddenFH performs similarly poor with no error
+slice features, it recovers nearly all of its accuracy after being given
+the error slice features.
+%
+Top-1 accuracy jumps from 28.9\% to 54.4\% when we add \InSlice, and only
+improves by 2\% when we filter out expressions that are not part of the
+error slice.
 
-% Note also the jump in Recall when we filter out expressions that are not
-% part of the error slice.
-% %
-% Reducing the search space not only improves our chances of making a
-% single correct prediction, it also allows us to make \emph{multiple}
-% correct predictions per program.
-
-Given the decisive benefits of filtering out expressions that do not
-belong to the type-error slice, we choose to filter all programs by
-\InSlice.
+Still, the 2\% boost comes at zero cost, and given the other benefits
+of filtering out expressions that do not belong to the type error slice,
+we choose to filter all programs by \InSlice.
 
@@
   \caption{
-    Impact of contextual features on accuracy.
-    \ES{TODO: expand caption}
+    Impact of Contextual Features on Accuracy.
     \ES{TODO: load these numbers from CSV}

### 2017-03-27 paper/oopsla17-submission/gallery.tex
more eval

@@
 \label{sec:failed-predictions}
-Of course, our classifiers are sometimes wrong, we focus next on the
-$\sim$10\% of programs where \emph{none} of our Top-3 predictions are correct.
+Of course, our classifiers are sometimes wrong, we focus next on programs
+where our classifier makes an incorrect prediction.
 
@@
 
+\paragraph{Currying Considered Harmful?}
+% data/sp14/0887.ml
+Our next example is another ill-fated attempt at |clone|.
+%
+\begin{ecode}
+  let rec clone x n =
+    let rec loop x n acc =
+      if n < 0 then
+        acc
+      else
+        ==loop== __(x, (n - 1), (x :: acc))__ in
+    loop (x, n, [])
+\end{ecode}
+The issue here is that \ocaml functions are \emph{curried} by default
+--- \ie they take their arguments one at a time --- but our student has
+called the inner |loop| with all three arguments in a tuple.
+%
+Many experienced functional programmers would choose to keep |loop|
+curried and rewrite the calls, however our student decides instead to
+\emph{uncurry} |loop|, making it take a tuple of arguments.
+%
+\sherrloc blames the recursive call to |loop| while our classifier
+blames the tuple of arguments --- a reasonable suggestion, but alas not
+the answer the student was looking for.
+\ES{WHY??}
+%
+
 \paragraph{Composing Functions}
-% data/sp14/0234.ml
+% data/sp14/2269.ml
 Next, let us consider the |pipe| function that composes a list of
 functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
-function |fun x -> f (g (h x))|.
+function |fun x -> h (g (f x))|.
 %
@@
   let pipe fs =
-    let f a x z = _=z=___ a__ in
-    let base y = y in
+    let f a x y = ==y== __(a y)__ in
+    let base x = x in
     List.fold_left f base fs
@@
 %
-Leaving aside issues of \emph{functional} correctness, the error in the
-student's code is that she has swapped function and argument in the
-application |z a|.
+The error in our student's code is that she has applied |y| rather than
+|x| to the result of |a y|.
 %
-\sherrloc correctly blames the use of |z|, while our classifier
-(incorrectly) blames the entire application (\ocaml blames the use of
-|f| on line 4).
-\ES{TODO: WHY??}
+\sherrloc correctly blames the first occurrence of |y|, while our
+classifier (incorrectly) blames the application |a y| (\ocaml blames
+the occurrence of |base| on line 4).
 %
-
-\paragraph{A Functional \texttt{while}-loop}
-% data/sp14/0255.ml
-Finally, consider the following attempt at the functional
-\texttt{while}-loop from before.
-%
-\begin{ecode}
-  let rec wwhile (f, b) =
-    let check = f b in
-    let (x, y) = (f, b) in
-    if __y__ = false then
-      x
-    else
-      wwhile (f, ==x==)
-\end{ecode}
-% ES: the let (x, y) = (f, b) is actually a match, but this seems easier to present..
-Our student
-
-\ES{marking ``y'' seems totally bogus, why is it even part of the error
-  slice? (other locations are ``f b'' and ``false'' though, which is not
-  any better)}
+\ES{TODO: WHY?}
 

### 2017-03-27 paper/oopsla17-submission/evaluation-accuracy.tex
minor

@@
 In this experiment we compare the accuracy of our predictions to the
-state of the art in type-error localization.
+state of the art in type error localization.
 
@@
 We provide two baselines for the comparison: a random choice of location
-from a minimized type-error slice, and the \ocaml compiler that
+from a minimized type error slice, and the \ocaml compiler that
 programmers typically interact with.

### 2017-03-27 paper/oopsla17-submission/evaluation-utility.tex
minor

@@
 Thus, our first experiment seeks to quantify the impact of \InSlice by
-comparing the accuracy of a linear model on three sets of features:
+comparing the accuracy of our classifiers on three sets of features:
 %
@@
     \midrule
-    Local Syntax       & 47          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
-    + \InSlice         & 48          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
-    Filter by \InSlice & 47          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
+    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
+    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
+    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
     \bottomrule
@@
 Still, the 2\% boost comes at zero cost, and given the other benefits
-of filtering out expressions that do not belong to the type error slice,
+of filtering out expressions that do not belong to the type error slice
+--- shrinking the search space and guaranteeing our predictions are actionable ---
 we choose to filter all programs by \InSlice.
@@
 $\lambda = 0.001$, and a mini-batch size of 200, but we
-train for a full 20 epochs.
-
+train for a full 20 epochs to make the differences more apparent.
+%
 \begin{table}[ht]
@@
 \end{table}
-
+%
 \paragraph{Results}

### 2017-03-28 paper/oopsla17-submission/related.tex
braindump related work on type errors

@@
 
-In this section we describe two relevant aspects of related work:
+In this section we describe three relevant aspects of related work:
+%
+programming languages approaches to diagnosing type errors,
+%
 software engineering approaches to fault localization, and
-programming languages approaches to pinpointing type errors.
+%
+applying machine learning techniques to problems in programming languages.
 
+\subsection{Type Error Diagnosis}
+\label{sec:type-error-diagnosis}
+Languages with static type systems and type inference produce type
+errors that novices often perceive as difficult to
+interpret~\citep{Wand1986-nw}.
+%
+For example, approaches based on Hindley-Milner or other constraint
+systems typically have the issue that when the constraint $\alpha=\beta$
+is violated, the system cannot know whether the user should intend to
+fix $\alpha$ or $\beta$ and must thus report the discrepancy in a
+generic manner.
+%
+As a result, a number of approaches have been proposed to
+more precisely localize the type error,
+give clearer error messages, or
+fix the error automatically.
+%
+% The technique most related to our work is \ES{TODO}.  It has been shown
+% to be quite good at \ES{TODO}. However, we choose to focus on \ES{TODO},
+% and our approach contains \ES{TODO}, which is not present in previous
+% work.
+
+\paragraph{Localizing Type Errors}
+It is well-known that the original Damas-Milner algorithm $\mathcal{W}$
+produces poor error locations~\citep{Wand1986-nw}, often reported far
+from their source.
+%
+The reason is that the type checker collects typing constraints as it
+traverses the program, and it crucially assumes that if a constraint can
+be safely added to the set of assumptions (\ie no type error has been
+found yet), then the constraint can be \emph{assumed} to be correct.
+%
+The type checker reports an error the moment it finds a constraint that
+contradicts one of the assumptions, blaming the new inconsistent
+constraint, and thus it is extremely sensitive to the order in which it
+traverses the source program, the infamous ``left-to-right''
+bias~\citep{McAdam1998-ub}.
+%
+Several alternative traversal strategies have been proposed, \eg
+top-down rather than bottom-up~\citep{Lee1998-ys}, or a \emph{symmetric}
+traversal that checks sub-expressions independently and only reports an
+error when two inconsistent sets of constraints are
+merged~\citep{McAdam1998-ub,Yang1999-yr}.
+
+Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
+extracts a complete and minimal subset of program expressions that
+contribute to the type error, \ie all of the expressions that are
+required for the error to manifest and no more.
+%
+Slicing techniques typically require rewriting the type checker with a
+specialized constraint language and solver, though
+\citet{Schilling2011-yf} shows how to turn any type checker into an
+error slicer by treating it as a black-box.
+
+While slicing techniques guarantee enough information to diagnose the
+error, they can fall into the trap of providing \emph{too much}
+information, producing a slice that is not significantly smaller than
+the input program.
+%
+Thus, recent work has focused on finding the \emph{most likely} source
+of a type error.
+%
+\citet{Zhang2014-lv} use Bayesian reasoning to search the constraint
+graph for constraints that participate in many unsatisfiable paths and
+relatively few satisfiable paths, based on the intuition that the
+program should be mostly correct.
+%
+\citet{Pavlinovic2014-mr} translate the localization problem into a
+MaxSMT problem, asking an off-the-shelf SMT solver to find the smallest
+set of constraints that can be removed such that the resulting system is
+satisfiable.
+%
+\citet{Loncaric2016-uk} improve the scalability of
+\citeauthor{Pavlinovic2014-mr} by reusing the existing type checker as
+an optimized theory solver in the Nelson-Oppen~\citeyear{Nelson1979-td}
+style, and thus require only a MaxSAT solver.
+%
+All three of these techniques support \emph{weighted} constraints to
+incorporate knowledge about the relative frequency of different errors,
+but only \citeauthor{Pavlinovic2014-mr} use the weights, setting them to
+the size of the sub-expression that induced the constraint.
+%
+In contrast, our classifiers learn a set of heuristics for predicting
+the source of type errors by observing a set of ill-typed programs and
+their subsequent fixes, in a sense using \emph{only} the weights and no
+constraint solver.
+%
+It may be profitable to combine both approaches, \ie learn a set of good
+weights for one of the above techniques from our training data.
+
+\paragraph{Explaining Type Errors}
+In this paper we have focused solely on the task of \emph{localizing} a
+type error, but a good error report should also \emph{explain} the
+error.
+%
+\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
+attempt to explain type errors by collecting the chain of inferences
+made by the type checker --- essentially the typing derivation --- and
+presenting them to the user.
+%
+Such explanations can be lengthy, as an attempt to compress the
+explanation \citet{Yang2000-kz} presents a visualization of the
+inference process.
+%
+\citet{Gast2004-zd} produces a type error slice enhanced by arrows
+showing the flow of data from sources with incompatible types to a
+shared sink, borrowing the insight of dataflows-as-explanations from
+\textsc{MrSpidey}~\citep{Flanagan1996-bu}.
+
+Such \emph{static} explanations of type errors run the risk of
+overwhelming the user with too much information, it may be preferable to
+treat type error diagnosis as an \emph{interactive} debugging session.
+%
+\citet{Bernstein1995-yj} extend the type inference procedure to handle
+\emph{open} expressions (\ie with unbound variables), allowing users to
+interactively query the type checker for the types of sub-expressions.
+%
+\citet{Chitil2001-td} proposes \emph{algorithmic debugging} of type
+errors, presenting the user with a sequence of yes-or-no questions about
+the inferred types of sub-expressions that guide the user to a specific
+explanation for the error.
+%
+\citet{Seidel2016-ul} explain type errors by searching for inputs that
+expose the \emph{run-time} error that the type system prevented, and
+present users with an interactive visualization of the erroneous
+computation.
+
+The best explanation of a type error, however, might be given by an
+expert, \eg the compiler or library author.
+%
+\citet{Hage2006-hc} catalog a set of heuristics for
+improving the quality of error messages by examining the errors made by
+novices.
+%
+\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
+\citet{Serrano2016-oo} extend the ability to customize error messages to
+library authors, enabling \emph{domain-specific} type errors.
+%
+The 8.0 release of the
+Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
+incorporates some of these ideas, allowing library authors to supply
+custom errors when type-class resolution or type-family reduction fail,
+but not for ordinary unification failures.
+
+\paragraph{Fixing Type Errors}
+Some techniques go beyond explaining a type error or locating its
+source, and actually attempt to \emph{fix} the error automatically.
+%
+\citet{Lerner2007-dt} searches for fixes to type errors by enumerating a
+set of local mutations to the program and querying the type checker to
+see if the error remains.
+%
+\citet{Chen2014-gd} use a notion of \emph{variation-based} typing to
+track choices made by the type checker and enumerate potential type (and
+expression) changes that would fix the error.
+%
+They also extend the algorithmic debugging technique of
+\citeauthor{Chitil2001-td} by allowing the user to enter the expected
+type of specific sub-expressions and suggesting fixes based on these
+desired types \citeyear{Chen2014-vm}.
+%
+Our classifiers do not attempt to suggest fixes to type errors, but it
+may be possible to do so by training a classifier to predict the
+syntactic class of each expression in the \emph{fixed} program --- we
+believe this is an exciting direction for future work.
+
+\subsection{Fault Localization}
 Given a defect, fault localization is the task of identifying
@@
 
-Languages with static type systems and type inference produce type errors
-that novices often perceive as difficult to
-interpret~\citep{Wand1986-nw,Beaven1993-hb,Duggan1996-by,Yang2000-yr,McAdam1998-ub}.
-For example, approaches based on Hindley-Milner or other constraint systems
-typically have the issue that when the constraint $A=B$ is violated, 
-the system cannot know whether the user should intend to fix $A$ or $B$ and
-must thus report the discrepancy in a generic manner. As a result, a number
-of approaches have been proposed to either give clearer error messages
-(e.g.,~\cite{FIXME}) or more precisely localize the type error
-(e.g.,~\cite{FIXME}). The technique most related to our work is \ES{TODO}.
-It has been shown to be quite good at \ES{TODO}. However, we choose to
-focus on \ES{TODO}, and our approach contains \ES{TODO}, which is not
-present in previous work.
+\subsection{Machine Learning for Programming Languages}
+\label{sec:ml-pl}
 

### 2017-03-28 paper/oopsla17-submission/discussion.tex
stub for discussion

\section{Discussion}
\label{sec:discussion}

\paragraph{Additional Features}
\begin{itemize}
\item def-use chains
\item connect function and arguments in application (right now each only know about the application itself)
\item ??
\end{itemize}

\paragraph{User-Defined Types}
\begin{itemize}
\item Hard to support in general, as each data and type constructor gets
  its own set of syntactic features.
\item Attempt to mitigate by adding generic ``is-datacon'' and
  ``is-match'' feature, unclear how much it will help
\end{itemize}

\paragraph{Independent vs Joint Predictions}
We treat each sub-expression as if it exists in a vacuum, in reality
making a positive prediction for one expression may make another
prediction much less likely \citep{Raychev2015-jg}. We compensate
somewhat by ranking predictions by ``confidence'', but would be
interesting to see how structured learning would perform.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2017-03-29 paper/oopsla17-submission/related.tex
related

@@
 
-In this section we describe three relevant aspects of related work:
+In this section we describe two relevant aspects of related work:
 %
-programming languages approaches to diagnosing type errors,
+programming languages approaches to diagnosing type errors, and
 %
-software engineering approaches to fault localization, and
+software engineering approaches to fault localization.
 %
-applying machine learning techniques to problems in programming languages.
+% applying machine learning techniques to problems in programming languages.
 
@@
 \subsection{Fault Localization}
-Given a defect, fault localization is the task of identifying
-``suspicious'' program elements (e.g., lines, statements) that are likely
-implicated in the defect (i.e., that should be changed to fix the defect).
-Perhaps the best-known is Tarantula, which uses a simple mathematical
-formula based on measured information from dynamic normal and buggy
-runs~\citep{Jones2002-us}. Other similar approaches, including those of
-\citet{Chen2002-qz} and \citet{Abreu2006-fn,Abreu2007-mu} consider alternate features
-of information or refined formulae and generally obtain more precise
-results; see \citet{Wong2009-pd} for a survey. While some researchers have
-approached such fault localization with an eye toward optimality (e.g.,
-\citet{Yoo2013-rw} determine optimal coefficients), in general
-such fault localization approaches are limited by their reliance on either
-running tests or including relevant features. For example, Tarantula-based
-techniques require a normal and a buggy run of the program. By contrast,
-we consider incomplete programs with type errors that may not be executed
-in any standard sense. Similarly, the features available influence the
-classes of defects that can be localized. A fault localization scheme based
-purely on control flow features lines will have difficulty with cross-site
-scripting or SQL code injection attacks, which follow the same control flow
-path on normal and buggy runs (differing only in the user-supplied data).
-By contrast, we \ES{TODO}.
+Given a defect, \emph{fault localization} is the task of identifying
+``suspicious'' program elements (\eg lines, statements) that are likely
+implicated in the defect (\ie that should be changed to fix the defect)
+%
+--- thus, type error localization can be viewed as an instance of fault
+localization.
+%
+The best-known fault localization technique is likely Tarantula, which
+uses a simple mathematical formula based on measured information from
+dynamic normal and buggy runs~\citep{Jones2002-us}.
+%
+Other similar approaches, including those of \citet{Chen2002-qz} and
+\citet{Abreu2006-fn,Abreu2007-mu} consider alternate features of
+information or refined formulae and generally obtain more precise
+results; see \citet{Wong2009-pd} for a survey.
+%
+While some researchers have approached such fault localization with an
+eye toward optimality (\eg \citet{Yoo2013-rw} determine optimal
+coefficients), in general such fault localization approaches are limited
+by their reliance on either running tests or including relevant
+features.
+%
+For example, Tarantula-based techniques require a normal and a buggy run
+of the program.
+%
+By contrast, we consider incomplete programs with type errors that may
+not be executed in any standard sense.
+%
+Similarly, the features available influence the classes of defects that
+can be localized.
+%
+A fault localization scheme based purely on control flow features
+will have difficulty with cross-site scripting or SQL code injection
+attacks, which follow the same control flow path on normal and buggy
+runs (differing only in the user-supplied data).
+%
+Our feature set is comprised entirely of syntactic and typing features,
+a natural choice for modeling type errors, but it would likely not
+generalize to other types of defects.
 
-\subsection{Machine Learning for Programming Languages}
-\label{sec:ml-pl}
+% \subsection{Machine Learning for Programming Languages}
+% \label{sec:ml-pl}
 

### 2017-03-29 paper/oopsla17-submission/related.tex
related

@@
 %
-The reason is that the type checker collects typing constraints as it
-traverses the program, and it crucially assumes that if a constraint can
-be safely added to the set of assumptions (\ie no type error has been
-found yet), then the constraint can be \emph{assumed} to be correct.
+The type checker collects typing constraints as it traverses the
+program, and it crucially assumes that if a constraint can be safely
+added to the set of assumptions (\ie no type error has been found yet),
+then the constraint can be \emph{assumed} to be correct.
 %
@@
 constraint, and thus it is extremely sensitive to the order in which it
-traverses the source program, the infamous ``left-to-right''
-bias~\citep{McAdam1998-ub}.
+traverses the source program (the infamous ``left-to-right''
+bias~\citep{McAdam1998-ub}).
 %
-Several alternative traversal strategies have been proposed, \eg
+Several alternative traversal have been proposed, \eg
 top-down rather than bottom-up~\citep{Lee1998-ys}, or a \emph{symmetric}
@@
 Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
-extracts a complete and minimal subset of program expressions that
-contribute to the type error, \ie all of the expressions that are
-required for the error to manifest and no more.
+overcomes the traversal bias by extracting a complete and minimal subset
+of program expressions that contribute to the type error, \ie all of the
+expressions that are required for the error to manifest and no more.
 %

### 2017-03-29 paper/oopsla17-submission/gallery.tex
minor
fix subsection title

@@
-\subsection{Qualitative Evaluation}
+\subsection{Explaining Predictions}
 \label{sec:qualitative}
@@
 %
-(2) \sherrloc's prediction (in \textbf{bold}), and
+(2) \sherrloc's prediction (in \textbf{bold}),
 %
-(3) the prediction of our Decision Tree (\underline{underlined}).
+(3) the prediction of our Decision Tree (\underline{underlined}), and
+%
+(4) an \emph{explanation} of why our classifier made its prediction.
 %
 We choose the Decision Tree classifier for this section as its model
-is more easily interpreted than the MLP.
+is more easily interpreted than the MLP, we will attempt to explain
+its predictions by analyzing the paths induced in the tree by the
+programs.
 %
-We will also attempt, for each program, to explain \emph{why} the
-Decision Tree made its prediction, by analyzing the paths induced
-by the programs.
+% We will also attempt, for each program, to explain \emph{why} the
+% Decision Tree made its prediction, by analyzing the paths induced
+% by the programs.
 
@@
 
-\paragraph{Cloning an Element}
+\paragraph{Constructing a List of Duplicates}
 % data/sp14/0148.ml

### 2017-03-29 paper/oopsla17-submission/discussion.tex
some discussion

@@
 
-\paragraph{Additional Features}
-\begin{itemize}
-\item def-use chains
-\item connect function and arguments in application (right now each only know about the application itself)
-\item ??
-\end{itemize}
+We have shown that we can outperform the state of the art in type error
+localization by learning a model of the errors that programmers make,
+using a set of features that closely resemble the information the type
+checker sees.
+%
+However, there is certainly room for improvement.
+%
+In this section we highlight some limitations of our approach and
+potential avenues for future work.
 
 \paragraph{User-Defined Types}
-\begin{itemize}
-\item Hard to support in general, as each data and type constructor gets
-  its own set of syntactic features.
-\item Attempt to mitigate by adding generic ``is-datacon'' and
-  ``is-match'' feature, unclear how much it will help
-\end{itemize}
+Probably the single biggest limitation of our technique is that we have
+features for specific data and type constructors.
+%
+Anything our models learn about errors made with the |::| constructor or
+the |list| type cannot easily be translated to new, user-defined
+datatypes the model has never encountered.
+%
+We can attempt to mitigate this to some extent by adding generic
+syntactic features for data constructors and match expressions, but it
+remains to be seen how much these help \ES{TODO: UW data!}.
+%
+Furthermore, there is no obvious analog for transferring knowledge to
+new type constructors, which we have seen are both more compact and
+helpful.
+
+As an alternative to encoding information about \emph{specific}
+constructors, we might investigate a more abstract representation.
+%
+For example, instead of modeling |x :: 2| as a |::| constructor with a
+right child of type |int|, we might model it as some (unknown) constructor
+whose right child has an incompatible type.
+%
+We might symmetrically model the |2| as an integer literal whose type is
+incompatible with its parent.
+%
+Anything we learn about |::| and |2| can now be transferred directly to
+yet unseen types, but we run the risk generalizing \emph{too much} ---
+\ie perhaps programmers make different mistakes with |list|s than they
+do with other types, and are thus likely to choose different fixes.
+%
+Balancing the trade-off between specificity and generalizability appears
+to be a challenging task.
+
+% \begin{itemize}
+% \item Hard to support in general, as each data and type constructor gets
+%   its own set of syntactic features.
+% \item Attempt to mitigate by adding generic ``is-datacon'' and
+%   ``is-match'' feature, unclear how much it will help
+% \end{itemize}
+
+\paragraph{Additional Features}
+There are a number of other features that could improve the model's
+ability to localize errors, that would be easier to add than
+user-defined types.
+%
+For example, each occurrence of a variable knows only its type and the
+immediate context in which it occurs, but it may be helpful to know
+about \emph{other} occurrences of the same variable.
+%
+If a variable is generally used in the context of a |float| but has a
+single occurrence in the context of an |int|, it seems likely that the
+latter occurrence (or context) is to blame.
+%
+Similarly, arguments to a function application are not aware of the
+typing constraints imposed on them by the function (and vice versa),
+they only know that they are occurring in the context of an application
+and the type of the application itself.
+%
+Finally, \emph{n-grams} on the token stream have proven effective for
+probabilistic modeling of programming languages, we may find that they
+aid in our task as well.
+%
+For example, if the observed tokens in an expression diverge from the
+n-gram model's predictions, that indicates that there is something
+unusual about the program at that point, and it may signal an error.
+
+% \begin{itemize}
+% \item def-use chains
+% \item connect function and arguments in application (right now each only know about the application itself)
+% \item ??
+% \end{itemize}
 
 \paragraph{Independent vs Joint Predictions}
-We treat each sub-expression as if it exists in a vacuum, in reality
-making a positive prediction for one expression may make another
-prediction much less likely \citep{Raychev2015-jg}. We compensate
-somewhat by ranking predictions by ``confidence'', but would be
-interesting to see how structured learning would perform.
+We treat each sub-expression as if it exists in a vacuum, but in reality
+the program has a rich \emph{graphical} structure, particularly if one adds
+edges connecting different occurrences of the same variable.
+%
+\citet{Raychev2015-jg} have used these richer models to great effect to
+make \emph{interdependent} predictions about programs, \eg
+de-obfuscating variable names or even inferring types.
+%
+One could even view our task of locating the source of an error as simply
+another property to be predicted over a graphical model of the program.
+%
+One of the key advantages of a graphical model is that the predictions
+made for one node can influence the predictions made for another node,
+this is known as \emph{structured learning}.
+%
+For example, if, given the expression |1 + true|, we predict |true| to
+be erroneous, we may be much less likely to predict |+| as erroneous.
+%
+We compensate somewhat for our lack of structure by ranking our
+predictions by ``confidence'', but it would be interesting to see how
+structured learning over graphical models would perform.
 

### 2017-03-30 paper/oopsla17-submission/overview.tex
init paper
some overview

--- 2017-02-14 paper/oopsla17-submission/evaluation.tex
@@
-\section{Evaluation}
-\label{sec:evaluation}
+\section{Overview}
+\label{sec:overview}
+
+We start with an overview of our approach to localizing type errors by
+learning a model of the mistakes programmers actually make.
+%
+We formulate the problem of type error localization as a
+\emph{supervised classification} problem.
+%
+A \emph{classification} problem entails learning a function that maps
+inputs to a discrete set of output \emph{labels}, in contrast to
+\emph{regression} where the output is typically a real number.
+%
+In a \emph{supervised} learning problem one is given a \emph{training}
+set where the inputs and labels are known, and the task is to learn a
+function that accurately maps the inputs to outputs and
+\emph{generalizes} to new, yet-unseen inputs.
+
+The key technical challenges we address are:
+%
+(1) formulating type error localization as a classification problem,
+%
+(2) generating correct labels for supervised learning, and
+%
+(3) presenting the predictions of the classifier in an intuitive manner.
+
+\paragraph{\textbf{Challenge (1):} How to represent programs?}
+
+\paragraph{\textbf{Solution:} Use a \textbf{set} of feature vectors}
+
+\paragraph{\textbf{Challenge (2):} What is the ``correct'' fix?}
+
+\paragraph{\textbf{Solution:} Use the students fixes}
+
+\paragraph{\textbf{Challenge (3):} What should be presented to the user?}
+
+\paragraph{\textbf{Solution:} Rank predictions and present most likely}
 

### 2017-03-31 paper/oopsla17-submission/introduction.tex
some intro

--- 2017-02-14 paper/oopsla17-submission/evaluation.tex
@@
-\section{Evaluation}
-\label{sec:evaluation}
+\section{Introduction}
+\label{sec:introduction}
+
+Type systems are one of the most widely used techniques for program
+verification, a type checker only accepts a program if it can
+\emph{prove} the absence of an entire class of bugs.
+%
+Typed functional languages like \ocaml and \haskell are equipped with
+particularly powerful type checkers that can operate without a single
+annotation, alleviating a prime concern with verification techniques.
+%
+Yet this power comes with a cost, type annotations signify the
+programmer's \emph{intent}, without them the type checker must
+make assumptions about what the programmer intended to do.
+%
+This results in the common phenomenon of type errors being reported far
+from their source, an annoyance to veterans of the type system, but a
+major hurdle for novices.
+
+\begin{figure}[ht]
+\begin{minipage}{0.45\linewidth}
+\begin{ecode}
+  let rec sumList xs =
+    match xs with
+    | []   -> __[]__
+    | h::t -> h __+__ ==sumList t==
+\end{ecode}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{verbatim}
+File "sumList.ml", line 4, characters 16-25:
+  This expression has type 'a list
+  but an expression was expected of type int
+\end{verbatim}
+\end{minipage}
+\caption{(left) An ill-typed \ocaml program that should sum the elements of a
+  list. The \underline{underlined} expressions are equally valid
+  locations to blame. The expression blamed by the \ocaml compiler
+  is in \textbf{bold}.
+  (right) The error reported by \ocaml.}
+\label{fig:sumList}
+\end{figure}
+%
+Consider the |sumList| program in \autoref{fig:sumList}, written by a
+student in an undergraduate Programming Languages course.
+%
+The program is meant to sum the integers in a list, but our student has
+accidentally given the empty list as the base case, rather than 0.
+%
+The \ocaml compiler collects typing constraints as it traverses the
+program, and reports an error the moment it finds an inconsistent
+constraint.
+%
+In this case it blames the recursive call to |sumList|, complaining that
+|sumList| returns a |list| where an |int| was expected by the |+|
+operator.
+
+Recent work on type error \emph{localization}
+%
+\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}
+%
+frames localization as an optimization problem, seeking a set of
+expressions to blame that minimize some cost function, \eg the number of
+constraints in the blame set, or the ratio of ``good'' paths through the
+constraint graph to ``bad'' paths.
+%
+However, even these techniques fail to distinguish between the two
+underlined expressions in \autoref{fig:sumList}, it would be equally
+valid to blame \emph{either} of them alone.
+%
+The |[]| on line 3 could be changed to |0|, or the |+| on line 4 could
+be changed to either |@| (list append) or |::|, all of which would give
+type-correct programs.
+%
+Thus, these state-of-the-art techniques are forced to either blame
+\emph{both} locations, or choose one \emph{arbitrarily}.
+
+It may appear obvious to the reader that |[]| is the correct expression
+to blame, but how is the type checker to know that?
+%
+Our solution is to \emph{learn} how to localize type errors by observing
+a large set of (novice) errors and their \emph{subsequent fixes}.
+%
+We accomplish this via three concrete contributions.
+
+\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
+Our first contribution is a large dataset of novice \ocaml programs,
+in particular, but not limited to, \emph{ill-typed} programs.
+%
+We used an instrumented version of the \ocaml compiler to collect
+student interactions over two instances of an undergraduate Programming
+Languages course at
+%
+\begin{anonsuppress}
+UC San Diego (IRB \#140608).
+\end{anonsuppress}
+\begin{noanonsuppress}
+AUTHOR's INSTITUTION.
+\end{noanonsuppress}
+%
+The data forms a time-series of programs submitted to the \ocaml
+compiler, from which we extracted a set of \emph{ill-typed} programs
+and their \emph{subsequent fixes}.
+%
+Our dataset comprises over 4,500 such pairs of programs, and many more
+programs that are either well-typed or lack a corresponding fix.
+%
+The entire dataset (both the raw time-series and our extracted pairs of
+ill-typed programs and fixes) is available at
+%
+\begin{anonsuppress}
+\ES{TODO: finally post the data somehwere..}.
+\end{anonsuppress}
+\begin{noanonsuppress}
+HIDDEN FOR DOUBLE-BLIND REVIEW.
+\end{noanonsuppress}
+%
+
+\paragraph{\textbf{(2) Data-Driven Diagnosis of Type Errors}}
+Our second contribution is a \emph{data-driven} technique for localizing
+type errors (\autoref{sec:learning}).
+%
+We use machine learning techniques to train a set of classifiers that can
+predict, given an ill-typed program, which expressions such be changed
+in order for the program to type check.
+%
+Our classifiers can be trained on a relatively small set of programs
+(\ie a single course's worth) and outperform the state of the art by
+15--30\% in blame accuracy.
+%
+The key technical challenge that we tackle is defining a suitable
+interface between the type checker and the classifier, specifically
+\emph{embedding} programs as feature vectors and presenting the
+classifier's predictions to the user.
+
+\paragraph{\textbf{(3) A Large-Scale Evaluation of Type Error Localization}}
+Our third contribution is an evaluation of our classifiers and the
+state-of-the-art techniques on our suite of over 4,500 student programs
+(\autoref{sec:evaluation}), the largest evaluation of type error
+localization techniques we are aware of.
+%
+We judge the correctness of a localization by comparing it to the
+student's eventual fix and find that our classifiers are 15--30\% more
+accurate.
+%
+We also examine which types of features contribute the most to our
+classifiers' predictions, and endeavor to \emph{explain} specific
+predictions that our classifiers get right (and wrong).
+
+% \begin{enumerate}
+% \item An large-scale evaluation of state-of-the-art type error
+%   localization techniques on over 4,500 student programs.
+%   %
+%   We collected a time-series of \ocaml programs from our students, and
+%   use their subsequent \emph{fixes} to ill-typed programs as
+%   \emph{oracles} for the location of the error.
+%   %
+%   This is the most extensive evaluation of the accuracy of type error
+%   localization techniques that we are aware of.
+%   % based not on expert \emph{opinion}, but
+%   % on the eventual \emph{fix} implemented by a novice user.
+%   %
+%   % \ES{need to explain \emph{why} this is superior..}
+% \item A machine learning approach to \emph{modeling} (novice) type
+%   errors that can outperform the state-of-the-art by 15--30\% in
+%   localizing errors.
+%   %
+%   Our classifiers can be trained on a relatively small amount of data --- we
+%   used a single class of around 50 students --- and appear to
+%   generalize to other instances of the same class.
+%   %
+%   \ES{and hopefully completely separate classes!}
+%   %
+%   In contrast to many approaches to \emph{fault localization} from the
+%   software engineering community, our model does not use any linguistic
+%   modeling techniques, \eg \emph{n-grams} over the token stream; rather,
+%   it relies entirely on features of the abstract syntax tree and a
+%   partial typing derivation.
+%   %
+%   \ES{if there's time, investigate adding Wes' n-gram model}
+% \end{enumerate}
 

### 2017-03-31 paper/oopsla17-submission/introduction.tex
some overview

@@
 %
-\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}
+\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr},
 %
+determining which expression(s) to blame for the error,
 frames localization as an optimization problem, seeking a set of
@@
 
+Overall, our results show that data-driven diagnosis is a practical
+technique for localizing type errors by learning from past mistakes.
+%
+% We believe, furthermore, that with the right set of features our
+% approach could be applied to other domains where there are multiple,
+% equally valid explanations for an error.
+% \ES{this sounds really dumb... need better phrasing}
+
 % \begin{enumerate}

### 2017-03-31 paper/oopsla17-submission/overview.tex
some overview

@@
 
-\paragraph{\textbf{Challenge (1):} How to represent programs?}
+\paragraph{\textbf{Challenge (1): How to represent programs?}}
 
-\paragraph{\textbf{Solution:} Use a \textbf{set} of feature vectors}
+\paragraph{\textbf{Solution: Use a set of feature vectors}}
 
-\paragraph{\textbf{Challenge (2):} What is the ``correct'' fix?}
+\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
+As we saw in \autoref{fig:sumList}, type errors admit many different
+fixes to different expressions, each resulting in a different program.
+%
+How is the type checker to judge which expressions are more likely to be
+incorrect than others?
+%
+Prior work has enlisted expert users to manually judge ill-typed
+programs and determine the ``correct'' fix, but this approach does not
+scale well to a dataset large enough to perform machine learning.
+%
+Furthermore, while the expert user has intimate knowledge of the type
+system, she cannot know what the student \emph{intended} to do.
 
-\paragraph{\textbf{Solution:} Use the students fixes}
+\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
+%
+Rather than relying on expert users, we allow our students to do the
+work of assigning blame themselves.
+%
+Software development is an iterative process and our students will
+eventually fix their own programs, perhaps after multiple ill-fated
+attempts.
+%
+For each ill-typed program in our dataset, we find the first
+\emph{subsequent} program that type checks and declare that to be the
+fix.
+%
+From this pair of an ill-typed program and its fix, we can extract a
+\emph{diff} of the abstract syntax trees.
+%
+For example, suppose our student fixed the |sumList| program in
+\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
+only the |[]| expression.
+%
+Thus we would determine that the |[]| expression (and \emph{not} the
+|+|) is to blame for the error.
 
-\paragraph{\textbf{Challenge (3):} What should be presented to the user?}
+\paragraph{\textbf{Challenge (3): What should be presented to the user?}}
+We now have the tools to train a classifier to predict, for each
+expression in an ill-typed program, whether it should be changed.
+%
+But this is not yet particularly suitable as \emph{user feedback}.
+%
+A recent survey of developers by \citet{Kochhar2016-oc} found that
+developers are unlikely to examine more than around five potentially
+erroneous locations before falling back to manual debugging.
+%
+Thus, we should limit our predictions to a select few to be presented to
+the user.
 
-\paragraph{\textbf{Solution:} Rank predictions and present most likely}
+\paragraph{\textbf{Solution: Rank expressions by ``confidence''}}
+Luckily, many machine learning classifiers produce not only a predicted
+label, but also a metric that can be interpreted as the classifier's
+\emph{confidence} in its prediction.
+%
+Thus, we \emph{rank} each expression by the classifier's confidence that
+it should be changed, and present only the top-$k$ predictions to the
+user (in practice $k=3$).
 

### 2017-03-31 paper/oopsla17-submission/overview.tex
more overview

@@
 \paragraph{\textbf{Challenge (1): How to represent programs?}}
+The first and prime challenge we address is how to translate programs
+into a format that machine learning techniques can understand.
+%
+Programs can be represented in many ways, from the raw text, to richly
+structured abstract syntax trees and control-flow graphs, but none
+of these representations are immediately amenable to machine learning.
+%
+Machine learning techniques typically take as input \emph{feature
+vectors}, fixed-length vectors of real numbers where each element
+represents some property of the input.
+%
+Thus, in order to apply machine learning to our problem we must
+determine:
+%
+(1) what are interesting features of a program, and
+%
+(2) how can we represent a variable-size program as a fixed-length vector?
 
-\paragraph{\textbf{Solution: Use a set of feature vectors}}
+\paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
+We start by assuming that we will extract features from the program's
+abstract syntax tree (AST), a natural choice since the type checker
+operates on ASTs.
+%
+An obvious set of interesting features is then the syntactic class of
+each expression, \ie the classifier should be able to distinguish
+between the |[]| and |+| expressions in \autoref{fig:sumList} because
+they are formed by different constructors in the AST.
+%
+A programming language has a fixed set of AST elements, which suggests a
+natural way to obtain fixed-size feature vectors: rather than translating
+a program in a single feature vector, we translate it into a \emph{set} of
+feature vectors, one for each expression in the program.
+%
+Thus, the |[]| and |+| expressions will each be given their own feature
+vector, and we will ask the classifier to predict for each
+\emph{independently} whether it should be blamed.
+
+Of course, each expression occurs in some \emph{context}, and we would
+like the classifier to be able make decisions based on the context as
+well.
+%
+For example, we might expect that the recursive call |sumList t| is
+unlikely to be at fault because |sumList| has a function type --- the
+user may have called the wrong function or supplied the wrong argument,
+but the application itself is probably correct.
+%
+Conversely, if the student wrote |h sumList t| (\ie forgot the |+|), we
+might wish to blame the application rather than |h| because |h|
+\emph{does not} have a function type.
+%
+Thus we \emph{extend} each expression's feature vector with
+\emph{contextual features} of its parent and children.
+%
+These contextual features include the syntactic class of the neighboring
+expressions, their inferred types, and, crucially, whether the
+expression occurs in a minimal type error \emph{slice}.
+%
+A minimal type error slice of a program includes all expressions that
+are necessary for the error to manifest and no more, this allows us to
+tell the classifier which expressions could potentially be blamed ---
+a change to an expression outside of the minimal slice cannot possibly
+fix the type error.
 

### 2017-03-31 paper/oopsla17-submission/overview.tex
minor

@@
 \paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
+\ES{name up for grabs...}
 We start by assuming that we will extract features from the program's
@@
 %
-An obvious set of interesting features is then the syntactic class of
-each expression, \ie the classifier should be able to distinguish
-between the |[]| and |+| expressions in \autoref{fig:sumList} because
-they are formed by different constructors in the AST.
+Features would then be \emph{queries} over the AST, an obvious set of
+interesting features to start with is the syntactic class of each
+expression.
 %
-A programming language has a fixed set of AST elements, which suggests a
-natural way to obtain fixed-size feature vectors: rather than translating
-a program in a single feature vector, we translate it into a \emph{set} of
-feature vectors, one for each expression in the program.
+That is, the classifier should be able to distinguish between the |[]|
+and |+| expressions in \autoref{fig:sumList} because they represent
+different types of expressions.
+%
+A programming language has a fixed set of AST elements\footnote{We will
+  discuss handling arbitrary, user-defined types in
+  \autoref{sec:discussion}.}, which suggests a natural way to obtain
+fixed-size feature vectors: rather than translating a program in a
+single feature vector, we translate it into a \emph{set} of feature
+vectors, one for each expression in the program.
 %
@@
 %
-Prior work has enlisted expert users to manually judge ill-typed
+Prior work has often enlisted expert users to manually judge ill-typed
 programs and determine the ``correct'' fix, but this approach does not
@@
 Furthermore, while the expert user has intimate knowledge of the type
-system, she cannot know what the student \emph{intended} to do.
+system, she cannot in general know what the student \emph{intended} to do.
 
@@
 Thus we would determine that the |[]| expression (and \emph{not} the
-|+|) is to blame for the error.
+|+| or the recursive call |sumList t|) is to blame for the error.
 
 \paragraph{\textbf{Challenge (3): What should be presented to the user?}}
-We now have the tools to train a classifier to predict, for each
-expression in an ill-typed program, whether it should be changed.
+We now have the features and labels needed to train a classifier to
+predict, for each expression in an ill-typed program, whether it should
+be blamed.
 %
@@
 
-\paragraph{\textbf{Solution: Rank expressions by ``confidence''}}
+\paragraph{\textbf{Solution: Rank blame locations by ``confidence''}}
 Luckily, many machine learning classifiers produce not only a predicted
@@
 Thus, we \emph{rank} each expression by the classifier's confidence that
-it should be changed, and present only the top-$k$ predictions to the
+it should be blamed, and present only the top-$k$ predictions to the
 user (in practice $k=3$).

### 2017-03-31 paper/oopsla17-submission/introduction.tex
minor

@@
 We use machine learning techniques to train a set of classifiers that can
-predict, given an ill-typed program, which expressions such be changed
+predict, given an ill-typed program, which expressions should be changed
 in order for the program to type check.

### 2017-04-02 paper/oopsla17-submission/introduction.tex
Wes pass over abstract and introduction.

--- 2017-03-31 paper/oopsla17-submission/introduction.tex
@@
 
-Type systems are one of the most widely used techniques for program
-verification, a type checker only accepts a program if it can
+Type systems are one of the most widely-used techniques for program
+verification. A type checker only accepts a program if it can
 \emph{prove} the absence of an entire class of bugs.
@@
 %
-Yet this power comes with a cost, type annotations signify the
-programmer's \emph{intent}, without them the type checker must
+Yet this power comes with a cost: since type annotations signify the
+programmer's \emph{intent}, without them the checker must
 make assumptions about what the programmer intended to do.
@@
 from their source, an annoyance to veterans of the type system, but a
-major hurdle for novices.
+major hurdle for novices~\cite{FIXME,FIXME}.
 
@@
   is in \textbf{bold}.
+
+  FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
+  ``with'' are also bolded (by \\ecode)! You need to find another way to
+  highlight what ocaml is yelling about.
+
   (right) The error reported by \ocaml.}
@@
 %
-The program is meant to sum the integers in a list, but our student has
+The program is meant to sum the integers in a list, but the student has
 accidentally given the empty list as the base case, rather than 0.
@@
 |sumList| returns a |list| where an |int| was expected by the |+|
-operator.
+operator. Note that this blame is inconsistent with the programmer's
+intention and thus may not be helpful debugging information to a novice.
 
@@
 frames localization as an optimization problem, seeking a set of
-expressions to blame that minimize some cost function, \eg the number of
+expressions to blame that minimize some cost function (\eg the number of
 constraints in the blame set, or the ratio of ``good'' paths through the
-constraint graph to ``bad'' paths.
+constraint graph to ``bad'' paths).
 %
@@
 Our first contribution is a large dataset of novice \ocaml programs,
-in particular, but not limited to, \emph{ill-typed} programs.
+including, but not limited to, \emph{ill-typed} programs.
 %
@@
 \begin{noanonsuppress}
-AUTHOR's INSTITUTION.
+AUTHOR's INSTITUTION (IRB HIDDEN).
 \end{noanonsuppress}
@@
 We use machine learning techniques to train a set of classifiers that can
-predict, given an ill-typed program, which expressions such be changed
-in order for the program to type check.
+predict, given an ill-typed program, which expressions should be changed
+for the program to type check.
 %
 Our classifiers can be trained on a relatively small set of programs
-(\ie a single course's worth) and outperform the state of the art by
-15--30\% in blame accuracy.
+(\ie a single course's worth). 
+% WRW notes that we say this below in contribution (3), which is only three
+% sentences away, so to a first-time reader it sounds like we are
+% double-counting:
+%
+% and outperform the state of the art by
+% 15--30\% in blame accuracy.
 %
@@
 student's eventual fix and find that our classifiers are 15--30\% more
-accurate.
+accurate than the state of the art.
 %
-We also examine which types of features contribute the most to our
+We also examine which features contribute the most to our
 classifiers' predictions, and endeavor to \emph{explain} specific

### 2017-04-02 paper/oopsla17-submission/overview.tex
Wes pass over Section 2

@@
 We formulate the problem of type error localization as a
-\emph{supervised classification} problem.
+\emph{supervised classification} problem~\cite{FIXME}. 
+% Give a generic citation for supervised classification, such as an ML
+% textbook or seminal ML paper or survey, above.
 %
@@
 inputs to a discrete set of output \emph{labels}, in contrast to
-\emph{regression} where the output is typically a real number.
+\emph{regression}, where the output is typically a real number.
 %
@@
 %
-(1) formulating type error localization as a classification problem,
+(1) formulating type error localization as a classification problem;
 %
-(2) generating correct labels for supervised learning, and
+(2) generating correct labels for supervised learning; and
 %
@@
 The first and prime challenge we address is how to translate programs
-into a format that machine learning techniques can understand.
+into a format favorable to machine learning. 
 %
-Programs can be represented in many ways, from the raw text, to richly
-structured abstract syntax trees and control-flow graphs, but none
-of these representations are immediately amenable to machine learning.
+Programs can be represented in many ways, from the raw text to 
+richly-structured abstract syntax trees and control-flow graphs, but 
+representations are not directly amenable to machine learning.
 %
@@
 %
-Thus, in order to apply machine learning to our problem we must
+Thus, to apply machine learning to our problem we must
 determine:
 %
-(1) what are interesting features of a program, and
+(1) what are interesting features of a program; and
 %
@@
 \ES{name up for grabs...}
-We start by assuming that we will extract features from the program's
+We propose to extract features from the program's
 abstract syntax tree (AST), a natural choice since the type checker
@@
 %
-Features would then be \emph{queries} over the AST, an obvious set of
-interesting features to start with is the syntactic class of each
-expression.
+Our features are \emph{queries} over the AST. Our initial set of
+interesting features includes the syntactic class of each expression.
 %
@@
 and |+| expressions in \autoref{fig:sumList} because they represent
-different types of expressions.
+different syntactic expression forms.
 %
-A programming language has a fixed set of AST elements\footnote{We will
+A programming language has a fixed set of AST elements,\footnote{We will
   discuss handling arbitrary, user-defined types in
-  \autoref{sec:discussion}.}, which suggests a natural way to obtain
-fixed-size feature vectors: rather than translating a program in a
+  \autoref{sec:discussion}.} which suggests a natural way to obtain
+fixed-size feature vectors. Rather than translating a program into a
 single feature vector, we translate it into a \emph{set} of feature
@@
 %
-Thus we \emph{extend} each expression's feature vector with
+Thus we extend each expression's feature vector with
 \emph{contextual features} of its parent and children.
@@
 These contextual features include the syntactic class of the neighboring
-expressions, their inferred types, and, crucially, whether the
-expression occurs in a minimal type error \emph{slice}.
+expressions, their inferred types (when available), and, crucially, whether
+the expression occurs in a minimal type error \emph{slice}.
 %
 A minimal type error slice of a program includes all expressions that
-are necessary for the error to manifest and no more, this allows us to
-tell the classifier which expressions could potentially be blamed ---
-a change to an expression outside of the minimal slice cannot possibly
-fix the type error.
+are necessary for the error to manifest and no more. We propose to
+use type error slices to communicate to the classifier which expressions
+could potentially be blamed --- a change to an expression outside of the
+minimal slice cannot possibly fix the type error. This is conceptually
+related to the use of fault localization in other program analysis
+and transformation techniques (see Section~\ref{sec:related-work}).
 
@@
 Prior work has often enlisted expert users to manually judge ill-typed
-programs and determine the ``correct'' fix, but this approach does not
-scale well to a dataset large enough to perform machine learning.
+programs and determine the ``correct'' fix (e.g.,~\cite{FIXME,FIXME}), but
+this approach does not scale well to a dataset large enough to support
+machine learning.
 %
-Furthermore, while the expert user has intimate knowledge of the type
-system, she cannot in general know what the student \emph{intended} to do.
+Furthermore, while expert users have intimate knowledge of the type system,
+they cannot know in general what novice users \emph{intended} to do.
 
@@
 %
-Software development is an iterative process and our students will
+Software development is an iterative process and our students 
 eventually fix their own programs, perhaps after multiple ill-fated
@@
 it should be blamed, and present only the top-$k$ predictions to the
-user (in practice $k=3$).
+user (in practice $k=3$). This use of ranking to report
+the results of a program analysis is popular in other problem domains
+(e.g., via statistics FIXME: cite Engler's z-ranking paper, etc.); we focus
+explicitly on the use of data-driven machine learning confidence as a
+ranking source.
 

### 2017-04-02 paper/oopsla17-submission/learning.tex
some intro
Wes pass over Section 3

@@
 \label{sec:learning}
-Next, we describe our approach to localizing type errors, in the context
+In this section, we describe our approach to localizing type errors, in the
+context
 of \lang (\autoref{fig:syntax}), a simple lambda calculus with integers,
@@
 
-\begin{figure}[ht]
-\begin{minipage}{0.6\linewidth}
-\begin{code}
-  let rec sumList xs =
-    match xs with
-    | []     -> []
-    | hd::tl -> hd + sumList tl
-\end{code}
-\end{minipage}
-\begin{minipage}{0.3\linewidth}
-\begin{code}[numbers=left]
-[]
-hd + sumList tl
-sumList tl
-tl
-\end{code}
-\end{minipage}
-\caption{An ill-typed program (left) and a selection of its
-  sub-expressions (right).}
-\label{fig:sumList}
-\end{figure}
+% \begin{figure}[ht]
+% \begin{minipage}{0.6\linewidth}
+% \begin{code}
+%   let rec sumList xs =
+%     match xs with
+%     | []     -> []
+%     | hd::tl -> hd + sumList tl
+% \end{code}
+% \end{minipage}
+% \begin{minipage}{0.3\linewidth}
+% \begin{code}[numbers=left]
+% []
+% hd + sumList tl
+% sumList tl
+% tl
+% \end{code}
+% \end{minipage}
+% \caption{An ill-typed program (left) and a selection of its
+%   sub-expressions (right).}
+% \label{fig:sumList}
+% \end{figure}
 \begin{table}[ht]
@@
 %
-This is not a \emph{fundamental} limitation though, one could easily add
-features that specify whether a match \emph{contains} a particular
-pattern, and thus have a match expression that enables multiple
+We note that this is not a \emph{fundamental} limitation, and one could
+easily add features that specify whether a match \emph{contains} a
+particular pattern, and thus have a match expression that enables multiple
 features.
@@
 \subsubsection{Contextual syntactic features}
-These are like local syntactic features, but lifted to describe the
+These are similar to local syntactic features, but lifted to describe the
 parent and children of the current expression.
@@
 This gives us a notion of the \emph{context} in which an expression
-occurs, similar to the \emph{n-grams} commonly found in linguistic
-models.
+occurs, similar to the \emph{n-grams} commonly used in linguistic
+models~\cite{FIXME,FIXME}.
+% FIXME: cite "on the naturalness of software" and "on the uniqueness of
+% software" at the very least for the above
 
@@
 \subsubsection{Expression size}
-We also add a feature representing the \emph{size} of each expression,
+We also propose a feature representing the \emph{size} of each expression,
 \ie how many sub-expressions does it contain?
 %
-For example, the \ExprSize feature in \autoref{tab:sumList} is set to 3
-for the expression |sumList tl| as it contains 3 expressions:
+For example, the \ExprSize feature in \autoref{tab:sumList} is set to three
+for the expression |sumList tl| as it contains three expressions:
 the two variables and the application itself.
@@
 
-Note that our use of typing features in an \emph{ill-typed} program
-subjects us to the infamous traversal
-bias~\cite{McAdam1998-ub,Yang2000-yr} --- in particular with the
-|sumList tl| expression, which might alternatively be assigned
+Note that our use of typing features in an ill-typed program
+subjects us to \emph{traversal
+bias}~\cite{McAdam1998-ub,Yang2000-yr}. For example, the 
+|sumList tl| expression might alternatively be assigned
 the type $\tint$.
 %
-We rely on the model to counteract this bias.
+We rely on the model to counteract this bias (see FIXME reference to later
+in the paper).
 
 \subsubsection{Type error slice}
-Finally, we would like the model to be able to distinguish between
+Finally, we desire a model that distinguishes between
 changes that could fix the error, and changes that
@@
 (\ie the set of expressions that contribute to the error), and add a
-feature that is enabled \emph{iff} an expression is part of the slice.
+feature that is enabled for expressions that are part of the slice.
 %
@@
 %
-If the program contains multiple type errors, we will compute
+If the program contains multiple type errors, we compute
 a minimal slice for each error.
@@
 %
-There are a large number of learning algorithms to choose from, existing
+There are many learning algorithms to choose from, existing
 on a spectrum that balances expressiveness with ease of training (and of
@@
 
-Linear models like logistic regression are popular for their ease of
+Linear models, like logistic regression, are popular for their ease of
+% FIXME: Are you sure you want to call LOGISTIC regression a LINEAR model? 
 training and of interpreting the resulting model; however, they can be

### 2017-04-02 paper/oopsla17-submission/evaluation-accuracy.tex
Wes pass through 4.2

@@
 We provide two baselines for the comparison: a random choice of location
-from a minimized type error slice, and the \ocaml compiler that
-programmers typically interact with.
+from a minimized type error slice, and the standard \ocaml compiler.
 
@@
 %
-In order to ensure a fair comparison when evaluating \mycroft and
-\sherrloc, we ignored program where they predicted locations that our
-oracle could not match with a program expression --- 6--8\% of programs
-for \mycroft and 3--4\% for \sherrloc.
+To ensure a fair comparison when evaluating \mycroft and
+\sherrloc, we removed from the dataset programs where they predicted
+locations that our oracle could not match with a program expression ---
+6--8\% of programs for \mycroft and 3--4\% for \sherrloc.
 %
-We also ignored programs where \mycroft or \sherrloc timed out (after
-one minute) or where they encountered an unsupported language feature ---
-another 5\% for \mycroft and 12--13\% for \sherrloc.
+We also do not consider programs where \mycroft or \sherrloc timed out
+(after one minute) or where they encountered an unsupported language
+feature --- another 5\% for \mycroft and 12--13\% for \sherrloc. This
+experimental design gives the state of the art tools the ``benefit of the
+doubt''.
 
@@
 %
-We preemptively discard expressions that are not part of the minimal
-type error slice --- we will explain the rationale for this in
-\autoref{sec:feature-utility} --- and thus the final feature count is
-276.
+% We preemptively discard expressions that are not part of the minimal
+% type error slice --- we will explain the rationale for this in
+% \autoref{sec:feature-utility} --- and thus the final feature count is
+% 276.
 %
@@
 \item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
-\item Two multi-layer perceptrons, both trained with $\alpha = 0.001$,
+\item Two multi-layer perceptron neural networks, both trained with $\alpha = 0.001$,
   $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
@@
   a hidden layer of 500 neurons. This allows us to investigate how well
-  the MLP can \emph{compress} its model.
+  the MLP can \emph{compress} its model (cf.~\cite{FIXME}).
 \end{enumerate}
@@
 %
-Localizing the type errors in our benchmarks amounted on average to
+Localizing the type errors in our benchmarks amounted, on average, to
 selecting one of 2--3 correct locations out of a slice of 10--11.
 %
-Our baseline of selecting at random is able to achieve 30\% Top-1
+Our baseline of selecting at random achieves 30\% Top-1
 accuracy (58\% Top-3), while \ocaml achieves a Top-1 accuracy of 45\%.
@@
 
-Surprisingly, there is little variation in accuracy between classifiers
-with the exception of the \linear model, they all achieve around 70\%
+Surprisingly, there is little variation in accuracy between classifiers.
+With the exception of the \linear model, they all achieve around 70\%
 Top-1 accuracy and around 90\% Top-3 accuracy.
 %
-This would appear to suggest that the model they learn is relatively
+This suggests that the model they learn is relatively
 simple.
@@
 vice versa, they appear to generalize better from the \FALL data. 
+FIXME: Why? What is your explanation for this? Is it just sizes of those 
+datasets or something qualitative about the program pairs in them?

### 2017-04-02 paper/oopsla17-submission/evaluation.tex
misc
minor
Wes pass through 4.2

@@
   %
-  Which feature sets are most important in general?
+  Which feature sets are \emph{most important} in general?
   % How much do the features described in \autoref{sec:features}
@@
   %
-  Can we explain \emph{individual} predictions (correct or not)?
+  Can we \emph{explain} individual predictions (correct or not)?
 \end{description}
@@
 undergraduate Programming Languages course at
-\if@anonymous
-AUTHOR INSTITUTION.
-\else
+\begin{anonsuppress}
 UC San Diego (IRB \#140608).
-\fi
+\end{anonsuppress}
+\begin{noanonsuppress}
+AUTHOR's INSTITUTION.
+\end{noanonsuppress}
 %
 We recorded each interaction with the \ocaml top-level system over the
-course of the first three assignments, which allowed us to capture
+course of the first three assignments, capturing
 ill-typed programs and, crucially, their subsequent fixes.
@@
 we extracted 2,333 pairs of programs, and the second comes from the Fall
-2015 class (\FALL), from which we extracted 2,273 pairs of programs.
+2015 class (\FALL), from which we extracted 2,273 pairs of programs. In
+total, programs from FIXME distinct student authors appear in the 
+dataset.
 %
@@
 (polymorphic) algebraic data types.
+% FIXME: This is a journal article. Why not just exhaustively list the
+% functional programming idioms? (cf. the course syllabus and the
+% assignment descriptions)
 
@@
   string literals, and a user-defined |expr| type of simple arithmetic
-  expressions. We will discuss the challenge of supporting other
+  expressions. We discuss the challenge of supporting other
   user-defined types in \autoref{sec:discussion}.
@@
   disabled. If an expression has more than three children, the
-  classifiers will be unable to learn anything about the additional
+  classifiers will receive no information about the additional
   children.
@@
   and for the contextual sub-expressions.
-\item 1 feature denoting the size of each sub-expression.
-\item 1 feature denoting whether each sub-expression is part of the
+\item one feature denoting the size of each sub-expression.
+\item one feature denoting whether each sub-expression is part of the
   minimal type error slice.
@@
 still retain the large majority of program pairs.
+% FIXME: Can you say that this dataset curation is similar to any other
+% datasets (e.g., the washington one)? Anything you could cite and discuss
+% here would take some of the pressure off.
 
 \subsubsection{Accuracy Metric}
-All of the tools we compare (with the exception of \ocaml) can produce a
-list of potential error locations, which might in the limit include the
-entire type error slice.
+All of the tools we compare (with the exception of the standard \ocaml
+compiler) can produce a list of potential error locations. 
 %
@@
 We evaluate each tool on whether a changed expression occurred in its
-top 1, top 2, or top 3 predictions.
+top one, top two, or top three predictions.
 

### 2017-04-02 paper/oopsla17-submission/evaluation-utility.tex
Wes pass to Section 5, including adding threats to validity ...

@@
 %
-In order to answer this question we investigate the performance of
+To answer this question we investigate the performance of
 classifiers trained on various subsets of the features.
@@
 does not have to learn the importance of \InSlice.
+
+% FIXME: Wes feels that there should be a sentence in the next paragraph
+% explaining to the reader why we didn't just use an ANOVA or the ReliefF
+% method or whatever to figure out feature importance. Feature overlap?
 
@@
 %
-As expected the baseline performs the worst, with a mere 23.6\% \linear
+As expected, the baseline performs the worst, with a mere 23.6\% \linear
 Top-1 accuracy.
@@
 
-Continuing on, we will investigate the relative impact of the other
+We investigate the relative impact of the other
 three classes of features discussed in \autoref{sec:features}, assuming
@@
 %
-(1) expression size,
-(2) contextual syntactic features, and
+(1) expression size;
+(2) contextual syntactic features; and
 (3) typing features.
 %
-As before we perform a 10-fold cross-validation with $\alpha = 0.001$,
+As before, we perform a 10-fold cross-validation with $\alpha = 0.001$,
 $\lambda = 0.001$, and a mini-batch size of 200, but we
@@
 \paragraph{Results}
-\autoref{tab:contextual-features} summarizes the results of our experiment.
+\autoref{tab:contextual-features} summarizes the results of this experiment.
 %
-As we can see, the \linear classifier and the \hiddenFH start off
+The \linear classifier and the \hiddenFH start off
 competitive when given only local syntactic features, but the \hiddenFH
-quickly begins to outperform as more features are added.
+quickly outperforms as more features are added.
 
-\ExprSize appears to be the weakest feature, improving \linear Top-1
+\ExprSize is the weakest feature, improving \linear Top-1
 accuracy by only 1\% and \hiddenFH by only 4\%.
@@
 in accuracy when added alongside the contextual or typing features.
+FIXME: Indicate that this is expected because of feature overlap: the
+contextual ``child'' features partially encode subtree size.
 
@@
 They improve Top-1 accuracy by an additional 2\%, and are much more
-compact --- we only have 55 typing features compared to 176 contextual
-syntactic features.
+compact --- requiring only have 55 typing features compared to 176
+contextual syntactic features.
 %

### 2017-04-02 paper/oopsla17-submission/evaluation.tex
Wes pass to Section 5, including adding threats to validity ...

@@
 
+\subsubsection{Threats to Validity}
+
+Although our experiments demonstrate that our technique can pinpoint type
+errors more accurately than the state of the art and that our features are
+relevant to blame assignment, our results may not generalize.
+
+One threat to validity associated with supervised machine learning is
+overfitting (\ie learning a model that is too complex with respect to the
+data). We mitigate this threat by having separate training and testing
+datasets (section FIXME) and via cross-validation (section
+FIXME)~\cite{FIXME}.
+
+FIXME threat associated with benchmarks that are not indicative. We
+mitigate this threat by including the largest empirical evaluation we are
+aware of (re-quote number of program pairs) from two different semesters of
+students (re-quote number of students). FIXME note that we ack that
+students are not indicative of industrial programmers, but the goal of this
+paper is to help novices, since they find type inference errors to be the
+most cryptic~\cite{FIXME}.
+
+FIXME threat associated with removing from the dataset program pairs
+that change everything or where our parser cannot find nodes corresponding
+to the other tools or whatever. FIXME Note that the former is
+standard or otherwise similar to previous work~\cite{FIXME}. FIXME note tha
+the latter actually helps the state of the art, and anyway our victory
+margins are so large that even if assumed they were perfect on all of those
+and we were not, we'd still win (double check that math).
+
+FIXME threat associated with using just OCaml instead of Haskell, Reason,
+whatever. FIXME admits comparison with previous work, FIXME our functional
+subset of ocaml is basically haskell anyway, FIXME Algorithm W is the same
+for all of them.
+
 \input{gallery}

### 2017-04-02 paper/oopsla17-submission/gallery.tex
Wes pass to Section 5, including adding threats to validity ...

@@
 We choose the Decision Tree classifier for this section as its model
-is more easily interpreted than the MLP, we will attempt to explain
-its predictions by analyzing the paths induced in the tree by the
-programs.
+is more easily interpreted than the MLP. We explain its predictions by
+analyzing the paths induced in the decision tree by the features of the
+input programs.
 %
@@
 |int list| returned by |clone| onto |l2|, which is another |int list|,
-but they decide to \emph{blame} the call to |clone|, while our Decision
+but they decide to blame the call to |clone|, while our Decision
 Tree correctly blames the |::| operator.
@@
 %
-As a hint we first have them write a |wwhile| function that performs the
+As a hint, we first have them write a |wwhile| function that performs the
 functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
@@
 \label{sec:failed-predictions}
-Of course, our classifiers are sometimes wrong, we focus next on programs
-where our classifier makes an incorrect prediction.
+We also describe indicative programs where our classifier makes an
+incorrect prediction.
 

### 2017-04-02 paper/oopsla17-submission/discussion.tex
Wes pass over rest of text

@@
 Probably the single biggest limitation of our technique is that we have
-features for specific data and type constructors.
+(a finite set of) features for specific data and type constructors.
 %
@@
 %
-We can attempt to mitigate this to some extent by adding generic
+We can mitigate this, to some extent, by adding generic
 syntactic features for data constructors and match expressions, but it
@@
 Finally, \emph{n-grams} on the token stream have proven effective for
-probabilistic modeling of programming languages, we may find that they
-aid in our task as well.
+probabilistic modeling of programming languages~\cite{FIXME,FIXME}, we may
+find that they aid in our task as well.
 %

### 2017-04-02 paper/oopsla17-submission/related.tex
Wes pass over rest of text

@@
 %
-A fault localization scheme based purely on control flow features
+For example, a fault localization scheme based purely on control flow features
 will have difficulty with cross-site scripting or SQL code injection

### 2017-04-04 paper/oopsla17-submission/evaluation-accuracy.tex
update some eval numbers, more to come

@@
 Localizing the type errors in our benchmarks amounted, on average, to
-selecting one of 2--3 correct locations out of a slice of 10--11.
+selecting one of 3 correct locations out of a slice of 10.
 %
@@
 Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
-ranging from 60--62\% Top-1 accuracy (85--88\% Top-3) for the \linear
-classifier to 71--73\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
+ranging from 59--63\% Top-1 accuracy (86--88\% Top-3) for the \linear
+classifier to 70--74\% Top-1 accuracy (91--92\% Top-3) for the \hiddenFH.
 
@@
 %
-This suggests that the model they learn is relatively
-simple.
+This suggests that the model they learn is relatively simple.
 %
@@
 trained on the \FALL programs and tested on the \SPRING programs than
-vice versa, they appear to generalize better from the \FALL data. 
-FIXME: Why? What is your explanation for this? Is it just sizes of those 
-datasets or something qualitative about the program pairs in them?
+vice versa.
+% , they appear to generalize better from the \FALL data.
+% FIXME: Why? What is your explanation for this? Is it just sizes of those
+% datasets or something qualitative about the program pairs in them?

### 2017-04-04 paper/oopsla17-submission/evaluation.tex
Wes change to SEction 4, including adding threats to validity

@@
 %
-The first dataset comes from the Spring 2014 class (\SPRING), from which
-we extracted 2,333 pairs of programs, and the second comes from the Fall
-2015 class (\FALL), from which we extracted 2,273 pairs of programs. In
-total, programs from FIXME distinct student authors appear in the 
-dataset.
+The first dataset comes from the Spring 2014 class (\SPRING), with a
+cohort of 46 students. The second comes from the Fall 2015 class
+(\FALL), with a cohort of 56 students.
 %
@@
 (polymorphic) algebraic data types.
-% FIXME: This is a journal article. Why not just exhaustively list the
-% functional programming idioms? (cf. the course syllabus and the
-% assignment descriptions)
 
-\subsubsection{Feature Selection}
-We extract a set of 277 features from each sub-expression in a student
+\paragraph{Feature Selection}
+We extract a set of 276 features from each sub-expression in a student
 program, including:
@@
 \item one feature denoting the size of each sub-expression.
-\item one feature denoting whether each sub-expression is part of the
-  minimal type error slice.
+% \item one feature denoting whether each sub-expression is part of the
+%   minimal type error slice.
 \end{enumerate}
 
-\subsubsection{Blame Oracle}
+\paragraph{Blame Oracle}
 We identified the fixes for each ill-typed program with an
@@
 \end{enumerate}
+
+Students may have made many, potentially unrelated, changes between
+compilations, but at some point the ``fix'' becomes a ``rewrite''.
 %
-As the students may have made many, potentially unrelated, changes
-between compilations, we discard any program pairs where more than 40\%
-of the sub-expressions have changed.
+We do not wish to consider the ``rewrites'' in our evaluation, so we
+discard outliers where the fraction of expressions that have changed is
+more than one standard deviation above the mean, or 44\%.
 %
-We picked 40\% as an estimate of the inflection point where we could
-still retain the large majority of program pairs.
-% FIXME: Can you say that this dataset curation is similar to any other
-% datasets (e.g., the washington one)? Anything you could cite and discuss
-% here would take some of the pressure off.
+This accounts for roughly 14\% of each dataset, leaving us with
+2,425 program pairs for \SPRING and 2,325 pairs for \FALL.
 
-\subsubsection{Accuracy Metric}
+% we discard any program pairs where more than 40\%
+% of the sub-expressions have changed.
+% %
+% We picked 40\% as an estimate of the inflection point where we could
+% still retain the large majority of program pairs.
+% % FIXME: Can you say that this dataset curation is similar to any other
+% % datasets (e.g., the washington one)? Anything you could cite and discuss
+% % here would take some of the pressure off.
+
+\paragraph{Accuracy Metric}
 All of the tools we compare (with the exception of the standard \ocaml
@@
 \input{evaluation-utility}
+\input{gallery}
 
-\subsubsection{Threats to Validity}
+\subsection{Threats to Validity}
 
@@
 
-\input{gallery}
-
 %%% Local Variables:

### 2017-04-04 paper/oopsla17-submission/discussion.tex
address some of wes' fixmes

@@
 Finally, \emph{n-grams} on the token stream have proven effective for
-probabilistic modeling of programming languages~\cite{FIXME,FIXME}, we may
-find that they aid in our task as well.
+probabilistic modeling of programming languages
+\citep{Hindle2012-hf,Gotterbarn1998-gc}, we may find that they aid in
+our task as well.
 %

### 2017-04-04 paper/oopsla17-submission/learning.tex
address some of wes' fixmes

@@
 \begin{table}[ht]
+\caption{Example Feature Vectors}\label{tab:sumList}
 \begin{tabular}{lrrrrrr}
@@
 \end{tabular}
-\caption{A selection of the features we would extract from the
+\bigskip
+\caption*{A selection of the features we would extract from the
 \lstinline!sumList! program in \autoref{fig:sumList}. A feature is
@@
 describes the parent of the current expression, a ``-C$n$'' suffix
-  indicates that the feature describes the $n$-th (left-to-right) child
-  of the current expression.  Note that, since we rely on a partial
-  typing derivation, we are subject to the well-known traversal bias and
-  label the expression \lstinline!sumList tl! as having type
-  $\tlist{\cdot}$. The model will have to learn to correct for this
-  bias.}
-\label{tab:sumList}
+indicates that the feature describes the $n$-th (left-to-right) child of
+the current expression.  Note that, since we rely on a partial typing
+derivation, we are subject to the well-known traversal bias and label
+the expression \lstinline!sumList tl! as having type
+$\tlist{\cdot}$. The model will have to learn to correct for this bias.}
 \end{table}
-
 The first issue we must tackle is formulating our learning task in
@@
 occurs, similar to the \emph{n-grams} commonly used in linguistic
-models~\cite{FIXME,FIXME}.
-% FIXME: cite "on the naturalness of software" and "on the uniqueness of
-% software" at the very least for the above
+models \citep{Hindle2012-hf,Gotterbarn1998-gc}.
 
@@
 %
-We rely on the model to counteract this bias (see FIXME reference to later
-in the paper).
+We rely on the model to counteract this bias (see \autoref{sec:evaluation}).
 

### 2017-04-04 paper/oopsla17-submission/overview.tex
address some of wes' fixmes

@@
 We formulate the problem of type error localization as a
-\emph{supervised classification} problem~\cite{FIXME}. 
-% Give a generic citation for supervised classification, such as an ML
-% textbook or seminal ML paper or survey, above.
+\emph{supervised classification} problem (see \citealt{Kotsiantis2007-pj}
+for a survey).
 %
@@
 Prior work has often enlisted expert users to manually judge ill-typed
-programs and determine the ``correct'' fix (e.g.,~\cite{FIXME,FIXME}), but
+programs and determine the ``correct'' fix \citep[e.g.][]{Loncaric2016-uk}, but
 this approach does not scale well to a dataset large enough to support
@@
 the results of a program analysis is popular in other problem domains
-(e.g., via statistics FIXME: cite Engler's z-ranking paper, etc.); we focus
+\citep[\eg via statistics][]{Kremenek2003-ck}; we focus
 explicitly on the use of data-driven machine learning confidence as a

### 2017-04-04 paper/oopsla17-submission/learning.tex
references

@@
 occurs, similar to the \emph{n-grams} commonly used in linguistic
-models \citep{Hindle2012-hf,Gotterbarn1998-gc}.
+models \citep{Hindle2012-hf,Gabel2010-el}.
 

### 2017-04-04 paper/oopsla17-submission/learning.tex
some (possibly) better text for logistic regression

@@
 \label{sec:models}
-\ES{TODO: mention that this is a SUPERVISED, CLASSIFICATION problem}
+\lstDeleteShortInline{|} % sigh...
+
+Recall that we formulate the localization problem as a supervised
+classification problem.
 %
+This means that we assume the existence of a training set
+$D = \{\langle \mathbf{x}_i, y_i \rangle\}_{i=1}^{n}$
+of sample inputs $\mathbf{x}$ and their corresponding labels $y$.
+%
+The learning task is to model the probability $Pr(y\ |\ \mathbf{x})$
+of a label being assigned to a given input.
+
 There are many learning algorithms to choose from, existing
@@
 %
-In this section we present the learning algorithms we will investigate,
-beginning with the simplest.
+In this section we present four standard learning algorithms that we
+will investigate:
 %
-For each algorithm, we describe \ES{TODO...}
+(1) logistic regression,
+(2) decision trees,
+(3) random forests, and
+(4) multi-layer perceptrons.
+%
+For each algorithm, we describe the form of the model it learns, the
+training procedure, and any hyper-parameters that must be supplied by a
+human.
 
 \paragraph{Logistic regression}
-The simplest model we investigate is logistic regression, which learns a
-linear function of the features.
+% The simplest model we investigate is logistic regression, which learns a
+% linear function of the features.
+% %
+% The goal is to learn a set of weights $W$ and biases $b$ such that the
+% function
+% %
+% $$
+% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-W\mathbf{x} - b)}
+% $$
+% %
+% effectively maps feature vectors $\mathbf{x}$ to labels $y$.
+
+% \ES{TODO: talk about logistic regression and the loss function?}
+
+The simplest model we investigate is logistic regression, which can be
+understood in two steps.
 %
-The goal is to learn a set of weights $W$ and biases $b$ such that the
-function
+First, we compute the \emph{evidence} for a label $y$ as a linear
+function of the input $\mathbf{x}$, with a set of weights $w_i$ and a
+bias $b$ that will be learned from the training set.
+$$
+\mathsf{ev} = \sum_i w_i \mathbf{x}_i + b
+$$
+Then we use the logistic function $f(x) = \frac{1}{1 + e^{-x}}$ to
+compress the evidence term into the $[0,1]$ interval, which can then be
+interpreted as the probability that the label $y$ should be applied.
 %
+This gives us the final model
 $$
-y = Wx + b
+Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_i w_i \mathbf{x}_i - b)}
 $$
+
+Training a logistic regression entails finding optimal values for the
+weights and bias, such that a \emph{cost} function is minimized.
 %
-effectively maps feature vectors $x$ to labels $y$.
+In practice, the cost function used for logistic regression is the
+\emph{cross entropy}, a measure of the similarity between two
+probability distributions, between the predicted labels and the ground
+truth.
+%
+The search for optimal values is typically done via \emph{stochastic
+  gradient descent}, an iterative method for optimizing a cost function.
+%
+Starting with an initial estimate of the weights (\eg a normal
+distribution), one repeatedly makes predictions for the training data,
+computes the cost function for the predictions, and updates the
+parameters according to the gradient of the cost function.
+%
+This process is repeated until the parameters converge, or until a time
+limit is exhausted.
+%
+Stochastic gradient descent requires a hyper-parameter $\eta$, often
+called the \emph{learning rate}, which controls how much the weights and
+bias are updated at each iteration.
 
-\ES{TODO: talk about logistic regression and the loss function?}
+It has been observed \ES{CITE??} that large weights often coincide with
+\emph{overfitting} of the model --- where the model performs well on the
+training samples but not on the testing samples.
+%
+Thus, it is common to add a \emph{regularization} term to the cost
+function, a popular choice is the L2 norm of the weights $\sum_i w_i^2$,
+which has the effect of penalizing the model for learning large weights.
+The contribution of the regularization term to the over cost function is
+controlled by another hyper-parameter $\lambda$, often called the
+regularization rate.
 
-Linear models, like logistic regression, are popular for their ease of
-% FIXME: Are you sure you want to call LOGISTIC regression a LINEAR model? 
-training and of interpreting the resulting model; however, they can be
-limited in their applicability.
+Logistic regression is popular for its ease of training and of
+interpreting the resulting model; however, it can be limited in its
+applicability.
 %
-As the name implies, linear models may perform poorly when asked to
+As a generalized linear model, it may perform poorly when asked to
 learn an inherently \emph{nonlinear} function.
@@
 
+% ES: keep this at the bottom... sigh
+\lstMakeShortInline{|}
+
 %%% Local Variables:

### 2017-04-04 paper/oopsla17-submission/learning.tex
some text about neural netwroks

@@
 
-Recall that we formulate the localization problem as a supervised
+Recall that we formulate type error localization as a supervised
 classification problem.
@@
 This means that we assume the existence of a training set
-$D = \{\langle \mathbf{x}_i, y_i \rangle\}_{i=1}^{n}$
+$S = \{\langle \mathbf{x}_i, y_i \rangle\}_{i=1}^{n}$
 of sample inputs $\mathbf{x}$ and their corresponding labels $y$.
@@
 (3) random forests, and
-(4) multi-layer perceptrons.
+(4) neural networks.
 %
@@
 
-\paragraph{Logistic regression}
+\paragraph{Logistic Regression}
 % The simplest model we investigate is logistic regression, which learns a
@@
 First, we compute the \emph{evidence} for a label $y$ as a linear
-function of the input $\mathbf{x}$, with a set of weights $w_i$ and a
-bias $b$ that will be learned from the training set.
+function of the feature vector $\mathbf{x}$, using a weight vector
+$\mathbf{w}$ and a bias $b$ that will be learned from the training set.
 $$
-\mathsf{ev} = \sum_i w_i \mathbf{x}_i + b
+\mathsf{ev} = \sum_j w_j x_j + b
 $$
-Then we use the logistic function $f(x) = \frac{1}{1 + e^{-x}}$ to
-compress the evidence term into the $[0,1]$ interval, which can then be
-interpreted as the probability that the label $y$ should be applied.
+Then we use the logistic function (also called a sigmoid function)
+$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
+$[0,1]$ interval, which can then be interpreted as the probability that
+the label $y$ should be applied.
 %
-This gives us the final model
+This gives us the final model:
 $$
-Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_i w_i \mathbf{x}_i - b)}
+Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j x_j - b)}
 $$
@@
 Thus, it is common to add a \emph{regularization} term to the cost
-function, a popular choice is the L2 norm of the weights $\sum_i w_i^2$,
+function, a popular choice is the $L_2$ norm of the weights $\sum_j w_j^2$,
 which has the effect of penalizing the model for learning large weights.
-The contribution of the regularization term to the over cost function is
+The contribution of the regularization term to the overall cost function is
 controlled by another hyper-parameter $\lambda$, often called the
@@
 
-Logistic regression is popular for its ease of training and of
-interpreting the resulting model; however, it can be limited in its
-applicability.
+As a generalized linear model, logistic regression is popular for its
+ease of training and of interpreting the resulting model; however, it
+can be limited in its applicability.
 %
-As a generalized linear model, it may perform poorly when asked to
-learn an inherently \emph{nonlinear} function.
+In particular, it may perform poorly when asked to learn an inherently
+\emph{nonlinear} function.
 %
@@
 
-\paragraph{Decision trees}
+\paragraph{Decision Trees}
 \ES{TODO: Maybe Huma?}
 
-\paragraph{Random forests}
+\paragraph{Random Forests}
 \ES{TODO: Maybe Huma?}
 
-\paragraph{Neural networks}
-The last (and most complex) model we discuss is Artificial Neural
-Networks (ANN).
+\paragraph{Neural Networks}
+The last (and most complex) model we discuss is a type of neural network
+called a \emph{multi-layer perceptron} \ES{CITE}.
+%
+Neural networks can be understood as a collection of linear models, just
+like the one used in logistic regression, arranged in a directed graph.
+%
+Each node in the graph corresponds to a single linear model, called a
+\emph{neuron}, and the edges propagate signals from the input features,
+through the neurons, to the output labels.
+%
+The linear model in each neuron is wrapped in an \emph{activation
+  function}, which controls whether (or how strongly) the signal will be
+propagated out.
+%
+These activation functions are what allow neural networks to accurately
+model nonlinear functions.
+%
+A common choice of activation function is the same sigmoid function
+used by logistic regression, alternatives include the hyperbolic $\tanh$
+function, and more recently the \emph{rectified linear unit} (ReLU)
+$f(x) = \mathsf{max}(0,x)$ \ES{CITE}.
 
-The fundamental building block of an ANN is the linear model, the
-nonlinearity arises from the \emph{activation functions} that govern
-whether an individual linear model propagates its signal to the next layer.
+The neurons are arranged in \emph{layers}; each layer contains a set of
+neurons that take in signals from the previous layer and propagate
+signals to the next layer.
+%
+The number of layers, the number of neurons per layer, and the
+connections between layers constitute the \emph{architecture} of a
+neural network, and are all important hyper-parameters.
+%
+We consider a particular class of neural networks called
+\emph{multi-layer perceptrons} (MLP), where the layers a fully connected to
+each other, \ie each neuron propagates its signal to every neuron in the
+subsequent layer.
+%
+We will only consider MLPs with a single layer of neurons, but we will
+vary the number of neurons in our experiments.
+
+Neural networks are trained similarly to logistic regression using
+stochastic gradient descent, and thus the learning rate $\eta$ and
+regularization rate $\lambda$ are also important hyper-parameters.
+
+\ES{should mention that logistic regression can be thought of as a degenerate MLP with 0 layers, and perhaps add a diagram?}
+
+\ES{need to say something about neurons corresponding to ``hidden'' features, give intuition..}
+
+% The fundamental building block of an ANN is the linear model, the
+% nonlinearity arises from the \emph{activation functions} that govern
+% whether an individual linear model propagates its signal to the next layer.
 

### 2017-04-04 paper/oopsla17-submission/evaluation-accuracy.tex

@@
 \item A logistic regression trained with a learning rate
-  $\alpha = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
+  $\eta = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
   mini-batch size of 200.
@@
 \item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
-\item Two multi-layer perceptron neural networks, both trained with $\alpha = 0.001$,
+\item Two multi-layer perceptron neural networks, both trained with $\eta = 0.001$,
   $\lambda = 0.001$, and a mini-batch size of 200. The first MLP

### 2017-04-04 paper/oopsla17-submission/evaluation-utility.tex

@@
 classifiers, cross-validated ($k=10$) over the combined SP14/FA15
-dataset. We used a learning rate $\alpha=0.001$, L2 regularization rate
+dataset. We used a learning rate $\eta=0.001$, L2 regularization rate
 $\lambda=0.001$, and mini-batch size of 200. We trained for a single
@@
 %
-As before, we perform a 10-fold cross-validation with $\alpha = 0.001$,
+As before, we perform a 10-fold cross-validation with $\eta = 0.001$,
 $\lambda = 0.001$, and a mini-batch size of 200, but we

### 2017-04-04 paper/oopsla17-submission/introduction.tex
Wes pass over Section 2.
address some of wes' fixmes

@@
 from their source, an annoyance to veterans of the type system, but a
-major hurdle for novices~\cite{FIXME,FIXME}.
+major hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.
 
@@
   is in \textbf{bold}.
-
-  FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
-  ``with'' are also bolded (by \\ecode)! You need to find another way to
-  highlight what ocaml is yelling about.
-
+  %
+  % FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
+  % ``with'' are also bolded (by \\ecode)! You need to find another way to
+  % highlight what ocaml is yelling about.
+  %
   (right) The error reported by \ocaml.}

### 2017-04-04 paper/oopsla17-submission/learning.tex
some decision tree text

@@
 \paragraph{Decision Trees}
-\ES{TODO: Maybe Huma?}
+Decision tree algorithms learn a tree of binary predicates over the
+features, recursively partitioning the input space until a final
+classification can be made.
+%
+Each node in the tree contains a single predicate of the form
+$x_j \leq t$ for some threshold $t$, which determines whether a given
+input should proceed down the left or right subtree.
+%
+Each leaf is labeled with a prediction and the fraction of training
+samples that would reach it, the latter quantity can be interpreted as
+the decision tree's confidence in its prediction.
+\ES{HUMA: is this correct?}
+
+Training a decision tree entails finding both a set of good partitioning
+predicates and a good ordering of the predicates.
+%
+This is usually done in a top-down greedy manner by selecting the
+predicate the best partitions the entire training set, and recursively
+partitioning the subsets.
+%
+Standard algorithms for training decision trees include C4.5 \ES{CITE} and
+CART \ES{CITE}.
+%
+Overfitting can also be a problem for decision trees; the countermeasure
+is usually either halting the learning process before the tree grows too
+deep, or pruning the resulting tree.
+
+The advantages of decision trees are that they are relatively easy to
+train and that they produce a \emph{white-box} model, making it trivial
+to analyze the entire model and explain individual predictions.
+%
+The disadvantages are that they are prone to overfitting and that the
+model can be unstable, \ie small changes to the training data may
+produce large changes in the model.
+%
+Both disadvantages are addressed by random forests.
 

### 2017-04-05 paper/oopsla17-submission/learning.tex
more learning algos

@@
 %
-For each algorithm, we describe the form of the model it learns, the
-training procedure, and any hyper-parameters that must be supplied by a
-human.
+For each algorithm, we describe the model it learns, the training
+procedure, any hyper-parameters that must be tuned by a human, and a
+summary of its advantages and disadvantages.
 
@@
 $$
+% $$
+% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + e^{-\mathsf{ev}}}
+% $$
 
@@
 Each node in the tree contains a single predicate of the form
-$x_j \leq t$ for some threshold $t$, which determines whether a given
-input should proceed down the left or right subtree.
+$x_j \leq t$ for some feature $x_j$ and threshold $t$, which determines
+whether a given input should proceed down the left or right subtree.
 %
@@
 \paragraph{Random Forests}
-\ES{TODO: Maybe Huma?}
+One technique for increasing the accuracy of a classifier is to train an
+\emph{ensemble} of distinct classifiers and use a majority vote
+to make the final prediction \ES{CITE?}.
+%
+Random forests \ES{CITE} are an instance of this method, with decision
+trees as the underlying classifiers.
+%
+The training process involves taking $N$ random subsets of the training
+data and training a separate decision tree on each subset --- the
+training process for the decision trees is often modified slightly to
+reduce correlation between trees, by forcing the tree to pick features
+from a random subset of all features at each node.
+%
+In order to make a prediction for an unseen sample, the random forest
+asks each decision tree to make a prediction, and then predicts the most
+common label.
+
+The diversity of the underlying models tends to make random forests less
+susceptible to the overfitting and instability of decision trees, but it
+also makes the learned model more difficult to interpret.
+%
+Furthermore, the number of trees in the forest, $N$, becomes an important
+hyper-parameter that must be set manually.
 
@@
 %
-A common choice of activation function is the same sigmoid function
-used by logistic regression, alternatives include the hyperbolic $\tanh$
-function, and more recently the \emph{rectified linear unit} (ReLU)
-$f(x) = \mathsf{max}(0,x)$ \ES{CITE}.
+A common choice of activation function is the same sigmoid function used
+by logistic regression, alternatives include the hyperbolic $\tanh$
+function, and more recently the \emph{rectified linear unit} (ReLU),
+defined as $f(x) = \mathsf{max}(0,x)$ \ES{CITE}.
 
@@
 regularization rate $\lambda$ are also important hyper-parameters.
+%
+They are, however, much more difficult to train as they require larger
+training sets.
+
+Neural networks are popular for their ability to uncover ``hidden''
+features in the data.
+%
+Each neuron learns a different linear combination of the signals from
+the previous layer; combined with the activation function this can be
+thought of as the neuron learning to respond to a new, \emph{compound}
+feature of its inputs.
+%
+This ability to discover interesting combinations of features makes
+neural networks easier to use in some respect; they can reduce the
+importance of manual \emph{feature engineering} (selecting interesting
+features to provide as inputs).
+%
+However, the sheer size and complexity of neural networks makes the
+model particularly difficult to interpret --- they are generally regarded
+as ``black boxes.''
+%
+Neural networks are also very susceptible to overfitting, so
+regularization methods like the $L_2$ norm used for logistic regression
+are important.
 

### 2017-04-05 paper/oopsla17-submission/evaluation-accuracy.tex
eval

@@
 \item A logistic regression trained with a learning rate
-  $\eta = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
+  $\eta = 0.001$, an $L_2$ regularization rate $\lambda = 0.001$, and a
   mini-batch size of 200.
@@
   a hidden layer of 500 neurons. This allows us to investigate how well
-  the MLP can \emph{compress} its model (cf.~\cite{FIXME}).
+  the MLP can \emph{compress} its model (cf.~\cite{FIXME}). The neurons
+  use rectified linear units (ReLU) as their activation function, a
+  common practice in modern neural networks.
 \end{enumerate}
@@
 before being evaluated on the other.
+%
+The logistic regression and MLPs were trained with the \textsc{Adam}
+optimizer \ES{CITE}, a variant of stochastic gradient descent that has
+been found to converge faster.
 

### 2017-04-05 paper/oopsla17-submission/overview.tex
some (possibly) better text for logistic regression
some notes from huma meeting

@@
 related to the use of fault localization in other program analysis
-and transformation techniques (see Section~\ref{sec:related-work}).
+and transformation techniques (see \autoref{sec:related-work}).
 
@@
 the results of a program analysis is popular in other problem domains
-\citep[\eg via statistics][]{Kremenek2003-ck}; we focus
+\citep[see, \eg][]{Kremenek2003-ck}; we focus
 explicitly on the use of data-driven machine learning confidence as a

### 2017-04-05 paper/oopsla17-submission/evaluation.tex
expand threats to validity

@@
 One threat to validity associated with supervised machine learning is
-overfitting (\ie learning a model that is too complex with respect to the
-data). We mitigate this threat by having separate training and testing
-datasets (section FIXME) and via cross-validation (section
-FIXME)~\cite{FIXME}.
+overfitting (\ie learning a model that is too complex with respect to
+the data).
+%
+We mitigate this threat by having separate training and testing datasets
+(\autoref{sec:quantitative}) and via cross-validation
+(~\autoref{sec:feature-utility})~\cite{FIXME}.
 
-FIXME threat associated with benchmarks that are not indicative. We
-mitigate this threat by including the largest empirical evaluation we are
-aware of (re-quote number of program pairs) from two different semesters of
-students (re-quote number of students). FIXME note that we ack that
-students are not indicative of industrial programmers, but the goal of this
-paper is to help novices, since they find type inference errors to be the
-most cryptic~\cite{FIXME}.
+Our benchmarks were drawn from students in an undergraduate course and
+may not be representative of other student populations.
+%
+We mitigate this threat by including the largest empirical evaluation of
+type error localization that we are aware of: over 4,500 pairs of
+ill-typed programs and fixes from two instances of the course, with
+programs from 102\ES{TODO: check} different students.
+%
+We acknowledge, of course, that students are not industrial programmers
+and our results may not translate to large-scale software development;
+however, we are particularly interested in aiding novice programmers
+as they learn to work inside the type system.
 
-FIXME threat associated with removing from the dataset program pairs
-that change everything or where our parser cannot find nodes corresponding
-to the other tools or whatever. FIXME Note that the former is
-standard or otherwise similar to previous work~\cite{FIXME}. FIXME note tha
-the latter actually helps the state of the art, and anyway our victory
-margins are so large that even if assumed they were perfect on all of those
-and we were not, we'd still win (double check that math).
+Our removal of program pairs that changed too much, where our oracle
+could not identify the blame of the other tools, or where the other
+tools timed out or encountered unsupported language features is another
+threat to validity.
+%
+It is possible that including the programs that changed excessively
+would hurt our models, or that the other tools would perform
+better on the programs with unsupported language features.
+%
+We note however that
+%
+(1) outlier removal is a standard technique in machine learning
+\ES{CITE?}; and
+%
+(2) our Top-1 accuracy margin is large enough that even if we assumed
+that \sherrloc were perfect on all of the excluded programs, it would
+only tie our \hiddenFH in Top-1 accuracy.
+%
+\ES{I think this is accurate, but should double check..}
 
-FIXME threat associated with using just OCaml instead of Haskell, Reason,
-whatever. FIXME admits comparison with previous work, FIXME our functional
-subset of ocaml is basically haskell anyway, FIXME Algorithm W is the same
-for all of them.
+Examining programs written in \ocaml as opposed to \haskell or any other
+language poses another threat to validity, common type errors may differ
+in different languages.
+%
+\ocaml is, however, a standard target for research in type error
+localization and thus our choice admits a direct comparison with prior
+work.
+%
+Furthermore, the functional core of \ocaml that we support does not
+differ significantly from the functional core of \haskell or SML, all of
+which are effectively lambda calculi with a Hindley-Milner-style type
+system.\footnote{\haskell's type classes are a notable exception, they
+  are also known to cause confusing type errors and would be interesting
+  to study in future work.}
+
+Finally, our use of student fixes as oracles for the source of type
+errors assumes that, on average, students are able to correctly identify
+the source of an error.
+%
+As the students are in the process of learning the language and type
+system, this assumption may be faulty.
+%
+It may be that \emph{expert} users would disagree with many of the
+student fixes, and that it is harder to learn a model of expert fixes,
+or that the state of the art would be better at predicting expert fixes.
+%
+As we have noted before, we believe it is reasonable to use student
+fixes as oracles because the student is the best judge of what she
+\emph{meant} to do.
 

### 2017-04-06 paper/oopsla17-submission/evaluation.tex
rework gallery with explanations

@@
 (\autoref{sec:quantitative}) and via cross-validation
-(~\autoref{sec:feature-utility})~\cite{FIXME}.
+(\autoref{sec:feature-utility})~\cite{FIXME}.
 

### 2017-04-06 paper/oopsla17-submission/discussion.tex
add some more citations

@@
 probabilistic modeling of programming languages
-\citep{Hindle2012-hf,Gotterbarn1998-gc}, we may find that they aid in
+\citep{Hindle2012-hf,Gabel2010-el}, we may find that they aid in
 our task as well.
@@
 %
-We compensate somewhat for our lack of structure by ranking our
-predictions by ``confidence'', but it would be interesting to see how
-structured learning over graphical models would perform.
+We compensate somewhat for our lack of structure by adding contextual
+features and by ranking our predictions by ``confidence'', but it would
+be interesting to see how structured learning over graphical models
+would perform.
 

### 2017-04-06 paper/oopsla17-submission/evaluation-accuracy.tex
add some more citations

@@
 The logistic regression and MLPs were trained with the \textsc{Adam}
-optimizer \ES{CITE}, a variant of stochastic gradient descent that has
-been found to converge faster.
+optimizer \citep{Kingma2014-ng}, a variant of stochastic gradient
+descent that has been found to converge faster.
 

### 2017-04-06 paper/oopsla17-submission/learning.tex
add some more citations

@@
 %
-For each algorithm, we describe the model it learns, the training
-procedure, any hyper-parameters that must be tuned by a human, and a
-summary of its advantages and disadvantages.
+A thorough introduction to these techniques can be found in introductory
+machine learning textbooks (\eg \ES{TODO: huma/kamalika?}).
+%
+We will briefly introduce each technique by describing the model it
+learns, the training procedure, any hyper-parameters that must be tuned
+by a human, and a summary of its advantages and disadvantages.
 
@@
 
-It has been observed \ES{CITE??} that large weights often coincide with
+It has been observed that large weights often coincide with
 \emph{overfitting} of the model --- where the model performs well on the
@@
 Thus, it is common to add a \emph{regularization} term to the cost
-function, a popular choice is the $L_2$ norm of the weights $\sum_j w_j^2$,
-which has the effect of penalizing the model for learning large weights.
-The contribution of the regularization term to the overall cost function is
-controlled by another hyper-parameter $\lambda$, often called the
+function, a popular choice is the $L_2$ norm of the weights
+$\sum_j w_j^2$, which has the effect of penalizing the model for
+learning large weights \citep{Park2008-no}.
+%
+The contribution of the regularization term to the overall cost function
+is controlled by another hyper-parameter $\lambda$, often called the
 regularization rate.
@@
 %
-Standard algorithms for training decision trees include C4.5 \ES{CITE} and
-CART \ES{CITE}.
+Standard algorithms for training decision trees include C4.5
+\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
 %
@@
 \emph{ensemble} of distinct classifiers and use a majority vote
-to make the final prediction \ES{CITE?}.
+to make the final prediction.
 %
-Random forests \ES{CITE} are an instance of this method, with decision
+Random forests are an instance of this method, with decision
 trees as the underlying classifiers.
@@
 The last (and most complex) model we discuss is a type of neural network
-called a \emph{multi-layer perceptron} \ES{CITE}.
+called a \emph{multi-layer perceptron} (see \citealt{Nielsen2015-pu} for
+a thorough introduction to neural networks).
 %
@@
 function, and more recently the \emph{rectified linear unit} (ReLU),
-defined as $f(x) = \mathsf{max}(0,x)$ \ES{CITE}.
+defined as $f(x) = \mathsf{max}(0,x)$ \citep{Nair2010-xg}.
 

### 2017-04-06 paper/oopsla17-submission/learning.tex
minor

@@
 %
-This limitation can be mitigated to some extent by adding more features
-to the model (\eg combinations of the existing features \ES{CITE?}), but
-the fundamental issue remains.
+This limitation can be mitigated by adding transformed or compound
+features to the model --- \eg if one knows that the label has a
+quadratic relationship with a feature $x$, the transformed feature $x^2$
+can be added --- but this typically requires knowledge of the underlying
+model.
+
+% This limitation can be mitigated to some extent by adding more features
+% to the model (\eg combinations of the existing features \ES{CITE?}), but
+% the fundamental issue remains.
 

### 2017-04-06 paper/oopsla17-submission/evaluation-accuracy.tex
explain from beginning that we use In-Slice as a hard filter

@@
 We provide two baselines for the comparison: a random choice of location
-from a minimized type error slice, and the standard \ocaml compiler.
+from the minimized type error slice, and the standard \ocaml compiler.
 

### 2017-04-06 paper/oopsla17-submission/evaluation.tex
add some more citations
explain from beginning that we use In-Slice as a hard filter

@@
 \end{description}
-\ES{may want to swap (1) and (2)..}
+%\ES{may want to swap (1) and (2)..}
 
@@
 \paragraph{Feature Selection}
-We extract a set of 276 features from each sub-expression in a student
+We extract a set of 277 features from each sub-expression in a student
 program, including:
@@
 \item one feature denoting the size of each sub-expression.
-% \item one feature denoting whether each sub-expression is part of the
-%   minimal type error slice.
+\item one feature denoting whether each sub-expression is part of the
+  minimal type error slice. We use this feature as a ``hard''
+  constraint, sub-expressions that are not part of the minimal slice
+  will be preemptively discarded, we justify this decision in
+  \autoref{sec:feature-utility}.
 \end{enumerate}

### 2017-04-06 paper/oopsla17-submission/learning.tex
explain from beginning that we use In-Slice as a hard filter

@@
 
-Note that our use of typing features in an ill-typed program
-subjects us to \emph{traversal
-bias}~\cite{McAdam1998-ub,Yang2000-yr}. For example, the 
-|sumList tl| expression might alternatively be assigned
-the type $\tint$.
+Note that our use of typing features in an ill-typed program subjects us
+to \emph{traversal bias} \citep{McAdam1998-ub}. For example, the
+|sumList tl| expression might alternatively be assigned the type
+$\tint$.
 %
-We rely on the model to counteract this bias (see \autoref{sec:evaluation}).
+Our models will have to learn good localizations in spite this bias (see
+\autoref{sec:evaluation}).
 
 \subsubsection{Type error slice}
-Finally, we desire a model that distinguishes between
-changes that could fix the error, and changes that
-\emph{cannot possibly} fix the error.
+Finally, we wish to distinguish between changes that could fix the
+error, and changes that \emph{cannot possibly} fix the error.
 %
@@
 a minimal slice for each error.
+
+In practice, we have found that \InSlice is a particularly important
+feature, and thus include a pre-processing step that discards all
+expressions where it is disabled.
+%
+As a result, the |tl| expression would never actually be shown to the
+classifier.
+%
+ We will demonstrate the importance of \InSlice empirically in
+\autoref{sec:feature-utility}.
 

### 2017-04-06 paper/oopsla17-submission/overview.tex
explain from beginning that we use In-Slice as a hard filter

@@
 %
-A minimal type error slice of a program includes all expressions that
-are necessary for the error to manifest and no more. We propose to
-use type error slices to communicate to the classifier which expressions
-could potentially be blamed --- a change to an expression outside of the
-minimal slice cannot possibly fix the type error. This is conceptually
-related to the use of fault localization in other program analysis
-and transformation techniques (see \autoref{sec:related-work}).
+A minimal type error slice \citep{Haack2003-vc} includes all expressions that
+are necessary for the error to manifest and no more.
+%
+We propose to use type error slices to communicate to the classifier
+which expressions could potentially be blamed --- a change to an
+expression outside of the minimal slice cannot possibly fix the type
+error.
+%
+We will show that the type error slice is so important
+(\autoref{sec:feature-utility}) that it is actually beneficial to
+automatically discard expressions that are not part of the slice, rather
+than letting the classifier learn to do so.
+%
+Our use of type error slices is conceptually related to the use of fault
+localization in other program analysis and transformation techniques
+(see \autoref{sec:related-work}).
 

### 2017-04-07 paper/oopsla17-submission/evaluation-utility.tex

@@
 \label{sec:type-error-slice}
-The \InSlice feature should be highly predictive --- a fix must change
-at least one expression in the type error slice.
+First we must justify our decision to automatically exclude expressions
+outside the minimal type error slice from consideration.
 %
-Thus, our first experiment seeks to quantify the impact of \InSlice by
-comparing the accuracy of our classifiers on three sets of features:
+% The \InSlice feature should be highly predictive --- a fix must change
+% at least one expression in the type error slice.
+% %
+% Thus, our first experiment seeks to quantify the impact of \InSlice by
+% comparing the accuracy of our classifiers on three sets of features:
+%
+Thus, we compare the accuracy of our classifiers on three sets of features:
 %
 \begin{enumerate}
-\item A baseline with \emph{only} local syntactic features.
+\item A baseline with only local syntactic features and no
+  preemptive filtering by \InSlice.
 \item The features of (1) extended with \InSlice.
 \item The same features as (1), but we preemptively discard samples
-  where \InSlice is \emph{disabled}.
+  where \InSlice is disabled.
 \end{enumerate}
@@
 classifiers, cross-validated ($k=10$) over the combined SP14/FA15
-dataset. We used a learning rate $\eta=0.001$, L2 regularization rate
+dataset. We used a learning rate $\eta=0.001$, $L_2$ regularization rate
 $\lambda=0.001$, and mini-batch size of 200. We trained for a single
@@
     \midrule
-    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
-    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
-    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
+    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 32.4\% & 49.3\% & 60.4\% & 39.6\% \\
+    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 56.1\% & 71.6\% & 82.7\% & 57.4\% \\
+    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 57.8\% & 72.8\% & 83.3\% & 57.8\% \\
     \bottomrule
@@
 %
-Top-1 accuracy jumps from 28.9\% to 54.4\% when we add \InSlice, and only
-improves by 2\% when we filter out expressions that are not part of the
+Top-1 accuracy jumps from 32.4\% to 56.1\% when we add \InSlice, and only
+improves by 1.7\% when we filter out expressions that are not part of the
 error slice.
 
-Still, the 2\% boost comes at zero cost, and given the other benefits
+Still, the accuracy gain comes at zero cost, and given the other benefits
 of filtering out expressions that do not belong to the type error slice
@@
     \midrule
-    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
+    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 57.7\% & 72.4\% & 83.0\% & 57.7\% \\
     \midrule
-    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
-    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
-    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
+    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.6\% & 75.0\% & 83.5\% & 57.2\% \\
+    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.5\% & 84.4\% & 90.5\% & 68.8\% \\
+    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 74.0\% & 85.5\% & 90.8\% & 69.1\% \\
     \midrule
-    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
-    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
-    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
+    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.4\% & 84.1\% & 90.8\% & 69.0\% \\
+    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.7\% & 86.3\% & 90.9\% & 69.5\% \\
+    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.6\% & 88.1\% & 92.9\% & 72.9\% \\
     \midrule
-    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
+    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 76.8\% & 87.9\% & 92.7\% & 72.5\% \\
     \bottomrule

### 2017-04-07 paper/oopsla17-submission/evaluation.tex

@@
 ill-typed programs and fixes from two instances of the course, with
-programs from 102\ES{TODO: check} different students.
+programs from 102 different students.
 %

### 2017-04-07 paper/oopsla17-submission/gallery.tex
regenerate ml files, hopefully for the last time!

@@
 %
-(3) the prediction of our Decision Tree (\underline{underlined}), and
+(3) the prediction of our \dectree classifier (\underline{underlined}), and
 %
@@
 %
-We choose the Decision Tree classifier for this section as its model
-is more easily interpreted than the MLP. We explain its predictions by
-analyzing the paths induced in the decision tree by the features of the
-input programs.
+We choose the \dectree classifier for this section as its model is more
+easily interpreted than the MLP.\@
+%
+We also exclude the \ExprSize feature from the model used in this
+section, as it makes the predictions harder to motivate, and as we saw
+in \autoref{sec:feature-utility} it does not appear to contribute
+significantly to the model's accuracy.
+%
+We explain the predictions by analyzing the paths induced in
+the decision tree by the features of the input programs.
 %
@@
 
+\subsubsection{Failed Predictions}
+\label{sec:failed-predictions}
+We begin with a few programs where our classifier fails to
+make the correct prediction.
+%
+For these programs we will additionally \colorbox{red!50}{highlight} the
+correct blame location.
+
+\paragraph{Constructing a List of Duplicates}
+% data/sp14/0219.ml
+Our first program is a simple recursive function |clone| that takes an
+item |x| and a count |n|, and produces a list containing |n| copies of
+|x|.
+%
+\begin{ecode}
+  let rec clone x n =
+    let loop acc n =
+      if n <= 0 then
+        acc
+      else
+        (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
+    loop [] n
+\end{ecode}
+% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
+%
+The student has defined a helper function |loop| with an accumulator
+|acc|, likely meant to call itself tail-recursively.
+%
+Unfortunately, she has called the top-level function |clone| rather than
+|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
+for the |x| argument to |clone|.
+
+Our classifier incorrectly predicts that the use of |x| in the recursive
+call is the most likely source of the error.
+%
+Our second prediction coincides with \sherrloc (and \ocaml), blaming the
+the first argument to |clone|.
+%
+This is also incorrect, but may be more helpful than our first
+prediction --- if our student decides that she has certainly provided
+the correct \emph{argument}, an alternative explanation is that
+perhaps she has called the wrong \emph{function}.
+
+We confess that both of these predictions are difficult to explain by
+examining the induced paths.
+%
+In particular, neither prediction references a property of the current
+expression, only its context, which is surprising.
+%
+Much clearer is why we fail to blame the occurrence of |clone|, the two
+enabled features on the path are:
+%
+(1) the parent is an application; and
+%
+(2) |clone| has a function type.
+%
+The model appears to have learned that programmers typically call the
+correct function.
+
+\paragraph{Currying Considered Harmful?}
+% data/sp14/1204.ml
+Our next example is another ill-fated attempt at |clone|.
+%
+\begin{ecode}
+  let rec clone x n =
+    let rec loop (*@\colorbox{red!50}{x n acc}@*) =
+      if n < 0 then
+        acc
+      else
+        ==loop== __(x, (n - 1), (x :: acc))__ in
+    loop (x, n, [])
+\end{ecode}
+The issue here is that \ocaml functions are \emph{curried} by default
+--- \ie they take their arguments one at a time --- but our student has
+called the inner |loop| with all three arguments in a tuple.
+%
+Many experienced functional programmers would choose to keep |loop|
+curried and rewrite the calls, however our student decides instead to
+\emph{uncurry} |loop|, making it take a tuple of arguments.
+%
+\sherrloc blames the recursive call to |loop| while our classifier
+blames the tuple of arguments --- a reasonable suggestion, but alas not
+the answer the student was looking for.
+
+We fail to blame the definition of |loop| because it is defining a
+function.
+%
+First, note that we represent |let f x y = e|\ \ as\ \ |let f = fun x -> fun y -> e|,
+thus a change to the pattern |x| would be treated as a change to the outer
+|fun| expression.
+%
+With this in mind, we can explain our failure to blame the definition of
+|loop| (the outer |fun|) as follows:
+%
+(1) it has a function type;
+%
+(2) its child is a |fun|; and
+%
+(3) its parent is a |let|.
+%
+Thus it appears to the model that the outer |fun| is simply part of a
+function definition, a common and innocuous phenomenon.
+
+% \paragraph{Composing Functions}
+% % data/sp14/3041.ml
+% Next, let us consider the |pipe| function that composes a list of
+% functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
+% function |fun x -> h (g (f x))|.
+% %
+% \begin{ecode}
+%   let pipe fs =
+%     let f a x y = ==y== __(a y)__ in
+%     let base x = x in
+%     List.fold_left f base fs
+% \end{ecode}
+% %
+% The error in our student's code is that she has applied |y| rather than
+% |x| to the result of |a y|.
+% %
+% \sherrloc correctly blames the first occurrence of |y|, while our
+% classifier (incorrectly) blames the application |a y| (\ocaml blames
+% the occurrence of |base| on line 4).
+% %
+% \ES{this has the same explanation as the 1st clone above}
+
+\subsubsection{Correct Predictions}
+\label{sec:correct-predictions}
+Next, we present a few indicative programs where our first prediction is
+correct, and all of the other tools' top three predictions are
+incorrect.
+
 \paragraph{Extracting the Digits of an Integer}
-% data/sp14/1655.ml
-Our first program is a simple recursive function |digitsOfInt| that
-extracts the digits of an |int|.
+% data/sp14/2176.ml
+Consider first a simple recursive function |digitsOfInt| that extracts
+the digits of an |int|.
 %
@@
 actually correct!
+
+As our \dectree correctly points out (with high confidence), the fault
+lies with the list literal \emph{surrounding} the recursive call, remove
+it and the type error disappears.
 %
-As our Decision Tree correctly points out, the fault lies with the list
-literal \emph{surrounding} the recursive call, remove it and the type
-error disappears.
+An examination of the path induced by the list literal reveals that our
+\dectree is basing its decision on the fact that
 %
-\ES{TODO: WHY?}
+(1) the expression is a list literal;
+%
+(2) the child expression is an application, whose return type mentions |int|; and
+%
+(3) the parent expression's type mentions |list|.
+%
+Interestingly, \dectree incorrectly predicts that the child application
+should change as well, but it is less confident of this prediction and
+ranks it below the correct blame assignment.
 
 \paragraph{Padding a list}
-% data/sp14/0306.ml
+% data/sp14/0442.ml
 Our next program, |padZero|, is given two |int list|s as input, and must
@@
 
-The student first defines a helper function |clone|, which takes an
-input element |x| and an |int| |n|, and returns a list containing |n|
-|x|s.
+The student first defines a helper function |clone|, as seen above.
 %
@@
 Then she defines |padZero| with a simple branch that determines which
-list is shorter, followed by a |clone| with the appropriate number of
+list is shorter, followed by a |clone| for the appropriate number of
 zeros.
@@
 |int list| returned by |clone| onto |l2|, which is another |int list|,
-but they decide to blame the call to |clone|, while our Decision
-Tree correctly blames the |::| operator.
+but they decide to blame the call to |clone|, while our
+\dectree correctly blames the |::| constructor.
+
+Examining the path induced by the |::| constructor, we can see that our
+\dectree is influenced by the fact that:
 %
-\ES{TODO: WHY? (might be too similar to previous?)}
+(1) the expression is a data constructor (not specifically |::|);
+%
+(2) the parent expression is a tuple; and
+%
+(3) the leftmost child is an application.
+%
+We note that first fact appears to be particularly significant; an
+examination of the training samples that reach that decision node
+reveals that before observing the \textsc{Is-Constructor} feature, the
+classifier is slightly in favor of predicting ``blame'', but afterwards
+it is heavily in favor of predicting ``blame''.
+%
+Many of the following decisions change the balance back towards ``no
+blame'' if the ``true'' path is taken, but the |::| constructor always
+takes the ``false'' path.
+%
+It would appear that our \dectree has learned that data constructors are
+particularly suspicious, and is looking for exceptions to this general
+rule.
+%
+% \ES{FIXME: this is probably super confusing, checking with Huma to see
+%   if we can get a nice graph to illustrate what's happening..}
+
+Our \dectree correctly predicts that the recursive call blamed by
+\sherrloc should not be blamed; a similar examination suggests that the
+crucial observation is that the parent of the recursive call is a data
+constructor application.
+%
 
@@
 
-\paragraph{Computing the Fixed Point of a Function}
-% data/sp14/0941.ml
-Finally, our students must write a |fixpoint| function that computes the
-fixed point of a given function |f|, starting from an initial value |b|.
-%
-As a hint, we first have them write a |wwhile| function that performs the
-functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
-output back in until it receives a (boolean) signal to stop.
-%
-\begin{ecode}
-  let rec wwhile (f, b) =
-    match f b with
-    | (x, false) -> x
-    | (x, true)  -> wwhile (f, x)
-\end{ecode}
-\lstset{firstnumber=last}
-%
-The |fixpoint| function can then be written as a clever instantiation of
-the arguments to |wwhile|.
-%
-\begin{ecode}
-  let fixpoint (f, b) =
-    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
-    wwhile (g, b)
-\end{ecode}
-\lstset{firstnumber=1}
-%
-Sadly, our student has forgotten that |g| should itself be a function.
-%
-As a result, she passes a pair of a \emph{pair} and a starting value to
-|wwhile|, rather than a pair of a \emph{function} and a starting value.
-%
-\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
-actually blames the use of |g|), but identifies the construction of the
-pair inside |g| as the most likely culprit, while our Decision Tree
-correctly identifies the \emph{definition} of |g| as the source of the
-error.
-\ES{TODO: WHY?}
-
-\subsubsection{Failed Predictions}
-\label{sec:failed-predictions}
-We also describe indicative programs where our classifier makes an
-incorrect prediction.
-
-\paragraph{Constructing a List of Duplicates}
-% data/sp14/0148.ml
-Consider the following implementation of the |clone| function from
-before.
-%
-\begin{ecode}
-  let rec clone x n =
-    let loop acc n =
-      if n <= 0 then
-        acc
-      else
-        clone ==([==_=x=_==] @ acc)== (n - 1) in
-    loop [] n
-\end{ecode}
-% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
-%
-The student has defined a helper function |loop| with an accumulator
-|acc|, likely meant to call itself tail-recursively.
-%
-Unfortunately, she has called the top-level function |clone| rather than
-|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
-for the |x| argument to |clone|.
-
-Our classifier incorrectly predicts that the use of |x| in the recursive
-call is the most likely source of the error.
-\ES{TOOD: WHY??}
-%
-Our second prediction coincides with \sherrloc (and \ocaml), blaming the
-the first argument to |clone|.
-%
-This is also incorrect, but may be more helpful than our first
-prediction --- if our student decides that she has certainly provided
-the correct \emph{argument}, an alternative explanation is that
-perhaps she has called the wrong \emph{function}.
-%
-Our final prediction is the entire body of |clone|, which is hardly
-helpful.
-
-\paragraph{Currying Considered Harmful?}
-% data/sp14/0887.ml
-Our next example is another ill-fated attempt at |clone|.
-%
-\begin{ecode}
-  let rec clone x n =
-    let rec loop x n acc =
-      if n < 0 then
-        acc
-      else
-        ==loop== __(x, (n - 1), (x :: acc))__ in
-    loop (x, n, [])
-\end{ecode}
-The issue here is that \ocaml functions are \emph{curried} by default
---- \ie they take their arguments one at a time --- but our student has
-called the inner |loop| with all three arguments in a tuple.
-%
-Many experienced functional programmers would choose to keep |loop|
-curried and rewrite the calls, however our student decides instead to
-\emph{uncurry} |loop|, making it take a tuple of arguments.
-%
-\sherrloc blames the recursive call to |loop| while our classifier
-blames the tuple of arguments --- a reasonable suggestion, but alas not
-the answer the student was looking for.
-\ES{WHY??}
-%
-
-\paragraph{Composing Functions}
-% data/sp14/2269.ml
-Next, let us consider the |pipe| function that composes a list of
-functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
-function |fun x -> h (g (f x))|.
-%
-\begin{ecode}
-  let pipe fs =
-    let f a x y = ==y== __(a y)__ in
-    let base x = x in
-    List.fold_left f base fs
-\end{ecode}
-%
-The error in our student's code is that she has applied |y| rather than
-|x| to the result of |a y|.
-%
-\sherrloc correctly blames the first occurrence of |y|, while our
-classifier (incorrectly) blames the application |a y| (\ocaml blames
-the occurrence of |base| on line 4).
-%
-\ES{TODO: WHY?}
+% \ES{I don't much like this example anymore..}
+% \paragraph{Computing the Fixed Point of a Function}
+% % data/sp14/1266.ml
+% Finally, our students must write a |fixpoint| function that computes the
+% fixed point of a given function |f|, starting from an initial value |b|.
+% %
+% As a hint, we first have them write a |wwhile| function that performs the
+% functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
+% output back in until it receives a (boolean) signal to stop.
+% %
+% \begin{ecode}
+%   let rec wwhile (f, b) =
+%     match f b with
+%     | (x, false) -> x
+%     | (x, true)  -> wwhile (f, x)
+% \end{ecode}
+% \lstset{firstnumber=last}
+% %
+% The |fixpoint| function can then be written as a clever instantiation of
+% the arguments to |wwhile|.
+% %
+% \begin{ecode}
+%   let fixpoint (f, b) =
+%     let g = __let bb = f b in ___=(bb, (bb = b))=_ in
+%     wwhile (g, b)
+% \end{ecode}
+% \lstset{firstnumber=1}
+% %
+% Sadly, our student has forgotten that |g| should itself be a function.
+% %
+% As a result, she passes a pair of a \emph{pair} and a starting value to
+% |wwhile|, rather than a pair of a \emph{function} and a starting value.
+% %
+% \sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
+% actually blames the use of |g|), but identifies the construction of the
+% pair inside |g| as the most likely culprit, while our \dectree correctly
+% identifies the \emph{definition} of |g| as the source of the error.
+% %
+% Our \dectree appears to base its prediction, with complete confidence,
+% on the fact that:
+% %
+% (1) the parent expression is a |let|;
+% %
+% (2) the right child is a tuple; and finally
+% %
+% (3) the current expression is a |let|.
+% %
 

### 2017-04-07 paper/oopsla17-submission/evaluation-accuracy.tex
update numbers

@@
 %
-\begin{enumerate}
-\item A logistic regression trained with a learning rate
+\begin{description}
+\item[\linear] A logistic regression trained with a learning rate
   $\eta = 0.001$, an $L_2$ regularization rate $\lambda = 0.001$, and a
   mini-batch size of 200.
-\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
-\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
-\item Two multi-layer perceptron neural networks, both trained with $\eta = 0.001$,
+\item[\dectree] A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
+\item[\forest] A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
+\item[\hiddenT and \hiddenFH] Two multi-layer perceptron neural networks, both trained with $\eta = 0.001$,
   $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
@@
   common practice in modern neural networks.
-\end{enumerate}
+\end{description}
 %

### 2017-04-07 paper/oopsla17-submission/evaluation-utility.tex
update numbers

@@
     \midrule
-    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 32.4\% & 49.3\% & 60.4\% & 39.6\% \\
-    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 56.1\% & 71.6\% & 82.7\% & 57.4\% \\
-    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 57.8\% & 72.8\% & 83.3\% & 57.8\% \\
+    Local Syntax       & 44          & & 27.7\% & 46.7\% & 58.6\% & 38.2\% & & 32.6\% & 49.5\% & 60.2\% & 39.4\% \\
+    + \InSlice         & 45          & & 46.4\% & 65.1\% & 76.1\% & 48.9\% & & 55.4\% & 71.0\% & 82.1\% & 57.1\% \\
+    Filter by \InSlice & 44          & & 55.9\% & 71.9\% & 82.9\% & 57.5\% & & 57.8\% & 72.7\% & 82.9\% & 57.6\% \\
     \bottomrule
@@
 %
-As expected, the baseline performs the worst, with a mere 23.6\% \linear
+As expected, the baseline performs the worst, with a mere 27.7\% \linear
 Top-1 accuracy.
@@
 However, filtering out expressions that are not part of the slice
-\emph{further} improves the results to 54.9\% \linear Top-1 accuracy.
+\emph{further} improves the results to 55.9\% \linear Top-1 accuracy.
 %
@@
 %
-Top-1 accuracy jumps from 32.4\% to 56.1\% when we add \InSlice, and only
-improves by 1.7\% when we filter out expressions that are not part of the
+Top-1 accuracy jumps from 32.4\% to 55.4\% when we add \InSlice, and only
+improves by 2\% when we filter out expressions that are not part of the
 error slice.
@@
     \midrule
-    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 57.7\% & 72.4\% & 83.0\% & 57.7\% \\
+    Local Syntax             &  44         & & 56.6\% & 72.6\% & 83.2\% & 57.7\% & & 57.6\% & 72.6\% & 83.1\% & 57.7\% \\
     \midrule
-    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.6\% & 75.0\% & 83.5\% & 57.2\% \\
-    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.5\% & 84.4\% & 90.5\% & 68.8\% \\
-    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 74.0\% & 85.5\% & 90.8\% & 69.1\% \\
+    + Size                   &  45         & & 57.2\% & 73.7\% & 83.0\% & 57.4\% & & 60.9\% & 75.5\% & 84.1\% & 57.9\% \\
+    + Context                & 220         & & 61.1\% & 78.7\% & 86.7\% & 63.0\% & & 71.5\% & 84.8\% & 90.8\% & 69.0\% \\
+    + Types                  & 102         & & 63.1\% & 77.8\% & 85.2\% & 61.6\% & & 73.4\% & 85.4\% & 90.7\% & 68.9\% \\
     \midrule
-    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.4\% & 84.1\% & 90.8\% & 69.0\% \\
-    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.7\% & 86.3\% & 90.9\% & 69.5\% \\
-    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.6\% & 88.1\% & 92.9\% & 72.9\% \\
+    + Context + Size         & 221         & & 61.1\% & 78.8\% & 86.3\% & 62.4\% & & 71.5\% & 84.4\% & 90.8\% & 69.0\% \\
+    + Types + Size           & 103         & & 61.9\% & 78.7\% & 85.6\% & 62.0\% & & 73.1\% & 85.2\% & 91.1\% & 69.5\% \\
+    + Context + Types        & 275         & & 63.1\% & 80.9\% & 88.0\% & 65.0\% & & 77.2\% & 88.3\% & 92.5\% & 72.6\% \\
     \midrule
-    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 76.8\% & 87.9\% & 92.7\% & 72.5\% \\
+    + Context + Types + Size & 276         & & 62.8\% & 80.5\% & 88.1\% & 65.1\% & & 77.3\% & 88.1\% & 92.7\% & 72.7\% \\
     \bottomrule
@@
 \ExprSize is the weakest feature, improving \linear Top-1
-accuracy by only 1\% and \hiddenFH by only 4\%.
+accuracy by less than 1\% and \hiddenFH by only 3\%.
 %
 In contrast, the contextual syntactic features improve \linear Top-1
-accuracy by 5\% (\resp 15\%), and the typing features improve
-Top-1 accuracy by 7\% (\resp 17\%).
+accuracy by 5\% (\resp 14\%), and the typing features improve
+Top-1 accuracy by 7\% (\resp 16\%).
 %
@@
 in accuracy when added alongside the contextual or typing features.
-FIXME: Indicate that this is expected because of feature overlap: the
-contextual ``child'' features partially encode subtree size.
+\ES{FIXME: Indicate that this is expected because of feature overlap: the
+contextual ``child'' features partially encode subtree size.}
 
@@
 %
-However, typing features do not \emph{subsume} contextual syntactic
-features, we can gain an additional 1\% Top-1 accuracy (\resp 4\%) by
-adding \emph{both}.
+However, typing features do not appear to \emph{subsume} contextual
+syntactic features, the \hiddenFH gains an additional 4\% Top-1 accuracy
+when both are added.

### 2017-04-07 paper/oopsla17-submission/evaluation.tex
update numbers

@@
   and for the contextual sub-expressions.
-\item one feature denoting the size of each sub-expression.
-\item one feature denoting whether each sub-expression is part of the
+\item One feature denoting the size of each sub-expression.
+\item One feature denoting whether each sub-expression is part of the
   minimal type error slice. We use this feature as a ``hard''

### 2017-04-09 paper/oopsla17-submission/evaluation-accuracy.tex
misc

@@
   mini-batch size of 200.
-\item[\dectree] A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
-\item[\forest] A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
-\item[\hiddenT and \hiddenFH] Two multi-layer perceptron neural networks, both trained with $\eta = 0.001$,
-  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
-  contains a single hidden layer of 10 neurons, and the second contains
-  a hidden layer of 500 neurons. This allows us to investigate how well
-  the MLP can \emph{compress} its model (cf.~\cite{FIXME}). The neurons
-  use rectified linear units (ReLU) as their activation function, a
-  common practice in modern neural networks.
+\item[\dectree] A decision tree trained with the CART algorithm
+  \citep{Breiman1984-qy} and an impurity threshold of $10^{-7}$ (used to
+  avoid overfitting via early stopping).
+\item[\forest] A random forest \citep{Breiman2001-wo} of 30
+  estimators, trained with an impurity threshold of $10^{-7}$.
+\item[\hiddenT and \hiddenFH] Two multi-layer perceptron neural
+  networks, both trained with $\eta = 0.001$, $\lambda = 0.001$, and a
+  mini-batch size of 200. The first MLP contains a single hidden layer
+  of 10 neurons, and the second contains a hidden layer of 500
+  neurons. This allows us to investigate how well the MLP can
+  \emph{compress} its model (cf.~\cite{FIXME}). The neurons use
+  rectified linear units (ReLU) as their activation function, a common
+  practice in modern neural networks.
 \end{description}
@@
 %
-\sherrloc outperforms both baselines with 56\% Top-1 accuracy (84\% Top
+\sherrloc outperforms both baselines with 56\% Top-1 accuracy (84--86\% Top
 3), while \mycroft actually underperforms \ocaml with 38--41\% Top-1
@@
 Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
-ranging from 59--63\% Top-1 accuracy (86--88\% Top-3) for the \linear
-classifier to 70--74\% Top-1 accuracy (91--92\% Top-3) for the \hiddenFH.
+ranging from 58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear
+classifier to 71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
 

### 2017-04-09 paper/oopsla17-submission/evaluation-utility.tex
misc

@@
 in accuracy when added alongside the contextual or typing features.
-\ES{FIXME: Indicate that this is expected because of feature overlap: the
-contextual ``child'' features partially encode subtree size.}
+%
+This is likely explained by \emph{feature overlap}, \ie the contextual
+features of ``child'' expressions additionally provide some information
+about the size of the subtree.
 

### 2017-04-09 paper/oopsla17-submission/evaluation.tex
api

@@
 \paragraph{Blame Oracle}
-We identified the fixes for each ill-typed program with an
-expression-level diff~\citep{Lempsink2009-xf}.
-%
-We consider two sources of changes:
-%
-\begin{enumerate}
-\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
-  is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
-  changed, as it has been replaced by $g$.
-\item If a new expression has been inserted \emph{around} an existing
-  expression, \eg if $\eapp{f}{x}$ is rewritten to
-  $\eplus{\eapp{f}{x}}{1}$, we will mark the application expression
-  $\eapp{f}{x}$ (but not $f$ or $x$) as changed, as the $+$ operator now
-  occupies the original location of the application.
-\end{enumerate}
 
-Students may have made many, potentially unrelated, changes between
-compilations, but at some point the ``fix'' becomes a ``rewrite''.
+A disadvantage of using diffs as blame oracles is that students may have
+made many, potentially unrelated, changes between compilations; at some
+point the ``fix'' becomes a ``rewrite''.
 %
@@
 discard outliers where the fraction of expressions that have changed is
-more than one standard deviation above the mean, or 44\%.
+more than one standard deviation above the mean, establishing a diff
+threshold of 44\%.
 %

### 2017-04-09 paper/oopsla17-submission/learning.tex
api

@@
 %
+Our goal is to instantiate the $\blamesym$ function of
+\autoref{fig:api}, which takes as input a $\Model$ of type errors and an
+ill-typed program $e$, and returns an ordered list of subexpressions
+from $e$ paired with the probability $\Runit$ that they should be
+blamed.
+
+A $\Model$ is produced by the $\trainsym$ function, which performs
+supervised learning on a training set of feature vectors $\V$ and
+(boolean) labels $\B$.
+%
+Once trained, a $\Model$ can be $\evalsym$uated on a new input vector,
+producing the probability $\Runit$ that the blame label should be
+applied.
+%
+We describe multiple types of $\Model$ and their instantiations of
+$\trainsym$ and $\evalsym$ in \autoref{sec:models}.
+
+Of course, the $\Model$ expects feature vectors $\V$ and blame labels
+$\B$, but we are given program pairs.
+%
+So our first step must be to define a suitable translation from program
+pairs to feature vectors and labels, \ie we must define the
+$\extractsym$ function in \autoref{fig:api}.
+%
+We accomplish this by modeling features as real-valued functions of
+expressions, and extract a feature vector for each \emph{subexpression}
+of the ill-typed program (\autoref{sec:features}).
+%
+Then we define the blame labels for the training set to be the
+subexpressions that changed between the ill-typed program and its
+subsequent fix, and model $\blamesym$ as a function from a program pair
+to the set of expressions that changed (\autoref{sec:labels}).
+%
+The $\extractsym$ function, then, extracts $\featuresym$ from each
+subexpression and computes the blamed expressions according to
+$\labelsym$.
+
+% First, we define the inputs to the model, a set of \emph{features}
+% (\autoref{sec:features}) that we will use to describe programs.
+% %
+% Second, we define the expected outputs of the model, a set of
+% \emph{labels} (\autoref{sec:labels}) that we will use to assign blame.
+% %
+% Third, we describe the actual \emph{models} (\autoref{sec:models}) that we
+% will evaluate.
+
 \input{syntax}
-%
-First, we define the inputs to the model, a set of \emph{features}
-(\autoref{sec:features}) that we will use to describe programs.
-%
-Second, we define the expected outputs of the model, a set of
-\emph{labels} (\autoref{sec:labels}) that we will use to assign blame.
-%
-Third, we describe the actual \emph{models} (\autoref{sec:models}) that we
-will evaluate.
 
@@
 We are given programs over \lang, but the learning algorithms expect to work
-with \emph{feature vectors} --- vectors of real numbers, where each
+with \emph{feature vectors} $\V$ --- vectors of real numbers, where each
 column describes a particular aspect of the input.
@@
 For example, the \IsNil feature in \autoref{tab:sumList} describes
-whether the current expression is the empty list $\enil$.
+whether an expression is the empty list $\enil$.
 
@@
 These are similar to local syntactic features, but lifted to describe the
-parent and children of the current expression.
+parent and children of an expression.
 %
 For example, the \IsCaseListP feature in \autoref{tab:sumList} describes
-whether current expression's \emph{parent} matches on a list.
+whether an expression's \emph{parent} matches on a list.
 %
@@
 In practice, we have found that \InSlice is a particularly important
-feature, and thus include a pre-processing step that discards all
+feature, and thus include a post-processing step that discards all
 expressions where it is disabled.
@@
 \label{sec:labels}
-We define the outputs of the model to be a simple ``good'' or ``bad''
-label, where ``good'' means the expression \emph{should not} change and
-``bad'' means the expression \emph{should} change. 
+We define the output of the model to be a boolean label, where ``false''
+means the expression \emph{should not} change and ``true'' means the
+expression \emph{should} change.
 %
@@
 %
-Thus, we add a post-processing step where we \emph{rank} the expressions
-by the model's confidence that they will change, and select only the top
-$k$ (in practice $k=3$) to present to the user.
+Thus, we \emph{rank} the expressions by the model's confidence that they
+will change, and select only the top $k$ (in practice $k=3$) to present
+to the user.
+
+We identify the fixes for each ill-typed program with an
+expression-level diff~\citep{Lempsink2009-xf}.
+%
+We consider two sources of changes:
+%
+\begin{enumerate}
+\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
+  is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
+  changed, as it has been replaced by $g$.
+\item If a new expression has been inserted \emph{around} an existing
+  expression, \eg if $\eapp{f}{x}$ is rewritten to
+  $\eplus{\eapp{f}{x}}{1}$, we will mark the application expression
+  $\eapp{f}{x}$ (but not $f$ or $x$) as changed, as the $+$ operator now
+  occupies the original location of the application.
+\end{enumerate}
 

### 2017-04-10 paper/oopsla17-submission/evaluation-accuracy.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

@@
 \sherrloc, we removed from the dataset programs where they predicted
-locations that our oracle could not match with a program expression ---
+locations that our oracle could not match with a program expression:
 6--8\% of programs for \mycroft and 3--4\% for \sherrloc.
@@
 (after one minute) or where they encountered an unsupported language
-feature --- another 5\% for \mycroft and 12--13\% for \sherrloc. This
+feature: another 5\% for \mycroft and 12--13\% for \sherrloc. This
 experimental design gives the state of the art tools the ``benefit of the

### 2017-04-10 paper/oopsla17-submission/evaluation.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

@@
   constraint, sub-expressions that are not part of the minimal slice
-  will be preemptively discarded, we justify this decision in
+  will be preemptively discarded. We justify this decision in
   \autoref{sec:feature-utility}.
@@
 
+A related threat to construct validity is our definition of the immedate
+next well-typed program as the intended ground truth answer (see
+\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
+intermediate well-typed program ``rewrites'' between the original ill-typed
+program and the final intended answer. Our approach to discarding outliers
+(see \autoref{sec:evaluation}) is designed to mitigate this threat.
+
 Our removal of program pairs that changed too much, where our oracle

### 2017-04-10 paper/oopsla17-submission/introduction.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

@@
 %
-The data forms a time-series of programs submitted to the \ocaml
+The data consist of a time-series of programs submitted to the \ocaml
 compiler, from which we extracted a set of \emph{ill-typed} programs

### 2017-04-10 paper/oopsla17-submission/learning.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

@@
 %
-In this section we present four standard learning algorithms that we
-will investigate:
+In this section we consider four standard learning algorithms: 
 %
@@
 by a human, and a summary of its advantages and disadvantages.
+
+FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'', 
+contains much more detail than an OOPSLA audience wants on
+introductory/undergraduate machine learning. (Especially since nothing
+after this point is a novel contribution we make.) Perhaps we can cut some
+space by focusing on the pros and cons of each technique rather than
+describing the details? This is a PL paper.
 

### 2017-04-10 paper/oopsla17-submission/overview.tex
Wes pass over the paper. Looking good from here, will send email to Eric.

@@
 %
-Programs can be represented in many ways, from the raw text to 
+Programs can be represented in various ways, from the raw text to 
 richly-structured abstract syntax trees and control-flow graphs, but 
-representations are not directly amenable to machine learning.
+many representations are not directly amenable to machine learning.
 %
@@
 Prior work has often enlisted expert users to manually judge ill-typed
-programs and determine the ``correct'' fix \citep[e.g.][]{Loncaric2016-uk}, but
+programs and determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk}, but
 this approach does not scale well to a dataset large enough to support

### 2017-04-10 paper/oopsla17-submission/learning.tex
tweak notiation for 3.3 to be consistent with fig 3

@@
 This means that we assume the existence of a training set
-$S = \{\langle \mathbf{x}_i, y_i \rangle\}_{i=1}^{n}$
-of sample inputs $\mathbf{x}$ and their corresponding labels $y$.
+$S = \{\langle \mathbf{v}_i, b_i \rangle\}_{i=1}^{n}$
+of sample inputs $\mathbf{v}$ and their corresponding labels $b$.
 %
-The learning task is to model the probability $Pr(y\ |\ \mathbf{x})$
+The learning task is to model the probability $Pr(b\ |\ \mathbf{v})$
 of a label being assigned to a given input.
@@
 
-FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'', 
+\ES{FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'',
 contains much more detail than an OOPSLA audience wants on
@@
 space by focusing on the pros and cons of each technique rather than
-describing the details? This is a PL paper.
+describing the details? This is a PL paper. }
 
@@
 %
-First, we compute the \emph{evidence} for a label $y$ as a linear
-function of the feature vector $\mathbf{x}$, using a weight vector
-$\mathbf{w}$ and a bias $b$ that will be learned from the training set.
+First, we compute the \emph{evidence} for a label $b$ as a linear
+function of the feature vector $\mathbf{v}$, using a weight vector
+$\mathbf{w}$ and a bias $w_0$ that will be learned from the training set.
 $$
-\mathsf{ev} = \sum_j w_j x_j + b
+\mathsf{ev} = \sum_j w_j v_j + w_0
 $$
@@
 $[0,1]$ interval, which can then be interpreted as the probability that
-the label $y$ should be applied.
+the label $b$ should be applied.
 %
@@
 $$
-Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j x_j - b)}
+Pr(b\ |\ \mathbf{v}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j v_j - w_0)}
 $$
@@
 features to the model --- \eg if one knows that the label has a
-quadratic relationship with a feature $x$, the transformed feature $x^2$
+quadratic relationship with a feature $v$, the transformed feature $v^2$
 can be added --- but this typically requires knowledge of the underlying
@@
 Each node in the tree contains a single predicate of the form
-$x_j \leq t$ for some feature $x_j$ and threshold $t$, which determines
+$v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
 whether a given input should proceed down the left or right subtree.

### 2017-04-10 paper/oopsla17-submission/gallery.tex
use colors instaed of bold/underline

@@
 \label{sec:qualitative}
+
+\definecolor{tree}{HTML}{66C2A5}
+\definecolor{sherrloc}{HTML}{FC8D62}
+\definecolor{fix}{HTML}{8DA0CB}
 
@@
 %
-(2) \sherrloc's prediction (in \textbf{bold}),
+(2) \colorbox{sherrloc!75}{\sherrloc's} prediction,
 %
-(3) the prediction of our \dectree classifier (\underline{underlined}), and
+(3) our \colorbox{tree!75}{\dectree's} prediction, and
 %
@@
 %
-For these programs we will additionally \colorbox{red!50}{highlight} the
+For these programs we will additionally \colorbox{fix!75}{highlight} the
 correct blame location.
@@
 %
+% \begin{ecode}
+%   let rec clone x n =
+%     let loop acc n =
+%       if n <= 0 then
+%         acc
+%       else
+%         (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
+%     loop [] n
+% \end{ecode}
 \begin{ecode}
@@
       else
-        (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
+        (*@\colorbox{fix!75}{clone} \colorbox{sherrloc!75}{([\colorbox{tree!75}{x}] @\ acc)}@*) (n - 1) in
     loop [] n
@@
   let rec clone x n =
-    let rec loop (*@\colorbox{red!50}{x n acc}@*) =
+    let rec loop (*@\colorbox{fix!75}{x n acc}@*) =
       if n < 0 then
@@
       else
-        ==loop== __(x, (n - 1), (x :: acc))__ in
+        (*@\colorbox{sherrloc!75}{loop} \colorbox{tree!75}{(x, (n - 1), (x :: acc))}@*) in
     loop (x, n, [])
 \end{ecode}
+%
 The issue here is that \ocaml functions are \emph{curried} by default
@@
     else
-      [n mod 10] @ __[ ___=digitsOfInt (n / 10)=___ ]__
+      [n mod 10] @ (*@\colorbox{tree!75}{[ \colorbox{sherrloc!75}{digitsOfInt (n / 10)} ]}@*)
 \end{ecode}
@@
     else
-      (l1, _=clone 0 n=___ :: l2__)
+      (l1, (*@\colorbox{tree!75}{\colorbox{sherrloc!75}{clone 0 n} :: l2}@*))
 \end{ecode}

### 2017-04-10 paper/oopsla17-submission/conclusion.tex
fiddling with graphs instead of tables..

--- 2017-04-02 paper/oopsla17-submission/discussion.tex
@@
-\section{Discussion}
-\label{sec:discussion}
-
-We have shown that we can outperform the state of the art in type error
-localization by learning a model of the errors that programmers make,
-using a set of features that closely resemble the information the type
-checker sees.
-%
-However, there is certainly room for improvement.
-%
-In this section we highlight some limitations of our approach and
-potential avenues for future work.
-
-\paragraph{User-Defined Types}
-Probably the single biggest limitation of our technique is that we have
-(a finite set of) features for specific data and type constructors.
-%
-Anything our models learn about errors made with the |::| constructor or
-the |list| type cannot easily be translated to new, user-defined
-datatypes the model has never encountered.
-%
-We can mitigate this, to some extent, by adding generic
-syntactic features for data constructors and match expressions, but it
-remains to be seen how much these help \ES{TODO: UW data!}.
-%
-Furthermore, there is no obvious analog for transferring knowledge to
-new type constructors, which we have seen are both more compact and
-helpful.
-
-As an alternative to encoding information about \emph{specific}
-constructors, we might investigate a more abstract representation.
-%
-For example, instead of modeling |x :: 2| as a |::| constructor with a
-right child of type |int|, we might model it as some (unknown) constructor
-whose right child has an incompatible type.
-%
-We might symmetrically model the |2| as an integer literal whose type is
-incompatible with its parent.
-%
-Anything we learn about |::| and |2| can now be transferred directly to
-yet unseen types, but we run the risk generalizing \emph{too much} ---
-\ie perhaps programmers make different mistakes with |list|s than they
-do with other types, and are thus likely to choose different fixes.
-%
-Balancing the trade-off between specificity and generalizability appears
-to be a challenging task.
-
-% \begin{itemize}
-% \item Hard to support in general, as each data and type constructor gets
-%   its own set of syntactic features.
-% \item Attempt to mitigate by adding generic ``is-datacon'' and
-%   ``is-match'' feature, unclear how much it will help
-% \end{itemize}
-
-\paragraph{Additional Features}
-There are a number of other features that could improve the model's
-ability to localize errors, that would be easier to add than
-user-defined types.
-%
-For example, each occurrence of a variable knows only its type and the
-immediate context in which it occurs, but it may be helpful to know
-about \emph{other} occurrences of the same variable.
-%
-If a variable is generally used in the context of a |float| but has a
-single occurrence in the context of an |int|, it seems likely that the
-latter occurrence (or context) is to blame.
-%
-Similarly, arguments to a function application are not aware of the
-typing constraints imposed on them by the function (and vice versa),
-they only know that they are occurring in the context of an application
-and the type of the application itself.
-%
-Finally, \emph{n-grams} on the token stream have proven effective for
-probabilistic modeling of programming languages~\cite{FIXME,FIXME}, we may
-find that they aid in our task as well.
-%
-For example, if the observed tokens in an expression diverge from the
-n-gram model's predictions, that indicates that there is something
-unusual about the program at that point, and it may signal an error.
-
-% \begin{itemize}
-% \item def-use chains
-% \item connect function and arguments in application (right now each only know about the application itself)
-% \item ??
-% \end{itemize}
-
-\paragraph{Independent vs Joint Predictions}
-We treat each sub-expression as if it exists in a vacuum, but in reality
-the program has a rich \emph{graphical} structure, particularly if one adds
-edges connecting different occurrences of the same variable.
-%
-\citet{Raychev2015-jg} have used these richer models to great effect to
-make \emph{interdependent} predictions about programs, \eg
-de-obfuscating variable names or even inferring types.
-%
-One could even view our task of locating the source of an error as simply
-another property to be predicted over a graphical model of the program.
-%
-One of the key advantages of a graphical model is that the predictions
-made for one node can influence the predictions made for another node,
-this is known as \emph{structured learning}.
-%
-For example, if, given the expression |1 + true|, we predict |true| to
-be erroneous, we may be much less likely to predict |+| as erroneous.
-%
-We compensate somewhat for our lack of structure by ranking our
-predictions by ``confidence'', but it would be interesting to see how
-structured learning over graphical models would perform.
+\section{Conclusion}
+\label{sec:conclusion}
 

### 2017-04-10 paper/oopsla17-submission/evaluation-utility-graph.tex
fiddling with graphs instead of tables..

\begin{figure}[t]
\centering
\begin{subfigure}[t]{\linewidth}
\centering
\begin{tikzpicture}
\begin{axis}[
  % name=slice,
  % scale only axis,
  %at=(feature.above north), anchor=below south east,
  ybar stacked,
  width=0.5\linewidth,
  height=5cm,
  %title={Impact of Type Error Slice},
  ylabel={Accuracy},
  bar width=0.5cm,
  ymin=0.2,
  ymax=1,
  ytick={0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  symbolic x coords={op+slice-no-slice, op+slice, op+slice-only-slice},
  enlarge x limits=0.25,
  xtick=data,
  xtick style={draw=none},
  xticklabels={\textsc{Local Syntax}, +\InSlice, \textsc{Filter \InSlice}},
  %x tick label style={rotate=45, anchor=north east},
  reverse legend,
  transpose legend,
  legend style={
    %legend pos = outer north east,
    at={(1.5,0.5)},
    anchor=center,
    legend columns=4
  },
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green1, bar shift=.25cm] table[x=features, y=top-1] {\SliceHiddenBench};
\addlegendentry{Top-1}
\addplot[draw=black, fill=green2, bar shift=.25cm] table[x=features, y expr=\thisrow{top-2} - \thisrow{top-1}] {\SliceHiddenBench};
\addlegendentry{Top-2}
\addplot[draw=black, fill=green3, bar shift=.25cm] table[x=features, y expr=\thisrow{top-3} - \thisrow{top-2}] {\SliceHiddenBench};
\addlegendentry{Top-3}
\addlegendimage{empty legend}
\addlegendentry{\hiddenFH}

\resetstackedplots

\addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=features, y=top-1] {\SliceLinearBench};
\addlegendentry{Top-1}
\addplot[draw=black, fill=blue2, bar shift=-.25cm] table[x=features, y expr=\thisrow{top-2} - \thisrow{top-1}] {\SliceLinearBench};
\addlegendentry{Top-2}
\addplot[draw=black, fill=blue3, bar shift=-.25cm] table[x=features, y expr=\thisrow{top-3} - \thisrow{top-2}] {\SliceLinearBench};
\addlegendentry{Top-3}
\addlegendimage{empty legend}
\addlegendentry{\linear}
\end{axis}
\end{tikzpicture}
\caption{Impact of type error slice on blame accuracy.}\label{fig:slice-utility}
\end{subfigure}

\begin{subfigure}[t]{\linewidth}
\begin{tikzpicture}
\begin{axis}[
  % name=feature,
  % scale only axis,
  ybar stacked,
  width=\linewidth,
  height=6cm,
  %title={Impact of Contextual Feature Classes},
  ylabel={Accuracy},
  bar width=0.5cm,
  ymin=0.5,
  ymax=1,
  ytick={0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  symbolic x coords={op, op+size, op+context, op+type, op+context+size, op+type+size, op+context+type, op+context+type+size},
  % enlarge x limits=0.5,
  xtick=data,
  xtick style={draw=none},
  xticklabels={
    \textsc{Local Syntax},
    \textsc{+Size}, \textsc{+Context}, \textsc{+Type},
    +C+S, +T+S, +C+T,
    +C+T+S
    % + Size, + Context, + Type,
    % + Context + Size, + Type + Size, + Context + Type,
    % + Context + Type + Size
  },
  %x tick label style={rotate=45, anchor=north east},
  % reverse legend,
  % transpose legend,
  % legend style={
  %   anchor = south east,
  %   at = {(1,1)},
  %   %legend pos = outer north east,
  %   legend columns=4},
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green1, bar shift=.25cm] table[x=features, y=top-1] {\FeatureHiddenBench};
% \addlegendentry{Top-1}
\addplot[draw=black, fill=green2, bar shift=.25cm] table[x=features, y expr=\thisrow{top-2} - \thisrow{top-1}] {\FeatureHiddenBench};
% \addlegendentry{Top-2}
\addplot[draw=black, fill=green3, bar shift=.25cm] table[x=features, y expr=\thisrow{top-3} - \thisrow{top-2}] {\FeatureHiddenBench};
% \addlegendentry{Top-3}
% \addlegendimage{empty legend}
% \addlegendentry{\hiddenFH}

\resetstackedplots

\addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=features, y=top-1] {\FeatureLinearBench};
% \addlegendentry{Top-1}
\addplot[draw=black, fill=blue2, bar shift=-.25cm] table[x=features, y expr=\thisrow{top-2} - \thisrow{top-1}] {\FeatureLinearBench};
% \addlegendentry{Top-2}
\addplot[draw=black, fill=blue3, bar shift=-.25cm] table[x=features, y expr=\thisrow{top-3} - \thisrow{top-2}] {\FeatureLinearBench};
% \addlegendentry{Top-3}
% \addlegendimage{empty legend}
% \addlegendentry{\linear}

%\legend{Top-1, Top-2, Top-3}
\end{axis}

\end{tikzpicture}
\caption{Impact of contextual features on blame accuracy.}\label{fig:context-utility}
\end{subfigure}

\caption{
  %
  Results of our experiments on feature utility.
%
}
\label{fig:slice-utility-results}
\end{figure}

### 2017-04-10 paper/oopsla17-submission/evaluation-utility.tex
fiddling with graphs instead of tables..

@@
 %
-\begin{table}[ht]
-  \caption{
-    Impact of Type Error Slice on Accuracy.
-    \ES{TODO: load these numbers from CSV}
-  }\label{tab:type-error-slice}
-  \centering
-  \begin{tabular}{lrcrrrrcrrrr}
-    \toprule
-                       &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
-                                         \cmidrule{4-7}                        \cmidrule{9-12}
-    Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
-    \midrule
-    Local Syntax       & 44          & & 27.7\% & 46.7\% & 58.6\% & 38.2\% & & 32.6\% & 49.5\% & 60.2\% & 39.4\% \\
-    + \InSlice         & 45          & & 46.4\% & 65.1\% & 76.1\% & 48.9\% & & 55.4\% & 71.0\% & 82.1\% & 57.1\% \\
-    Filter by \InSlice & 44          & & 55.9\% & 71.9\% & 82.9\% & 57.5\% & & 57.8\% & 72.7\% & 82.9\% & 57.6\% \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+\input{evaluation-utility-graph}
+% \begin{table}[ht]
+%   \caption{
+%     Impact of Type Error Slice on Accuracy.
+%     \ES{TODO: load these numbers from CSV}
+%   }\label{tab:type-error-slice}
+%   \centering
+%   \begin{tabular}{lrcrrrrcrrrr}
+%     \toprule
+%                        &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
+%                                          \cmidrule{4-7}                        \cmidrule{9-12}
+%     Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
+%     \midrule
+%     Local Syntax       & 44          & & 27.7\% & 46.7\% & 58.6\% & 38.2\% & & 32.6\% & 49.5\% & 60.2\% & 39.4\% \\
+%     + \InSlice         & 45          & & 46.4\% & 65.1\% & 76.1\% & 48.9\% & & 55.4\% & 71.0\% & 82.1\% & 57.1\% \\
+%     Filter by \InSlice & 44          & & 55.9\% & 71.9\% & 82.9\% & 57.5\% & & 57.8\% & 72.7\% & 82.9\% & 57.6\% \\
+%     \bottomrule
+%   \end{tabular}
+% \end{table}
 
-\autoref{tab:type-error-slice} shows the results of our experiment.
+\autoref{fig:slice-utility} shows the results of our experiment.
 %
@@
 %
-\begin{table}[ht]
-  \caption{
-    Impact of Contextual Features on Accuracy.
-    \ES{TODO: load these numbers from CSV}
-  }\label{tab:contextual-features}
-  \centering
-  \begin{tabular}{lrcrrrrcrrrr}
-    \toprule
-                             &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
-                                               \cmidrule{4-7}                        \cmidrule{9-12}
-    Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
-    \midrule
-    Local Syntax             &  44         & & 56.6\% & 72.6\% & 83.2\% & 57.7\% & & 57.6\% & 72.6\% & 83.1\% & 57.7\% \\
-    \midrule
-    + Size                   &  45         & & 57.2\% & 73.7\% & 83.0\% & 57.4\% & & 60.9\% & 75.5\% & 84.1\% & 57.9\% \\
-    + Context                & 220         & & 61.1\% & 78.7\% & 86.7\% & 63.0\% & & 71.5\% & 84.8\% & 90.8\% & 69.0\% \\
-    + Types                  & 102         & & 63.1\% & 77.8\% & 85.2\% & 61.6\% & & 73.4\% & 85.4\% & 90.7\% & 68.9\% \\
-    \midrule
-    + Context + Size         & 221         & & 61.1\% & 78.8\% & 86.3\% & 62.4\% & & 71.5\% & 84.4\% & 90.8\% & 69.0\% \\
-    + Types + Size           & 103         & & 61.9\% & 78.7\% & 85.6\% & 62.0\% & & 73.1\% & 85.2\% & 91.1\% & 69.5\% \\
-    + Context + Types        & 275         & & 63.1\% & 80.9\% & 88.0\% & 65.0\% & & 77.2\% & 88.3\% & 92.5\% & 72.6\% \\
-    \midrule
-    + Context + Types + Size & 276         & & 62.8\% & 80.5\% & 88.1\% & 65.1\% & & 77.3\% & 88.1\% & 92.7\% & 72.7\% \\
-    \bottomrule
-  \end{tabular}
-  % \begin{minipage}{0.49\linewidth}
+% \begin{table}[ht]
+%   \caption{
+%     Impact of Contextual Features on Accuracy.
+%     \ES{TODO: load these numbers from CSV}
+%   }\label{tab:contextual-features}
 %   \centering
-  % \hiddenF
-  % \begin{tabular}{lrrrr}
+%   \begin{tabular}{lrcrrrrcrrrr}
 %     \toprule
-  %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+%                              &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
+%                                                \cmidrule{4-7}                        \cmidrule{9-12}
+%     Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
 %     \midrule
-  %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
+%     Local Syntax             &  44         & & 56.6\% & 72.6\% & 83.2\% & 57.7\% & & 57.6\% & 72.6\% & 83.1\% & 57.7\% \\
 %     \midrule
-  %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
-  %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
-  %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
+%     + Size                   &  45         & & 57.2\% & 73.7\% & 83.0\% & 57.4\% & & 60.9\% & 75.5\% & 84.1\% & 57.9\% \\
+%     + Context                & 220         & & 61.1\% & 78.7\% & 86.7\% & 63.0\% & & 71.5\% & 84.8\% & 90.8\% & 69.0\% \\
+%     + Types                  & 102         & & 63.1\% & 77.8\% & 85.2\% & 61.6\% & & 73.4\% & 85.4\% & 90.7\% & 68.9\% \\
 %     \midrule
-  %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
-  %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
-  %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
+%     + Context + Size         & 221         & & 61.1\% & 78.8\% & 86.3\% & 62.4\% & & 71.5\% & 84.4\% & 90.8\% & 69.0\% \\
+%     + Types + Size           & 103         & & 61.9\% & 78.7\% & 85.6\% & 62.0\% & & 73.1\% & 85.2\% & 91.1\% & 69.5\% \\
+%     + Context + Types        & 275         & & 63.1\% & 80.9\% & 88.0\% & 65.0\% & & 77.2\% & 88.3\% & 92.5\% & 72.6\% \\
 %     \midrule
-  %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
+%     + Context + Types + Size & 276         & & 62.8\% & 80.5\% & 88.1\% & 65.1\% & & 77.3\% & 88.1\% & 92.7\% & 72.7\% \\
 %     \bottomrule
 %   \end{tabular}
-  % \end{minipage}
-\end{table}
+%   % \begin{minipage}{0.49\linewidth}
+%   % \centering
+%   % \hiddenF
+%   % \begin{tabular}{lrrrr}
+%   %   \toprule
+%   %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
+%   %   \midrule
+%   %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
+%   %   \midrule
+%   %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
+%   %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
+%   %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
+%   %   \midrule
+%   %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
+%   %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
+%   %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
+%   %   \midrule
+%   %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
+%   %   \bottomrule
+%   % \end{tabular}
+%   % \end{minipage}
+% \end{table}
 %
 \paragraph{Results}
-\autoref{tab:contextual-features} summarizes the results of this experiment.
+\autoref{fig:context-utility} summarizes the results of this experiment.
 %

### 2017-04-10 paper/oopsla17-submission/evaluation-utility-graph.tex
adding recall to utility graph

@@
     anchor=center,
-    legend columns=4
+    legend columns=5
   },
@@
 
+\addplot+[stack plots=false, draw=black, fill=none, thick, bar shift=.25cm] table[x=features, y=recall] {\SliceHiddenBench};
+\addlegendentry{Recall}
 \addplot[draw=black, fill=green1, bar shift=.25cm] table[x=features, y=top-1] {\SliceHiddenBench};
@@
 
+\addplot+[stack plots=false, draw=black, fill=none, thick, bar shift=-.25cm] table[x=features, y=recall] {\SliceLinearBench};
+\addlegendentry{Recall}
 \addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=features, y=top-1] {\SliceLinearBench};
@@
 \addlegendentry{\linear}
+
 \end{axis}
@@
 
+\addplot+[stack plots=false, draw=black, fill=none, thick, bar shift=.25cm] table[x=features, y=recall] {\FeatureHiddenBench};
+%\addlegendentry{Recall}
 \addplot[draw=black, fill=green1, bar shift=.25cm] table[x=features, y=top-1] {\FeatureHiddenBench};
@@
 
+\addplot+[stack plots=false, draw=black, fill=none, thick, bar shift=-.25cm] table[x=features, y=recall] {\FeatureLinearBench};
 \addplot[draw=black, fill=blue1, bar shift=-.25cm] table[x=features, y=top-1] {\FeatureLinearBench};

### 2017-04-10 paper/oopsla17-submission/conclusion.tex
small conclusion

@@
 \label{sec:conclusion}
+
+We have presented a new data-driven approach to localizing type errors,
+using machine learning to assign blame based on previously seen novice
+mistakes.
+%
+We translate ill-typed programs into sets of feature vectors, predict
+for each sub-expression whether it should be blamed, rank the
+predictions based on the model's confidence, and select the top three
+blame assignments to present to the user.
+%
+We represent expressions solely by their syntactic class and type, and
+those of their parent and children.
+%
+We have evaluated our approach with five different learning algorithms
+on a suite of over 4,500 student programs, and found that all of our
+models outperform the state of the art, by up to a 16\% margin.
+%
 

### 2017-04-10 paper/oopsla17-submission/introduction.tex
small conclusion

@@
 %
-Yet this power comes with a cost: since type annotations signify the
+Yet this power comes with a cost: type annotations signify the
 programmer's \emph{intent}, without them the checker must
@@
 from their source, an annoyance to veterans of the type system, but a
-major hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.
+significant hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.
 

### 2017-04-10 paper/oopsla17-submission/evaluation-utility-graph.tex
fix recall computation in `features/bin/Evaluate.hs`
better caption

@@
   xtick style={draw=none},
+  xticklabel style={align=center},
   xticklabels={
-    \textsc{Local Syntax},
-    \textsc{+Size}, \textsc{+Context}, \textsc{+Type},
-    +C+S, +T+S, +C+T,
-    +C+T+S
+    \textsc{Local Syntax}\\(44),
+    \textsc{+Size}\\(45), \textsc{+Context}\\(220), \textsc{+Type}\\(102),
+    +C+S\\(221), +T+S\\(103), +C+T\\(275),
+    +C+T+S\\(276)
     % + Size, + Context, + Type,
@@
 \end{tikzpicture}
-\caption{Impact of contextual features on blame accuracy.}\label{fig:context-utility}
+\caption{Impact of contextual features on blame accuracy. Starting from
+  a baseline of local syntactic features, we add each combination of
+  expression size, contextual syntactic, and typing features. The total
+  number of features is given in parentheses.}
+\label{fig:context-utility}
 \end{subfigure}

### 2017-04-11 paper/oopsla17-submission/conclusion.tex
minor conclusion

@@
 %
-We have evaluated our approach with five different learning algorithms
+We have evaluated our approach with five different classifiers
 on a suite of over 4,500 student programs, and found that all of our
@@
 %
+We are particularly excited that our classifiers can be trained on a
+modest amount of data, \eg a single course's worth, and envision a
+future where each course has a model of its students' errors.
 

### 2017-04-11 paper/oopsla17-submission/introduction.tex
use colors in opening example

@@
     match xs with
-    | []   -> __[]__
-    | h::t -> h __+__ ==sumList t==
+    | []   -> (*@\colorbox{tree!75}{\colorbox{sherrloc!75}{[]}}@*)
+    | h::t -> h (*@\colorbox{tree!75}{+}@*) (*@\colorbox{fix!75}{sumList t}@*)
 \end{ecode}
@@
 \caption{(left) An ill-typed \ocaml program that should sum the elements of a
-  list. The \underline{underlined} expressions are equally valid
-  locations to blame. The expression blamed by the \ocaml compiler
-  is in \textbf{bold}.
+  list, with highlights indicating three possible blame assignments based on:
+  %
+  (1) the \colorbox{fix!75}{\ocaml} compiler;
+  %
+  (2) the \colorbox{sherrloc!75}{fix} made by the programmer; and
+  %
+  (3) \colorbox{tree!75}{minimizing} the number of edits required.
+  % The \underline{underlined} expressions are equally valid
+  % locations to blame. The expression blamed by the \ocaml compiler
+  % is in \textbf{bold}.
   %

### 2017-04-11 paper/oopsla17-submission/evaluation-utility-graph.tex
fiddle
clearly give recall axis, looks a bit busy now..

@@
 \begin{axis}[
-  % name=slice,
+  name=slice,
   % scale only axis,
@@
     %legend pos = outer north east,
-    at={(1.5,0.5)},
+    at={(1.75,0.5)},
     anchor=center,
@@
 \addlegendentry{\linear}
-
+\end{axis}
+\begin{axis}[
+  ybar stacked,
+  width=0.5\linewidth,
+  height=5cm,
+  ylabel={Recall},
+  axis y line*=right,
+  ymin=0.2,
+  ymax=1,
+  ytick={0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = false,
+  xmin=0, xmax=1,
+  hide x axis,
+]
 \end{axis}
@@
 \end{subfigure}
+
+%\hfill\mbox{}
+
+\vspace{1\baselineskip}
 
@@
   ybar stacked,
-  width=\linewidth,
+  width=0.9\linewidth,
   height=6cm,
@@
 \end{axis}
+\begin{axis}[
+  ybar stacked,
+  width=0.9\linewidth,
+  height=6cm,
+  ylabel={Recall},
+  axis y line*=right,
+  ymin=0.5,
+  ymax=1,
+  ytick={0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = false,
+  xmin=0, xmax=1,
+  hide x axis,
+]
+\end{axis}
 

### 2017-04-11 paper/oopsla17-submission/evaluation-utility.tex
prune some of the more obnoxious orphans

@@
 %
-Thus, we compare the accuracy of our classifiers on three sets of features:
+Thus, we compare our classifiers on three sets of features:
 %
@@
 competitive when given only local syntactic features, but the \hiddenFH
-quickly outperforms as more features are added.
+quickly outperforms as we add features.
 

### 2017-04-11 paper/oopsla17-submission/gallery.tex
fiddling with graphs instead of tables..
use colors in opening example
prune some of the more obnoxious orphans

@@
 \label{sec:qualitative}
-
-\definecolor{tree}{HTML}{66C2A5}
-\definecolor{sherrloc}{HTML}{FC8D62}
-\definecolor{fix}{HTML}{8DA0CB}
 
@@
 \sherrloc blames the recursive call to |loop| while our classifier
-blames the tuple of arguments --- a reasonable suggestion, but alas not
-the answer the student was looking for.
+blames the tuple of arguments --- a reasonable suggestion, but not
+the answer the student expected.
 
@@
 returning a |int list| rather than |int|, but the recursive call is
-actually correct!
+correct!
 
@@
 \sherrloc should not be blamed; a similar examination suggests that the
-crucial observation is that the parent of the recursive call is a data
+crucial observation is that the recursive call's parent is a data
 constructor application.

### 2017-04-12 paper/oopsla17-submission/learning.tex
learning edits

@@
 
-Recall that we formulate type error localization as a supervised
-classification problem.
+Recall that we formulate type error detection at a single expression as a supervised classification problem.
+This means that there is a training data set $S = \{ (x_i, y_i) \}_{i=1}^n$ of labeled examples, and our goal is to use it to build a classifier or a prediction rule that can predict a label for an input $x$.
+Because we run the classifier on each expression in order to determine the most likely type error location, we also require the classifier to output a {\em{confidence score}} that indicates how certain it is about its prediction.
+
+%Recall that we formulate type error localization as a supervised
+%classification problem.
 %
-This means that we assume the existence of a training set
-$S = \{\langle \mathbf{v}_i, b_i \rangle\}_{i=1}^{n}$
-of sample inputs $\mathbf{v}$ and their corresponding labels $b$.
+%This means that we assume the existence of a training set
+%$S = \{\langle \mathbf{v}_i, b_i \rangle\}_{i=1}^{n}$
+%of sample inputs $\mathbf{v}$ and their corresponding labels $b$.
 %
-The learning task is to model the probability $Pr(b\ |\ \mathbf{v})$
-of a label being assigned to a given input.
+%The learning task is to model the probability $Pr(b\ |\ \mathbf{v})$
+%of a label being assigned to a given input.
 
@@
 interpreting the learned model).
+In this section we consider four standard learning algorithms: (1) logistic regression, (2) decision trees, (3) random forests, and (4) neural networks.
+A thorough introduction to these techniques can be found in introductory
+machine learning textbooks (\eg \cite{FriedmanHastieTibshirani2000}).
+% \ES{TODO: huma/kamalika?}).
 %
-In this section we consider four standard learning algorithms: 
-%
-(1) logistic regression,
-(2) decision trees,
-(3) random forests, and
-(4) neural networks.
-%
-A thorough introduction to these techniques can be found in introductory
-machine learning textbooks (\eg \ES{TODO: huma/kamalika?}).
-%
-We will briefly introduce each technique by describing the model it
-learns, the training procedure, any hyper-parameters that must be tuned
-by a human, and a summary of its advantages and disadvantages.
+
+Below we briefly introduce each technique by describing the model it learns, and summarize its advantages and disadvantages.
+For our application, we are particularly interested in three properties -- expressiveness, interpretability and proneness to overfitting.
+Expressiveness measures how complex the prediction rule is allowed to be, and interpretability measures how easy it is to explain the cause of prediction to a human.
+Finally proneness to overfitting measures how easily the rule generalizes to unseen examples that are not in the training set; a rule that is highly prone to overfitting might perform poorly on a test set even though its training performance is very high.
+
+%We will briefly introduce each technique by describing the model it
+%learns, the training procedure, any hyper-parameters that must be tuned
+%by a human, and a summary of its advantages and disadvantages.
 
@@
 space by focusing on the pros and cons of each technique rather than
-describing the details? This is a PL paper. }
+describing the details?
+This is a PL paper.
+} \ES{KC: I completely agree.
+Perhaps we can add a table here.
+Judging from what we use, the features we need are: (a) expressivity (b) proneness to overfitting and (c) interpretability.
+We can rate the four classifiers in these three areas.
+}
 
-\paragraph{Logistic Regression}
+\paragraph{Logistic Regression.} The simplest classifier we investigate is logistic regression.
+This is a linear model, where the goal is to learn a set of weights $W$ and a bias $b$ that describe the following model for predicting a label $y$ from a feature vector $x$:
+\[ \Pr(y = 1 | x) = \frac{1}{1 + e^{-W^{\top} x - b}} \]
+The weights $W$ and $b$ are learnt from training data, and the value of $\Pr(y | x)$ naturally leads to a confidence score.
+Logistic regression is a widely used classification algorithm, which is largely preferred for its simplicity, relative ease of generalization as well as ease of human interpretability.
+The main limitation is that the prediction rule is constrained to be a linear combination of the features, and hence relatively simple.
+While this can be somewhat mitigated by adding higher order (quadratic or cubic) features, this often requires thorough domain knowledge.
+
 % The simplest model we investigate is logistic regression, which learns a
@@
 
-The simplest model we investigate is logistic regression, which can be
-understood in two steps.
+%The simplest model we investigate is logistic regression, which can be
+%understood in two steps.
 %
-First, we compute the \emph{evidence} for a label $b$ as a linear
-function of the feature vector $\mathbf{v}$, using a weight vector
-$\mathbf{w}$ and a bias $w_0$ that will be learned from the training set.
-$$
-\mathsf{ev} = \sum_j w_j v_j + w_0
-$$
-Then we use the logistic function (also called a sigmoid function)
-$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
-$[0,1]$ interval, which can then be interpreted as the probability that
-the label $b$ should be applied.
+%First, we compute the \emph{evidence} for a label $b$ as a linear
+%function of the feature vector $\mathbf{v}$, using a weight vector
+%$\mathbf{w}$ and a bias $w_0$ that will be learned from the training set.
+%$$
+%\mathsf{ev} = \sum_j w_j v_j + w_0
+%$$
+%Then we use the logistic function (also called a sigmoid function)
+%$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
+%$[0,1]$ interval, which can then be interpreted as the probability that
+%the label $b$ should be applied.
 %
-This gives us the final model:
-$$
-Pr(b\ |\ \mathbf{v}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j v_j - w_0)}
-$$
+%This gives us the final model:
+%$$
+%Pr(b\ |\ \mathbf{v}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j v_j - w_0)}
+%$$
 % $$
@@
 
-Training a logistic regression entails finding optimal values for the
-weights and bias, such that a \emph{cost} function is minimized.
+%Training a logistic regression entails finding optimal values for the
+%weights and bias, such that a \emph{cost} function is minimized.
 %
-In practice, the cost function used for logistic regression is the
-\emph{cross entropy}, a measure of the similarity between two
-probability distributions, between the predicted labels and the ground
-truth.
+%In practice, the cost function used for logistic regression is the
+%\emph{cross entropy}, a measure of the similarity between two
+%probability distributions, between the predicted labels and the ground
+%truth.
 %
-The search for optimal values is typically done via \emph{stochastic
-  gradient descent}, an iterative method for optimizing a cost function.
+%The search for optimal values is typically done via \emph{stochastic
+%  gradient descent}, an iterative method for optimizing a cost function.
 %
-Starting with an initial estimate of the weights (\eg a normal
-distribution), one repeatedly makes predictions for the training data,
-computes the cost function for the predictions, and updates the
-parameters according to the gradient of the cost function.
+%Starting with an initial estimate of the weights (\eg a normal
+%distribution), one repeatedly makes predictions for the training data,
+%computes the cost function for the predictions, and updates the
+%parameters according to the gradient of the cost function.
 %
-This process is repeated until the parameters converge, or until a time
-limit is exhausted.
+%This process is repeated until the parameters converge, or until a time
+%limit is exhausted.
 %
-Stochastic gradient descent requires a hyper-parameter $\eta$, often
-called the \emph{learning rate}, which controls how much the weights and
-bias are updated at each iteration.
+%Stochastic gradient descent requires a hyper-parameter $\eta$, often
+%called the \emph{learning rate}, which controls how much the weights and
+%bias are updated at each iteration.
 
-It has been observed that large weights often coincide with
-\emph{overfitting} of the model --- where the model performs well on the
-training samples but not on the testing samples.
+%It has been observed that large weights often coincide with
+%\emph{overfitting} of the model --- where the model performs well on the
+%training samples but not on the testing samples.
 %
-Thus, it is common to add a \emph{regularization} term to the cost
-function, a popular choice is the $L_2$ norm of the weights
-$\sum_j w_j^2$, which has the effect of penalizing the model for
-learning large weights \citep{Park2008-no}.
+%Thus, it is common to add a \emph{regularization} term to the cost
+%function, a popular choice is the $L_2$ norm of the weights
+%$\sum_j w_j^2$, which has the effect of penalizing the model for
+%learning large weights \citep{Park2008-no}.
 %
-The contribution of the regularization term to the overall cost function
-is controlled by another hyper-parameter $\lambda$, often called the
-regularization rate.
+%The contribution of the regularization term to the overall cost function
+%is controlled by another hyper-parameter $\lambda$, often called the
+%regularization rate.
 
-As a generalized linear model, logistic regression is popular for its
-ease of training and of interpreting the resulting model; however, it
-can be limited in its applicability.
+%As a generalized linear model, logistic regression is popular for its
+%ease of training and of interpreting the resulting model; however, it
+%can be limited in its applicability.
 %
-In particular, it may perform poorly when asked to learn an inherently
-\emph{nonlinear} function.
+%In particular, it may perform poorly when asked to learn an inherently
+%\emph{nonlinear} function.
 %
-This limitation can be mitigated by adding transformed or compound
-features to the model --- \eg if one knows that the label has a
-quadratic relationship with a feature $v$, the transformed feature $v^2$
-can be added --- but this typically requires knowledge of the underlying
-model.
+%This limitation can be mitigated by adding transformed or compound
+%features to the model --- \eg if one knows that the label has a
+%quadratic relationship with a feature $v$, the transformed feature $v^2$
+%can be added --- but this typically requires knowledge of the underlying
+%model.
 
@@
 
-\paragraph{Decision Trees}
-Decision tree algorithms learn a tree of binary predicates over the
-features, recursively partitioning the input space until a final
-classification can be made.
-%
-Each node in the tree contains a single predicate of the form
+\paragraph{Decision Trees.} Decision tree algorithms learn a tree of binary predicates over the features, recursively partitioning the input space until a final
+classification can be made. Each node in the tree contains a single predicate of the form
 $v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
 whether a given input should proceed down the left or right subtree.
-%
 Each leaf is labeled with a prediction and the fraction of training
 samples that would reach it, the latter quantity can be interpreted as
-the decision tree's confidence in its prediction.
 \ES{HUMA: is this correct?}
+the decision tree's confidence in its prediction. This leads to a prediction rule
+that can be quite complex depending on the input data.
 
-Training a decision tree entails finding both a set of good partitioning
-predicates and a good ordering of the predicates.
+Training a decision tree entails finding both a set of good partitioning and a good ordering of the predicates.
+This is usually done in a top-down greedy manner, and there are several standard training algorithms such as C4.5
+\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
+
+Another advantage of decision trees is their ease of interpretation --  the decision rule is a white-box model that can be readily described to a human, especially when the tree is small.
+The main limitation is that these trees are often prone to overfitting; while this is somewhat mitigated in practice by pruning the tree, this is still a well-known problem.
+
+%predicates and a good ordering of the predicates.
 %
-This is usually done in a top-down greedy manner by selecting the
-predicate the best partitions the entire training set, and recursively
-partitioning the subsets.
+%This is usually done in a top-down greedy manner by selecting the
+%predicate the best partitions the entire training set, and recursively
+%partitioning the subsets.
 %
-Standard algorithms for training decision trees include C4.5
-\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
+%Standard algorithms for training decision trees include C4.5
+%\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
 %
-Overfitting can also be a problem for decision trees; the countermeasure
-is usually either halting the learning process before the tree grows too
-deep, or pruning the resulting tree.
+%Overfitting can also be a problem for decision trees; the countermeasure
+%is usually either halting the learning process before the tree grows too
+%deep, or pruning the resulting tree.
 
-The advantages of decision trees are that they are relatively easy to
-train and that they produce a \emph{white-box} model, making it trivial
-to analyze the entire model and explain individual predictions.
+%The advantages of decision trees are that they are relatively easy to
+%train and that they produce a \emph{white-box} model, making it trivial
+%to analyze the entire model and explain individual predictions.
 %
-The disadvantages are that they are prone to overfitting and that the
-model can be unstable, \ie small changes to the training data may
-produce large changes in the model.
+%The disadvantages are that they are prone to overfitting and that the
+%model can be unstable, \ie small changes to the training data may
+%produce large changes in the model.
 %
-Both disadvantages are addressed by random forests.
+%Both disadvantages are addressed by random forests.
 
-\paragraph{Random Forests}
-One technique for increasing the accuracy of a classifier is to train an
-\emph{ensemble} of distinct classifiers and use a majority vote
-to make the final prediction.
+\paragraph{Random Forests.} Random forests mitigate the overfitting problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
+The agreement among the distinct trees in the ensemble forms a natural confidence score.
+Since each classifier in the ensemble is a decision tree, this still allows for complex and highly expressive classifiers.
+\ES{Huma: is this the confidence score that we are using?}
+
+%One technique for increasing the accuracy of a classifier is to train an
+%\emph{ensemble} of distinct classifiers and use a majority vote
+%to make the final prediction.
 %
-Random forests are an instance of this method, with decision
-trees as the underlying classifiers.
+%Random forests are an instance of this method, with decision
+%trees as the underlying classifiers.
 %
@@
 %
-In order to make a prediction for an unseen sample, the random forest
-asks each decision tree to make a prediction, and then predicts the most
-common label.
+%In order to make a prediction for an unseen sample, the random forest
+%asks each decision tree to make a prediction, and then predicts the most
+%common label.
 
 The diversity of the underlying models tends to make random forests less
-susceptible to the overfitting and instability of decision trees, but it
+susceptible to the overfitting, but it
 also makes the learned model more difficult to interpret. 
 %
-Furthermore, the number of trees in the forest, $N$, becomes an important
-hyper-parameter that must be set manually.
+%Furthermore, the number of trees in the forest, $N$, becomes an important
+%hyper-parameter that must be set manually.
 

### 2017-04-12 paper/oopsla17-submission/learning.tex
editing 3.3

@@
 Recall that we formulate type error detection at a single expression as a supervised classification problem.
-This means that there is a training data set $S = \{ (x_i, y_i) \}_{i=1}^n$ of labeled examples, and our goal is to use it to build a classifier or a prediction rule that can predict a label for an input $x$.
-Because we run the classifier on each expression in order to determine the most likely type error location, we also require the classifier to output a {\em{confidence score}} that indicates how certain it is about its prediction.
+This means that we are given a training data set $S = \{ (x_i, y_i) \}_{i=1}^n$ of labeled examples, and our goal is to use it to build a {\em{classifier}}, or a prediction rule that can predict a label for an input $x$.
+Since we apply the classifier on each expression in the program to determine those that are the most likely to be type errors, we also require the classifier to output a {\em{confidence score}} that measures how sure the classifier is about its prediction.
 
@@
 Below we briefly introduce each technique by describing the model it learns, and summarize its advantages and disadvantages.
-For our application, we are particularly interested in three properties -- expressiveness, interpretability and proneness to overfitting.
-Expressiveness measures how complex the prediction rule is allowed to be, and interpretability measures how easy it is to explain the cause of prediction to a human.
-Finally proneness to overfitting measures how easily the rule generalizes to unseen examples that are not in the training set; a rule that is highly prone to overfitting might perform poorly on a test set even though its training performance is very high.
+For our application, we are particularly interested in three properties -- expressiveness, interpretability and ease of generalization.
+Expressiveness measures how complex prediction rules are allowed to be, and interpretability measures how easy it is to explain the cause of prediction to a human.
+Finally ease of generalization measures how easily the rule generalizes to unseen examples that are not in the training set; a rule that is not easily generalizable might perform poorly on an unseen test set even when its training performance is high.
 
@@
 The weights $W$ and $b$ are learnt from training data, and the value of $\Pr(y | x)$ naturally leads to a confidence score.
-Logistic regression is a widely used classification algorithm, which is largely preferred for its simplicity, relative ease of generalization as well as ease of human interpretability.
-The main limitation is that the prediction rule is constrained to be a linear combination of the features, and hence relatively simple.
-While this can be somewhat mitigated by adding higher order (quadratic or cubic) features, this often requires thorough domain knowledge.
+Logistic regression is a widely used classification algorithm, which is largely preferred for its simplicity, ease of generalization as well as interpretability.
+Its main limitation is that the prediction rule is constrained to be a linear combination of the features, and hence relatively simple.
+While this can be somewhat mitigated by adding higher order (quadratic or cubic) features, this often requires substantial domain knowledge.
 
@@
 the decision tree's confidence in its prediction. This leads to a prediction rule
-that can be quite complex depending on the input data.
+that can be quite expressive depending on the data used to build it.
 
-Training a decision tree entails finding both a set of good partitioning and a good ordering of the predicates.
+Training a decision tree entails finding both a set of good partitioning and a good ordering of the predicates based on data.
 This is usually done in a top-down greedy manner, and there are several standard training algorithms such as C4.5
@@
 Another advantage of decision trees is their ease of interpretation --  the decision rule is a white-box model that can be readily described to a human, especially when the tree is small.
-The main limitation is that these trees are often prone to overfitting; while this is somewhat mitigated in practice by pruning the tree, this is still a well-known problem.
+However, main limitation is that these trees often do not generalize well; while this is somewhat mitigated in practice by pruning the tree, this is still a well-known problem.
 
@@
 
-\paragraph{Random Forests.} Random forests mitigate the overfitting problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
+\paragraph{Random Forests.} Random forests mitigate the generalization problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
 The agreement among the distinct trees in the ensemble forms a natural confidence score.
@@
 \paragraph{Neural Networks}
-The last (and most complex) model we discuss is a type of neural network
+The last (and most complex) model we use is a type of neural network
 called a \emph{multi-layer perceptron} (see \citealt{Nielsen2015-pu} for
-a thorough introduction to neural networks).
+an introduction to neural networks).
+
+A multi-layered perceptron can be represented as a directed acyclic graph whose nodes are arranged in layers that are fully connected by weighted edges.
+The first layer corresponds to the input features, and the final to the output.
+The output of a node $v$ in layer $l$ is given by:
+\[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
+where $N(v)$ is the set of nodes in the previous layer that are adjacent to $v$, $W_{jv}$ is the weight of the $(j, v)$ edge and $h_j$ is the output of node $j$ in the previous layer.
+Finally $g$ is a non-linear function, called the activation function, which in recent work is commonly chosen to be the \emph{rectified linear unit} (ReLU),
+defined as $g(x) = \mathsf{max}(0,x)$ \citep{Nair2010-xg}.
+The number of layers, the number of neurons per layer, and the connections between layers constitute the \emph{architecture} of a neural network.
+In our work, we use relatively simple neural networks which have an input layer, a single hidden layer and an output layer.
+
+HEREHERE
 %
-Neural networks can be understood as a collection of linear models, just
-like the one used in logistic regression, arranged in a directed graph.
+%Neural networks can be understood as a collection of linear models, just
+%like the one used in logistic regression, arranged in a directed graph.
 %
-Each node in the graph corresponds to a single linear model, called a
-\emph{neuron}, and the edges propagate signals from the input features,
-through the neurons, to the output labels.
+%Each node in the graph corresponds to a single linear model, called a
+%\emph{neuron}, and the edges propagate signals from the input features,
+%through the neurons, to the output labels.
 %
-The linear model in each neuron is wrapped in an \emph{activation
-  function}, which controls whether (or how strongly) the signal will be
-propagated out.
+%The linear model in each neuron is wrapped in an \emph{activation
+%  function}, which controls whether (or how strongly) the signal will be
+%propagated out.
 %
-These activation functions are what allow neural networks to accurately
-model nonlinear functions.
+%These activation functions are what allow neural networks to accurately
+%model nonlinear functions.
 %
-A common choice of activation function is the same sigmoid function used
-by logistic regression, alternatives include the hyperbolic $\tanh$
-function, and more recently the \emph{rectified linear unit} (ReLU),
-defined as $f(x) = \mathsf{max}(0,x)$ \citep{Nair2010-xg}.
+%A common choice of activation function is the same sigmoid function used
+%by logistic regression, alternatives include the hyperbolic $\tanh$
+%function, and more recently the \emph{rectified linear unit} (ReLU),
+%defined as $f(x) = \mathsf{max}(0,x)$ \citep{Nair2010-xg}.
 
-The neurons are arranged in \emph{layers}; each layer contains a set of
-neurons that take in signals from the previous layer and propagate
-signals to the next layer.
+%The neurons are arranged in \emph{layers}; each layer contains a set of
+%neurons that take in signals from the previous layer and propagate
+%signals to the next layer.
 %
-The number of layers, the number of neurons per layer, and the
-connections between layers constitute the \emph{architecture} of a
-neural network, and are all important hyper-parameters.
+% and are all important hyper-parameters.
 %
-We consider a particular class of neural networks called
-\emph{multi-layer perceptrons} (MLP), where the layers a fully connected to
-each other, \ie each neuron propagates its signal to every neuron in the
-subsequent layer.
+%We consider a particular class of neural networks called
+%\emph{multi-layer perceptrons} (MLP), where the layers a fully connected to
+%each other, \ie each neuron propagates its signal to every neuron in the
+%subsequent layer.
 %
-We will only consider MLPs with a single layer of neurons, but we will
-vary the number of neurons in our experiments.
+%We will only consider MLPs with a single layer of neurons, but we will
+%vary the number of neurons in our experiments.
 
-Neural networks are trained similarly to logistic regression using
-stochastic gradient descent, and thus the learning rate $\eta$ and
-regularization rate $\lambda$ are also important hyper-parameters.
+%Neural networks are trained similarly to logistic regression using
+%stochastic gradient descent, and thus the learning rate $\eta$ and
+%regularization rate $\lambda$ are also important hyper-parameters.
 %
-They are, however, much more difficult to train as they require larger
-training sets.
+%They are, however, much more difficult to train as they require larger
+%training sets.
 
-Neural networks are popular for their ability to uncover ``hidden''
-features in the data.
+%Neural networks are popular for their ability to uncover ``hidden''
+%features in the data.
 %

### 2017-04-12 paper/oopsla17-submission/learning.tex

@@
 The first layer corresponds to the input features, and the final to the output.
-The output of a node $v$ in layer $l$ is given by:
+The output of a node $v$ in an internal layer is given by:
 \[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
@@
 The number of layers, the number of neurons per layer, and the connections between layers constitute the \emph{architecture} of a neural network.
-In our work, we use relatively simple neural networks which have an input layer, a single hidden layer and an output layer.
+In this work, we use relatively simple neural networks which have an input layer, a single hidden layer and an output layer.
 
-HEREHERE
+A major advantage of neural networks is its ability to discover interesting combinations of features through non-linearity, which significantly reduces the need for manual feature engineering, and allows high expressivity.
+On the other hand, this makes the networks particularly difficult to interpret and also difficult to generalize under vast amounts of training data are available.
+
+%HEREHERE
 %
@@
 %
-Each neuron learns a different linear combination of the signals from
-the previous layer; combined with the activation function this can be
-thought of as the neuron learning to respond to a new, \emph{compound}
-feature of its inputs.
+%Each neuron learns a different linear combination of the signals from
+%the previous layer; combined with the activation function this can be
+%thought of as the neuron learning to respond to a new, \emph{compound}
+%feature of its inputs.
 %
-This ability to discover interesting combinations of features makes
-neural networks easier to use in some respect; they can reduce the
-importance of manual \emph{feature engineering} (selecting interesting
-features to provide as inputs).
+%This ability to discover interesting combinations of features makes
+%neural networks easier to use in some respect; they can reduce the
+%importance of manual \emph{feature engineering} (selecting interesting
+%features to provide as inputs).
 %
-However, the sheer size and complexity of neural networks makes the
-model particularly difficult to interpret --- they are generally regarded
-as ``black boxes.''
+%However, the sheer size and complexity of neural networks makes the
+%model particularly difficult to interpret --- they are generally regarded
+%as ``black boxes.''
 %
-Neural networks are also very susceptible to overfitting, so
-regularization methods like the $L_2$ norm used for logistic regression
-are important.
+%Neural networks are also very susceptible to overfitting, so
+%regularization methods like the $L_2$ norm used for logistic regression
+%are important.
 
-\ES{should mention that logistic regression can be thought of as a degenerate MLP with 0 layers, and perhaps add a diagram?}
+%\ES{should mention that logistic regression can be thought of as a degenerate MLP with 0 layers, and perhaps add a diagram?}
 
-\ES{need to say something about neurons corresponding to ``hidden'' features, give intuition..}
+%\ES{need to say something about neurons corresponding to ``hidden'' features, give intuition..}
 

### 2017-04-12 paper/oopsla17-submission/gallery.tex
a more printer-friendly colorscheme, should probably add legends..

@@
 %
-(2) \colorbox{sherrloc!75}{\sherrloc's} prediction,
+(2) \hlSherrloc{\sherrloc's} prediction,
 %
-(3) our \colorbox{tree!75}{\dectree's} prediction, and
+(3) our \hlTree{\dectree's} prediction, and
 %
@@
 %
-For these programs we will additionally \colorbox{fix!75}{highlight} the
+For these programs we will additionally \hlFix{highlight} the
 correct blame location.
@@
       else
-        (*@\colorbox{fix!75}{clone} \colorbox{sherrloc!75}{([\colorbox{tree!75}{x}] @\ acc)}@*) (n - 1) in
+        (*@\hlFix{clone} \hlSherrloc{([\hlTree{x}] @\ acc)}@*) (n - 1) in
     loop [] n
@@
   let rec clone x n =
-    let rec loop (*@\colorbox{fix!75}{x n acc}@*) =
+    let rec loop (*@\hlFix{x n acc}@*) =
       if n < 0 then
@@
       else
-        (*@\colorbox{sherrloc!75}{loop} \colorbox{tree!75}{(x, (n - 1), (x :: acc))}@*) in
+        (*@\hlSherrloc{loop} \hlTree{(x, (n - 1), (x :: acc))}@*) in
     loop (x, n, [])
@@
     else
-      [n mod 10] @ (*@\colorbox{tree!75}{[ \colorbox{sherrloc!75}{digitsOfInt (n / 10)} ]}@*)
+      [n mod 10] @ (*@\hlTree{[ \hlSherrloc{digitsOfInt (n / 10)} ]}@*)
 \end{ecode}
@@
     else
-      (l1, (*@\colorbox{tree!75}{\colorbox{sherrloc!75}{clone 0 n} :: l2}@*))
+      (l1, (*@\hlTree{\hlSherrloc{clone 0 n} :: l2}@*))
 \end{ecode}

### 2017-04-12 paper/oopsla17-submission/introduction.tex
prune some of the more obnoxious orphans
a more printer-friendly colorscheme, should probably add legends..

@@
     match xs with
-    | []   -> (*@\colorbox{tree!75}{\colorbox{sherrloc!75}{[]}}@*)
-    | h::t -> h (*@\colorbox{tree!75}{+}@*) (*@\colorbox{fix!75}{sumList t}@*)
+    | []   -> (*@\hlTree{\hlSherrloc{[]}}@*)
+    | h::t -> h (*@\hlTree{+}@*) (*@\hlFix{sumList t}@*)
 \end{ecode}
@@
   %
-  (1) the \colorbox{fix!75}{\ocaml} compiler;
+  (1) the \hlFix{\ocaml} compiler;
   %
-  (2) the \colorbox{sherrloc!75}{fix} made by the programmer; and
+  (2) the \hlSherrloc{fix} made by the programmer; and
   %
-  (3) \colorbox{tree!75}{minimizing} the number of edits required.
+  (3) \hlTree{minimizing} the number of edits required.
   % The \underline{underlined} expressions are equally valid
@@
 |sumList| returns a |list| where an |int| was expected by the |+|
-operator. Note that this blame is inconsistent with the programmer's
-intention and thus may not be helpful debugging information to a novice.
+operator.
+%
+Note that this blame is inconsistent with the programmer's intention and
+may not be helpful debugging information to a novice.
 

### 2017-04-12 paper/oopsla17-submission/overview.tex
prune some of the more obnoxious orphans
a more printer-friendly colorscheme, should probably add legends..

@@
 %
-(2) generating correct labels for supervised learning; and
+%(2) generating correct labels for supervised learning; and
+(2) identifying ``correct'' blame assignments in our dataset; and
 %
@@
 The first and prime challenge we address is how to translate programs
-into a format favorable to machine learning. 
+into a format favorable to learning a model of type errors.
 %
@@
 %
+Furthermore, program representations that are amenable to one learning
+task, \eg code completion, may not be favorable to our problem of
+assigning blame for type errors.
+
 Machine learning techniques typically take as input \emph{feature
@@
 %
-(1) what are interesting features of a program; and
+(1) which features of a program would be useful for modeling type errors; and
 %
@@
 %
-Our features are \emph{queries} over the AST. Our initial set of
-interesting features includes the syntactic class of each expression.
+We model features as \emph{queries} over the AST.\@
+%
+Our initial set of interesting features includes the syntactic class of
+each expression.
 %
@@
 %
+Modeling the syntactic class of an expression gives the classifier a
+basic notion of the relative frequency of blame assignment for the
+various program elements, \ie perhaps |[]| is \emph{empirically} more
+likely to be blamed than |+|.
+% % ES: (2) doens't really make sense here
+% (2) the type checker uses the syntactic class of an expression to
+%     impose typing constraints on the neighboring expressions.
+
 A programming language has a fixed set of AST elements,\footnote{We will
@@
 %
-For example, we might expect that the recursive call |sumList t| is
+The context is particularly important for our task as each expression
+imposes typing constraints on its neighbors, \eg a |+| operator tells
+the type checker that both children must have type |int| and that
+the parent must accept an |int|.
+%
+We might expect that the recursive call |sumList t| is
 unlikely to be at fault because |sumList| has a function type --- the
@@
 \emph{contextual features} of its parent and children.
-%
-These contextual features include the syntactic class of the neighboring
+
+Our contextual features include the syntactic class of the neighboring
 expressions, their inferred types (when available), and, crucially, whether
@@
 Prior work has often enlisted expert users to manually judge ill-typed
-programs and determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk}, but
-this approach does not scale well to a dataset large enough to support
-machine learning.
+programs and determine the ``correct'' fix
+\citep[\eg][]{Loncaric2016-uk}, but this approach does not scale well to
+a dataset large enough to support machine learning.
 %
@@
 Thus we would determine that the |[]| expression (and \emph{not} the
-|+| or the recursive call |sumList t|) is to blame for the error.
+|+| or the recursive call |sumList t|) is to blame.
 
-\paragraph{\textbf{Challenge (3): What should be presented to the user?}}
+\paragraph{\textbf{Challenge (3): How can we intuitively present our feedback?}}
 We now have the features and labels needed to train a classifier to

### 2017-04-13 paper/oopsla17-submission/evaluation-accuracy.tex
complexity rather than compression
misc

@@
 %
-To ensure a fair comparison when evaluating \mycroft and
-\sherrloc, we removed from the dataset programs where they predicted
-locations that our oracle could not match with a program expression:
-6--8\% of programs for \mycroft and 3--4\% for \sherrloc.
+Our experimental design gives the state of the art tools the ``benefit
+of the doubt'' in two ways.
+% To ensure a fair comparison when evaluating \mycroft and
+% \sherrloc, 
+First, when evaluating \mycroft and \sherrloc, we did not consider
+programs where they predicted locations that our oracle could not match
+with a program expression: 6--8\% of programs for \mycroft and 3--4\%
+for \sherrloc.
 %
-We also do not consider programs where \mycroft or \sherrloc timed out
-(after one minute) or where they encountered an unsupported language
-feature: another 5\% for \mycroft and 12--13\% for \sherrloc. This
-experimental design gives the state of the art tools the ``benefit of the
-doubt''.
+Second, we similarly ignored programs where \mycroft or \sherrloc timed
+out (after one minute) or where they encountered an unsupported language
+feature: another 5\% for \mycroft and 12--13\% for \sherrloc.
+%
 
@@
   networks, both trained with $\eta = 0.001$, $\lambda = 0.001$, and a
-  mini-batch size of 200. The first MLP contains a single hidden layer
-  of 10 neurons, and the second contains a hidden layer of 500
-  neurons. This allows us to investigate how well the MLP can
-  \emph{compress} its model (cf.~\cite{FIXME}). The neurons use
-  rectified linear units (ReLU) as their activation function, a common
-  practice in modern neural networks.
+  mini-batch size of 200.
+  %
+  The first MLP contains a single hidden layer of 10 neurons, and the
+  second contains a hidden layer of 500 neurons.
+  %
+  This gives us a measure of the complexity of the MLP's model, \ie
+  if the model requires many compound features, one would expect \hiddenFH
+  to outperform \hiddenT.
+  % This allows us to investigate how well the MLP can \emph{compress} its
+  % model (cf.~\cite{FIXME}).
+  %
+  The neurons use rectified linear units (ReLU) as their activation
+  function, a common practice in modern neural networks.
 \end{description}

### 2017-04-13 paper/oopsla17-submission/evaluation.tex
misc

@@
   %
-  Which feature sets are \emph{most important} in general?
+  Which feature sets are \emph{most important} for accurate predictions?
   % How much do the features described in \autoref{sec:features}
@@
 \paragraph{Blame Oracle}
-
-A disadvantage of using diffs as blame oracles is that students may have
+Recall from \autoref{sec:labels} that we automatically extract a blame
+oracle for each ill-typed program from the (AST) diff between it and the
+student's eventual fix.
+%
+A disadvantage of using diffs in this manner is that students may have
 made many, potentially unrelated, changes between compilations; at some

### 2017-04-13 paper/oopsla17-submission/evaluation-accuracy.tex
more notes from kamalika

@@
 %
+Our classifiers consistently outperform the competition, ranging from
+58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear classifier to
+71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
+%
 Our baseline of selecting at random achieves 30\% Top-1
@@
 %
-\sherrloc outperforms both baselines with 56\% Top-1 accuracy (84--86\% Top
-3), while \mycroft actually underperforms \ocaml with 38--41\% Top-1
-accuracy.
+\sherrloc outperforms both baselines, and comes close to our \linear
+classifier, with 56\% Top-1 accuracy (84--86\% Top 3), while \mycroft
+actually underperforms \ocaml with 38--41\% Top-1 accuracy.
 %
-Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
-ranging from 58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear
-classifier to 71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
+% Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
+% ranging from 58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear
+% classifier to 71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
 
-Surprisingly, there is little variation in accuracy between classifiers.
+Surprisingly, there is little variation in accuracy between our
+classifiers.
+%
 With the exception of the \linear model, they all achieve around 70\%

### 2017-04-13 paper/oopsla17-submission/evaluation-utility.tex
more notes from kamalika

@@
 We have shown that we can train a classifier to effectively localize
-type errors, but which of the 276 features that we use are contributing
-the most?
+type errors, but which of the feature classes from
+\autoref{sec:features} are contributing the most to our accuracy?
 %
-To answer this question we investigate the performance of
-classifiers trained on various subsets of the features.
+We focus specifically on feature \emph{classes} rather than individual
+features as our 276 features are conceptually grouped into a much
+smaller number of \emph{categorical} features.
+%
+For example, the syntactic class of an expression is conceptually a
+feature but there are 44 possible values it could take; to encode this
+feature for learning we split it into 44 distinct binary features.
+%
+Analyses that focus on individual features, \eg \textsc{ANOVA},
+are difficult to interpret in our setting, as they will tell us the
+importance of the binary features but not the higher-level categorical
+features.
+%
+Thus, to answer our question we investigate the performance of
+classifiers trained on various subsets of the feature classes.
 
@@
 where $\mathsf{predicted}$ is limited to the top 3 predictions of the
-classier, and $\mathsf{oracle}$ is limited to the expressions that
-changed \emph{and} were part of the type error slice. We make the
-latter distinction as (1) changes that are not part of the type error
-slice are noise in the data set, and (2) it makes the comparison between
-feature sets easier to interpret since $\mathsf{oracle}$ never changes.
+classier, and $\mathsf{oracle}$ is the student's fix, limited to changes
+that part of the type error slice.
+%
+We make the latter distinction as (1) changes that are not part of the
+type error slice are noise in the data set, and (2) it makes the
+comparison between feature sets easier to interpret since
+$\mathsf{oracle}$ never changes.
 % NOTE: keep this at the end of the para or it screws up spacing...
@@
 
+\paragraph{Results}
 \autoref{fig:slice-utility} shows the results of our experiment.

### 2017-04-13 paper/oopsla17-submission/overview.tex
misc
more notes from kamalika

@@
 programs and determine the ``correct'' fix
-\citep[\eg][]{Loncaric2016-uk}, but this approach does not scale well to
-a dataset large enough to support machine learning.
+\citep[\eg][]{Lerner2007-dt,Loncaric2016-uk}, but this approach does not
+scale well to a dataset large enough to support machine learning.
 %
@@
 %
-Rather than relying on expert users, we allow our students to do the
-work of assigning blame themselves.
+Rather than relying on expert users, we follow \citet{Zhang2014-lv} and
+allow our students to do the work of assigning blame themselves.
 %

### 2017-04-13 paper/oopsla17-submission/abstract.tex
asd

\begin{abstract}
  Localizing type errors is challenging in languages with global type
  inference, as the type checker must make assumptions about what the
  programmer intended to do.
  %

  We present a data-driven approach to type error localization that uses
  machine learning to assign blame based on previous
  ill-typed programs and their fixes.
  %
  First, we gather a large set of over 4,500 ill-typed student programs
  and their subsequent fixes.
  %
  Second, we frame type error localization as a supervised learning
  problem: to predict, for each expression in the
  program, whether it should be blamed.
  %
  We represent each expression by its syntactic class and type and those
  of its parent and children, associating blame with expressions that
  were changed in the human-fixed program.
  %
  Finally, we train and evaluate a series of off-the-shelf classifiers
  on our suite of programs,
  and find that they improve localization accuracy from a state-of-the-art
  56\% to 72\%.
\end{abstract}

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

--- 2017-04-12 paper/oopsla17-submission/introduction.tex
@@
 
-Type systems are one of the most widely-used techniques for program
-verification. A type checker only accepts a program if it can
-\emph{prove} the absence of an entire class of bugs.
+In Robin Milner's memorable words
 %
-Typed functional languages like \ocaml and \haskell are equipped with
-particularly powerful type checkers that can operate without a single
-annotation, alleviating a prime concern with verification techniques.
+``types are the leaven of computer programming;
+  they make it digestible.''
 %
-Yet this power comes with a cost: type annotations signify the
-programmer's \emph{intent}, without them the checker must
-make assumptions about what the programmer intended to do.
+Languages like \ocaml and \haskell make
+the value-proposition for types even more
+appealing by using constraints to (mostly)
+automatically synthesize the types for all
+program terms without troubling the
+programmer for any annotations.
+
+Unfortunately, this automation has come at a price.
+Verbose type annotations (as in, \eg Java) signify
+the programmer's \emph{intent} and help to correctly
+pinpoint the error when the code is ill-typed.
 %
-This results in the common phenomenon of type errors being reported far
-from their source, an annoyance to veterans of the type system, but a
-significant hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.
+In contrast, lacking such signifiers, automatic
+type inference algorithms are prone to report
+type errors quite far from their actual source.
+%
+While this can seem like a minor annoyance to
+veteran programmers, it turns out to be a significant
+stumbling block for novices~\citep{Wand1986-nw,Joosten1993-yx}.
+
+\mypara{Localizing Errors}
+%
+Several recent papers have proposed ways
+to improve feedback via error \emph{localization}.
+%
+At a high-level these techniques analyze
+the set of typing constraints to find
+the minimum (weighted) subset that,
+if removed would make the constraints
+satisfiable and hence, memory-safe~\citep{Jose:2011}
+or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-vm,Pavlinovic2014-mr}.
+The finger of blame is then pointed at the
+sub-terms that yielded those constraints.
+
+This minimum-weight approach suffers
+from several drawbacks.
+%
+First, they are not \emph{portable}:
+the algorithms for computing
+the minimum weighted subset must be
+designed afresh for different kinds
+of type systems and constraints.
+%
+Second, they are not \emph{extensible}:
+the weights and analyses are limited
+to attributes of the constraints, and
+not other \emph{signals} from the
+program's syntax that may be more
+germane to error localization.
+%
+Third, and perhaps most importantly,
+they are not \emph{adaptable}: the
+actual ``weights'' are fixed in
+an ad-hoc fashion, based on the
+\emph{analysis designers} notion
+of what kinds of errors are more
+or less likely, rather than
+capturing and adapting to the
+ground truth about the kinds
+of errors programmers actually
+make in practice.
+
+\mypara{A Data-Driven Approach}
+%
+In this paper, we introduce \toolname
+a \emph{data-driven} approach to error
+localization.
+%
+\toolname works by using a large corpus
+of training data --- pairs of ill-typed
+programs and their ``fixed'' versions ---
+to automatically \emph{learn a model}
+of where the error is most likely.
+%
+When given a new ill-typed program,
+\toolname simply executes the model
+to generate a list of likely error
+locations ranked by likelihood.
+%
+We evaluate \toolname by comparing its
+precision against the state-of-the-art
+on a set of over 4,500 ill-typed Ocaml
+programs drawn from introductory
+programming classes.
+%
+We show that \toolname's data-driven
+model is able to correctly predict
+the exact sub-expression that needs to
+be changed, 74\% of the time compared
+to \ocaml's 45\% and \sherrloc's 56\%.
+%
+When the \emph{top} ranked location
+is considered. \toolname's accuracy jumps
+to above 85\% (resp. 90\%) when we consider
+the \emph{top-2} and \emph{top-3} locations.
+
+% - 45
+% - 56
+% - 74
+%
+% - 85
+% - 91
+
+AST-Dispersion
+
+AST
+
+\toolname addresses both of the limitations
+of previous
+
+* portable   [use different features]
+* extensible [add more features]
+* adaptable  [add more data]
+
+%
+Our solution is to \emph{learn} how to
+localize type errors by observing
+a large set of (novice) errors and their
+\emph{subsequent fixes}.
+%
+
+\mypara{Contributions}
+%
+We accomplish this via three concrete contributions.
+
+\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
+Our first contribution is a large dataset of novice \ocaml programs,
+including, but not limited to, \emph{ill-typed} programs.
+%
+We used an instrumented version of the \ocaml compiler
+to collect student interactions over two instances of
+an undergraduate Programming Languages course at
+%
+\begin{anonsuppress}
+UC San Diego (IRB \#140608).
+\end{anonsuppress}
+\begin{noanonsuppress}
+AUTHOR's INSTITUTION (IRB HIDDEN).
+\end{noanonsuppress}
+%
+The data consist of a time-series of programs submitted to the \ocaml
+compiler, from which we extracted a set of \emph{ill-typed} programs
+and their \emph{subsequent fixes}.
+%
+Our dataset comprises over 4,500 such pairs of programs, and many more
+programs that are either well-typed or lack a corresponding fix.
+%
+The entire dataset (both the raw time-series and our extracted pairs of
+ill-typed programs and fixes) is available at
+%
+\begin{anonsuppress}
+\ES{TODO: finally post the data somehwere..}.
+\end{anonsuppress}
+\begin{noanonsuppress}
+HIDDEN FOR DOUBLE-BLIND REVIEW.
+\end{noanonsuppress}
+%
+
+\paragraph{\textbf{(2) Data-Driven Diagnosis of Type Errors}}
+Our second contribution is a \emph{data-driven} technique for localizing
+type errors (\autoref{sec:learning}).
+%
+We use machine learning techniques to train a set of classifiers that can
+predict, given an ill-typed program, which expressions should be changed
+for the program to type check.
+%
+Our classifiers can be trained on a relatively small set of programs
+(\ie a single course's worth).
+%
+% and outperform the state of the art by
+% 15--30\% in blame accuracy.
+%
+The key technical challenge that we tackle is defining a suitable
+interface between the type checker and the classifier, specifically
+\emph{embedding} programs as feature vectors and presenting the
+classifier's predictions to the user.
+
+\paragraph{\textbf{(3) A Large-Scale Evaluation of Type Error Localization}}
+Our third contribution is an evaluation of our classifiers and the
+state-of-the-art techniques on our suite of over 4,500 student programs
+(\autoref{sec:evaluation}), the largest evaluation of type error
+localization techniques we are aware of.
+%
+We judge the correctness of a localization by comparing it to the
+student's eventual fix and find that our classifiers are 15--30\% more
+accurate than the state of the art.
+%
+We also examine which features contribute the most to our
+classifiers' predictions, and endeavor to \emph{explain} specific
+predictions that our classifiers get right (and wrong).
+
+Overall, our results show that data-driven diagnosis is a practical
+technique for localizing type errors by learning from past mistakes.
+
+%
+% We believe, furthermore, that with the right set of features our
+% approach could be applied to other domains where there are multiple,
+% equally valid explanations for an error.
+% \ES{this sounds really dumb... need better phrasing}
+
+Consider the |sumList| program in \autoref{fig:sumList},
+written by a student in an undergraduate Programming Languages
+course.
+%
+The program is meant to sum the integers in a list, but
+the student has accidentally given the empty list as the
+base case, rather than 0.
+%
+The \ocaml compiler collects typing constraints as it
+traverses the program, and reports an error the moment
+it finds an inconsistent constraint.
+%
+In this case it blames the recursive call to |sumList|,
+complaining that |sumList| returns a |list| where an |int|
+was expected by the |+| operator.
+%
+Note that this \emph{blame} is inconsistent with the
+programmer's intention and may not be helpful debugging
+information to a novice.
+
+It may appear obvious to the reader that
+|[]| is the correct expression to blame,
+but how is the type checker to know that?
+
+Indeed, both the venerable
+Ocaml compiler as well as
+recent techniques~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
+fail to distinguish between
+the two underlined expressions
+in \autoref{fig:sumList};
+it would be equally valid
+to blame \emph{either}
+of them alone.
+%
+The |[]| on line 3 could be changed to |0|,
+or the |+| on line 4 could be changed to
+either |@| (list append) or |::|, all of
+which would give type-correct programs.
+%
+Thus, these state-of-the-art techniques
+are forced to either blame \emph{both}
+locations, or choose one \emph{arbitrarily}.
 
@@
 \end{figure}
-%
-Consider the |sumList| program in \autoref{fig:sumList}, written by a
-student in an undergraduate Programming Languages course.
-%
-The program is meant to sum the integers in a list, but the student has
-accidentally given the empty list as the base case, rather than 0.
-%
-The \ocaml compiler collects typing constraints as it traverses the
-program, and reports an error the moment it finds an inconsistent
-constraint.
-%
-In this case it blames the recursive call to |sumList|, complaining that
-|sumList| returns a |list| where an |int| was expected by the |+|
-operator.
-%
-Note that this blame is inconsistent with the programmer's intention and
-may not be helpful debugging information to a novice.
-
-Recent work on type error \emph{localization}
-%
-\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr},
-%
-determining which expression(s) to blame for the error,
-frames localization as an optimization problem, seeking a set of
-expressions to blame that minimize some cost function (\eg the number of
-constraints in the blame set, or the ratio of ``good'' paths through the
-constraint graph to ``bad'' paths).
-%
-However, even these techniques fail to distinguish between the two
-underlined expressions in \autoref{fig:sumList}, it would be equally
-valid to blame \emph{either} of them alone.
-%
-The |[]| on line 3 could be changed to |0|, or the |+| on line 4 could
-be changed to either |@| (list append) or |::|, all of which would give
-type-correct programs.
-%
-Thus, these state-of-the-art techniques are forced to either blame
-\emph{both} locations, or choose one \emph{arbitrarily}.
-
-It may appear obvious to the reader that |[]| is the correct expression
-to blame, but how is the type checker to know that?
-%
-Our solution is to \emph{learn} how to localize type errors by observing
-a large set of (novice) errors and their \emph{subsequent fixes}.
-%
-We accomplish this via three concrete contributions.
-
-\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
-Our first contribution is a large dataset of novice \ocaml programs,
-including, but not limited to, \emph{ill-typed} programs.
-%
-We used an instrumented version of the \ocaml compiler to collect
-student interactions over two instances of an undergraduate Programming
-Languages course at
-%
-\begin{anonsuppress}
-UC San Diego (IRB \#140608).
-\end{anonsuppress}
-\begin{noanonsuppress}
-AUTHOR's INSTITUTION (IRB HIDDEN).
-\end{noanonsuppress}
-%
-The data consist of a time-series of programs submitted to the \ocaml
-compiler, from which we extracted a set of \emph{ill-typed} programs
-and their \emph{subsequent fixes}.
-%
-Our dataset comprises over 4,500 such pairs of programs, and many more
-programs that are either well-typed or lack a corresponding fix.
-%
-The entire dataset (both the raw time-series and our extracted pairs of
-ill-typed programs and fixes) is available at
-%
-\begin{anonsuppress}
-\ES{TODO: finally post the data somehwere..}.
-\end{anonsuppress}
-\begin{noanonsuppress}
-HIDDEN FOR DOUBLE-BLIND REVIEW.
-\end{noanonsuppress}
-%
-
-\paragraph{\textbf{(2) Data-Driven Diagnosis of Type Errors}}
-Our second contribution is a \emph{data-driven} technique for localizing
-type errors (\autoref{sec:learning}).
-%
-We use machine learning techniques to train a set of classifiers that can
-predict, given an ill-typed program, which expressions should be changed
-for the program to type check.
-%
-Our classifiers can be trained on a relatively small set of programs
-(\ie a single course's worth). 
-% WRW notes that we say this below in contribution (3), which is only three
-% sentences away, so to a first-time reader it sounds like we are
-% double-counting:
-%
-% and outperform the state of the art by
-% 15--30\% in blame accuracy.
-%
-The key technical challenge that we tackle is defining a suitable
-interface between the type checker and the classifier, specifically
-\emph{embedding} programs as feature vectors and presenting the
-classifier's predictions to the user.
-
-\paragraph{\textbf{(3) A Large-Scale Evaluation of Type Error Localization}}
-Our third contribution is an evaluation of our classifiers and the
-state-of-the-art techniques on our suite of over 4,500 student programs
-(\autoref{sec:evaluation}), the largest evaluation of type error
-localization techniques we are aware of.
-%
-We judge the correctness of a localization by comparing it to the
-student's eventual fix and find that our classifiers are 15--30\% more
-accurate than the state of the art.
-%
-We also examine which features contribute the most to our
-classifiers' predictions, and endeavor to \emph{explain} specific
-predictions that our classifiers get right (and wrong).
-
-Overall, our results show that data-driven diagnosis is a practical
-technique for localizing type errors by learning from past mistakes.
-%
-% We believe, furthermore, that with the right set of features our
-% approach could be applied to other domains where there are multiple,
-% equally valid explanations for an error.
-% \ES{this sounds really dumb... need better phrasing}
 

### 2017-04-13 paper/oopsla17-submission/learning.tex
more kamalika notes

@@
 \label{sec:labels}
-We define the output of the model to be a boolean label, where ``false''
-means the expression \emph{should not} change and ``true'' means the
-expression \emph{should} change.
+Recall that we make predictions in two stages.
+%
+First, we use $\evalsym$ to predict for each subexpression whether it
+should be blamed, and extract a confidence score $\mathcal{P}$ from the
+$\Model$.
+%
+Thus, we define the output of the $\Model$ to be a boolean label, where
+``false'' means the expression \emph{should not} change and ``true''
+means the expression \emph{should} change.
 %
@@
 %
-Many learning algorithms produce not only a prediction, but also a
-metric that can be interpreted as a \emph{confidence} in the prediction.
-%
-Thus, we \emph{rank} the expressions by the model's confidence that they
-will change, and select only the top $k$ (in practice $k=3$) to present
-to the user.
+Second, we \emph{rank} each subexpression by the confidence $\mathcal{P}$
+that it should be blamed, and return to the user the top $k$
+most likely blame assignments (in practice $k=3$).
+
+% Many learning algorithms produce not only a prediction, but also a
+% metric that can be interpreted as a \emph{confidence} in the prediction.
+% %
+% Thus, we \emph{rank} the expressions by the model's confidence that they
+% will change, and select only the top $k$ (in practice $k=3$) to present
+% to the user.
 

### 2017-04-13 paper/oopsla17-submission/abstract.tex
asd

@@
 \begin{abstract}
-  Localizing type errors is challenging in languages with global type
-  inference, as the type checker must make assumptions about what the
-  programmer intended to do.
+Localizing type errors is challenging in
+languages with global type inference, as
+the type checker must make assumptions
+about what the programmer intended to do.
 %
-
-  We present a data-driven approach to type error localization that uses
-  machine learning to assign blame based on previous
-  ill-typed programs and their fixes.
+We introduce \toolname, a \emph{data-driven}
+approach to error localization based on
+supervised learning.
 %
-  First, we gather a large set of over 4,500 ill-typed student programs
-  and their subsequent fixes.
+\toolname works by using a large corpus
+of training data --- pairs of ill-typed
+programs and their ``fixed'' versions ---
+to automatically \emph{learn a model}
+of where the error is most likely.
 %
-  Second, we frame type error localization as a supervised learning
-  problem: to predict, for each expression in the
-  program, whether it should be blamed.
+When given a new ill-typed program,
+\toolname executes the model to
+generate a list of likely error
+locations ranked by likelihood.
 %
-  We represent each expression by its syntactic class and type and those
-  of its parent and children, associating blame with expressions that
-  were changed in the human-fixed program.
+We evaluate \toolname by comparing its
+precision against the state-of-the-art
+on a set of over 4,500 ill-typed Ocaml
+programs drawn from introductory
+programming classes.
 %
-  Finally, we train and evaluate a series of off-the-shelf classifiers
-  on our suite of programs,
-  and find that they improve localization accuracy from a state-of-the-art
-  56\% to 72\%.
+We show that when the top-ranked location
+is considered, \toolname's data-driven
+model is able to correctly predict
+the exact sub-expression that needs to
+be changed, 74\% of the time, which is
+29 points higher than \ocaml, and
+18 points higher than the state-of-the-art
+\sherrloc tool.
+%
+Further, \toolname's accuracy surpasses
+85\% when we consider the \emph{top-2}
+locations and reaches 91\% if consider
+the \emph{top-3}.
 \end{abstract}

### 2017-04-13 paper/oopsla17-submission/conclusion.tex
asd

--- 2017-04-10 paper/oopsla17-submission/conclusion.tex
@@
 \label{sec:conclusion}
+
+We have presented \toolname which
+combining modern statistical methods
+with domain-specific feature engineering
+to open the door to a new data-driven
+path to precise error localization.
+%
+\toolname
+translates pairs of
+blame-labeled ill-typed
+programs into bags-of-abstracted-terms,
+and then trains a classifier that
+predicts the value of the
+whether a feature vector gets
+a blame label.
+%
+Given a new ill-typed program,
+we extract its bag-of-extracted-terms
+to compute the blame label for each
+term and rank the predictions based
+on the model's confidence, to select
+the top 1-3 blame assignments to
+present to the user.
+
+We have evaluated our approach with
+five different classifiers on a suite
+of over 4,500 student programs.
+We find that while machine
+learning on its own performs slightly worse
+than existing purely constraint based
+approaches (\eg \ocaml, \sherrloc), once
+we augment the data with feature abstraction
+corresponding to the \emph{type error slice}~\cite{Tip2001-qp},
+the resulting models models outperform
+the state of the art by up to 18
+percentage points.
+
+As with other forms of machine learning,
+a key concern is that of \emph{data-set bias}: are
+the models learned by \toolname specific
+to our data set, and would they fail on
+\emph{other} \ocaml programs?
+%
+We address this concern in two ways.
+%
+First, we partition the data by year,
+and show that models learned from one
+year generalize to, \ie perform nearly
+as well on the programs from the other
+year.
+%
+Second, we argue that in our setting
+this bias is \emph{feature} (and not
+a bug): it allows \toolname to \emph{adapt}
+to the kinds of errors that programmers
+(specifically, novices who are in greatest
+need of precise feedback) actually make
+rather than hardwiring the biases of
+compiler authors who, by dint of their
+training and experience, may suffer from
+blind spots with regards to such problems.
+%
+In this regard, we are particularly pleased
+that our classifiers can be trained on a
+modest amount of data, \eg a single course's
+worth, and envision a future where each course
+has a model of its students' errors.
 

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

@@
 
-In Robin Milner's memorable words
+% In Robin Milner's memorable words
+% %
+% ``types are the leaven of computer programming;
+  % they make it digestible''~\cite{tapl}.
 %
-``types are the leaven of computer programming;
-  they make it digestible.''
+Types are wonderful.
 %
@@
 the value-proposition for types even more
-appealing by using constraints to (mostly)
-automatically synthesize the types for all
-program terms without troubling the
-programmer for any annotations.
-
+appealing by using constraints to automatically
+synthesize the types for all program terms
+without troubling the programmer for any
+annotations.
+%
 Unfortunately, this automation has come at a price.
-Verbose type annotations (as in, \eg Java) signify
-the programmer's \emph{intent} and help to correctly
-pinpoint the error when the code is ill-typed.
+Verbose type annotations (as in, \eg, Java) signify
+the programmer's intent and help to correctly
+blame the erroneous sub-term when the code is
+ill-typed.
 %
@@
 %
-First, they are not \emph{portable}:
+First, they are not \emph{extensible}:
+the weights and analyses are limited
+to attributes of the constraints, and
+not other \emph{signals} from the
+program's syntax that may be more
+germane to error localization. Furthermore,
 the algorithms for computing
@@
 %
-Second, they are not \emph{extensible}:
-the weights and analyses are limited
-to attributes of the constraints, and
-not other \emph{signals} from the
-program's syntax that may be more
-germane to error localization.
-%
-Third, and perhaps most importantly,
+Second, and perhaps most importantly,
 they are not \emph{adaptable}: the
-actual ``weights'' are fixed in
-an ad-hoc fashion, based on the
+weights are fixed in an ad-hoc fashion, based on the
 \emph{analysis designers} notion
@@
 %
-In this paper, we introduce \toolname
+In this paper, we introduce \toolname \footnote{Numeric Analysis of Type Errors}
 a \emph{data-driven} approach to error
-localization.
+localization based on supervised
+classification~\citep{Kotsiantis2007-pj}.
 %
@@
 the \emph{top-2} and \emph{top-3} locations.
+%
+We achieve these advances by identifying
+and then solving three key challenges.
 
-% - 45
-% - 56
-% - 74
+\mypara{Challenge 1: Acquiring Labeled Programs}
 %
-% - 85
-% - 91
+The first challenge for supervised learning
+is to acquire a corpus of training data, here,
+a set of ill-typed programs \emph{labeled}
+with the exact sub-terms that are the actual
+cause of the type error.
+%
+Prior work has often enlisted expert users
+to manually judge ill-typed programs and
+determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk},
+but this approach does not scale well to
+a dataset large enough to support machine
+learning.
+%
+Worse, while expert users have intimate
+knowledge of the type system, they may
+have a blind spot with regards to the
+kinds of mistakes novices make, and
+cannot know in general what novice users
+intended to do.
 
-AST-Dispersion
-
-AST
-
-\toolname addresses both of the limitations
-of previous
-
-* portable   [use different features]
-* extensible [add more features]
-* adaptable  [add more data]
-
+Our \emph{first contribution} \autoref{sec:overview}
+is a set of more than 4,500 labeled programs,
+gives us an accurate ground truth of
+the kinds of errors and the (locations
+of the) fixes that novices make in
+practice.
 %
-Our solution is to \emph{learn} how to
-localize type errors by observing
-a large set of (novice) errors and their
-\emph{subsequent fixes}.
+We obtain this set by observing that
+software development is an iterative
+process and programmers eventually
+fix their own ill-typed programs,
+perhaps after multiple incorrect
+exploratory attempts.
 %
-
-\mypara{Contributions}
-%
-We accomplish this via three concrete contributions.
-
-\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
-Our first contribution is a large dataset of novice \ocaml programs,
-including, but not limited to, \emph{ill-typed} programs.
-%
-We used an instrumented version of the \ocaml compiler
-to collect student interactions over two instances of
-an undergraduate Programming Languages course at
+To exploit this observation we instrumented
+the \ocaml compiler to collect fine-grained
+traces of student interactions over two instances
+of an undergraduate Programming Languages course at
 %
@@
 %
-The data consist of a time-series of programs submitted to the \ocaml
-compiler, from which we extracted a set of \emph{ill-typed} programs
-and their \emph{subsequent fixes}.
+We can then post-process the resulting time-series
+of programs submitted to the \ocaml compiler into
+a set of pairs of ill-typed programs and their
+\emph{fixed} versions, the \emph{first} (type-)correct
+program in the trace suffix.
 %
-Our dataset comprises over 4,500 such pairs of programs, and many more
-programs that are either well-typed or lack a corresponding fix.
+We obtain the location labels by computing a
+\emph{tree-diff} between the two terms to find
+the exact sub-terms that were changed in the fix.
+
+\mypara{Challenge 2: Modeling Programs as Vectors}
 %
-The entire dataset (both the raw time-series and our extracted pairs of
-ill-typed programs and fixes) is available at
+Modern supervised learning algorithms work on sets of
+\emph{feature vectors}: real-valued points in an
+$n$-dimensional space. While it is easy to compute such
+vectors for documents, images and sound (respectively
+word-counts, pixel-values and frequencies), it has
+hitherto been unclear how to precisely model program
+structure as numeric data.
+
+Our \emph{second contribution} (\autoref{sec:learning})
+solves this problem with a new representation called
+a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
 %
-\begin{anonsuppress}
-\ES{TODO: finally post the data somehwere..}.
-\end{anonsuppress}
-\begin{noanonsuppress}
-HIDDEN FOR DOUBLE-BLIND REVIEW.
-\end{noanonsuppress}
+(1)~each program is represented by the \emph{bag}
+    or multiset of (sub-)terms that appears inside
+    it, and
 %
+(2)~each (sub-)term $t$ is \emph{abstracted} as
+    a vector $[f_1(t), \ldots, f_n(t)]$ where
+    each $f_i$ is a \emph{feature abstraction}
+    that maps a term to a numeric value.
+%
+We can model the error labels with a simple feature
+abstraction (\eg is-a-diff-subterm).
+%
+We can even account for \emph{contextual} information like
+the features of the parent- and children- terms by
+concatenating the feature vectors of each term with those
+of its parent and children.
 
-\paragraph{\textbf{(2) Data-Driven Diagnosis of Type Errors}}
-Our second contribution is a \emph{data-driven} technique for localizing
-type errors (\autoref{sec:learning}).
+\mypara{Challenge 3: Training Precise Classifiers}
 %
-We use machine learning techniques to train a set of classifiers that can
-predict, given an ill-typed program, which expressions should be changed
-for the program to type check.
+Finally, the last and most important challenge is to
+use the labeled BOAT data to train classifiers that
+are capable of \emph{precisely} pinpointing the error
+locations in real programs.
 %
-Our classifiers can be trained on a relatively small set of programs
-(\ie a single course's worth).
+The key here is to find the right set of feature
+abstractions and classification algorithms, that
+lead to precise predictions.
 %
-% and outperform the state of the art by
-% 15--30\% in blame accuracy.
+Fortunately, our BOAT model allows us a great deal of
+latitude in our choice of features.
 %
-The key technical challenge that we tackle is defining a suitable
-interface between the type checker and the classifier, specifically
-\emph{embedding} programs as feature vectors and presenting the
-classifier's predictions to the user.
+We can use abstraction functions to capture different
+aspects of a term ranging from
+%
+syntactic features (\eg is-a-data constructor, is-a-literal,
+is-an-arithmetic-operation, is-a-function-application \etc),
+%
+to semantic features captured by the type system (\eg is-a-list,
+is-an-integer, is-a-function \etc).
 
-\paragraph{\textbf{(3) A Large-Scale Evaluation of Type Error Localization}}
-Our third contribution is an evaluation of our classifiers and the
-state-of-the-art techniques on our suite of over 4,500 student programs
-(\autoref{sec:evaluation}), the largest evaluation of type error
-localization techniques we are aware of.
+Our \emph{third contribution} \autoref{sec:evaluation}
+is a systematic evaluation of our data-driven approach
+using different classes of features like the above, and
+with three different classification algorithms: linear
+classifiers, decision forests and deep neural networks.
 %
-We judge the correctness of a localization by comparing it to the
-student's eventual fix and find that our classifiers are 15--30\% more
-accurate than the state of the art.
+Our evaluation lets us identify the best features and
+classifier and empirically characterize the importance
+of different features for error localization.
 %
-We also examine which features contribute the most to our
-classifiers' predictions, and endeavor to \emph{explain} specific
-predictions that our classifiers get right (and wrong).
+In particular, we find that while machine learning
+on its own performs slightly worse than existing
+purely constraint based approaches (\eg \ocaml, \sherrloc),
+once we augment the data with feature abstraction
+corresponding to the \emph{type error slice} \cite{Tip2001-qp},
+the resulting classifier significantly outperforms
+state-of-the-art by 18 percentage points.
 
-Overall, our results show that data-driven diagnosis is a practical
-technique for localizing type errors by learning from past mistakes.
-
+Thus, by combining modern statistical methods
+with domain-specific feature engineering
+to open the door to a new data-driven
+path to precise error localization.
 %
-% We believe, furthermore, that with the right set of features our
-% approach could be applied to other domains where there are multiple,
-% equally valid explanations for an error.
-% \ES{this sounds really dumb... need better phrasing}
-
-Consider the |sumList| program in \autoref{fig:sumList},
-written by a student in an undergraduate Programming Languages
-course.
+We can easily \emph{extend} \toolname to
+new languages or forms of correctness
+checks by swapping in a different
+set of feature abstraction functions.
 %
-The program is meant to sum the integers in a list, but
-the student has accidentally given the empty list as the
-base case, rather than 0.
+Furthermore, our data-driven approach
+allows \toolname to \emph{adapt} to
+the kinds of errors that programmers
+(specifically, novices who are in greatest
+need of precise feedback) actually make
+rather than hardwiring the biases of
+compiler authors who, by dint of their
+training and experience, may suffer from
+blind spots with regards to such problems.
 %
-The \ocaml compiler collects typing constraints as it
-traverses the program, and reports an error the moment
-it finds an inconsistent constraint.
-%
-In this case it blames the recursive call to |sumList|,
-complaining that |sumList| returns a |list| where an |int|
-was expected by the |+| operator.
-%
-Note that this \emph{blame} is inconsistent with the
-programmer's intention and may not be helpful debugging
-information to a novice.
-
-It may appear obvious to the reader that
-|[]| is the correct expression to blame,
-but how is the type checker to know that?
-
-Indeed, both the venerable
-Ocaml compiler as well as
-recent techniques~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
-fail to distinguish between
-the two underlined expressions
-in \autoref{fig:sumList};
-it would be equally valid
-to blame \emph{either}
-of them alone.
-%
-The |[]| on line 3 could be changed to |0|,
-or the |+| on line 4 could be changed to
-either |@| (list append) or |::|, all of
-which would give type-correct programs.
-%
-Thus, these state-of-the-art techniques
-are forced to either blame \emph{both}
-locations, or choose one \emph{arbitrarily}.
-
-\begin{figure}[ht]
-\begin{minipage}{0.45\linewidth}
-\begin{ecode}
-  let rec sumList xs =
-    match xs with
-    | []   -> (*@\hlTree{\hlSherrloc{[]}}@*)
-    | h::t -> h (*@\hlTree{+}@*) (*@\hlFix{sumList t}@*)
-\end{ecode}
-\end{minipage}
-\begin{minipage}{0.49\linewidth}
-\begin{verbatim}
-File "sumList.ml", line 4, characters 16-25:
-  This expression has type 'a list
-  but an expression was expected of type int
-\end{verbatim}
-\end{minipage}
-\caption{(left) An ill-typed \ocaml program that should sum the elements of a
-  list, with highlights indicating three possible blame assignments based on:
-  %
-  (1) the \hlFix{\ocaml} compiler;
-  %
-  (2) the \hlSherrloc{fix} made by the programmer; and
-  %
-  (3) \hlTree{minimizing} the number of edits required.
-  % The \underline{underlined} expressions are equally valid
-  % locations to blame. The expression blamed by the \ocaml compiler
-  % is in \textbf{bold}.
-  %
-  % FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
-  % ``with'' are also bolded (by \\ecode)! You need to find another way to
-  % highlight what ocaml is yelling about.
-  %
-  (right) The error reported by \ocaml.}
-\label{fig:sumList}
-\end{figure}
-
-% \begin{enumerate}
-% \item An large-scale evaluation of state-of-the-art type error
-%   localization techniques on over 4,500 student programs.
-%   %
-%   We collected a time-series of \ocaml programs from our students, and
-%   use their subsequent \emph{fixes} to ill-typed programs as
-%   \emph{oracles} for the location of the error.
-%   %
-%   This is the most extensive evaluation of the accuracy of type error
-%   localization techniques that we are aware of.
-%   % based not on expert \emph{opinion}, but
-%   % on the eventual \emph{fix} implemented by a novice user.
-%   %
-%   % \ES{need to explain \emph{why} this is superior..}
-% \item A machine learning approach to \emph{modeling} (novice) type
-%   errors that can outperform the state-of-the-art by 15--30\% in
-%   localizing errors.
-%   %
-%   Our classifiers can be trained on a relatively small amount of data --- we
-%   used a single class of around 50 students --- and appear to
-%   generalize to other instances of the same class.
-%   %
-%   \ES{and hopefully completely separate classes!}
-%   %
-%   In contrast to many approaches to \emph{fault localization} from the
-%   software engineering community, our model does not use any linguistic
-%   modeling techniques, \eg \emph{n-grams} over the token stream; rather,
-%   it relies entirely on features of the abstract syntax tree and a
-%   partial typing derivation.
-%   %
-%   \ES{if there's time, investigate adding Wes' n-gram model}
-% \end{enumerate}
+In contrast, our results show that \toolname's
+data-driven diagnosis can be an effective
+technique for localizing errors by collectively
+learning from past mistakes.
 

### 2017-04-13 paper/oopsla17-submission/scratch.tex
asd

--- 2017-04-12 paper/oopsla17-submission/introduction.tex
@@
-\section{Introduction}
-\label{sec:introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Type systems are one of the most widely-used techniques for program
-verification. A type checker only accepts a program if it can
-\emph{prove} the absence of an entire class of bugs.
+Consider the |sumList| program in \autoref{fig:sumList},
+written by a student in an undergraduate Programming Languages
+course.
 %
-Typed functional languages like \ocaml and \haskell are equipped with
-particularly powerful type checkers that can operate without a single
-annotation, alleviating a prime concern with verification techniques.
+The program is meant to sum the integers in a list, but
+the student has accidentally given the empty list as the
+base case, rather than 0.
 %
-Yet this power comes with a cost: type annotations signify the
-programmer's \emph{intent}, without them the checker must
-make assumptions about what the programmer intended to do.
+The \ocaml compiler collects typing constraints as it
+traverses the program, and reports an error the moment
+it finds an inconsistent constraint.
 %
-This results in the common phenomenon of type errors being reported far
-from their source, an annoyance to veterans of the type system, but a
-significant hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.
+In this case it blames the recursive call to |sumList|,
+complaining that |sumList| returns a |list| where an |int|
+was expected by the |+| operator.
+%
+Note that this \emph{blame} is inconsistent with the
+programmer's intention and may not be helpful debugging
+information to a novice.
+
+It may appear obvious to the reader that
+|[]| is the correct expression to blame,
+but how is the type checker to know that?
+
+Indeed, both the venerable
+Ocaml compiler as well as
+recent techniques~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
+fail to distinguish between
+the two underlined expressions
+in \autoref{fig:sumList};
+it would be equally valid
+to blame \emph{either}
+of them alone.
+%
+The |[]| on line 3 could be changed to |0|,
+or the |+| on line 4 could be changed to
+either |@| (list append) or |::|, all of
+which would give type-correct programs.
+%
+Thus, these state-of-the-art techniques
+are forced to either blame \emph{both}
+locations, or choose one \emph{arbitrarily}.
 
@@
 \end{figure}
-%
-Consider the |sumList| program in \autoref{fig:sumList}, written by a
-student in an undergraduate Programming Languages course.
-%
-The program is meant to sum the integers in a list, but the student has
-accidentally given the empty list as the base case, rather than 0.
-%
-The \ocaml compiler collects typing constraints as it traverses the
-program, and reports an error the moment it finds an inconsistent
-constraint.
-%
-In this case it blames the recursive call to |sumList|, complaining that
-|sumList| returns a |list| where an |int| was expected by the |+|
-operator.
-%
-Note that this blame is inconsistent with the programmer's intention and
-may not be helpful debugging information to a novice.
-
-Recent work on type error \emph{localization}
-%
-\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr},
-%
-determining which expression(s) to blame for the error,
-frames localization as an optimization problem, seeking a set of
-expressions to blame that minimize some cost function (\eg the number of
-constraints in the blame set, or the ratio of ``good'' paths through the
-constraint graph to ``bad'' paths).
-%
-However, even these techniques fail to distinguish between the two
-underlined expressions in \autoref{fig:sumList}, it would be equally
-valid to blame \emph{either} of them alone.
-%
-The |[]| on line 3 could be changed to |0|, or the |+| on line 4 could
-be changed to either |@| (list append) or |::|, all of which would give
-type-correct programs.
-%
-Thus, these state-of-the-art techniques are forced to either blame
-\emph{both} locations, or choose one \emph{arbitrarily}.
-
-It may appear obvious to the reader that |[]| is the correct expression
-to blame, but how is the type checker to know that?
-%
-Our solution is to \emph{learn} how to localize type errors by observing
-a large set of (novice) errors and their \emph{subsequent fixes}.
-%
-We accomplish this via three concrete contributions.
-
-\paragraph{\textbf{(1) A Dataset of Novice \ocaml Programs}}
-Our first contribution is a large dataset of novice \ocaml programs,
-including, but not limited to, \emph{ill-typed} programs.
-%
-We used an instrumented version of the \ocaml compiler to collect
-student interactions over two instances of an undergraduate Programming
-Languages course at
-%
-\begin{anonsuppress}
-UC San Diego (IRB \#140608).
-\end{anonsuppress}
-\begin{noanonsuppress}
-AUTHOR's INSTITUTION (IRB HIDDEN).
-\end{noanonsuppress}
-%
-The data consist of a time-series of programs submitted to the \ocaml
-compiler, from which we extracted a set of \emph{ill-typed} programs
-and their \emph{subsequent fixes}.
-%
-Our dataset comprises over 4,500 such pairs of programs, and many more
-programs that are either well-typed or lack a corresponding fix.
-%
-The entire dataset (both the raw time-series and our extracted pairs of
-ill-typed programs and fixes) is available at
-%
-\begin{anonsuppress}
-\ES{TODO: finally post the data somehwere..}.
-\end{anonsuppress}
-\begin{noanonsuppress}
-HIDDEN FOR DOUBLE-BLIND REVIEW.
-\end{noanonsuppress}
-%
-
-\paragraph{\textbf{(2) Data-Driven Diagnosis of Type Errors}}
-Our second contribution is a \emph{data-driven} technique for localizing
-type errors (\autoref{sec:learning}).
-%
-We use machine learning techniques to train a set of classifiers that can
-predict, given an ill-typed program, which expressions should be changed
-for the program to type check.
-%
-Our classifiers can be trained on a relatively small set of programs
-(\ie a single course's worth). 
-% WRW notes that we say this below in contribution (3), which is only three
-% sentences away, so to a first-time reader it sounds like we are
-% double-counting:
-%
-% and outperform the state of the art by
-% 15--30\% in blame accuracy.
-%
-The key technical challenge that we tackle is defining a suitable
-interface between the type checker and the classifier, specifically
-\emph{embedding} programs as feature vectors and presenting the
-classifier's predictions to the user.
-
-\paragraph{\textbf{(3) A Large-Scale Evaluation of Type Error Localization}}
-Our third contribution is an evaluation of our classifiers and the
-state-of-the-art techniques on our suite of over 4,500 student programs
-(\autoref{sec:evaluation}), the largest evaluation of type error
-localization techniques we are aware of.
-%
-We judge the correctness of a localization by comparing it to the
-student's eventual fix and find that our classifiers are 15--30\% more
-accurate than the state of the art.
-%
-We also examine which features contribute the most to our
-classifiers' predictions, and endeavor to \emph{explain} specific
-predictions that our classifiers get right (and wrong).
-
-Overall, our results show that data-driven diagnosis is a practical
-technique for localizing type errors by learning from past mistakes.
-%
-% We believe, furthermore, that with the right set of features our
-% approach could be applied to other domains where there are multiple,
-% equally valid explanations for an error.
-% \ES{this sounds really dumb... need better phrasing}
-
-% \begin{enumerate}
-% \item An large-scale evaluation of state-of-the-art type error
-%   localization techniques on over 4,500 student programs.
-%   %
-%   We collected a time-series of \ocaml programs from our students, and
-%   use their subsequent \emph{fixes} to ill-typed programs as
-%   \emph{oracles} for the location of the error.
-%   %
-%   This is the most extensive evaluation of the accuracy of type error
-%   localization techniques that we are aware of.
-%   % based not on expert \emph{opinion}, but
-%   % on the eventual \emph{fix} implemented by a novice user.
-%   %
-%   % \ES{need to explain \emph{why} this is superior..}
-% \item A machine learning approach to \emph{modeling} (novice) type
-%   errors that can outperform the state-of-the-art by 15--30\% in
-%   localizing errors.
-%   %
-%   Our classifiers can be trained on a relatively small amount of data --- we
-%   used a single class of around 50 students --- and appear to
-%   generalize to other instances of the same class.
-%   %
-%   \ES{and hopefully completely separate classes!}
-%   %
-%   In contrast to many approaches to \emph{fault localization} from the
-%   software engineering community, our model does not use any linguistic
-%   modeling techniques, \eg \emph{n-grams} over the token stream; rather,
-%   it relies entirely on features of the abstract syntax tree and a
-%   partial typing derivation.
-%   %
-%   \ES{if there's time, investigate adding Wes' n-gram model}
-% \end{enumerate}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "main"
-%%% End:

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

@@
 %
-Types are wonderful.
+Types are awesome.
 %
@@
 %
-In this paper, we introduce \toolname \footnote{Numeric Analysis of Type Errors}
+In this paper, we introduce \toolname
+\footnote{``Numeric Analysis of Type Errors''; any resemblance to persons living or dead is purely coincidental.}
 a \emph{data-driven} approach to error
@@
 of training data --- pairs of ill-typed
-programs and their ``fixed'' versions ---
+programs and their fixed versions ---
 to automatically \emph{learn a model}
@@
 the exact sub-expression that needs to
-be changed, 74\% of the time compared
-to \ocaml's 45\% and \sherrloc's 56\%.
+be changed, 74\% of the time, which is
+29 points higher than \ocaml, and
+18 points higher than the state-of-the-art
+\sherrloc tool.
 %
-When the \emph{top} ranked location
-is considered. \toolname's accuracy jumps
-to above 85\% (resp. 90\%) when we consider
-the \emph{top-2} and \emph{top-3} locations.
+Further, \toolname's accuracy surpasses
+85\% when we consider the \emph{top-2}
+locations and reaches 91\% if consider
+the \emph{top-3}.
 %

### 2017-04-13 paper/oopsla17-submission/evaluation.tex
more notes from kamalika

@@
   %
-  How often are our predictions \emph{correct}?
+  How often do we \emph{correctly} assign blame for the type error?
+  %
+  % We compare our technique with a variety of off-the-shelf classifiers
+  % and find that our top-ranked blame assignments have an accuracy of
+  % 72\%, compared to a state-of-the-art 56\%.
   % For how many ill-typed programs can we accurately predict the source
@@
 %
-We mitigate this threat by having separate training and testing datasets
-(\autoref{sec:quantitative}) and via cross-validation
-(\autoref{sec:feature-utility})~\cite{FIXME}.
+A similar issue that arises in machine learning is model stability (\ie
+can small changes to the training set produce large changes in the model?).
+%
+We mitigate these threats by:
+%
+(1) using separate training and testing datasets drawn from distinct
+student populations (\autoref{sec:quantitative}), demonstrating the
+generality of our models; and
+%
+(2) via cross-validation on the joint dataset
+(\autoref{sec:feature-utility}), which demonstrates the stability of our
+models by averaging the accuracy of 10 models trained on distinct
+subsets of the data.
 
@@
 Examining programs written in \ocaml as opposed to \haskell or any other
-language poses another threat to validity, common type errors may differ
-in different languages.
+typed functional language poses yet another threat, common type errors
+may differ in different languages.
 %

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

@@
 cannot know in general what novice users
-intended to do.
+intended.
 
-Our \emph{first contribution} \autoref{sec:overview}
+Our \emph{first contribution} (\autoref{sec:overview})
 is a set of more than 4,500 labeled programs,
@@
 
-Our \emph{third contribution} \autoref{sec:evaluation}
+Our \emph{third contribution} (\autoref{sec:evaluation})
 is a systematic evaluation of our data-driven approach

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

@@
 sub-terms that yielded those constraints.
-
+%
 This minimum-weight approach suffers
-from several drawbacks.
+from two drawbacks.
 %
@@
 or less likely, rather than
-capturing and adapting to the
-ground truth about the kinds
-of errors programmers actually
-make in practice.
+adapting to the kinds of mistakes
+programmers actually make in practice.
 
@@
 traces of student interactions over two instances
-of an undergraduate Programming Languages course at
+of an undergraduate Programming Languages course
 %
-\begin{anonsuppress}
-UC San Diego (IRB \#140608).
-\end{anonsuppress}
-\begin{noanonsuppress}
-AUTHOR's INSTITUTION (IRB HIDDEN).
-\end{noanonsuppress}
+\footnote{
+% \begin{anonsuppress}
+% UC San Diego (IRB \#140608).
+% \end{anonsuppress}
+% \begin{noanonsuppress}
+at AUTHOR's INSTITUTION (IRB HIDDEN).
+% \end{noanonsuppress}
+}
 %
@@
 Thus, by combining modern statistical methods
-with domain-specific feature engineering
-to open the door to a new data-driven
-path to precise error localization.
+with domain-specific feature engineering, \toolname
+opens the door to a new data-driven path to
+precise error localization.
 %
-We can easily \emph{extend} \toolname to
-new languages or forms of correctness
-checks by swapping in a different
-set of feature abstraction functions.
+In the future, we could \emph{extend}
+\toolname to new languages or forms
+of correctness checks by swapping in
+a different set of feature abstraction
+functions.
 %

### 2017-04-13 paper/oopsla17-submission/intro2.tex
minor fixes

--- 2017-04-13 paper/oopsla17-submission/intro2.tex
@@
 Unfortunately, this automation has come at a price.
-Verbose type annotations (as in, \eg, Java) signify
+Verbose type annotations (as in, \eg, \java) signify
 the programmer's intent and help to correctly
@@
 %
-In contrast, lacking such signifiers, automatic
+In the absence of such signifiers, automatic
 type inference algorithms are prone to report
-type errors quite far from their actual source.
+type errors far from their actual source
+\citep{Wand1986-nw}.
 %
 While this can seem like a minor annoyance to
-veteran programmers, it turns out to be a significant
-stumbling block for novices~\citep{Wand1986-nw,Joosten1993-yx}.
+veteran programmers, \citet{Joosten1993-yx} have found
+that novices often focus their attention on the \emph{location}
+reported and disregard the \emph{message}.
 
@@
 satisfiable and hence, memory-safe~\citep{Jose:2011}
-or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-vm,Pavlinovic2014-mr}.
+or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-gd,Pavlinovic2014-mr}.
 The finger of blame is then pointed at the
@@
 First, they are not \emph{extensible}:
-the weights and analyses are limited
-to attributes of the constraints, and
-not other \emph{signals} from the
-program's syntax that may be more
-germane to error localization. Furthermore,
-the algorithms for computing
+% the weights and analyses are limited
+% to attributes of the constraints, and
+% not other \emph{signals} from the
+% program's syntax that may be more
+% germane to error localization. Furthermore,
+the constraint languages and algorithms for computing
 the minimum weighted subset must be
 designed afresh for different kinds
-of type systems and constraints.
+of type systems and constraints \citep{Loncaric2016-uk}.
 %
@@
 weights are fixed in an ad-hoc fashion, based on the
-\emph{analysis designers} notion
+\emph{analysis designer's} notion
 of what kinds of errors are more
@@
 localization based on supervised
-classification~\citep{Kotsiantis2007-pj}.
+learning (see \citealt{Kotsiantis2007-pj} for a survey).
 %
-\toolname works by using a large corpus
+\toolname works by analyzing a large corpus
 of training data --- pairs of ill-typed
-programs and their fixed versions ---
+programs and their subsequent fixes ---
 to automatically \emph{learn a model}
-of where the error is most likely.
+of where errors are most likely to
+be found.
 %
-When given a new ill-typed program,
+Given a new ill-typed program,
 \toolname simply executes the model
-to generate a list of likely error
+to generate a list of potential error
 locations ranked by likelihood.
@@
 precision against the state-of-the-art
-on a set of over 4,500 ill-typed Ocaml
+on a set of over 4,500 ill-typed \ocaml
 programs drawn from introductory
@@
 the exact sub-expression that needs to
-be changed, 74\% of the time, which is
-29 points higher than \ocaml, and
+be changed 74\% of the time,
+29 points higher than \ocaml and
 18 points higher than the state-of-the-art
@@
 Further, \toolname's accuracy surpasses
-85\% when we consider the \emph{top-2}
+85\% when we consider the \emph{top two}
 locations and reaches 91\% if consider
-the \emph{top-3}.
+the \emph{top three}.
 %
@@
 The first challenge for supervised learning
-is to acquire a corpus of training data, here,
+is to acquire a corpus of training data, in our setting
 a set of ill-typed programs \emph{labeled}
@@
 to manually judge ill-typed programs and
-determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk},
+determine the ``correct'' fix
+\citep[\eg][]{Lerner2007-dt,Loncaric2016-uk},
 but this approach does not scale well to
@@
 software development is an iterative
-process and programmers eventually
+process; programmers eventually
 fix their own ill-typed programs,
@@
 %
-We can then post-process the resulting time-series
+We then post-process the resulting time-series
 of programs submitted to the \ocaml compiler into
 a set of pairs of ill-typed programs and their
-\emph{fixed} versions, the \emph{first} (type-)correct
+subsequent \emph{fixes}, the first (type-)correct
 program in the trace suffix.
 %
-We obtain the location labels by computing a
+Finally, we label the changed locations by computing a
 \emph{tree-diff} between the two terms to find
@@
 \emph{feature vectors}: real-valued points in an
-$n$-dimensional space. While it is easy to compute such
-vectors for documents, images and sound (respectively
-word-counts, pixel-values and frequencies), it has
+$n$-dimensional space. While there are standard
+techniques for computing such
+vectors for documents, images, and sound (respectively
+word-counts, pixel-values, and frequencies), it has
 hitherto been unclear how to precisely model program

### 2017-04-13 paper/oopsla17-submission/intro2.tex
pass over intro2

@@
 traces of student interactions over two instances
-of an undergraduate Programming Languages course
-%
-\footnote{
+of an undergraduate Programming Languages
+course.\footnote{
 % \begin{anonsuppress}
@@
 a set of pairs of ill-typed programs and their
-subsequent \emph{fixes}, the first (type-)correct
+subsequent \emph{fixes}, the first (type-)~correct
 program in the trace suffix.
@@
 %
-Modern supervised learning algorithms work on sets of
+Modern supervised learning algorithms work on %sets of
 \emph{feature vectors}: real-valued points in an
@@
 vectors for documents, images, and sound (respectively
-word-counts, pixel-values, and frequencies), it has
-hitherto been unclear how to precisely model program
-structure as numeric data.
+word-counts, pixel-values, and frequencies),
+% it has
+% hitherto been unclear how to precisely model program
+% structure as numeric data.
+there are no similarly standard representations for
+programs.
 
 Our \emph{second contribution} (\autoref{sec:learning})
-solves this problem with a new representation called
+solves this problem with a simple, yet expressive, representation called
 a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
@@
 (1)~each program is represented by the \emph{bag}
-    or multiset of (sub-)terms that appears inside
-    it, and
+    or multiset of (sub-)~terms that appears inside
+    it; and
 %
-(2)~each (sub-)term $t$ is \emph{abstracted} as
-    a vector $[f_1(t), \ldots, f_n(t)]$ where
+(2)~each (sub-)~term $t$ is \emph{abstracted} as
+    a vector $[f_1(t), \ldots, f_n(t)]$, where
     each $f_i$ is a \emph{feature abstraction}
@@
 We can model the error labels with a simple feature
-abstraction (\eg is-a-diff-subterm).
+abstraction (\eg does-change-in-fix).
 %
 We can even account for \emph{contextual} information like
-the features of the parent- and children- terms by
+the features of the parent- and child- terms by
 concatenating the feature vectors of each term with those
 of its parent and children.
+%
+We have found this representation to be particularly
+convenient as it gives us flexibility
+in modeling the syntactic and semantic structure of programs
+while retaining compatibility with off-the-shelf classifiers,
+in contrast to, \eg, \citet{Raychev2015-jg}, who had to develop
+their own variants of classifiers to obtain their results.
 
@@
 syntactic features (\eg is-a-data constructor, is-a-literal,
-is-an-arithmetic-operation, is-a-function-application \etc),
+is-an-arithmetic-operation, is-a-function-application, \etc),
 %
 to semantic features captured by the type system (\eg is-a-list,
-is-an-integer, is-a-function \etc).
+is-an-integer, is-a-function, \etc).
 
@@
 using different classes of features like the above, and
-with three different classification algorithms: linear
-classifiers, decision forests and deep neural networks.
+with three different classification algorithms: logistic
+regression, decision trees, and neural networks.
 %
@@
 In particular, we find that while machine learning
-on its own performs slightly worse than existing
+over syntactic features of each term in isolation
+performs worse than existing
 purely constraint based approaches (\eg \ocaml, \sherrloc),
-once we augment the data with feature abstraction
-corresponding to the \emph{type error slice} \cite{Tip2001-qp},
-the resulting classifier significantly outperforms
-state-of-the-art by 18 percentage points.
+augmenting the data with a single feature corresponding to
+the \emph{type error slice} \citep{Tip2001-qp} brings our
+classifiers up to par with the state of the art,
+and further augmenting the data with \emph{contextual}
+features allows our classifiers to outperform
+the state of the art by 18 percentage points.
+% In particular, we find that while machine learning
+% on its own performs slightly worse than existing
+% purely constraint based approaches (\eg \ocaml, \sherrloc),
+% once we augment the data with feature abstraction
+% corresponding to the \emph{type error slice} \citep{Tip2001-qp},
+% the resulting classifier significantly outperforms the
+% state of the art by 18 percentage points.
 

### 2017-04-13 paper/oopsla17-submission/abstract.tex
more intro2, plus abstract and conclusion

--- 2017-04-13 paper/oopsla17-submission/abstract.tex
@@
 \begin{abstract}
-  Localizing type errors is challenging in languages with global type
-  inference, as the type checker must make assumptions about what the
-  programmer intended to do.
+Localizing type errors is challenging in
+languages with global type inference, as
+the type checker must make assumptions
+about what the programmer intended to do.
 %
-
-  We present a data-driven approach to type error localization that uses
-  machine learning to assign blame based on previous
-  ill-typed programs and their fixes.
+We introduce \toolname, a \emph{data-driven}
+approach to error localization based on
+supervised learning.
 %
-  First, we gather a large set of over 4,500 ill-typed student programs
-  and their subsequent fixes.
+\toolname works by using a large corpus
+of training data --- pairs of ill-typed
+programs and their ``fixed'' versions ---
+to automatically \emph{learn a model}
+of where the error is most likely
+to be found.
 %
-  Second, we frame type error localization as a supervised learning
-  problem: to predict, for each expression in the
-  program, whether it should be blamed.
+Given a new ill-typed program,
+\toolname executes the model to
+generate a list of potential blame
+assignments ranked by likelihood.
 %
-  We represent each expression by its syntactic class and type and those
-  of its parent and children, associating blame with expressions that
-  were changed in the human-fixed program.
+We evaluate \toolname by comparing its
+precision to the state of the art
+on a set of over 4,500 ill-typed Ocaml
+programs drawn from introductory
+programming classes.
 %
-  Finally, we train and evaluate a series of off-the-shelf classifiers
-  on our suite of programs,
-  and find that they improve localization accuracy from a state-of-the-art
-  56\% to 72\%.
+We show that when the top-ranked blame assignment
+is considered, \toolname's data-driven
+model is able to correctly predict
+the exact sub-expression that should
+be changed 74\% of the time, %which is
+29 points higher than \ocaml and
+18 points higher than the state-of-the-art
+\sherrloc tool.
+%
+Furthermore, \toolname's accuracy surpasses
+85\% when we consider the top \emph{two}
+locations and reaches 91\% if consider
+the top \emph{three}.
 \end{abstract}

### 2017-04-13 paper/oopsla17-submission/conclusion.tex
more intro2, plus abstract and conclusion

--- 2017-04-10 paper/oopsla17-submission/conclusion.tex
@@
 
-We have presented a new data-driven approach to localizing type errors,
-using machine learning to assign blame based on previously seen novice
-mistakes.
+We have presented \toolname, which
+combines modern statistical methods
+with domain-specific feature engineering
+to open the door to a new data-driven
+path to precise error localization.
 %
-We translate ill-typed programs into sets of feature vectors, predict
-for each sub-expression whether it should be blamed, rank the
-predictions based on the model's confidence, and select the top three
-blame assignments to present to the user.
+\toolname
+translates pairs of
+blame-labeled ill-typed
+programs into bags-of-abstracted-terms,
+and then trains a classifier
+to predict whether a feature vector
+should get a blame label.
+% that
+% predicts the value of the
+% whether a feature vector gets
+% a blame label.
 %
-We represent expressions solely by their syntactic class and type, and
-those of their parent and children.
+Given a new ill-typed program,
+we extract its bag-of-extracted-terms
+to compute the blame label for each
+term and rank the predictions based
+on the model's confidence, to select
+the top 1-3 blame assignments to
+present to the user.
+
+We have evaluated our approach with
+five different classifiers on a suite
+of over 4,500 student programs.
 %
-We have evaluated our approach with five different learning algorithms
-on a suite of over 4,500 student programs, and found that all of our
-models outperform the state of the art, by up to a 16\% margin.
+We found that while machine learning
+over syntactic features of each term in isolation
+performs worse than existing
+purely constraint based approaches, % (\eg \ocaml, \sherrloc),
+augmenting the data with a single feature corresponding to
+the type error slice brings our
+classifiers up to par with the state of the art,
+and further augmenting the data with
+features of an expression's parent and children
+allows our classifiers to outperform
+the state of the art by 18 percentage points.
+
+% We find that while machine
+% learning on its own performs slightly worse
+% than existing purely constraint based
+% approaches (\eg \ocaml, \sherrloc), once
+% we augment the data with feature abstraction
+% corresponding to the \emph{type error slice}~\cite{Tip2001-qp},
+% the resulting models models outperform
+% the state of the art by up to 18
+% percentage points.
+
+As with other forms of machine learning,
+a key concern is that of \emph{data-set bias}: are
+the models learned by \toolname specific
+to our data set, and would they fail on
+\emph{other} \ocaml programs?
 %
+We address this concern in two ways.
+%
+First, we partition the data by year,
+and show that models learned from one
+year generalize to, \ie perform nearly
+as well on, the programs from the other
+year.
+%
+Second, we argue that in our setting
+this bias is a \emph{feature} (and not
+a bug): it allows \toolname to \emph{adapt}
+to the kinds of errors that programmers
+(specifically, novices who are in greatest
+need of precise feedback) actually make
+rather than hardwiring the biases of
+compiler authors who, by dint of their
+training and experience, may suffer from
+blind spots with regards to such problems.
+%
+In this regard, we are particularly pleased
+that our classifiers can be trained on a
+modest amount of data, \ie a single course's
+worth, and envision a future where each course
+comes equipped with a model of its students' errors.
 

### 2017-04-13 paper/oopsla17-submission/intro2.tex
more intro2, plus abstract and conclusion

@@
 \toolname simply executes the model
-to generate a list of potential error
-locations ranked by likelihood.
+to generate a list of potential
+blame assignments ranked by likelihood.
 %
@@
 %
-We show that \toolname's data-driven
+We show that, when restricted to a
+\emph{single} prediction, \toolname's data-driven
 model is able to correctly predict
-the exact sub-expression that needs to
+the exact sub-expression that should
 be changed 74\% of the time,
@@
 %
-Further, \toolname's accuracy surpasses
+Furthermore, \toolname's accuracy surpasses
 85\% when we consider the \emph{top two}
@@
 is a set of more than 4,500 labeled programs,
-gives us an accurate ground truth of
+giving us an accurate ground truth of
 the kinds of errors and the (locations
@@
 %
-Finally, we label the changed locations by computing a
+Finally, we compute the blame labels using a
 \emph{tree-diff} between the two terms to find
@@
 solves this problem with a simple, yet expressive, representation called
-a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
+a \emph{Bag-of-Abstracted-Terms (BOAT)} wherein
 %
@@
 %
-We can model the error labels with a simple feature
+We can model the blame labels with a simple feature
 abstraction (\eg does-change-in-fix).
@@
 use the labeled BOAT data to train classifiers that
-are capable of \emph{precisely} pinpointing the error
-locations in real programs.
+are capable of \emph{precisely} pinpointing the errors
+in real programs.
 %

### 2017-04-13 paper/oopsla17-submission/intro2.tex
emph
minor

@@
 %
-\toolname works by analyzing a large corpus
+\toolname analyzes a large corpus
 of training data --- pairs of ill-typed
@@
 Furthermore, \toolname's accuracy surpasses
-85\% when we consider the \emph{top two}
+85\% when we consider the top \emph{two}
 locations and reaches 91\% if consider
-the \emph{top three}.
+the top \emph{three}.
 %

### 2017-04-13 paper/oopsla17-submission/intro2.tex
asd

--- 2017-04-13 paper/oopsla17-submission/intro2.tex
@@
-\section{Introduction}
+\section{\textbf{Introduction}}
 \label{sec:introduction}
@@
 Unfortunately, this automation has come at a price.
-Verbose type annotations (as in, \eg, \java) signify
+Verbose type annotations (as in, \eg, Java) signify
 the programmer's intent and help to correctly
@@
 %
-In the absence of such signifiers, automatic
+In contrast, lacking such signifiers, automatic
 type inference algorithms are prone to report
-type errors far from their actual source
-\citep{Wand1986-nw}.
+type errors quite far from their actual source.
 %
 While this can seem like a minor annoyance to
-veteran programmers, \citet{Joosten1993-yx} have found
-that novices often focus their attention on the \emph{location}
-reported and disregard the \emph{message}.
+veteran programmers, it turns out to be a significant
+stumbling block for novices~\citep{Wand1986-nw,Joosten1993-yx}.
 
@@
 satisfiable and hence, memory-safe~\citep{Jose:2011}
-or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-gd,Pavlinovic2014-mr}.
+or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-vm,Pavlinovic2014-mr}.
 The finger of blame is then pointed at the
@@
 First, they are not \emph{extensible}:
-% the weights and analyses are limited
-% to attributes of the constraints, and
-% not other \emph{signals} from the
-% program's syntax that may be more
-% germane to error localization. Furthermore,
-the constraint languages and algorithms for computing
+the weights and analyses are limited
+to attributes of the constraints, and
+not other \emph{signals} from the
+program's syntax that may be more
+germane to error localization. Furthermore,
+the algorithms for computing
 the minimum weighted subset must be
 designed afresh for different kinds
-of type systems and constraints \citep{Loncaric2016-uk}.
+of type systems and constraints.
 %
@@
 weights are fixed in an ad-hoc fashion, based on the
-\emph{analysis designer's} notion
+\emph{analysis designers} notion
 of what kinds of errors are more
@@
 localization based on supervised
-learning (see \citealt{Kotsiantis2007-pj} for a survey).
+classification~\citep{Kotsiantis2007-pj}.
 %
-\toolname works by analyzing a large corpus
+\toolname works by using a large corpus
 of training data --- pairs of ill-typed
-programs and their subsequent fixes ---
+programs and their fixed versions ---
 to automatically \emph{learn a model}
-of where errors are most likely to
-be found.
+of where the error is most likely.
 %
-Given a new ill-typed program,
+When given a new ill-typed program,
 \toolname simply executes the model
-to generate a list of potential error
+to generate a list of likely error
 locations ranked by likelihood.
@@
 precision against the state-of-the-art
-on a set of over 4,500 ill-typed \ocaml
+on a set of over 4,500 ill-typed Ocaml
 programs drawn from introductory
@@
 the exact sub-expression that needs to
-be changed 74\% of the time,
-29 points higher than \ocaml and
+be changed, 74\% of the time, which is
+29 points higher than \ocaml, and
 18 points higher than the state-of-the-art
@@
 Further, \toolname's accuracy surpasses
-85\% when we consider the \emph{top two}
+85\% when we consider the \emph{top-2}
 locations and reaches 91\% if consider
-the \emph{top three}.
+the \emph{top-3}.
 %
@@
 The first challenge for supervised learning
-is to acquire a corpus of training data, in our setting
+is to acquire a corpus of training data, here,
 a set of ill-typed programs \emph{labeled}
@@
 to manually judge ill-typed programs and
-determine the ``correct'' fix
-\citep[\eg][]{Lerner2007-dt,Loncaric2016-uk},
+determine the ``correct'' fix \citep[\eg][]{Loncaric2016-uk},
 but this approach does not scale well to
@@
 software development is an iterative
-process; programmers eventually
+process and programmers eventually
 fix their own ill-typed programs,
@@
 traces of student interactions over two instances
-of an undergraduate Programming Languages
-course.\footnote{
+of an undergraduate Programming Languages course
+%
+\footnote{
 % \begin{anonsuppress}
@@
 %
-We then post-process the resulting time-series
+We can then post-process the resulting time-series
 of programs submitted to the \ocaml compiler into
 a set of pairs of ill-typed programs and their
-subsequent \emph{fixes}, the first (type-)~correct
+\emph{fixed} versions, the \emph{first} (type-)correct
 program in the trace suffix.
 %
-Finally, we label the changed locations by computing a
+We obtain the location labels by computing a
 \emph{tree-diff} between the two terms to find
@@
 %
-Modern supervised learning algorithms work on %sets of
+Modern supervised learning algorithms work on sets of
 \emph{feature vectors}: real-valued points in an
-$n$-dimensional space. While there are standard
-techniques for computing such
-vectors for documents, images, and sound (respectively
-word-counts, pixel-values, and frequencies),
-% it has
-% hitherto been unclear how to precisely model program
-% structure as numeric data.
-there are no similarly standard representations for
-programs.
+$n$-dimensional space. While it is easy to compute such
+vectors for documents, images and sound (respectively
+word-counts, pixel-values and frequencies), it has
+hitherto been unclear how to precisely model program
+structure as numeric data.
 
 Our \emph{second contribution} (\autoref{sec:learning})
-solves this problem with a simple, yet expressive, representation called
+solves this problem with a new representation called
 a \emph{Bag-of-Abstract-Terms (BOAT)} wherein
@@
 (1)~each program is represented by the \emph{bag}
-    or multiset of (sub-)~terms that appears inside
-    it; and
+    or multiset of (sub-)terms that appears inside
+    it, and
 %
-(2)~each (sub-)~term $t$ is \emph{abstracted} as
-    a vector $[f_1(t), \ldots, f_n(t)]$, where
+(2)~each (sub-)term $t$ is \emph{abstracted} as
+    a vector $[f_1(t), \ldots, f_n(t)]$ where
     each $f_i$ is a \emph{feature abstraction}
@@
 We can model the error labels with a simple feature
-abstraction (\eg does-change-in-fix).
+abstraction (\eg is-a-diff-subterm).
 %
 We can even account for \emph{contextual} information like
-the features of the parent- and child- terms by
+the features of the parent- and children- terms by
 concatenating the feature vectors of each term with those
 of its parent and children.
-%
-We have found this representation to be particularly
-convenient as it gives us flexibility
-in modeling the syntactic and semantic structure of programs
-while retaining compatibility with off-the-shelf classifiers,
-in contrast to, \eg, \citet{Raychev2015-jg}, who had to develop
-their own variants of classifiers to obtain their results.
 
@@
 syntactic features (\eg is-a-data constructor, is-a-literal,
-is-an-arithmetic-operation, is-a-function-application, \etc),
+is-an-arithmetic-operation, is-a-function-application \etc),
 %
 to semantic features captured by the type system (\eg is-a-list,
-is-an-integer, is-a-function, \etc).
+is-an-integer, is-a-function \etc).
 
@@
 using different classes of features like the above, and
-with three different classification algorithms: logistic
-regression, decision trees, and neural networks.
+with three different classification algorithms: linear
+classifiers, decision forests and deep neural networks.
 %
@@
 In particular, we find that while machine learning
-over syntactic features of each term in isolation
-performs worse than existing
+on its own performs slightly worse than existing
 purely constraint based approaches (\eg \ocaml, \sherrloc),
-augmenting the data with a single feature corresponding to
-the \emph{type error slice} \citep{Tip2001-qp} brings our
-classifiers up to par with the state of the art,
-and further augmenting the data with \emph{contextual}
-features allows our classifiers to outperform
-the state of the art by 18 percentage points.
-% In particular, we find that while machine learning
-% on its own performs slightly worse than existing
-% purely constraint based approaches (\eg \ocaml, \sherrloc),
-% once we augment the data with feature abstraction
-% corresponding to the \emph{type error slice} \citep{Tip2001-qp},
-% the resulting classifier significantly outperforms the
-% state of the art by 18 percentage points.
+once we augment the data with feature abstraction
+corresponding to the \emph{type error slice} \cite{Tip2001-qp},
+the resulting classifier significantly outperforms
+state-of-the-art by 18 percentage points.
 

### 2017-04-13 paper/oopsla17-submission/overview.tex
asd

--- 2017-04-12 paper/oopsla17-submission/overview.tex
@@
-\section{Overview}
-\label{sec:overview}
+\mysection{Overview} \label{sec:overview}
 
-We start with an overview of our approach to localizing type errors by
-learning a model of the mistakes programmers actually make.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[t!]
+\begin{minipage}{0.45\linewidth}
+\begin{ecode}
+  let rec sumList xs =
+    match xs with
+    | []   -> (*@\hlTree{\hlSherrloc{[]}}@*)
+    | h::t -> h (*@\hlTree{+}@*) (*@\hlFix{sumList t}@*)
+\end{ecode}
+\end{minipage}
+\begin{minipage}{0.49\linewidth}
+\begin{verbatim}
+File "sumList.ml", line 4, characters 16-25:
+  This expression has type 'a list
+  but an expression was expected of type int
+\end{verbatim}
+\end{minipage}
+\caption{(left) An ill-typed \ocaml program that should sum the elements of a
+  list, with highlights indicating three possible blame assignments based on:
   %
-We formulate the problem of type error localization as a
-\emph{supervised classification} problem (see \citealt{Kotsiantis2007-pj}
-for a survey).
+  (1) the \hlFix{\ocaml} compiler;
   %
-A \emph{classification} problem entails learning a function that maps
-inputs to a discrete set of output \emph{labels}, in contrast to
-\emph{regression}, where the output is typically a real number.
+  (2) the \hlSherrloc{fix} made by the programmer; and
   %
-In a \emph{supervised} learning problem one is given a \emph{training}
-set where the inputs and labels are known, and the task is to learn a
-function that accurately maps the inputs to outputs and
-\emph{generalizes} to new, yet-unseen inputs.
+  (3) \hlTree{minimizing} the number of edits required.
+  % The \underline{underlined} expressions are equally valid
+  % locations to blame. The expression blamed by the \ocaml compiler
+  % is in \textbf{bold}.
+  %
+  % FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
+  % ``with'' are also bolded (by \\ecode)! You need to find another way to
+  % highlight what ocaml is yelling about.
+  %
+  (right) The error reported by \ocaml.}
+\label{fig:sumList}
+\end{figure}
 
-The key technical challenges we address are:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Lets start with an overview of \toolname's
+approach to localizing type errors by
+collectively learning from the mistakes
+programmers actually make.
+
+\mypara{The Problem}
 %
-(1) formulating type error localization as a classification problem;
+Consider the |sumList| program in
+\autoref{fig:sumList}, written by
+a student in an undergraduate
+Programming Languages course.
 %
-%(2) generating correct labels for supervised learning; and
-(2) identifying ``correct'' blame assignments in our dataset; and
+The program is meant to add up the
+integers in a list, but the student
+has accidentally given the empty
+list as the base case, rather than |0|.
 %
-(3) presenting the predictions of the classifier in an intuitive manner.
+The \ocaml compiler collects typing
+constraints as it traverses the program,
+and reports an error the moment it finds
+an inconsistent constraint.
+%
+In this case it blames the recursive call
+to |sumList|, complaining that |sumList|
+returns a |list| where an |int| was
+expected by the |+| operator.
+%
+This \emph{blame} assignment is inconsistent
+with the programmer's intention and may
+not be helpful debugging information to
+a novice.
+
+It may appear obvious to the reader that
+|[]| is the correct expression to blame,
+but how is a type checker to know that?
+%
+Indeed, both the venerable
+\ocaml compiler as well as
+recent techniques like
+\sherrloc and \mycroft~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
+fail to distinguish between
+the two underlined expressions
+in \autoref{fig:sumList};
+it would be equally valid
+to blame \emph{either}
+of them alone.
+%
+The |[]| on line 3 could be changed to |0|,
+or the |+| on line 4 could be changed to
+either |@| (list append) or |::|, all of
+which would give type-correct programs.
+%
+Thus, these state-of-the-art techniques
+are forced to either blame \emph{both}
+locations, or choose one \emph{arbitrarily}.
+
+\mypara{Error Localization via Supervised Classification}
+%
+Our approach is to view error localization as a
+\emph{supervised classification}
+problem~\citep{Kotsiantis2007-pj}.
+%
+A \emph{classification} problem entails learning
+a function that maps \emph{inputs} to a discrete
+set of output \emph{labels} (in contrast to, say,
+\emph{regression}, where the output is typically
+a real number.)
+%
+A \emph{supervised} learning problem is one where
+we are given a \emph{training set} where the
+inputs and labels are known, and the task is to
+learn a function that accurately maps the inputs
+to output labels and \emph{generalizes} to new,
+yet-unseen inputs.
+%
+To realize the above approach for error localization
+as a practical tool, we have to solve four sub-problems.
+%
+\begin{enumerate}
+  \item How can we acquire a \emph{training set} of
+        blame-labeled ill-typed programs?
+
+  \item How can we \emph{represent} blame-labeled programs
+        as numeric vectors amenable to machine learning?
+
+  \item How can we find \emph{features} that yield predictive
+        models?
+
+  \item How can we use the models to give localized
+        \emph{feedback} to the programmer?
+\end{enumerate}
+
+\mysubsection{Step 1: Acquiring a Blame-Labeled Training Set}
+
+The first step is to create a training
+set of ill-typed programs, where each
+erroneous sub-term is explicitly labeled.
+%
+Prior work has often enlisted
+expert users to curate a set of
+ill-typed programs and then
+\emph{manually} determine the
+correct fix~\citep[\eg][]{Lerner2007-dt,Loncaric2016-uk}.
+%
+This method is suitable for evaluating
+the quality of a localization (or repair)
+algorithm on a small number (\eg, 10-100s)
+of programs.
+%
+However, in general it requires a great
+deal of effort for the expert to divine
+the original programmers' intentions.
+%
+Consequently, is difficult to scale the
+expert-labeling to yield a dataset large
+enough (\eg, 1000s of programs) needed to
+facilitate machine learning.
+%
+More importantly, this approach fails to
+capture the \emph{frequency} with which
+errors occur in practice.
+
+We solve both the scale and frequency problems
+by instead extracting blame-labeled data sets
+from \emph{interaction traces}.
+%
+Software development is an iterative process.
+Programmers, perhaps after a lengthy (and
+sometimes frustrating) back-and-forth with
+the type checker, eventually end up fixing
+their own programs.
+%
+Thus, we implemented an instrumented
+version of the \ocaml compiler that records
+this conversation, \ie records the sequence
+of programs submitted by each programmer and
+whether or not it was deemed type-correct.
+%
+For each ill-typed program in
+a particular programmer's trace,
+we find the \emph{first subsequent}
+program in the trace that type checks
+and declare it to be the fixed version.
+%
+From this pair of an ill-typed program
+and its fix, we can extract a \emph{diff}
+of the abstract syntax trees, and then assign
+the blame labels to the \emph{smallest}
+sub-tree in the diff.
+
+For example, suppose our student
+fixed the |sumList| program in
+\autoref{fig:sumList} by replacing
+|[]| with |0|, the diff would
+include only the |[]| expression.
+%
+Thus we would determine that the
+|[]| expression (and \emph{not} the
+|+| or the recursive call |sumList t|)
+is to blame.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\mysubsection{Step 2: Representing Programs as Vectors}
+
+\mysubsection{Step 3: Feature Discovery}
+
+\mysubsection{Step 4: Generating Feedback}
+
+% Evaluation
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \paragraph{\textbf{Challenge (1): How to represent programs?}}
+
 The first and prime challenge we address is how to translate programs
@@
 \paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
-As we saw in \autoref{fig:sumList}, type errors admit many different
-fixes to different expressions, each resulting in a different program.
 %
-How is the type checker to judge which expressions are more likely to be
-incorrect than others?
-%
-Prior work has often enlisted expert users to manually judge ill-typed
-programs and determine the ``correct'' fix
-\citep[\eg][]{Loncaric2016-uk}, but this approach does not scale well to
-a dataset large enough to support machine learning.
-%
-Furthermore, while expert users have intimate knowledge of the type system,
-they cannot know in general what novice users \emph{intended} to do.
+%% As we saw in \autoref{fig:sumList},
+%% type errors admit many different
+%% fixes to different expressions,
+%% each resulting in a different program.
+%% %
+%% How is the type checker to judge
+%% which expressions are more likely
+%% to be incorrect than others?
 
@@
 %
-Rather than relying on expert users, we allow our students to do the
-work of assigning blame themselves.
+Rather than relying on expert users, we follow
+\citet{Zhang2014-lv} \RJ{CHECK} and allow our students to
+do the work of assigning blame themselves.
 %
@@
 attempts.
-%
-For each ill-typed program in our dataset, we find the first
-\emph{subsequent} program that type checks and declare that to be the
-fix.
-%
-From this pair of an ill-typed program and its fix, we can extract a
-\emph{diff} of the abstract syntax trees.
-%
-For example, suppose our student fixed the |sumList| program in
-\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
-only the |[]| expression.
-%
-Thus we would determine that the |[]| expression (and \emph{not} the
-|+| or the recursive call |sumList t|) is to blame.
 

### 2017-04-13 paper/oopsla17-submission/learning.tex
done with overview

@@
 % \end{figure}
-\begin{table}[ht]
-\caption{Example Feature Vectors}\label{tab:sumList}
-\begin{tabular}{lrrrrrr}
-\toprule
-Expression
-  & \IsNil & \IsCaseListP & \ExprSize
-  & \HasTypeIntCOne & \HasTypeList & \InSlice \\
-\midrule
-|[]|
-  & 1 & 1 & 1 & 0 & 1 & 1 \\
-|hd + sumList tl|
-  & 0 & 1 & 5 & 1 & 0 & 1 \\
-|sumList tl|
-  & 0 & 0 & 3 & 0 & 1 & 1 \\
-|tl|
-  & 0 & 0 & 1 & 0 & 1 & 0 \\
-\bottomrule
-\end{tabular}
-\bigskip
-\caption*{A selection of the features we would extract from the
-\lstinline!sumList! program in \autoref{fig:sumList}. A feature is
-considered \emph{enabled} if it has a non-zero value, and
-\emph{disabled} otherwise. A ``-P'' suffix indicates that the feature
-describes the parent of the current expression, a ``-C$n$'' suffix
-indicates that the feature describes the $n$-th (left-to-right) child of
-the current expression.  Note that, since we rely on a partial typing
-derivation, we are subject to the well-known traversal bias and label
-the expression \lstinline!sumList tl! as having type
-$\tlist{\cdot}$. The model will have to learn to correct for this bias.}
-\end{table}
+
+\input{feature-table}
+
 The first issue we must tackle is formulating our learning task in

### 2017-04-13 paper/oopsla17-submission/overview.tex
done with overview

@@
 %
-This method is suitable for evaluating
-the quality of a localization (or repair)
-algorithm on a small number (\eg, 10-100s)
-of programs.
+This method is suitable for
+\emph{evaluating} the quality
+of a localization (or repair)
+algorithm on a small number
+(\eg, 10-100s) of programs.
 %
-However, in general it requires a great
-deal of effort for the expert to divine
-the original programmers' intentions.
+However, in general it requires
+a great deal of effort for the
+expert to divine the original
+programmers' intentions.
 %
-Consequently, is difficult to scale the
-expert-labeling to yield a dataset large
-enough (\eg, 1000s of programs) needed to
-facilitate machine learning.
+Consequently, is difficult to
+scale the expert-labeling to
+yield a dataset large enough
+(\eg, 1000s of programs) needed
+to facilitate machine learning.
 %
-More importantly, this approach fails to
-capture the \emph{frequency} with which
-errors occur in practice.
+More importantly, this approach
+fails to capture the \emph{frequency}
+with which errors occur in practice.
 
-We solve both the scale and frequency problems
-by instead extracting blame-labeled data sets
+\mypara{Solution: Interaction Traces}
+%
+We solve both the scale and
+frequency problems by instead
+extracting blame-labeled data sets
 from \emph{interaction traces}.
@@
 
+\mypara{Example}
 For example, suppose our student
@@
 
+Next, we must find a way to translate
+highly structured and variable sized
+\emph{programs} into fixed size
+$n$-dimensional numeric \emph{vectors}
+that are needed for supervised
+classification.
+%
+While the PL literature is full
+of different kinds of ways to
+represent programs, from raw
+sequences of tokens, to
+richly-structured
+abstract syntax trees (AST) or
+control-flow graphs, it is
+unclear how to embed the
+above into a vector space.
+%
+Furthermore, it is unclear whether
+recent program representations that
+are amenable to one learning task,
+\eg code completion \cite{Devanbu:2012,Raychev:2014}
+or decompilation \cite{Raychev2015-jg,Bielik2016-br}
+are suitable for our problem of
+assigning blame for type errors.
+
+\mypara{Solution: Bags-of-Abstracted-Terms}
+%
+We present a new representation of programs
+that draws inspiration from the theory of
+abstract interpretation~\cite{CousotCousot77}.
+%
+Our representation as parameterized by a
+set of \emph{feature abstraction} functions,
+(abbreviated to feature abstractions)
+$f_1, \ldots, f_n$, that map terms to a
+numeric value (or just $\{0, 1\}$ to
+encode a Boolean property.)
+%
+Given a set of feature abstractions, we
+can represent a single program's AST as
+a \emph{bag-of-abstracted-terms} (BOAT)
+by:
+%
+(1)~Decomposing the AST (term) $t$ into
+    a \emph{bag} of its constituent sub-trees
+    (terms) $\{t_1,\ldots,t_m\}$, and then
+%
+(2)~Representing each sub-term $t_i$
+    with the $n$-dimensional
+    vector $[f_1(t_i),\ldots, f_n(t_i)]$.
+%
+Our choice of working with ASTs is natural
+as that is the representation over which the
+type-checker operates.
+
+\mypara{Modeling Contexts}
+%
+Each expression occurs in some surrounding
+\emph{context}, and we would like the
+classifier to be able make decisions based
+on the context as well.
+%
+The context is particularly important
+for our task as each expression
+imposes typing constraints on its
+neighbors.
+%
+For example, \eg a |+| operator tells
+the type checker that both \emph{children}
+must have type |int| and that the \emph{parent}
+must accept an |int|.
+%
+Similarly, if the student wrote
+|h sumList t| \ie forgot the |+|,
+we might wish to blame the application
+rather than |h| because |h|
+\emph{does not} have a function type.
+%
+The BOAT representation makes it
+easy to incorporate contexts: we
+simply \emph{concatenate} each
+term's feature vector with the
+\emph{contextual features} of
+its parent and children.
+
+%% \RJ{confusing -- delete}
+%% We might expect that the recursive
+%% call |sumList t| is unlikely to be
+%% at fault because |sumList| has a
+%% function type --- the user may have
+%% called the wrong function or
+%% supplied the wrong argument, but the
+%% application itself is probably correct.
+
 \mysubsection{Step 3: Feature Discovery}
+
+Next, we must find a \emph{good}
+set of features, that is, a set
+of features that yields predictive
+models. Our BOAT representation
+lets us iteratively solve this
+problem by just starting with
+a simple set of features, and
+then repeatedly adding more
+and more to capture important
+aspects needed to improve precision.
+%
+Our set feature abstractions
+includes those that capture the
+\emph{syntax}, \emph{types} and
+\emph{context} of each expression.
+
+\mypara{Syntax and Type Features}
+%
+We start by observing that
+at the very least, the
+classifier should be able
+to distinguish between the
+|[]| and |+| expressions
+in \autoref{fig:sumList}
+because they represent
+different \emph{syntactic}
+expression forms.
+%
+We model this by
+introducing feature
+abstractions of the form
+is-nil, is-plus, \etc for
+each of a fixed number of
+primitive literal terms.
+%
+Modeling the syntactic class of an
+expression gives the classifier a
+basic notion of the relative
+frequency of blame assignment
+for the various program elements,
+\ie perhaps |[]| is
+\emph{empirically} more
+likely to be blamed than |+|.
+%
+Similarly, we can model
+the \emph{type} of each
+sub-expression with features
+of the form is-int, is-bool, \etc.
+%
+We will discuss handling
+arbitrary, user-defined types
+in~\autoref{sec:discussion}.
+
+\mypara{Contextual Features: Error Slices}
+%
+Our contextual features include the
+syntactic class of the neighboring
+expressions, their inferred types
+(when available).
+%
+However, in we have found that
+the most important contextual
+signal is whether or not the
+expression occurs in a
+\emph{minimal type error slice}~\citep{Haack2003-vc}
+which includes the smallest
+subset of all expressions
+that are necessary for the
+error to manifest.
+%
+We propose to use type error slices
+to communicate to the classifier
+which expressions could
+\emph{potentially} be blamed --- a
+change to an expression outside of
+the minimal slice cannot possibly
+fix the type error.
+%
+We empirically demonstrate that
+the type error slice is so
+important (\autoref{sec:feature-utility})
+that it is actually beneficial to
+automatically discard expressions
+that are not part of the slice,
+rather than letting the classifier
+learn to do so.
+%
+Indeed, this domain-specific
+insight is crucial for learning
+classifiers that significantly
+outperform the state-of-the-art.
+
+\mypara{Example}
+%
+When \toolname is tasked with localizing
+the error in the example program of \autoref{fig:sumList},
+the |[]| and |+| sub-terms will each be given
+their own feature vector, and we will ask the
+classifier to predict for each \emph{independently}
+whether it should be blamed.
+%
+\autoref{tab:sumList} lists some
+of the sub-expressions of the example
+from \autoref{fig:sumList}, and their
+corresponding feature vectors.
+
+% Our use of type error slices is
+% conceptually related to the use
+% of fault localization in other
+% program analysis and transformation
+% techniques (see \autoref{sec:related-work}).
 
@@
 
-% Evaluation
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{\textbf{Challenge (1): How to represent programs?}}
-
-The first and prime challenge we address is how to translate programs
-into a format favorable to learning a model of type errors.
+Finally, after we have trained a classifier
+using the labeled data set, we need to use
+it to help users localize type errors in
+their programs.
 %
-Programs can be represented in various ways, from the raw text to
-richly-structured abstract syntax trees and control-flow graphs, but
-many representations are not directly amenable to machine learning.
+The classifier tells us whether or not
+a given sub-expression \emph{should be}
+blamed (\ie has the blame label) but this
+is not yet particularly suitable as
+\emph{user feedback}.
 %
-Furthermore, program representations that are amenable to one learning
-task, \eg code completion, may not be favorable to our problem of
-assigning blame for type errors.
-
-Machine learning techniques typically take as input \emph{feature
-vectors}, fixed-length vectors of real numbers where each element
-represents some property of the input.
+A recent survey of developers by
+\citet{Kochhar2016-oc} found that
+developers are unlikely to examine
+more than around five potentially
+erroneous locations before falling
+back to manual debugging.
 %
-Thus, to apply machine learning to our problem we must
-determine:
-%
-(1) which features of a program would be useful for modeling type errors; and
-%
-(2) how can we represent a variable-size program as a fixed-length vector?
-
-\paragraph{\textbf{Solution: Vectorized Abstract Syntax Trees}}
-\ES{name up for grabs...}
-We propose to extract features from the program's
-abstract syntax tree (AST), a natural choice since the type checker
-operates on ASTs.
-%
-We model features as \emph{queries} over the AST.\@
-%
-Our initial set of interesting features includes the syntactic class of
-each expression.
-%
-That is, the classifier should be able to distinguish between the |[]|
-and |+| expressions in \autoref{fig:sumList} because they represent
-different syntactic expression forms.
-%
-Modeling the syntactic class of an expression gives the classifier a
-basic notion of the relative frequency of blame assignment for the
-various program elements, \ie perhaps |[]| is \emph{empirically} more
-likely to be blamed than |+|.
-% % ES: (2) doens't really make sense here
-% (2) the type checker uses the syntactic class of an expression to
-%     impose typing constraints on the neighboring expressions.
-
-A programming language has a fixed set of AST elements,\footnote{We will
-  discuss handling arbitrary, user-defined types in
-  \autoref{sec:discussion}.} which suggests a natural way to obtain
-fixed-size feature vectors. Rather than translating a program into a
-single feature vector, we translate it into a \emph{set} of feature
-vectors, one for each expression in the program.
-%
-Thus, the |[]| and |+| expressions will each be given their own feature
-vector, and we will ask the classifier to predict for each
-\emph{independently} whether it should be blamed.
-
-Of course, each expression occurs in some \emph{context}, and we would
-like the classifier to be able make decisions based on the context as
-well.
-%
-The context is particularly important for our task as each expression
-imposes typing constraints on its neighbors, \eg a |+| operator tells
-the type checker that both children must have type |int| and that
-the parent must accept an |int|.
-%
-We might expect that the recursive call |sumList t| is
-unlikely to be at fault because |sumList| has a function type --- the
-user may have called the wrong function or supplied the wrong argument,
-but the application itself is probably correct.
-%
-Conversely, if the student wrote |h sumList t| (\ie forgot the |+|), we
-might wish to blame the application rather than |h| because |h|
-\emph{does not} have a function type.
-%
-Thus we extend each expression's feature vector with
-\emph{contextual features} of its parent and children.
-
-Our contextual features include the syntactic class of the neighboring
-expressions, their inferred types (when available), and, crucially, whether
-the expression occurs in a minimal type error \emph{slice}.
-%
-A minimal type error slice \citep{Haack2003-vc} includes all expressions that
-are necessary for the error to manifest and no more.
-%
-We propose to use type error slices to communicate to the classifier
-which expressions could potentially be blamed --- a change to an
-expression outside of the minimal slice cannot possibly fix the type
-error.
-%
-We will show that the type error slice is so important
-(\autoref{sec:feature-utility}) that it is actually beneficial to
-automatically discard expressions that are not part of the slice, rather
-than letting the classifier learn to do so.
-%
-Our use of type error slices is conceptually related to the use of fault
-localization in other program analysis and transformation techniques
-(see \autoref{sec:related-work}).
-
-\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
-%
-%% As we saw in \autoref{fig:sumList},
-%% type errors admit many different
-%% fixes to different expressions,
-%% each resulting in a different program.
-%% %
-%% How is the type checker to judge
-%% which expressions are more likely
-%% to be incorrect than others?
-
-\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
-%
-Rather than relying on expert users, we follow
-\citet{Zhang2014-lv} \RJ{CHECK} and allow our students to
-do the work of assigning blame themselves.
-%
-Software development is an iterative process and our students
-eventually fix their own programs, perhaps after multiple ill-fated
-attempts.
-
-\paragraph{\textbf{Challenge (3): How can we intuitively present our feedback?}}
-We now have the features and labels needed to train a classifier to
-predict, for each expression in an ill-typed program, whether it should
-be blamed.
-%
-But this is not yet particularly suitable as \emph{user feedback}.
-%
-A recent survey of developers by \citet{Kochhar2016-oc} found that
-developers are unlikely to examine more than around five potentially
-erroneous locations before falling back to manual debugging.
-%
-Thus, we should limit our predictions to a select few to be presented to
+Thus, we should limit our predictions
+to a select few to be presented to
 the user.
 
-\paragraph{\textbf{Solution: Rank blame locations by ``confidence''}}
-Luckily, many machine learning classifiers produce not only a predicted
-label, but also a metric that can be interpreted as the classifier's
+\mypara{Solution: Rank Locations by ``Confidence''}
+%
+Fortunately, many machine learning
+classifiers produce not only a predicted
+label, but also a metric that can be
+interpreted as the classifier's
 \emph{confidence} in its prediction.
 %
-Thus, we \emph{rank} each expression by the classifier's confidence that
-it should be blamed, and present only the top-$k$ predictions to the
-user (in practice $k=3$). This use of ranking to report
-the results of a program analysis is popular in other problem domains
-\citep[see, \eg][]{Kremenek2003-ck}; we focus
-explicitly on the use of data-driven machine learning confidence as a
-ranking source.
+Thus, we \emph{rank} each expression
+by the classifier's confidence that
+it should be blamed, and present only
+the top-$k$ predictions to the
+user (in practice $k=3$).
+%
+This use of ranking to report the
+results of a program analysis is
+popular in other problem domains
+\citep[see, \eg][]{Kremenek2003-ck};
+we focus explicitly on the use of
+data-driven machine learning
+confidence as a ranking source.
+%
+In \autoref{sec:evaluation} we show
+that \toolname's ranking approach
+yields a high-precision localizer:
+when the top-3 locations are considered,
+one of them is the real location
+more than 90\% of the time.
+
+%%% HOOK TO EVALUATION: Evaluation
 

### 2017-04-14 paper/oopsla17-submission/abstract.tex

@@
 %
-\toolname works by using a large corpus
+\toolname analyzes a large corpus
 of training data --- pairs of ill-typed

### 2017-04-14 paper/oopsla17-submission/overview.tex
minor tweaks to overview

@@
-\mysection{Overview} \label{sec:overview}
+\mysection{Overview}\label{sec:overview}
 
@@
 
-Lets start with an overview of \toolname's
+Let us start with an overview of \toolname's
 approach to localizing type errors by
@@
 recent techniques like
-\sherrloc and \mycroft~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
+\sherrloc and \mycroft
+\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}
 fail to distinguish between
-the two underlined expressions
+the |[]| and |+| expressions
 in \autoref{fig:sumList};
@@
 a function that maps \emph{inputs} to a discrete
-set of output \emph{labels} (in contrast to, say,
+set of output \emph{labels} (in contrast to %, say,
 \emph{regression}, where the output is typically
-a real number.)
+a real number).
 %
@@
 
-The first step is to create a training
+The first step is to gather a training
 set of ill-typed programs, where each
@@
 algorithm on a small number
-(\eg, 10-100s) of programs.
+(\eg 10s--100s) of programs.
 %
@@
 expert to divine the original
-programmers' intentions.
+programmer's intentions.
 %
@@
 yield a dataset large enough
-(\eg, 1000s of programs) needed
+(\eg 1000s of programs) needed
 to facilitate machine learning.
@@
 %
-Thus, we implemented an instrumented
-version of the \ocaml compiler that records
-this conversation, \ie records the sequence
+Thus, we instrumented
+the \ocaml compiler to record
+this conversation, \ie record the sequence
 of programs submitted by each programmer and
@@
 While the PL literature is full
-of different kinds of ways to
-represent programs, from raw
+of different program
+representations, from raw
 sequences of tokens, to
@@
 are amenable to one learning task,
-\eg code completion \cite{Devanbu:2012,Raychev:2014}
-or decompilation \cite{Raychev2015-jg,Bielik2016-br}
+\eg code completion \citep{Devanbu:2012,Raychev:2014}
+or decompilation \citep{Raychev2015-jg,Bielik2016-br}
 are suitable for our problem of
@@
 that draws inspiration from the theory of
-abstract interpretation~\cite{CousotCousot77}.
+abstract interpretation \citep{CousotCousot77}.
 %
@@
 numeric value (or just $\{0, 1\}$ to
-encode a Boolean property.)
+encode a boolean property).
 %
@@
 %
-(1)~Decomposing the AST (term) $t$ into
+(1)~decomposing the AST (term) $t$ into
     a \emph{bag} of its constituent sub-trees
-    (terms) $\{t_1,\ldots,t_m\}$, and then
+    (terms) $\{t_1,\ldots,t_m\}$; and then
 %
-(2)~Representing each sub-term $t_i$
+(2)~representing each sub-term $t_i$
     with the $n$-dimensional
@@
 %
-For example, \eg a |+| operator tells
+For example, a |+| operator tells
 the type checker that both \emph{children}
@@
 %
-Our set feature abstractions
-includes those that capture the
-\emph{syntax}, \emph{types} and
+Our set of feature abstractions
+%includes those that capture the
+captures the
+\emph{syntax}, \emph{types}, and
 \emph{context} of each expression.
@@
 syntactic class of the neighboring
-expressions, their inferred types
+expressions and their inferred types
 (when available).
 %
-However, in we have found that
+However, we have found that
 the most important contextual
@@
 expression occurs in a
-\emph{minimal type error slice}~\citep{Haack2003-vc}
+\emph{minimal type error slice}
+\citep{Tip2001-qp,Haack2003-vc}
 which includes the smallest
@@
 
-Finally, after we have trained a classifier
+Finally, having trained a classifier
 using the labeled data set, we need to use
@@
 The classifier tells us whether or not
-a given sub-expression \emph{should be}
+a sub-term \emph{should be}
 blamed (\ie has the blame label) but this
@@
 when the top-3 locations are considered,
-one of them is the real location
+at least one matches an actual student fix
 more than 90\% of the time.

### 2017-04-14 paper/oopsla17-submission/overview.tex
some notes

@@
 
+\ES{FIXME: this feels a bit redundant with CHALLENGE 1. the main thing i
+  get is that we use student fixes to extract AST-diffs, which is
+  already said in CHALLENGE 1}
+
 \mypara{Example}
@@
 
+\ES{FIXME: this feels a bit redundant with CHALLENGE 2. the main things
+  i get here are that we use sets of feature vectors, and that we can
+  concatenate vectors to add context. but this is all said in CHALLENGE
+  2 already.}
+
 %% \RJ{confusing -- delete}
@@
 error to manifest.
+%
+\ES{FIXME: this is a bit imprecise. we compute \textbf{a} minimal slice (ie if
+  you remove any more terms the error disappears) but not \textbf{the} minimal
+  slice. there may be other slices that are smaller, but computing all
+  slices is exponential, so we don’t want to do that..}
 %

### 2017-04-14 paper/oopsla17-submission/intro2.tex
tweak intro

--- 2017-04-13 paper/oopsla17-submission/intro2.tex
@@
 a \emph{Bag-of-Abstracted-Terms (BOAT)} wherein
-%
-(1)~each program is represented by the \emph{bag}
+each program is represented by the \emph{bag}
 or multiset of (sub-)~terms that appears inside
-    it; and
-%
-(2)~each (sub-)~term $t$ is \emph{abstracted} as
-    a vector $[f_1(t), \ldots, f_n(t)]$, where
-    each $f_i$ is a \emph{feature abstraction}
-    that maps a term to a numeric value.
+it; and further, each (sub-)~term is \emph{abstracted}
+as a feature vector comprising the numeric values
+returned by \emph{feature abstraction} functions
+that applied to the term.
 %
 We can model the blame labels with a simple feature
-abstraction (\eg does-change-in-fix).
+abstraction (\eg is-changed-in-fix).
 %
-We can even account for \emph{contextual} information like
-the features of the parent- and child- terms by
-concatenating the feature vectors of each term with those
-of its parent and children.
+We can even account for \emph{contextual} information
+like the features of the parent- and child- terms by
+\emph{concatenating} the feature vectors of each term
+with those of its parent and children.
 %
 We have found this representation to be particularly
-convenient as it gives us flexibility
-in modeling the syntactic and semantic structure of programs
-while retaining compatibility with off-the-shelf classifiers,
-in contrast to, \eg, \citet{Raychev2015-jg}, who had to develop
-their own variants of classifiers to obtain their results.
+convenient as it gives us flexibility in modeling the
+syntactic and semantic structure of programs while
+retaining compatibility with off-the-shelf classifiers,
+in contrast to, \eg, \citet{Raychev2015-jg}, who had
+to develop their own variants of classifiers to obtain
+their results.
 

### 2017-04-14 paper/oopsla17-submission/discussion.tex
shuffle

@@
-\section{Discussion}
+\section{Limitations and Threats to Validity}
 \label{sec:discussion}
+
+\subsection{Limitations}
+\label{sec:limitations}
 
@@
 
+\subsection{Threats to Validity}
+\label{sec:validity}
+
+Although our experiments demonstrate that our technique can pinpoint type
+errors more accurately than the state of the art and that our features are
+relevant to blame assignment, our results may not generalize.
+
+One threat to validity associated with supervised machine learning is
+overfitting (\ie learning a model that is too complex with respect to
+the data).
+%
+A similar issue that arises in machine learning is model stability (\ie
+can small changes to the training set produce large changes in the model?).
+%
+We mitigate these threats by:
+%
+(1) using separate training and testing datasets drawn from distinct
+student populations (\autoref{sec:quantitative}), demonstrating the
+generality of our models; and
+%
+(2) via cross-validation on the joint dataset
+(\autoref{sec:feature-utility}), which demonstrates the stability of our
+models by averaging the accuracy of 10 models trained on distinct
+subsets of the data.
+
+Our benchmarks were drawn from students in an undergraduate course and
+may not be representative of other student populations.
+%
+We mitigate this threat by including the largest empirical evaluation of
+type error localization that we are aware of: over 4,500 pairs of
+ill-typed programs and fixes from two instances of the course, with
+programs from 102 different students.
+%
+We acknowledge, of course, that students are not industrial programmers
+and our results may not translate to large-scale software development;
+however, we are particularly interested in aiding novice programmers
+as they learn to work inside the type system.
+
+A related threat to construct validity is our definition of the immedate
+next well-typed program as the intended ground truth answer (see
+\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
+intermediate well-typed program ``rewrites'' between the original ill-typed
+program and the final intended answer. Our approach to discarding outliers
+(see \autoref{sec:evaluation}) is designed to mitigate this threat.
+
+Our removal of program pairs that changed too much, where our oracle
+could not identify the blame of the other tools, or where the other
+tools timed out or encountered unsupported language features is another
+threat to validity.
+%
+It is possible that including the programs that changed excessively
+would hurt our models, or that the other tools would perform
+better on the programs with unsupported language features.
+%
+We note however that
+%
+(1) outlier removal is a standard technique in machine learning
+\ES{CITE?}; and
+%
+(2) our Top-1 accuracy margin is large enough that even if we assumed
+that \sherrloc were perfect on all of the excluded programs, it would
+only tie our \hiddenFH in Top-1 accuracy.
+%
+\ES{I think this is accurate, but should double check..}
+
+Examining programs written in \ocaml as opposed to \haskell or any other
+typed functional language poses yet another threat, common type errors
+may differ in different languages.
+%
+\ocaml is, however, a standard target for research in type error
+localization and thus our choice admits a direct comparison with prior
+work.
+%
+Furthermore, the functional core of \ocaml that we support does not
+differ significantly from the functional core of \haskell or SML, all of
+which are effectively lambda calculi with a Hindley-Milner-style type
+system.\footnote{\haskell's type classes are a notable exception, they
+  are also known to cause confusing type errors and would be interesting
+  to study in future work.}
+
+Finally, our use of student fixes as oracles for the source of type
+errors assumes that, on average, students are able to correctly identify
+the source of an error.
+%
+As the students are in the process of learning the language and type
+system, this assumption may be faulty.
+%
+It may be that \emph{expert} users would disagree with many of the
+student fixes, and that it is harder to learn a model of expert fixes,
+or that the state of the art would be better at predicting expert fixes.
+%
+As we have noted before, we believe it is reasonable to use student
+fixes as oracles because the student is the best judge of what she
+\emph{meant} to do.
+
 %%% Local Variables:

### 2017-04-14 paper/oopsla17-submission/overview.tex
shuffle

--- 2017-04-14 paper/oopsla17-submission/overview.tex
@@
 
+% \ES{FIXME: this feels a bit redundant with CHALLENGE 1. the main thing i
+%  get is that we use student fixes to extract AST-diffs, which is
+%  already said in CHALLENGE 1}
+
 \mypara{Example}
@@
 
+%% \ES{FIXME: this feels a bit redundant with CHALLENGE 2. the main things
+  %% i get here are that we use sets of feature vectors, and that we can
+  %% concatenate vectors to add context. but this is all said in CHALLENGE
+  %% 2 already.}
+
 %% \RJ{confusing -- delete}
@@
 signal is whether or not the
-expression occurs in a
-\emph{minimal type error slice}
+expression occurs in
+a minimal type error slice
 \citep{Tip2001-qp,Haack2003-vc}
-which includes the smallest
-subset of all expressions
-that are necessary for the
-error to manifest.
+which includes \emph{a minimal subset}
+of all expressions that are
+necessary for the error to manifest.
+%
+(That is, replacing any subterm
+with |undefined| or |assert false|
+would yield a well-typed program.)
 %
@@
 
+\input{feature-table}
+
 % Our use of type error slices is
@@
 
-\mypara{Solution: Rank Locations by ``Confidence''}
+\mypara{Solution: Rank Locations by Confidence}
 %

### 2017-04-14 paper/oopsla17-submission/evaluation.tex
done with overview
shuffle

--- 2017-04-13 paper/oopsla17-submission/evaluation.tex
@@
 \input{evaluation-utility}
-\input{gallery}
-
-\subsection{Threats to Validity}
-
-Although our experiments demonstrate that our technique can pinpoint type
-errors more accurately than the state of the art and that our features are
-relevant to blame assignment, our results may not generalize.
-
-One threat to validity associated with supervised machine learning is
-overfitting (\ie learning a model that is too complex with respect to
-the data).
-%
-We mitigate this threat by having separate training and testing datasets
-(\autoref{sec:quantitative}) and via cross-validation
-(\autoref{sec:feature-utility})~\cite{FIXME}.
-
-Our benchmarks were drawn from students in an undergraduate course and
-may not be representative of other student populations.
-%
-We mitigate this threat by including the largest empirical evaluation of
-type error localization that we are aware of: over 4,500 pairs of
-ill-typed programs and fixes from two instances of the course, with
-programs from 102 different students.
-%
-We acknowledge, of course, that students are not industrial programmers
-and our results may not translate to large-scale software development;
-however, we are particularly interested in aiding novice programmers
-as they learn to work inside the type system.
-
-A related threat to construct validity is our definition of the immedate
-next well-typed program as the intended ground truth answer (see
-\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
-intermediate well-typed program ``rewrites'' between the original ill-typed
-program and the final intended answer. Our approach to discarding outliers
-(see \autoref{sec:evaluation}) is designed to mitigate this threat.
-
-Our removal of program pairs that changed too much, where our oracle
-could not identify the blame of the other tools, or where the other
-tools timed out or encountered unsupported language features is another
-threat to validity.
-%
-It is possible that including the programs that changed excessively
-would hurt our models, or that the other tools would perform
-better on the programs with unsupported language features.
-%
-We note however that
-%
-(1) outlier removal is a standard technique in machine learning
-\ES{CITE?}; and
-%
-(2) our Top-1 accuracy margin is large enough that even if we assumed
-that \sherrloc were perfect on all of the excluded programs, it would
-only tie our \hiddenFH in Top-1 accuracy.
-%
-\ES{I think this is accurate, but should double check..}
-
-Examining programs written in \ocaml as opposed to \haskell or any other
-language poses another threat to validity, common type errors may differ
-in different languages.
-%
-\ocaml is, however, a standard target for research in type error
-localization and thus our choice admits a direct comparison with prior
-work.
-%
-Furthermore, the functional core of \ocaml that we support does not
-differ significantly from the functional core of \haskell or SML, all of
-which are effectively lambda calculi with a Hindley-Milner-style type
-system.\footnote{\haskell's type classes are a notable exception, they
-  are also known to cause confusing type errors and would be interesting
-  to study in future work.}
-
-Finally, our use of student fixes as oracles for the source of type
-errors assumes that, on average, students are able to correctly identify
-the source of an error.
-%
-As the students are in the process of learning the language and type
-system, this assumption may be faulty.
-%
-It may be that \emph{expert} users would disagree with many of the
-student fixes, and that it is harder to learn a model of expert fixes,
-or that the state of the art would be better at predicting expert fixes.
-%
-As we have noted before, we believe it is reasonable to use student
-fixes as oracles because the student is the best judge of what she
-\emph{meant} to do.
 

### 2017-04-14 paper/oopsla17-submission/qualitative.tex
shuffle

--- 2017-04-12 paper/oopsla17-submission/gallery.tex
@@
 %
+We also take this opportunity to examine some of the specific features
+our classifiers use to assign blame.
+%
 For each example, we provide
 %
-(1) the code,
+(1) the code;
 %
-(2) \hlSherrloc{\sherrloc's} prediction,
+(2) \hlSherrloc{\sherrloc's} prediction;
 %
-(3) our \hlTree{\dectree's} prediction, and
+(3) our \hlTree{\dectree's} prediction; and
 %
-(4) an \emph{explanation} of why our classifier made its prediction.
+(4) an \emph{explanation} of why our classifier made its prediction, in
+terms of the features used and their values.
 %
@@
 significantly to the model's accuracy.
+
+We explain the predictions by analyzing the paths induced in
+the decision tree by the features of the input expressions.
 %
-We explain the predictions by analyzing the paths induced in
-the decision tree by the features of the input programs.
+Recall that each node in a decision tree contains a simple predicate of
+the features, \eg ``is feature $v_j$ enabled?'', which determines
+whether a sample will continue down the left or right subtree.
+%
+Thus, we can examine the predicates used and the values of the
+corresponding features to explain \emph{why} our \dectree made its
+prediction.
+%
+We will focus particularly on the enabled features, as they generally
+provide more information than the disabled features.
+%
+Furthermore, each node is additionally labeled with the ratio of
+``blamed'' vs ``not-blamed'' training expressions that passed through
+it.
+%
+We can use this information to identify particularly important
+decisions, \ie we consider a decision that changes the ratio to be more
+interesting than a decision that does not.
+
 %

### 2017-04-14 paper/oopsla17-submission/evaluation.tex
asd

@@
-\section{Evaluation}
+\section{Quantitative Evaluation}
 \label{sec:evaluation}
@@
 
-\subsection{Methodology}
+\mysubsection{Methodology}
 \label{sec:methodology}
@@
 
-\paragraph{Feature Selection}
+\mypara{Feature Selection}
 We extract a set of 277 features from each sub-expression in a student
@@
 
-\paragraph{Blame Oracle}
+\mypara{Blame Oracle}
 Recall from \autoref{sec:labels} that we automatically extract a blame
@@
 
-\paragraph{Accuracy Metric}
+\mypara{Accuracy Metric}
 All of the tools we compare (with the exception of the standard \ocaml

### 2017-04-14 paper/oopsla17-submission/learning.tex
shuffle
asd

@@
 % \end{figure}
-
-\input{feature-table}
 
@@
 
-\paragraph{Logistic Regression.} The simplest classifier we investigate is logistic regression.
+\mypara{Logistic Regression.} The simplest classifier we investigate is logistic regression.
 This is a linear model, where the goal is to learn a set of weights $W$ and a bias $b$ that describe the following model for predicting a label $y$ from a feature vector $x$:
@@
 
-\paragraph{Decision Trees.} Decision tree algorithms learn a tree of binary predicates over the features, recursively partitioning the input space until a final
+\mypara{Decision Trees.} Decision tree algorithms learn a tree of binary predicates over the features, recursively partitioning the input space until a final
 classification can be made. Each node in the tree contains a single predicate of the form
@@
 
-\paragraph{Random Forests.} Random forests mitigate the generalization problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
+\mypara{Random Forests.} Random forests mitigate the generalization problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
 The agreement among the distinct trees in the ensemble forms a natural confidence score.
@@
 
-\paragraph{Neural Networks}
+\mypara{Neural Networks}
 The last (and most complex) model we use is a type of neural network

### 2017-04-14 paper/oopsla17-submission/conclusion.tex
shuffle
misc tweaks

@@
 Given a new ill-typed program,
-we extract its bag-of-extracted-terms
+we extract its bag-of-abstracted-terms
 to compute the blame label for each
@@
 on the model's confidence, to select
-the top 1-3 blame assignments to
+the top 1--3 blame assignments to
 present to the user.

### 2017-04-14 paper/oopsla17-submission/intro2.tex
misc tweaks

--- 2017-04-13 paper/oopsla17-submission/intro2.tex
@@
 Unfortunately, this automation has come at a price.
-Verbose type annotations (as in, \eg, \java) signify
+Type annotations signify
 the programmer's intent and help to correctly
@@
 The key here is to find the right set of feature
-abstractions and classification algorithms, that
+abstractions and classification algorithms that
 lead to precise predictions.

### 2017-04-14 paper/oopsla17-submission/learning.tex
misc tweaks

--- 2017-04-13 paper/oopsla17-submission/learning.tex
@@
-\section{Learning to Blame}
+\mysection{Learning to Blame}
 \label{sec:learning}
@@
 
-\subsection{Features}
+\mysubsection{Features}
 \label{sec:features}
@@
 
-\subsubsection{Local syntactic features}
+\mypara{Local syntactic features}
 These features describe the syntactic category of each expression $e$.
@@
 
-\subsubsection{Contextual syntactic features}
+\mypara{Contextual syntactic features}
 These are similar to local syntactic features, but lifted to describe the
@@
 This gives us a notion of the \emph{context} in which an expression
-occurs, similar to the \emph{n-grams} commonly used in linguistic
+occurs, similar to the \emph{n-grams} used in linguistic
 models \citep{Hindle2012-hf,Gabel2010-el}.
@@
 
-\subsubsection{Expression size}
+\mypara{Expression size}
 We also propose a feature representing the \emph{size} of each expression,
@@
 
-\subsubsection{Typing features}
+\mypara{Typing features}
 A natural way of summarizing the context in which an expression occurs
@@
 For example, the expressions |tl| and |sumList tl|
-in \autoref{tab:sumList} both enable the feature \HasTypeList, as they
-are both inferred to have a type that mentions the $\tlist{\cdot}$
-constructor.
+in \autoref{tab:sumList} both enable \HasTypeList, as they
+are both inferred to have a type that mentions $\tlist{\cdot}$.
+%constructor.
 
@@
 
-\subsubsection{Type error slice}
+\mypara{Type error slice}
 Finally, we wish to distinguish between changes that could fix the
@@
 
-\subsection{Labels}
+\mysubsection{Labels}
 \label{sec:labels}
@@
 
-\subsection{Learning Algorithms}
+\mysubsection{Learning Algorithms}
 \label{sec:models}
@@
 
-Recall that we formulate type error detection at a single expression as a supervised classification problem.
-This means that we are given a training data set $S = \{ (x_i, y_i) \}_{i=1}^n$ of labeled examples, and our goal is to use it to build a {\em{classifier}}, or a prediction rule that can predict a label for an input $x$.
-Since we apply the classifier on each expression in the program to determine those that are the most likely to be type errors, we also require the classifier to output a {\em{confidence score}} that measures how sure the classifier is about its prediction.
+Recall that we formulate type error detection at a single expression as
+a supervised classification problem.
+%
+This means that we are given a training data set
+$S = \{ \langle v_i, b_i \rangle \}_{i=1}^n$ of labeled examples, and
+our goal is to use it to build a \emph{classifier}, \ie a rule
+that can predict a label $b$ for an input $v$.
+%
+Since we apply the classifier on each expression in the program to
+determine those that are the most likely to be type errors, we also
+require the classifier to output a \emph{confidence score} that measures
+how sure the classifier is about its prediction.
 
@@
 
-There are many learning algorithms to choose from, existing
-on a spectrum that balances expressiveness with ease of training (and of
+There are many learning algorithms to choose from, existing on a
+spectrum that balances expressiveness with ease of training (and of
 interpreting the learned model).
-In this section we consider four standard learning algorithms: (1) logistic regression, (2) decision trees, (3) random forests, and (4) neural networks.
+%
+In this section we consider four standard learning algorithms: (1)
+logistic regression, (2) decision trees, (3) random forests, and (4)
+neural networks.
+%
 A thorough introduction to these techniques can be found in introductory
-machine learning textbooks (\eg \cite{FriedmanHastieTibshirani2000}).
+machine learning textbooks \citep[\eg][]{FriedmanHastieTibshirani2000}.
 % \ES{TODO: huma/kamalika?}).
@@
 
-Below we briefly introduce each technique by describing the model it learns, and summarize its advantages and disadvantages.
-For our application, we are particularly interested in three properties -- expressiveness, interpretability and ease of generalization.
-Expressiveness measures how complex prediction rules are allowed to be, and interpretability measures how easy it is to explain the cause of prediction to a human.
-Finally ease of generalization measures how easily the rule generalizes to unseen examples that are not in the training set; a rule that is not easily generalizable might perform poorly on an unseen test set even when its training performance is high.
+Below we briefly introduce each technique by describing the rules it
+learns, and summarize its advantages and disadvantages.
+%
+For our application, we are particularly interested in three properties
+-- expressiveness, interpretability and ease of generalization.
+%
+Expressiveness measures how complex prediction rules are allowed to be,
+and interpretability measures how easy it is to explain the cause of
+prediction to a human.
+%
+Finally ease of generalization measures how easily the rule generalizes
+to examples that are not in the training set; a rule that is not
+easily generalizable might perform poorly on an unseen test set even
+when its training performance is high.
 
@@
 
-\ES{FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'',
-contains much more detail than an OOPSLA audience wants on
-introductory/undergraduate machine learning. (Especially since nothing
-after this point is a novel contribution we make.) Perhaps we can cut some
-space by focusing on the pros and cons of each technique rather than
-describing the details?
+% \ES{FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'',
+% contains much more detail than an OOPSLA audience wants on
+% introductory/undergraduate machine learning. (Especially since nothing
+% after this point is a novel contribution we make.) Perhaps we can cut some
+% space by focusing on the pros and cons of each technique rather than
+% describing the details?
 This is a PL paper.
@@
 We can rate the four classifiers in these three areas.
-}
+% }
 
-\paragraph{Logistic Regression.} The simplest classifier we investigate is logistic regression.
-This is a linear model, where the goal is to learn a set of weights $W$ and a bias $b$ that describe the following model for predicting a label $y$ from a feature vector $x$:
-\[ \Pr(y = 1 | x) = \frac{1}{1 + e^{-W^{\top} x - b}} \]
-The weights $W$ and $b$ are learnt from training data, and the value of $\Pr(y | x)$ naturally leads to a confidence score.
-Logistic regression is a widely used classification algorithm, which is largely preferred for its simplicity, ease of generalization as well as interpretability.
-Its main limitation is that the prediction rule is constrained to be a linear combination of the features, and hence relatively simple.
-While this can be somewhat mitigated by adding higher order (quadratic or cubic) features, this often requires substantial domain knowledge.
+\mypara{Logistic Regression}
+The simplest classifier we investigate is logistic regression:
+a linear model where the goal is to learn a set of weights $W$
+that describe the following model for predicting a label
+$b$ from a feature vector $v$:
+%
+\[ \Pr(b = 1 | v) = \frac{1}{1 + e^{-W^{\top} v}} \]
+%
+The weights $W$ are learnt from training data, and the value of
+$\Pr(b | v)$ naturally leads to a confidence score $\Runit$.
+%
+Logistic regression is a widely used classification algorithm, preferred
+for its simplicity, ease of generalization, and interpretability.
+%
+Its main limitation is that the prediction rule is constrained to be a
+linear combination of the features, and hence relatively simple.
+%
+While this can be somewhat mitigated by adding higher order (quadratic
+or cubic) features, this often requires substantial domain knowledge.
 
@@
 
-\paragraph{Decision Trees.} Decision tree algorithms learn a tree of binary predicates over the features, recursively partitioning the input space until a final
-classification can be made. Each node in the tree contains a single predicate of the form
+\mypara{Decision Trees}
+Decision tree algorithms learn a tree of binary predicates over the
+features, recursively partitioning the input space until a final
+classification can be made.
+%
+Each node in the tree contains a single predicate of the form
 $v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
 whether a given input should proceed down the left or right subtree.
+%
 Each leaf is labeled with a prediction and the fraction of training
-samples that would reach it, the latter quantity can be interpreted as
-\ES{HUMA: is this correct?}
-the decision tree's confidence in its prediction. This leads to a prediction rule
-that can be quite expressive depending on the data used to build it.
+samples that would reach it; the latter quantity can be interpreted as
+the decision tree's confidence in its prediction.
+%
+This leads to a prediction rule that can be quite expressive depending
+on the data used to build it.
 
-Training a decision tree entails finding both a set of good partitioning and a good ordering of the predicates based on data.
-This is usually done in a top-down greedy manner, and there are several standard training algorithms such as C4.5
-\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
+Training a decision tree entails finding both a set of good partitioning
+predicates and a good ordering of the predicates based on data.
+%
+This is usually done in a top-down greedy manner, and there are several
+standard training algorithms such as C4.5 \citep{Quinlan1993-de} and
+CART \citep{Breiman1984-qy}.
 
-Another advantage of decision trees is their ease of interpretation --  the decision rule is a white-box model that can be readily described to a human, especially when the tree is small.
-However, main limitation is that these trees often do not generalize well; while this is somewhat mitigated in practice by pruning the tree, this is still a well-known problem.
+Another advantage of decision trees is their ease of interpretation ---
+the decision rule is a white-box model that can be readily described to
+a human, especially when the tree is small.
+%
+However, the main limitation is that these trees often do not generalize
+well, though this can be somewhat mitigated by \emph{pruning} the tree.
+% this is still a well-known problem.
 
@@
 
-\paragraph{Random Forests.} Random forests mitigate the generalization problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction.
-The agreement among the distinct trees in the ensemble forms a natural confidence score.
-Since each classifier in the ensemble is a decision tree, this still allows for complex and highly expressive classifiers.
-\ES{Huma: is this the confidence score that we are using?}
+\mypara{Random Forests}
+%
+Random forests mitigate the generalization problem by training an
+\emph{ensemble} of distinct decision trees and then using a majority
+vote to make a prediction.
+%
+The agreement among the distinct trees in the ensemble forms a natural
+confidence score.
+%
+Since each classifier in the ensemble is a decision tree, this still
+allows for complex and highly expressive classifiers.
+%\ES{Huma: is this the confidence score that we are using?}
 
@@
 training process for the decision trees is often modified slightly to
-reduce correlation between trees, by forcing the tree to pick features
+reduce correlation between trees, by forcing each tree to pick features
 from a random subset of all features at each node.
@@
 
-\paragraph{Neural Networks}
+\mypara{Neural Networks}
+%
 The last (and most complex) model we use is a type of neural network
@@
 an introduction to neural networks).
+%
+A multi-layer perceptron can be represented as a directed acyclic
+graph whose nodes are arranged in layers that are fully connected by
+weighted edges.
+%
+The first layer corresponds to the input features, and the final to the
+output.
+%
+%The output of a node $v$ in an internal layer is given by:
+The output of an internal node $v$ is
+%
+\[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
+%
+where $N(v)$ is the set of nodes in the previous layer that are adjacent
+to $v$, $W_{jv}$ is the weight of the $(j, v)$ edge and $h_j$ is the
+output of node $j$ in the previous layer.
+%
+Finally $g$ is a non-linear function, called the activation function,
+which in recent work is commonly chosen to be the \emph{rectified linear
+  unit} (ReLU), defined as $g(x) = \mathsf{max}(0,x)$
+\citep{Nair2010-xg}.
+%
+The number of layers, the number of neurons per layer, and the
+connections between layers constitute the \emph{architecture} of a
+neural network.
+%
+In this work, we use relatively simple neural networks which have an
+input layer, a single hidden layer and an output layer.
 
-A multi-layered perceptron can be represented as a directed acyclic graph whose nodes are arranged in layers that are fully connected by weighted edges.
-The first layer corresponds to the input features, and the final to the output.
-The output of a node $v$ in an internal layer is given by:
-\[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
-where $N(v)$ is the set of nodes in the previous layer that are adjacent to $v$, $W_{jv}$ is the weight of the $(j, v)$ edge and $h_j$ is the output of node $j$ in the previous layer.
-Finally $g$ is a non-linear function, called the activation function, which in recent work is commonly chosen to be the \emph{rectified linear unit} (ReLU),
-defined as $g(x) = \mathsf{max}(0,x)$ \citep{Nair2010-xg}.
-The number of layers, the number of neurons per layer, and the connections between layers constitute the \emph{architecture} of a neural network.
-In this work, we use relatively simple neural networks which have an input layer, a single hidden layer and an output layer.
-
-A major advantage of neural networks is its ability to discover interesting combinations of features through non-linearity, which significantly reduces the need for manual feature engineering, and allows high expressivity.
-On the other hand, this makes the networks particularly difficult to interpret and also difficult to generalize under vast amounts of training data are available.
+A major advantage of neural networks is their ability to discover
+interesting combinations of features through non-linearity, which
+significantly reduces the need for manual feature engineering, and
+allows high expressivity.
+%
+On the other hand, this makes the networks particularly difficult to
+interpret and also difficult to generalize under vast amounts of
+training data are available.
 

### 2017-04-14 paper/oopsla17-submission/discussion.tex

@@
-\section{Limitations and Threats to Validity}
+\mysection{Limitations and Threats to Validity}
 \label{sec:discussion}
-
-\subsection{Limitations}
-\label{sec:limitations}
 
@@
 
-\paragraph{User-Defined Types}
+\mysubsection{Limitations}
+\label{sec:limitations}
+
+\mypara{User-Defined Types}
 Probably the single biggest limitation of our technique is that we have
@@
 
-\paragraph{Additional Features}
+\mypara{Additional Features}
 There are a number of other features that could improve the model's
@@
 
-\paragraph{Independent vs Joint Predictions}
+\mypara{Independent vs Joint Predictions}
 We treat each sub-expression as if it exists in a vacuum, but in reality
@@
 
-\subsection{Threats to Validity}
+\mysubsection{Threats to Validity}
 \label{sec:validity}

### 2017-04-14 paper/oopsla17-submission/evaluation-accuracy.tex

@@
-\subsection{Blame Accuracy}
+\mysubsection{Blame Accuracy}
 \label{sec:quantitative}
@@
 
-\paragraph{Baseline}
+\mypara{Baseline}
 We provide two baselines for the comparison: a random choice of location
@@
 
-\paragraph{State of the Art}
+\mypara{State of the Art}
 \mycroft~\citep{Loncaric2016-uk} localizes type errors by searching for
@@
 
-\paragraph{Our Classifiers}
+\mypara{Our Classifiers}
 We evaluate five classifiers, each trained on the full set of features.
@@
 
-\paragraph{Results}
+\mypara{Results}
 \autoref{fig:accuracy-results} shows the results of our experiment.

### 2017-04-14 paper/oopsla17-submission/qualitative.tex

@@
-\subsection{Explaining Predictions}
+\mysection{Qualitative Evaluation}
 \label{sec:qualitative}
 
-Next, we present a \emph{qualitative} evaluation that compares the
-predictions made by our classifiers with those of \sherrloc.
+Next, we present a \emph{qualitative} evaluation
+that compares the predictions made by our classifiers
+with those of \sherrloc.
 %
@@
 
-\subsubsection{Failed Predictions}
+\mysubsection{Failed Predictions}
 \label{sec:failed-predictions}
+
 We begin with a few programs where our classifier fails to
@@
 
-\paragraph{Constructing a List of Duplicates}
+\mypara{Constructing a List of Duplicates}
 % data/sp14/0219.ml
@@
 
-\paragraph{Currying Considered Harmful?}
+\mypara{Currying Considered Harmful?}
 % data/sp14/1204.ml
@@
 
-% \paragraph{Composing Functions}
+% \mypara{Composing Functions}
 % % data/sp14/3041.ml
@@
 
-\subsubsection{Correct Predictions}
+\mysubsection{Correct Predictions}
 \label{sec:correct-predictions}
+
 Next, we present a few indicative programs where our first prediction is
@@
 
-\paragraph{Extracting the Digits of an Integer}
+\mypara{Extracting the Digits of an Integer}
 % data/sp14/2176.ml
@@
 
-\paragraph{Padding a list}
+\mypara{Padding a list}
 % data/sp14/0442.ml
@@
 % \ES{I don't much like this example anymore..}
-% \paragraph{Computing the Fixed Point of a Function}
+% \mypara{Computing the Fixed Point of a Function}
 % % data/sp14/1266.ml

### 2017-04-14 paper/oopsla17-submission/related.tex

@@
-\section{Related Work}
+\mysection{Related Work}
 \label{sec:related-work}
@@
 
-\subsection{Type Error Diagnosis}
+\mysubsection{Type Error Diagnosis}
 \label{sec:type-error-diagnosis}
@@
 
-\paragraph{Localizing Type Errors}
+\mypara{Localizing Type Errors}
 It is well-known that the original Damas-Milner algorithm $\mathcal{W}$
@@
 
+\mypara{Slicing}
 Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
@@
 the input program.
-%
+
+\mypara{Finding Likely Errors}
 Thus, recent work has focused on finding the \emph{most likely} source
@@
 
-\paragraph{Explaining Type Errors}
+\mypara{Explaining Type Errors}
 In this paper we have focused solely on the task of \emph{localizing} a
@@
 
+\mypara{Interactive Explanations}
 Such \emph{static} explanations of type errors run the risk of
@@
 
+\mypara{Programmatic Explanations}
+%
 The best explanation of a type error, however, might be given by an
@@
 
-\paragraph{Fixing Type Errors}
+\mypara{Fixing Type Errors}
 Some techniques go beyond explaining a type error or locating its
@@
 
-\subsection{Fault Localization}
+\mysubsection{Fault Localization}
 Given a defect, \emph{fault localization} is the task of identifying

### 2017-04-14 paper/oopsla17-submission/learning.tex
bad orphans

@@
 
-\input{feature-table}
-
 The first issue we must tackle is formulating our learning task in
@@
 %
-We are given programs over \lang, but the learning algorithms expect to work
+We are given programs over \lang, but learning algorithms expect to work
 with \emph{feature vectors} $\V$ --- vectors of real numbers, where each

### 2017-04-14 paper/oopsla17-submission/overview.tex
misc tweaks
bad orphans

@@
 %
-Indeed, both the venerable
-\ocaml compiler as well as
+Indeed, % both the venerable
+% \ocaml compiler as well as
 recent techniques like
@@
 \mypara{Example}
-For example, suppose our student
+Suppose our student
 fixed the |sumList| program in
@@
 at least one matches an actual student fix
-more than 90\% of the time.
+over 90\% of the time.
 

### 2017-04-14 paper/oopsla17-submission/abstract.tex
minor

@@
 precision to the state of the art
-on a set of over 4,500 ill-typed Ocaml
+on a set of over 4,500 ill-typed \ocaml
 programs drawn from introductory

### 2017-04-14 paper/oopsla17-submission/evaluation-utility-graph.tex
compress graphs a bit

--- 2017-04-10 paper/oopsla17-submission/evaluation-utility-graph.tex
@@
   width=0.5\linewidth,
-  height=5cm,
+  height=4cm,
   %title={Impact of Type Error Slice},
@@
 \addlegendentry{\linear}
-
+\end{axis}
+\begin{axis}[
+  ybar stacked,
+  width=0.5\linewidth,
+  height=4cm,
+  ylabel={Recall},
+  axis y line*=right,
+  ymin=0.2,
+  ymax=1,
+  ytick={0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = false,
+  xmin=0, xmax=1,
+  hide x axis,
+]
 \end{axis}
@@
   ybar stacked,
-  width=\linewidth,
-  height=6cm,
+  width=0.9\linewidth,
+  height=5cm,
   %title={Impact of Contextual Feature Classes},
@@
 \end{axis}
+\begin{axis}[
+  ybar stacked,
+  width=0.9\linewidth,
+  height=5cm,
+  ylabel={Recall},
+  axis y line*=right,
+  ymin=0.5,
+  ymax=1,
+  ytick={0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
+  ytick style={draw=none},
+  ymajorgrids = false,
+  xmin=0, xmax=1,
+  hide x axis,
+]
+\end{axis}
 

### 2017-04-15 paper/oopsla17-submission/discussion.tex
trimming

@@
 syntactic features for data constructors and match expressions, but it
-remains to be seen how much these help \ES{TODO: UW data!}.
+remains to be seen how much these help. % \ES{TODO: UW data!}.
 %
@@
 As an alternative to encoding information about \emph{specific}
-constructors, we might investigate a more abstract representation.
+constructors, we might use a more abstract representation.
 %

### 2017-04-15 paper/oopsla17-submission/evaluation-accuracy.tex
trimming

@@
 Our classifiers consistently outperform the competition, ranging from
-58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear classifier to
-71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
+60\% Top-1 accuracy (87\% Top-3) for the \linear classifier to
+72\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
 %
@@
 \sherrloc outperforms both baselines, and comes close to our \linear
-classifier, with 56\% Top-1 accuracy (84--86\% Top 3), while \mycroft
-actually underperforms \ocaml with 38--41\% Top-1 accuracy.
+classifier, with 56\% Top-1 accuracy (85\% Top 3), while \mycroft
+underperforms \ocaml at 39\% Top-1 accuracy.
 %

### 2017-04-15 paper/oopsla17-submission/evaluation-utility.tex
asd
trimming

@@
-\subsection{Feature Utility}
+\mysubsection{Feature Utility}
 \label{sec:feature-utility}
@@
 We focus specifically on feature \emph{classes} rather than individual
-features as our 276 features are conceptually grouped into a much
+features as our 277 features are conceptually grouped into a much
 smaller number of \emph{categorical} features.
@@
 
-\subsubsection{Type Error Slice}
+\mysubsubsection{Type Error Slice}
 \label{sec:type-error-slice}
@@
 
-% FIXME: Wes feels that there should be a sentence in the next paragraph
+% FIXME: Wes feels that there should be a sentence in the next mypara
 % explaining to the reader why we didn't just use an ANOVA or the ReliefF
@@
 classifiers, cross-validated ($k=10$) over the combined SP14/FA15
-dataset. We used a learning rate $\eta=0.001$, $L_2$ regularization rate
-$\lambda=0.001$, and mini-batch size of 200. We trained for a single
-epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
-total number of training samples would be roughly equal for each feature
-set.
+dataset.
+% We used a learning rate $\eta=0.001$, $L_2$ regularization rate
+%$\lambda=0.001$, and mini-batch size of 200.
+%
+We trained for a single epoch on feature sets (1) and (2), and for 8
+epochs on (3), so that the total number of training samples would be
+roughly equal for each feature set.
 %
 \lstDeleteShortInline{|} % sigh...
-In addition to the accuracy metric from before, we report each
+In addition to accuracy, we report each
 classifier's \emph{recall} --- \ie ``How many true changes can we
@@
 $$
-where $\mathsf{predicted}$ is limited to the top 3 predictions of the
-classier, and $\mathsf{oracle}$ is the student's fix, limited to changes
-that part of the type error slice.
+where $\mathsf{predicted}$ is limited to the top 3 predictions, and
+$\mathsf{oracle}$ is the student's fix, limited to changes that are in
+the type error slice.
 %
-We make the latter distinction as (1) changes that are not part of the
-type error slice are noise in the data set, and (2) it makes the
-comparison between feature sets easier to interpret since
-$\mathsf{oracle}$ never changes.
+We make the latter distinction as:
+%
+(1) changes that are not part of the type error slice are noise in the
+data set; and
+%
+(2) it makes the comparison easier to interpret since $\mathsf{oracle}$
+never changes.
 % NOTE: keep this at the end of the para or it screws up spacing...
@@
 
-\paragraph{Results}
+\mypara{Results}
 \autoref{fig:slice-utility} shows the results of our experiment.
@@
 error slice.
-
+%
 Still, the accuracy gain comes at zero cost, and given the other benefits
-of filtering out expressions that do not belong to the type error slice
+of filtering by \InSlice % out expressions that do not belong to the type error slice
 --- shrinking the search space and guaranteeing our predictions are actionable ---
@@
 
-\subsubsection{Contextual Features}
+\mysubsubsection{Contextual Features}
 \label{sec:contextual-features}
@@
 %
-As before, we perform a 10-fold cross-validation with $\eta = 0.001$,
-$\lambda = 0.001$, and a mini-batch size of 200, but we
-train for a full 20 epochs to make the differences more apparent.
+As before, we perform a 10-fold cross-validation,
+% with $\eta = 0.001$,
+% $\lambda = 0.001$, and a mini-batch size of 200
+but we train for a full 20 epochs to make the differences more apparent.
 %
@@
 %
-\paragraph{Results}
+\mypara{Results}
 \autoref{fig:context-utility} summarizes the results of this experiment.
@@
 They improve Top-1 accuracy by an additional 2\%, and are much more
-compact --- requiring only have 55 typing features compared to 176
+compact --- requiring only 55 typing features compared to 176
 contextual syntactic features.

### 2017-04-15 paper/oopsla17-submission/learning.tex
add Hastie2009 reference
trimming

@@
 ill-typed program $e$, and returns an ordered list of subexpressions
-from $e$ paired with the probability $\Runit$ that they should be
+from $e$ paired with the confidence score $\Runit$ that they should be
 blamed.
@@
 %
-We consider two sources of changes:
+We consider two sources of changes.
 %
-\begin{enumerate}
-\item If an expression has been removed wholesale, \eg if $\eapp{f}{x}$
+%\begin{enumerate}
+%\item 
+First, if an expression has been removed wholesale, \eg if $\eapp{f}{x}$
   is rewritten to $\eapp{g}{x}$, we will mark the expression $f$ as
   changed, as it has been replaced by $g$.
-\item If a new expression has been inserted \emph{around} an existing
+Second, if a new expression has been inserted \emph{around} an existing
   expression, \eg if $\eapp{f}{x}$ is rewritten to
@@
   occupies the original location of the application.
-\end{enumerate}
+%\end{enumerate}
 
@@
 A thorough introduction to these techniques can be found in introductory
-machine learning textbooks \citep[\eg][]{FriedmanHastieTibshirani2000}.
+machine learning textbooks \citep[\eg][]{Hastie2009-bn}.
 % \ES{TODO: huma/kamalika?}).
@@
 %
-Random forests mitigate the generalization problem by training an
-\emph{ensemble} of distinct decision trees and then using a majority
+Random forests improve generalization by training an
+\emph{ensemble} of distinct decision trees and using a majority
 vote to make a prediction.
 %
-The agreement among the distinct trees in the ensemble forms a natural
+The agreement among the trees forms a natural
 confidence score.
@@
 Since each classifier in the ensemble is a decision tree, this still
-allows for complex and highly expressive classifiers.
+allows for complex and expressive classifiers.
 %\ES{Huma: is this the confidence score that we are using?}
@@
 On the other hand, this makes the networks particularly difficult to
-interpret and also difficult to generalize under vast amounts of
+interpret and also difficult to generalize unless vast amounts of
 training data are available.

### 2017-04-15 paper/oopsla17-submission/discussion.tex
more trimming

@@
 %
-For example, each occurrence of a variable knows only its type and the
-immediate context in which it occurs, but it may be helpful to know
+For example, each occurrence of a variable knows only its type and its
+immediate neighbors, but it may be helpful to know
 about \emph{other} occurrences of the same variable.
 %
-If a variable is generally used in the context of a |float| but has a
-single occurrence in the context of an |int|, it seems likely that the
+If a variable is generally used as a |float| but has a
+single use as an |int|, it seems likely that the
 latter occurrence (or context) is to blame.
@@
 Similarly, arguments to a function application are not aware of the
-typing constraints imposed on them by the function (and vice versa),
-they only know that they are occurring in the context of an application
-and the type of the application itself.
+constraints imposed on them by the function (and vice versa),
+they only know that they are occurring in the context of an application.
+%and the type of the application itself.
 %
@@
   are also known to cause confusing type errors and would be interesting
-  to study in future work.}
+  to study as well.}
 

### 2017-04-15 paper/oopsla17-submission/qualitative.tex
more trimming

@@
 %
-In particular, neither prediction references a property of the current
-expression, only its context, which is surprising.
+In particular, both predictions only reference the expression's context,
+which is surprising.
 %
@@
 %
-The model appears to have learned that programmers typically call the
+The model seems to have learned that programmers typically call the
 correct function.

### 2017-04-15 paper/oopsla17-submission/discussion.tex
even more pruning

@@
-\mysection{Limitations and Threats to Validity}
+\mysection{Limitations}
 \label{sec:discussion}
@@
 
-\mysubsection{Limitations}
-\label{sec:limitations}
+% \mysubsection{Limitations}
+% \label{sec:limitations}
 
@@
 
-\mysubsection{Threats to Validity}
-\label{sec:validity}
-
-Although our experiments demonstrate that our technique can pinpoint type
-errors more accurately than the state of the art and that our features are
-relevant to blame assignment, our results may not generalize.
-
-One threat to validity associated with supervised machine learning is
-overfitting (\ie learning a model that is too complex with respect to
-the data).
-%
-A similar issue that arises in machine learning is model stability (\ie
-can small changes to the training set produce large changes in the model?).
-%
-We mitigate these threats by:
-%
-(1) using separate training and testing datasets drawn from distinct
-student populations (\autoref{sec:quantitative}), demonstrating the
-generality of our models; and
-%
-(2) via cross-validation on the joint dataset
-(\autoref{sec:feature-utility}), which demonstrates the stability of our
-models by averaging the accuracy of 10 models trained on distinct
-subsets of the data.
-
-Our benchmarks were drawn from students in an undergraduate course and
-may not be representative of other student populations.
-%
-We mitigate this threat by including the largest empirical evaluation of
-type error localization that we are aware of: over 4,500 pairs of
-ill-typed programs and fixes from two instances of the course, with
-programs from 102 different students.
-%
-We acknowledge, of course, that students are not industrial programmers
-and our results may not translate to large-scale software development;
-however, we are particularly interested in aiding novice programmers
-as they learn to work inside the type system.
-
-A related threat to construct validity is our definition of the immedate
-next well-typed program as the intended ground truth answer (see
-\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
-intermediate well-typed program ``rewrites'' between the original ill-typed
-program and the final intended answer. Our approach to discarding outliers
-(see \autoref{sec:evaluation}) is designed to mitigate this threat.
-
-Our removal of program pairs that changed too much, where our oracle
-could not identify the blame of the other tools, or where the other
-tools timed out or encountered unsupported language features is another
-threat to validity.
-%
-It is possible that including the programs that changed excessively
-would hurt our models, or that the other tools would perform
-better on the programs with unsupported language features.
-%
-We note however that
-%
-(1) outlier removal is a standard technique in machine learning
-\ES{CITE?}; and
-%
-(2) our Top-1 accuracy margin is large enough that even if we assumed
-that \sherrloc were perfect on all of the excluded programs, it would
-only tie our \hiddenFH in Top-1 accuracy.
-%
-\ES{I think this is accurate, but should double check..}
-
-Examining programs written in \ocaml as opposed to \haskell or any other
-typed functional language poses yet another threat, common type errors
-may differ in different languages.
-%
-\ocaml is, however, a standard target for research in type error
-localization and thus our choice admits a direct comparison with prior
-work.
-%
-Furthermore, the functional core of \ocaml that we support does not
-differ significantly from the functional core of \haskell or SML, all of
-which are effectively lambda calculi with a Hindley-Milner-style type
-system.\footnote{\haskell's type classes are a notable exception, they
-  are also known to cause confusing type errors and would be interesting
-  to study as well.}
-
-Finally, our use of student fixes as oracles for the source of type
-errors assumes that, on average, students are able to correctly identify
-the source of an error.
-%
-As the students are in the process of learning the language and type
-system, this assumption may be faulty.
-%
-It may be that \emph{expert} users would disagree with many of the
-student fixes, and that it is harder to learn a model of expert fixes,
-or that the state of the art would be better at predicting expert fixes.
-%
-As we have noted before, we believe it is reasonable to use student
-fixes as oracles because the student is the best judge of what she
-\emph{meant} to do.
-
 %%% Local Variables:

### 2017-04-15 paper/oopsla17-submission/evaluation.tex
even more pruning

@@
   floating point), equality and comparison operators, character and
-  string literals, and a user-defined |expr| type of simple arithmetic
+  string literals, and a user-defined % |expr| type of simple
+  arithmetic
   expressions. We discuss the challenge of supporting other
-  user-defined types in \autoref{sec:discussion}.
+  % user-defined
+  types in \autoref{sec:discussion}.
 \item 176 contextual syntactic features. For each sub-expression we
@@
   mentioned above. These features are extracted for each sub-expression
-  and for the contextual sub-expressions.
+  and its context. % for the contextual sub-expressions.
 \item One feature denoting the size of each sub-expression.

### 2017-04-15 paper/oopsla17-submission/learning.tex
even more pruning

@@
 supervised learning on a training set of feature vectors $\V$ and
-(boolean) labels $\B$.
+%(boolean)
+labels $\B$.
 %
-Once trained, a $\Model$ can be $\evalsym$uated on a new input vector,
-producing the probability $\Runit$ that the blame label should be
-applied.
+Once trained, we can $\evalsym$uate a $\Model$ on a new input vector,
+producing the confidence $\Runit$ that the blame label should apply. % be
+%applied.
 %
-We describe multiple types of $\Model$ and their instantiations of
-$\trainsym$ and $\evalsym$ in \autoref{sec:models}.
+We describe multiple $\Model$s and their instantiations of
+$\trainsym$ and $\evalsym$ (\autoref{sec:models}).
 
@@
 %
-We accomplish this by modeling features as real-valued functions of
-expressions, and extract a feature vector for each \emph{subexpression}
+We model features as real-valued functions of
+terms, and extract a feature vector for each \emph{subterm}
 of the ill-typed program (\autoref{sec:features}).

### 2017-04-15 paper/oopsla17-submission/related.tex
even more pruning

@@
 Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
-overcomes the traversal bias by extracting a complete and minimal subset
-of program expressions that contribute to the type error, \ie all of the
-expressions that are required for the error to manifest and no more.
+overcomes this bias by extracting a complete and minimal subset
+of terms that contribute to the error, \ie all of the
+terms that are required for it to manifest and no more.
 %
-Slicing techniques typically require rewriting the type checker with a
+Slicing typically requires rewriting the type checker with a
 specialized constraint language and solver, though
-\citet{Schilling2011-yf} shows how to turn any type checker into an
-error slicer by treating it as a black-box.
-
+\citet{Schilling2011-yf} shows how to turn any type checker into a
+slicer by treating it as a black-box.
+%
 While slicing techniques guarantee enough information to diagnose the
 error, they can fall into the trap of providing \emph{too much}
-information, producing a slice that is not significantly smaller than
-the input program.
+information, producing a slice that is not much smaller than
+the input. %original. % input program.
 
@@
 \citet{Pavlinovic2014-mr} translate the localization problem into a
-MaxSMT problem, asking an off-the-shelf SMT solver to find the smallest
+MaxSMT problem, asking an off-the-shelf solver to find the smallest
 set of constraints that can be removed such that the resulting system is
@@
 \citeauthor{Pavlinovic2014-mr} by reusing the existing type checker as
-an optimized theory solver in the Nelson-Oppen~\citeyear{Nelson1979-td}
+a theory solver in the Nelson-Oppen~\citeyear{Nelson1979-td}
 style, and thus require only a MaxSAT solver.
@@
 All three of these techniques support \emph{weighted} constraints to
-incorporate knowledge about the relative frequency of different errors,
+incorporate knowledge about the frequency of different errors,
 but only \citeauthor{Pavlinovic2014-mr} use the weights, setting them to
-the size of the sub-expression that induced the constraint.
+the size of the term that induced the constraint.
 %
@@
 attempt to explain type errors by collecting the chain of inferences
-made by the type checker --- essentially the typing derivation --- and
-presenting them to the user.
+made by the type checker %--- essentially the typing derivation ---
+and presenting them to the user.
 %
-Such explanations can be lengthy, as an attempt to compress the
-explanation \citet{Yang2000-kz} presents a visualization of the
-inference process.
+% Such explanations can be lengthy, as an attempt to compress the
+% explanation \citet{Yang2000-kz} presents a visualization of the
+% inference process.
 %
-\citet{Gast2004-zd} produces a type error slice enhanced by arrows
-showing the flow of data from sources with incompatible types to a
+\citet{Gast2004-zd} produces a slice enhanced by arrows
+showing the dataflow from sources with different types to a
 shared sink, borrowing the insight of dataflows-as-explanations from
@@
 \citet{Hage2006-hc} catalog a set of heuristics for
-improving the quality of error messages by examining the errors made by
+improving the quality of error messages by examining errors made by
 novices.
@@
 \citet{Serrano2016-oo} extend the ability to customize error messages to
-library authors, enabling \emph{domain-specific} type errors.
+library authors, enabling \emph{domain-specific} errors.
 %
@@
 Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
-incorporates some of these ideas, allowing library authors to supply
+incorporates these ideas, allowing library authors to supply
 custom errors when type-class resolution or type-family reduction fail,
@@
 \mypara{Fixing Type Errors}
-Some techniques go beyond explaining a type error or locating its
-source, and actually attempt to \emph{fix} the error automatically.
+Some techniques go beyond explaining or locating a type error,
+and actually attempt to \emph{fix} the error automatically.
 %
-\citet{Lerner2007-dt} searches for fixes to type errors by enumerating a
+\citet{Lerner2007-dt} searches for fixes by enumerating a
 set of local mutations to the program and querying the type checker to
@@
 \citet{Chen2014-gd} use a notion of \emph{variation-based} typing to
-track choices made by the type checker and enumerate potential type (and
-expression) changes that would fix the error.
+track choices made by the type checker and enumerate potential
+% type (and expression)
+changes that would fix the error.
 %
@@
 ``suspicious'' program elements (\eg lines, statements) that are likely
-implicated in the defect (\ie that should be changed to fix the defect)
+implicated in the defect %(\ie that should be changed to fix the defect)
 %

### 2017-04-15 paper/oopsla17-submission/conclusion.tex
20 pages

@@
-\section{Conclusion}
+\mysection{Conclusion}
 \label{sec:conclusion}
@@
 As with other forms of machine learning,
-a key concern is that of \emph{data-set bias}: are
-the models learned by \toolname specific
-to our data set, and would they fail on
-\emph{other} \ocaml programs?
+a key concern is that of \emph{bias}: are
+\toolname's models specific
+to our data set, would they fail on
+\emph{other} programs?
 %
@@
 (specifically, novices who are in greatest
-need of precise feedback) actually make
+need of precise feedback) actually make,
 rather than hardwiring the biases of
-compiler authors who, by dint of their
-training and experience, may suffer from
-blind spots with regards to such problems.
+experts who % , by dint of their
+% training and experience,
+may suffer from blind spots. % with regards to such problems.
 %

### 2017-04-15 paper/oopsla17-submission/qualitative.tex
20 pages

@@
 have equal length.
-
-The student first defines a helper function |clone|, as seen above.
+%
+The student first defines a helper |clone|.
 %
@@
 %
-Then she defines |padZero| with a simple branch that determines which
-list is shorter, followed by a |clone| for the appropriate number of
-zeros.
+Then she defines |padZero| with a branch to determine which
+list is shorter, followed by a |clone| to zero-pad it.
 %
@@
 
-Examining the path induced by the |::| constructor, we can see that our
+Examining the path induced by the |::|, we can see that our
 \dectree is influenced by the fact that:
 %
-(1) the expression is a data constructor (not specifically |::|);
+(1) |::| is a constructor; %the expression is a constructor (not specifically |::|);
 %
-(2) the parent expression is a tuple; and
+(2) the parent is a tuple; and
 %
@@
 We note that first fact appears to be particularly significant; an
-examination of the training samples that reach that decision node
-reveals that before observing the \textsc{Is-Constructor} feature, the
+examination of the training samples that reach that decision
+reveals that, before observing the \textsc{Is-Constructor} feature the
 classifier is slightly in favor of predicting ``blame'', but afterwards
@@
 %
-It would appear that our \dectree has learned that data constructors are
+It would appear that our \dectree has learned that constructors are
 particularly suspicious, and is looking for exceptions to this general

### 2017-04-15 paper/oopsla17-submission/related.tex
20 pages

@@
 %
-The 8.0 release of the
-Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
-incorporates these ideas, allowing library authors to supply
-custom errors when type-class resolution or type-family reduction fail,
-but not for ordinary unification failures.
+% The 8.0 release of the
+% Glasgow Haskell Compiler\footnote{\url{https://ghc.haskell.org/trac/ghc/wiki/Proposal/CustomTypeErrors}}
+% incorporates these ideas, allowing library authors to supply
+% custom errors when type-class resolution or type-family reduction fail,
+% but not for ordinary unification failures.
 

### 2017-04-15 paper/oopsla17-submission/conclusion.tex
better comma

@@
 to the kinds of errors that programmers
-(specifically, novices who are in greatest
+(specifically novices, who are in greatest
 need of precise feedback) actually make,

### 2017-04-15 paper/oopsla17-submission/intro2.tex
asd

@@
 if removed would make the constraints
-satisfiable and hence, memory-safe~\citep{Jose:2011}
+satisfiable and hence, assertion-safe~\citep{Jose:2011}
 or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-gd,Pavlinovic2014-mr}.

### 2017-04-15 paper/oopsla17-submission/threats.tex

\mysection{Threats to Validity}
\label{sec:validity}

Although our experiments demonstrate that our technique can pinpoint type
errors more accurately than the state of the art and that our features are
relevant to blame assignment, our results may not generalize.

One threat to validity associated with supervised machine learning is
overfitting (\ie learning a model that is too complex with respect to
the data).
%
A similar issue that arises in machine learning is model stability (\ie
can small changes to the training set produce large changes in the model?).
%
We mitigate these threats by:
%
(1) using separate training and testing datasets drawn from distinct
student populations (\autoref{sec:quantitative}), demonstrating the
generality of our models; and
%
(2) via cross-validation on the joint dataset
(\autoref{sec:feature-utility}), which demonstrates the stability of our
models by averaging the accuracy of 10 models trained on distinct
subsets of the data.

Our benchmarks were drawn from students in an undergraduate course and
may not be representative of other student populations.
%
We mitigate this threat by including the largest empirical evaluation of
type error localization that we are aware of: over 4,500 pairs of
ill-typed programs and fixes from two instances of the course, with
programs from 102 different students.
%
We acknowledge, of course, that students are not industrial programmers
and our results may not translate to large-scale software development;
however, we are particularly interested in aiding novice programmers
as they learn to work inside the type system.

A related threat to construct validity is our definition of the immedate
next well-typed program as the intended ground truth answer (see
\autoref{sec:overview}, Challenge 2). Students may, in theory, submit
intermediate well-typed program ``rewrites'' between the original ill-typed
program and the final intended answer. Our approach to discarding outliers
(see \autoref{sec:evaluation}) is designed to mitigate this threat.

Our removal of program pairs that changed too much, where our oracle
could not identify the blame of the other tools, or where the other
tools timed out or encountered unsupported language features is another
threat to validity.
%
It is possible that including the programs that changed excessively
would hurt our models, or that the other tools would perform
better on the programs with unsupported language features.
%
We note however that
%
(1) outlier removal is a standard technique in machine learning%
%\ES{CITE?}
; and
%
(2) our Top-1 accuracy margin is large enough that even if we assumed
that \sherrloc were perfect on all excluded programs, it would
only tie our \hiddenFH. % in Top-1 accuracy.
%
%\ES{I think this is accurate, but should double check..}

Examining programs written in \ocaml as opposed to \haskell or any other
typed functional language poses yet another threat, common type errors
may differ in different languages.
%
\ocaml is, however, a standard target for research in type error
localization and thus our choice admits a direct comparison with prior
work.
%
Furthermore, the functional core of \ocaml that we support does not
differ significantly from the functional core of \haskell or SML, all of
which are effectively lambda calculi with a Hindley-Milner-style type
system.
% \footnote{\haskell's type classes are a notable exception, they
%   are also known to cause confusing type errors and would be interesting
%   to study as well.}

Finally, our use of student fixes as oracles
% for the source of type errors
assumes that students are able to correctly identify
the source of an error.
%
As the students are in the process of learning the language and type
system, this assumption may be faulty.
%
It may be that \emph{expert} users would disagree with many of the
student fixes, and that it is harder to learn a model of expert fixes,
or that the state of the art would be better at predicting expert fixes.
%
As we have noted before, we believe it is reasonable to use student
fixes as oracles because the student is the best judge of what she
\emph{intended}.

### 2017-04-16 paper/oopsla17-submission/evaluation.tex
asd

@@
-\section{Quantitative Evaluation}
+\mysection{Quantitative Evaluation}
 \label{sec:evaluation}
@@
 %
-\begin{description}
-\item[1. Blame Accuracy (\autoref{sec:quantitative})]
+\begin{itemize}
+\item \textbf{How accurate is \toolname?}
   %
-  How often do we \emph{correctly} assign blame for the type error?
+  How often does \toolname
+  blame a \emph{correct}
+  location
+  for the type error?
+  (\autoref{sec:quantitative})
   %
@@
   % of the error?
-\item[2. Feature Utility (\autoref{sec:feature-utility})]
+\item \textbf{What features are used by \toolname?}
   %
-  Which feature sets are \emph{most important} for accurate predictions?
+  Which program features are the most
+  \emph{most important} for
+  localizing errors?
+   (\autoref{sec:feature-utility})
   % How much do the features described in \autoref{sec:features}
   % contribute to our predictions?
-\item[3. Explaining Predictions (\autoref{sec:qualitative})]
+\item[Why does \toolname work? (\autoref{sec:qualitative})]
   %
-  Can we \emph{explain} individual predictions (correct or not)?
-\end{description}
+  Do the models learned by \toolname
+  correspond to our intuition about
+  the real causes of errors? That is,
+  can we \emph{interpret} the individual
+  predictions (correct or not)?
+\end{itemize}
 %\ES{may want to swap (1) and (2)..}
+
+1. MLP-500 (fancy) is 72% vs Sherrloc 56% for top-1
+2. Logistic (simple) is 60% vs Sherrloc 56% for top-1
+3. MLP-500 reaches 91% for top-3
+4. InSlice is crucial, just InSlice + LocalSyntax gets Logistic 10pts higher than OCaml, on par with Sherrloc
+5. Typing features provide more bang/buck than contextual syntax
+6. Tree learns some simple and intuitive rules for predicting no-blame (e.g. defining a function is a fine thing to do, calling a function likewise)
+7.
 

### 2017-04-16 paper/oopsla17-submission/related.tex
asd

@@
 \label{sec:related-work}
+\label{sec:type-error-diagnosis}
 
@@
 % applying machine learning techniques to problems in programming languages.
-
-\mysubsection{Type Error Diagnosis}
-\label{sec:type-error-diagnosis}
+%
+% \mysubsection{Type Error Diagnosis}
 Languages with static type systems and type inference produce type
@@
 
-\mysubsection{Fault Localization}
+\mypara{Fault Localization}
+%
 Given a defect, \emph{fault localization} is the task of identifying

### 2017-04-16 paper/oopsla17-submission/learning.tex
C for confidence score

@@
 First, we use $\evalsym$ to predict for each subexpression whether it
-should be blamed, and extract a confidence score $\mathcal{P}$ from the
+should be blamed, and extract a confidence score $\Runit$ from the
 $\Model$.
@@
 %
-Second, we \emph{rank} each subexpression by the confidence $\mathcal{P}$
+Second, we \emph{rank} each subexpression by the confidence $\Runit$
 that it should be blamed, and return to the user the top $k$

### 2017-04-16 paper/oopsla17-submission/conclusion.tex
asd

@@
 to open the door to a new data-driven
-path to precise error localization.
+path to precise error localization,
+that significantly outperforms the
+state of the art on a new benchmark
+suite comprising 4500 student programs.
 %
-\toolname
-translates pairs of
-blame-labeled ill-typed
-programs into bags-of-abstracted-terms,
-and then trains a classifier
-to predict whether a feature vector
-should get a blame label.
-% that
-% predicts the value of the
-% whether a feature vector gets
-% a blame label.
-%
-Given a new ill-typed program,
-we extract its bag-of-abstracted-terms
-to compute the blame label for each
-term and rank the predictions based
-on the model's confidence, to select
-the top 1--3 blame assignments to
-present to the user.
-
-We have evaluated our approach with
-five different classifiers on a suite
-of over 4,500 student programs.
-%
-We found that while machine learning
-over syntactic features of each term in isolation
-performs worse than existing
-purely constraint based approaches, % (\eg \ocaml, \sherrloc),
-augmenting the data with a single feature corresponding to
-the type error slice brings our
-classifiers up to par with the state of the art,
-and further augmenting the data with
-features of an expression's parent and children
-allows our classifiers to outperform
-the state of the art by 18 percentage points.
+%%% RJNUKEDME   \toolname
+%%% RJNUKEDME   translates pairs of
+%%% RJNUKEDME   blame-labeled ill-typed
+%%% RJNUKEDME   programs into bags-of-abstracted-terms,
+%%% RJNUKEDME   and then trains a classifier
+%%% RJNUKEDME   to predict whether a feature vector
+%%% RJNUKEDME   should get a blame label.
+%%% RJNUKEDME   % that
+%%% RJNUKEDME   % predicts the value of the
+%%% RJNUKEDME   % whether a feature vector gets
+%%% RJNUKEDME   % a blame label.
+%%% RJNUKEDME   %
+%%% RJNUKEDME   Given a new ill-typed program,
+%%% RJNUKEDME   we extract its bag-of-abstracted-terms
+%%% RJNUKEDME   to compute the blame label for each
+%%% RJNUKEDME   term and rank the predictions based
+%%% RJNUKEDME   on the model's confidence, to select
+%%% RJNUKEDME   the top 1--3 blame assignments to
+%%% RJNUKEDME   present to the user.
+%%% RJNUKEDME
+%%% RJNUKEDME   We have evaluated our approach with
+%%% RJNUKEDME   five different classifiers on a suite
+%%% RJNUKEDME   of over 4,500 student programs.
+%%% RJNUKEDME   %
+%%% RJNUKEDME   We found that while machine learning
+%%% RJNUKEDME   over syntactic features of each term in isolation
+%%% RJNUKEDME   performs worse than existing
+%%% RJNUKEDME   purely constraint based approaches, % (\eg \ocaml, \sherrloc),
+%%% RJNUKEDME   augmenting the data with a single feature corresponding to
+%%% RJNUKEDME   the type error slice brings our
+%%% RJNUKEDME   classifiers up to par with the state of the art,
+%%% RJNUKEDME   and further augmenting the data with
+%%% RJNUKEDME   features of an expression's parent and children
+%%% RJNUKEDME   allows our classifiers to outperform
+%%% RJNUKEDME   the state of the art by 18 percentage points.
 
@@
 As with other forms of machine learning,
-a key concern is that of \emph{bias}: are
+a key concern is that of \emph{data-set bias}: are
 \toolname's models specific

### 2017-04-16 paper/oopsla17-submission/evaluation.tex
asd

--- 2017-04-15 paper/oopsla17-submission/evaluation.tex
@@
 %
-\begin{description}
-\item[1. Blame Accuracy (\autoref{sec:quantitative})]
+\begin{itemize}
+\item \textbf{Blame Accuracy}
   %
-  How often do we \emph{correctly} assign blame for the type error?
+  How often does \toolname
+  blame a \emph{correct}
+  location for the type error?
+  (\autoref{sec:quantitative})
   %
@@
   % of the error?
-\item[2. Feature Utility (\autoref{sec:feature-utility})]
+\item \textbf{Feature Utility}
   %
-  Which feature sets are \emph{most important} for accurate predictions?
+  Which program \emph{features are required}
+  to localize errors?
+   (\autoref{sec:feature-utility})
   % How much do the features described in \autoref{sec:features}
   % contribute to our predictions?
-\item[3. Explaining Predictions (\autoref{sec:qualitative})]
+\item \textbf{Interpretability}
   %
-  Can we \emph{explain} individual predictions (correct or not)?
-\end{description}
-%\ES{may want to swap (1) and (2)..}
+  %% Do the models learned by \toolname
+  %% correspond to our intuition about
+  %% the real causes of errors?
+  Are the models \emph{interpretable} using
+  our intuition about the causes of type errors?
+  (\autoref{sec:qualitative})
+\end{itemize}
+%
+%\mypara{Summary of Results}
+%
+In the sequel we present our experimental
+methodology \autoref{sec:methodology} and
+then drill into how we evaluated each of
+the questions above.
+%
+However, for the impatient reader, we begin
+with a quick summary of our main results:
+%
+%
+%\begin{itemize}
+%
+%\item \textbf
+
+\mypara{1. Data beats Algorithms}
+Our main result is that for type error
+localization, data is indeed unreasonably
+effective~\citep{halevy09}.
+%
+\toolname's most sophisticated
+\emph{neural network}
+based classifier's top-1 error
+location blames the correct
+sub-term 74\% of the time
+which, a good 18 points
+higher than the state of
+the art \sherrloc's 56\%.
+%
+However, even \toolname's simple
+\emph{logistic regression} based
+classifier is correct 60\%,
+\ie 4 points better than \sherrloc.
+%
+When the top-3 locations are considered,
+\toolname is correct 91\% of the time.
+
+% \item \textbf
+\mypara{2. Slicing is Critical}
+%
+However, data is effective \emph{only}
+when irrelevant sub-terms have been
+sliced out of consideration.
+%
+In fact, perhaps our most surprising
+result is that type-error-slicing and
+local syntax alone yields
+a neural classifier that is 10 points
+better than \ocaml and on par with
+\sherrloc.
+%
+That is once we focus classifiers on
+slices, purely syntactic features
+alone perform as well as the
+state-of-the-art.
+
+%\item \textbf
+\mypara{3. Size doesn't matter, types do}
+%
+We find that (after slices)
+contextual typing features
+provide the biggest
+improvement in accuracy.
+%
+Furthermore we find contextual syntax
+features to be relatively unimportant,
+which would support the hypothesis that
+the context's \emph{type} nicely
+summarizes the properties of the
+surrounding expressions.
+%
+Finally, we found that the \emph{size}
+of the sub-expression was not a useful
+feature. This was unexpected as we thought
+smaller expressions would be simpler and
+hence, more likely causes.
+
+% \item \textbf
+\mypara{4. Models Learn Typing Rules}
+%
+Finally, by investigating a few of the
+predictions made by the \emph{decision tree}
+based models, we found that the models
+do indeed capture some simple and intuitive
+rules for predicting well-typededness.
+%
+For example, in an application, if the
+left argument is a function then the
+error is likely on the right sub-term;
+in a function definition \RJ{fixme: orig
+defining a function is a fine thing
+to do?}
+%
+%\end{itemize}
+
+\RJ{eric please check the numbers -- we say 71, 72, 74, 91, 92, 94?}
 

### 2017-04-16 paper/oopsla17-submission/related.tex
asd

@@
 % \mysubsection{Type Error Diagnosis}
-Languages with static type systems and type inference produce type
-errors that novices often perceive as difficult to
-interpret~\citep{Wand1986-nw}.
-%
-For example, approaches based on Hindley-Milner or other constraint
-systems typically have the issue that when the constraint $\alpha=\beta$
-is violated, the system cannot know whether the user should intend to
-fix $\alpha$ or $\beta$ and must thus report the discrepancy in a
-generic manner.
-%
-As a result, a number of approaches have been proposed to
-more precisely localize the type error,
-give clearer error messages, or
-fix the error automatically.
+%% Languages with static type systems and type inference produce type
+%% errors that novices often perceive as difficult to
+%% interpret~\citep{Wand1986-nw}.
+%% %
+%% For example, approaches based on Hindley-Milner or other constraint
+%% systems typically have the issue that when the constraint $\alpha=\beta$
+%% is violated, the system cannot know whether the user should intend to
+%% fix $\alpha$ or $\beta$ and must thus report the discrepancy in a
+%% generic manner.
+%% %
+%% As a result, a number of approaches have been proposed to
+%% more precisely localize the type error,
+%% give clearer error messages, or
+%% fix the error automatically.
 %
@@
 \mypara{Localizing Type Errors}
-It is well-known that the original Damas-Milner algorithm $\mathcal{W}$
-produces poor error locations~\citep{Wand1986-nw}, often reported far
-from their source.
+It is well-known that the original Damas-Milner
+algorithm $\mathcal{W}$ produces errors far
+from their source, that novices percieve as
+difficult to interpret~\citep{Wand1986-nw}.
 %
-The type checker collects typing constraints as it traverses the
-program, and it crucially assumes that if a constraint can be safely
-added to the set of assumptions (\ie no type error has been found yet),
-then the constraint can be \emph{assumed} to be correct.
+%% The type checker collects typing constraints as it traverses the
+%% program, and it crucially assumes that if a constraint can be safely
+%% added to the set of assumptions (\ie no type error has been found yet),
+%% then the constraint can be \emph{assumed} to be correct.
 %
-The type checker reports an error the moment it finds a constraint that
-contradicts one of the assumptions, blaming the new inconsistent
-constraint, and thus it is extremely sensitive to the order in which it
-traverses the source program (the infamous ``left-to-right''
+The type checker reports an error the moment
+it finds a constraint that contradicts one
+of the assumptions, blaming the new inconsistent
+constraint, and thus it is extremely sensitive
+to the order in which it traverses the source
+program (the infamous ``left-to-right''
 bias~\citep{McAdam1998-ub}).
 %
-Several alternative traversal have been proposed, \eg
-top-down rather than bottom-up~\citep{Lee1998-ys}, or a \emph{symmetric}
-traversal that checks sub-expressions independently and only reports an
+Several alternative traversal have been proposed,
+\eg top-down rather than bottom-up~\citep{Lee1998-ys},
+or a \emph{symmetric} traversal that checks
+sub-expressions independently and only reports an
 error when two inconsistent sets of constraints are
 merged~\citep{McAdam1998-ub,Yang1999-yr}.
-
-\mypara{Slicing}
+%
+%\mypara{Slicing}
 Type error \emph{slicing}~\citep{Haack2003-vc,Tip2001-qp,Rahli2010-ps}
-overcomes this bias by extracting a complete and minimal subset
+overcomes the constraint-order bias by extracting a
+complete and minimal subset
 of terms that contribute to the error, \ie all of the
@@
 \textsc{MrSpidey}~\citep{Flanagan1996-bu}.
-
-\mypara{Interactive Explanations}
+%
+\citet{Hage2006-hc} catalog a set of heuristics for
+improving the quality of error messages by examining errors made by
+novices.
+%
+\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
+\citet{Serrano2016-oo} extend the ability to customize error messages to
+library authors, enabling \emph{domain-specific} errors.
+%
+%\mypara{Interactive Explanations}
 Such \emph{static} explanations of type errors run the risk of
@@
 
-\mypara{Programmatic Explanations}
-%
-The best explanation of a type error, however, might be given by an
-expert, \eg the compiler or library author.
-%
-\citet{Hage2006-hc} catalog a set of heuristics for
-improving the quality of error messages by examining errors made by
-novices.
-%
-\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
-\citet{Serrano2016-oo} extend the ability to customize error messages to
-library authors, enabling \emph{domain-specific} errors.
+%% RJ-CUT-ME \mypara{Programmatic Explanations}
+%% RJ-CUT-ME %
+%% RJ-CUT-ME The best explanation of a type error, however, might be given by an
+%% RJ-CUT-ME expert, \eg the compiler or library author.
+%% RJ-CUT-ME %
+%% RJ-CUT-ME \citet{Hage2006-hc} catalog a set of heuristics for
+%% RJ-CUT-ME improving the quality of error messages by examining errors made by
+%% RJ-CUT-ME novices.
+%% RJ-CUT-ME %
+%% RJ-CUT-ME \citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
+%% RJ-CUT-ME \citet{Serrano2016-oo} extend the ability to customize error messages to
+%% RJ-CUT-ME library authors, enabling \emph{domain-specific} errors.
 %

### 2017-04-16 paper/oopsla17-submission/evaluation.tex
tweak

@@
-\mysection{Quantitative Evaluation}
+\mysection{Evaluation}
 \label{sec:evaluation}
@@
 localization, data is indeed unreasonably
-effective~\citep{halevy09}.
+effective \citep{halevy09}.
 %
@@
 \emph{logistic regression} based
-classifier is correct 60\%,
+classifier is correct 60\% of the time,
 \ie 4 points better than \sherrloc.
@@
 That is once we focus classifiers on
-slices, purely syntactic features
+slices, local syntactic features
 alone perform as well as the
@@
 Furthermore we find contextual syntax
-features to be relatively unimportant,
-which would support the hypothesis that
+features to be mostly (but not entirely)
+redundant with typing features,
+which supports the hypothesis that
 the context's \emph{type} nicely
@@
 Finally, we found that the \emph{size}
-of the sub-expression was not a useful
-feature. This was unexpected as we thought
-smaller expressions would be simpler and
+of the sub-expression was not very useful.
+This was unexpected, as we thought
+smaller expressions would be simpler, and
 hence, more likely causes.
@@
 Finally, by investigating a few of the
-predictions made by the \emph{decision tree}
-based models, we found that the models
+predictions made by the \emph{decision tree}-based
+models, we found that the models
 do indeed capture some simple and intuitive
-rules for predicting well-typededness.
+rules for predicting well-typedness.
 %
-For example, in an application, if the
-left argument is a function then the
-error is likely on the right sub-term;
-in a function definition \RJ{fixme: orig
-defining a function is a fine thing
-to do?}
+For example, if the left child of an application
+is a function, then the application is likely
+correct.
+
+% in an application, if the
+% left argument is a function then the
+% error is likely on the right sub-term;
+% in a function definition
+% \RJ{fixme: orig
+% defining a function is a fine thing
+% to do?}
 %
 %\end{itemize}
-
-\RJ{eric please check the numbers -- we say 71, 72, 74, 91, 92, 94?}
+% \RJ{eric please check the numbers -- we say 71, 72, 74, 91, 92, 94?}
 

### 2017-04-17 paper/oopsla17-submission/evaluation.tex
minor tweaks

@@
 
-\mypara{1. Data beats Algorithms}
+\mypara{1. Data Beats Algorithms}
 Our main result is that for type error
@@
 \toolname's most sophisticated
-\emph{neural network}
-based classifier's top-1 error
-location blames the correct
+\emph{neural network}-based
+classifier's top-ranked
+prediction blames the correct
 sub-term 74\% of the time
@@
 %
-When the top-3 locations are considered,
+When the top three predictions are considered,
 \toolname is correct 91\% of the time.
@@
 % \item \textbf
-\mypara{2. Slicing is Critical}
+\mypara{2. Slicing Is Critical}
 %
@@
 In fact, perhaps our most surprising
-result is that type-error-slicing and
+result is that type error slicing and
 local syntax alone yields
-a neural classifier that is 10 points
+a classifier that is 10 points
 better than \ocaml and on par with
@@
 %
-That is once we focus classifiers on
-slices, local syntactic features
-alone perform as well as the
+That is, once we focus our classifiers on
+slices, purely syntactic features
+perform as well as the
 state-of-the-art.
@@
 %\item \textbf
-\mypara{3. Size doesn't matter, types do}
+\mypara{3. Size Doesn't Matter, Types Do}
 %
 We find that (after slices)
-contextual typing features
+typing features
 provide the biggest
@@
 %
-Furthermore we find contextual syntax
+Furthermore, we find contextual syntax
 features to be mostly (but not entirely)

### 2017-04-17 paper/oopsla17-submission/conclusion.tex
more minor

@@
 to open the door to a new data-driven
-path to precise error localization,
-that significantly outperforms the
+path towards precise error localization,
+significantly outperforming the
 state of the art on a new benchmark
-suite comprising 4500 student programs.
+suite comprising 4,500 student programs.
 %

### 2017-04-17 paper/oopsla17-submission/intro2.tex
Wes pass over abstract and intro. I am wary of "Types are awesome" as the lead, will send email.

--- 2017-04-14 paper/oopsla17-submission/intro2.tex
@@
 the minimum (weighted) subset that,
-if removed would make the constraints
-satisfiable and hence, memory-safe~\citep{Jose:2011}
+if removed, would make the constraints
+satisfiable and hence, assertion-safe~\citep{Jose:2011}
 or well-typed~\citep{Zhang2014-lv,Loncaric2016-uk,Chen2014-gd,Pavlinovic2014-mr}.

### 2017-04-17 paper/oopsla17-submission/abstract.tex
Wes pass over abstract and intro. I am wary of "Types are awesome" as the lead, will send email.
extract repeated performance numbers

@@
 the exact sub-expression that should
-be changed 74\% of the time, %which is
-29 points higher than \ocaml and
-18 points higher than the state-of-the-art
+be changed \HiddenFhTopOne\% of the time, %which is
+\ToolnameWinOcaml points higher than \ocaml and
+\ToolnameWinSherrloc points higher than the state-of-the-art
 \sherrloc tool.
@@
 Furthermore, \toolname's accuracy surpasses
-85\% when we consider the top \emph{two}
-locations and reaches 91\% if consider
+\HiddenFhTopTwo\% when we consider the top \emph{two}
+locations and reaches \HiddenFhTopThree\% if we consider
 the top \emph{three}.

### 2017-04-17 paper/oopsla17-submission/evaluation-accuracy.tex
extract repeated performance numbers

@@
 Our classifiers consistently outperform the competition, ranging from
-60\% Top-1 accuracy (87\% Top-3) for the \linear classifier to
-72\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
+\LinearTopOne\% Top-1 accuracy (\LinearTopThree\% Top-3)
+for the \linear classifier to
+\HiddenFhTopOne\% Top-1 accuracy (\HiddenFhTopThree\% Top-3)
+for the \hiddenFH.\@
 %
-Our baseline of selecting at random achieves 30\% Top-1
-accuracy (58\% Top-3), while \ocaml achieves a Top-1 accuracy of 45\%.
+Our baseline of selecting at random achieves \BaselineTopOne\% Top-1
+accuracy (\BaselineTopThree\% Top-3),
+while \ocaml achieves a Top-1 accuracy of \OcamlTopOne\%.
 %
 Interestingly, one only needs two \emph{random} guesses to outperform
-\ocaml, with 47\% accuracy.
+\ocaml, with \BaselineTopTwo\% accuracy.
 %
-\sherrloc outperforms both baselines, and comes close to our \linear
-classifier, with 56\% Top-1 accuracy (85\% Top 3), while \mycroft
-underperforms \ocaml at 39\% Top-1 accuracy.
+\sherrloc outperforms both baselines, and comes close to our \linear classifier,
+with \SherrlocTopOne\% Top-1 accuracy (\SherrlocTopThree\% Top 3),
+while \mycroft underperforms \ocaml at \MycroftTopOne\% Top-1 accuracy.
 %
@@
 In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
-hidden neurons than the \hiddenFH, it only loses around 2\% accuracy.
+hidden neurons than the \hiddenFH, it only loses around 4\% accuracy.
 % In particular, notice that the \hiddenT only loses around 2\% accuracy

### 2017-04-17 paper/oopsla17-submission/evaluation-utility.tex
extract repeated performance numbers

@@
 %
-As expected, the baseline performs the worst, with a mere 27.7\% \linear
+As expected, the baseline performs the worst, with a mere 28\% \linear
 Top-1 accuracy.
 %
-Adding \InSlice improves the results substantially with a 46.4\% \linear Top-1
+Adding \InSlice improves the results substantially with a 46\% \linear Top-1
 accuracy, demonstrating the importance of a minimal error slice.
@@
 However, filtering out expressions that are not part of the slice
-\emph{further} improves the results to 55.9\% \linear Top-1 accuracy.
+\emph{further} improves the results to 56\% \linear Top-1 accuracy.
 %
@@
 %
-Top-1 accuracy jumps from 32.4\% to 55.4\% when we add \InSlice, and only
-improves by 2\% when we filter out expressions that are not part of the
+Top-1 accuracy jumps from 33\% to 55\% when we add \InSlice, and only
+improves by 3\% when we filter out expressions that are not part of the
 error slice.
@@
 In contrast, the contextual syntactic features improve \linear Top-1
-accuracy by 5\% (\resp 14\%), and the typing features improve
-Top-1 accuracy by 7\% (\resp 16\%).
+accuracy by 4\% (\resp 14\%), and the typing features improve
+Top-1 accuracy by 6\% (\resp 16\%).
 %

### 2017-04-17 paper/oopsla17-submission/evaluation.tex
extract repeated performance numbers

@@
 prediction blames the correct
-sub-term 74\% of the time
-which, a good 18 points
+sub-term \HiddenFhTopOne\% of the time
+which, a good \ToolnameWinSherrloc points
 higher than the state of
-the art \sherrloc's 56\%.
+the art \sherrloc's \SherrlocTopOne\%.
 %
@@
 \emph{logistic regression} based
-classifier is correct 60\% of the time,
-\ie 4 points better than \sherrloc.
+classifier is correct \LinearTopOne\% of the time,
+\ie \LinearWinSherrloc points better than \sherrloc.
 %
 When the top three predictions are considered,
-\toolname is correct 91\% of the time.
+\toolname is correct \HiddenFhTopThree\% of the time.
 
@@
 local syntax alone yields
-a classifier that is 10 points
+a classifier that is \SlicingWinOcaml points
 better than \ocaml and on par with

### 2017-04-17 paper/oopsla17-submission/intro2.tex
extract repeated performance numbers

@@
 the exact sub-expression that should
-be changed 74\% of the time,
-29 points higher than \ocaml and
-18 points higher than the state-of-the-art
+be changed \HiddenFhTopOne\% of the time,
+\ToolnameWinOcaml points higher than \ocaml and
+\ToolnameWinSherrloc points higher than the state-of-the-art
 \sherrloc tool.
@@
 Furthermore, \toolname's accuracy surpasses
-85\% when we consider the top \emph{two}
-locations and reaches 91\% if consider
+\HiddenFhTopTwo\% when we consider the top \emph{two}
+locations and reaches \HiddenFhTopThree\% if consider
 the top \emph{three}.
@@
 features allows our classifiers to outperform
-the state of the art by 18 percentage points.
+the state of the art by \ToolnameWinSherrloc percentage points.
 % In particular, we find that while machine learning

### 2017-04-17 paper/oopsla17-submission/overview.tex
asd
extract repeated performance numbers

@@
 
-\mypara{Error Localization via Supervised Classification}
+\mypara{Solution: Localization via Supervised Classification}
 %
@@
 yields a high-precision localizer:
-when the top-3 locations are considered,
+when the top three locations are considered,
 at least one matches an actual student fix
-over 90\% of the time.
+\HiddenFhTopThree\% of the time.
 

### 2017-04-17 paper/oopsla17-submission/evaluation-utility.tex
forgot to account for list-literal feature

@@
 We focus specifically on feature \emph{classes} rather than individual
-features as our 277 features are conceptually grouped into a much
+features as our 282 features are conceptually grouped into a much
 smaller number of \emph{categorical} features.
@@
 For example, the syntactic class of an expression is conceptually a
-feature but there are 44 possible values it could take; to encode this
-feature for learning we split it into 44 distinct binary features.
+feature but there are 45 possible values it could take; to encode this
+feature for learning we split it into 45 distinct binary features.
 %
@@
 They improve Top-1 accuracy by an additional 2\%, and are much more
-compact --- requiring only 55 typing features compared to 176
+compact --- requiring only 55 typing features compared to 180
 contextual syntactic features.

### 2017-04-17 paper/oopsla17-submission/conclusion.tex
re-add a summary sentence to conclusion

@@
 %%% RJNUKEDME   of over 4,500 student programs.
-%%% RJNUKEDME   %
-%%% RJNUKEDME   We found that while machine learning
-%%% RJNUKEDME   over syntactic features of each term in isolation
-%%% RJNUKEDME   performs worse than existing
-%%% RJNUKEDME   purely constraint based approaches, % (\eg \ocaml, \sherrloc),
-%%% RJNUKEDME   augmenting the data with a single feature corresponding to
-%%% RJNUKEDME   the type error slice brings our
-%%% RJNUKEDME   classifiers up to par with the state of the art,
-%%% RJNUKEDME   and further augmenting the data with
-%%% RJNUKEDME   features of an expression's parent and children
-%%% RJNUKEDME   allows our classifiers to outperform
-%%% RJNUKEDME   the state of the art by 18 percentage points.
+%
+We found that while machine learning
+over syntactic features of each term in isolation
+performs worse than existing
+purely constraint based approaches, % (\eg \ocaml, \sherrloc),
+augmenting the data with a single feature corresponding to
+the type error slice brings our
+classifiers up to par with the state of the art,
+and further augmenting the data with
+features of an expression's parent and children
+allows our classifiers to outperform
+the state of the art by \ToolnameWinSherrloc
+percentage points.
 

### 2017-04-17 paper/oopsla17-submission/evaluation-utility-graph.tex
forgot to account for list-literal feature
re-add a summary sentence to conclusion

@@
   xticklabels={
-    \textsc{Local Syntax}\\(44),
-    \textsc{+Size}\\(45), \textsc{+Context}\\(220), \textsc{+Type}\\(102),
-    +C+S\\(221), +T+S\\(103), +C+T\\(275),
-    +C+T+S\\(276)
+    \textsc{Local Syntax}\\(45),
+    \textsc{+Size}\\(46), \textsc{+Context}\\(225), \textsc{+Type}\\(100),
+    +C+S\\(226), +T+S\\(101), +C+T\\(281),
+    +C+T+S\\(282)
     % + Size, + Context, + Type,
@@
 \end{tikzpicture}
-\caption{Impact of contextual features on blame accuracy. Starting from
-  a baseline of local syntactic features, we add each combination of
-  expression size, contextual syntactic, and typing features. The total
-  number of features is given in parentheses.}
+\caption{
+  %
+  Impact of contextual features on blame accuracy.
+  %
+  % Starting from a baseline of local syntactic features,
+  % we add each combination of
+  % expression size, contextual syntactic, and typing features.
+  %
+  The total number of features for each set is given in parentheses.
+}
 \label{fig:context-utility}

### 2017-04-17 paper/oopsla17-submission/related.tex
re-add a summary sentence to conclusion

@@
 the inferred types of sub-expressions that guide the user to a specific
-explanation for the error.
+explanation.
 %
@@
 Our feature set is comprised entirely of syntactic and typing features,
-a natural choice for modeling type errors, but it would likely not
-generalize to other types of defects.
+a natural choice for type errors, but it would likely not
+generalize to other defects.
 

### 2017-04-17 paper/oopsla17-submission/intro2.tex
colorbox borders to degrade to b/w better
hyphen

@@
 %
-Types are awesome.
+%Types are awesome.
+When viewed fully, types are awe-inspiring.
 %
@@
 type inference algorithms are prone to report
-type errors far from their actual source
+type errors far from their source
 \citep{Wand1986-nw}.
@@
 performs worse than existing
-purely constraint based approaches (\eg \ocaml, \sherrloc),
+purely constraint-based approaches (\eg \ocaml, \sherrloc),
 augmenting the data with a single feature corresponding to

### 2017-04-17 paper/oopsla17-submission/intro2.tex
small tweaks thanks to abaskt, gleissen, and atondawal

@@
 \HiddenFhTopTwo\% when we consider the top \emph{two}
-locations and reaches \HiddenFhTopThree\% if consider
+locations and reaches \HiddenFhTopThree\% if we consider
 the top \emph{three}.
@@
 returned by \emph{feature abstraction} functions
-that applied to the term.
+applied to the term.
 %
-We can model the blame labels with a simple feature
-abstraction (\eg is-changed-in-fix).
+% We can model the blame labels with a simple feature
+% abstraction (\eg is-changed-in-fix).
 %
-We can even account for \emph{contextual} information
-like the features of the parent- and child- terms by
+We can even recover \emph{contextual} information
+from the parent and child terms by
 \emph{concatenating} the feature vectors of each term
-with those of its parent and children.
+with those of its parent and children
+(within a fixed window).
 %
@@
 Finally, the last and most important challenge is to
-use the labeled BOAT data to train classifiers that
+use our BOAT representation to train classifiers that
 are capable of \emph{precisely} pinpointing the errors
@@
 The key here is to find the right set of feature
-abstractions and classification algorithms that
-lead to precise predictions.
+abstractions to model type errors,
+and classification algorithms that
+lead to precise blame assignments.
 %
@@
 %
-syntactic features (\eg is-a-data constructor, is-a-literal,
+syntactic features (\eg is-a-data-constructor, is-a-literal,
 is-an-arithmetic-operation, is-a-function-application, \etc),
@@
 is-an-integer, is-a-function, \etc).
+%
+We can similarly model the blame labels with a simple feature
+abstraction (\eg is-changed-in-fix).
 

### 2017-04-17 paper/oopsla17-submission/overview.tex
small tweaks thanks to abaskt, gleissen, and atondawal

@@
 %
-Our representation as parameterized by a
+Our representation is parameterized by a
 set of \emph{feature abstraction} functions,
@@
 abstractions of the form
-is-nil, is-plus, \etc for
+is-|[]|, is-|+|, \etc, for
 each of a fixed number of
-primitive literal terms.
+expression forms.
 %
@@
 sub-expression with features
-of the form is-int, is-bool, \etc.
+of the form is-|int|, is-|bool|, \etc.
 %

### 2017-04-24 paper/oopsla17-submission/conclusion.tex

@@
 performs worse than existing
-purely constraint based approaches, % (\eg \ocaml, \sherrloc),
+purely constraint-based approaches, % (\eg \ocaml, \sherrloc),
 augmenting the data with a single feature corresponding to

### 2017-04-24 paper/oopsla17-submission/evaluation.tex
forgot to account for list-literal feature

@@
 \mypara{Feature Selection}
-We extract a set of 277 features from each sub-expression in a student
+We extract a set of 282 features from each sub-expression in a student
 program, including:
@@
 \begin{enumerate}
-\item 44 local syntactic features. In addition to the syntax of \lang,
+\item 45 local syntactic features. In addition to the syntax of \lang,
   we support the full range of arithmetic operators (integer and
@@
   types in \autoref{sec:discussion}.
-\item 176 contextual syntactic features. For each sub-expression we
+\item 180 contextual syntactic features. For each sub-expression we
   additionally extract the local syntactic features of its parent and
@@
 more than one standard deviation above the mean, establishing a diff
-threshold of 44\%.
+threshold of 45\%.
 %

### 2017-04-24 paper/oopsla17-submission/learning.tex

@@
 
-A $\Model$ is produced by the $\trainsym$ function, which performs
+A $\Model$ is produced by $\trainsym$, which performs
 supervised learning on a training set of feature vectors $\V$ and
-%(boolean)
-labels $\B$.
+(boolean) labels $\B$.
 %
-Once trained, we can $\evalsym$uate a $\Model$ on a new input vector,
-producing the confidence $\Runit$ that the blame label should apply. % be
-%applied.
+Once trained, we can $\evalsym$uate a $\Model$ on a new input,
+producing the confidence $\Runit$ that the blame label should be
+applied.
 %
 We describe multiple $\Model$s and their instantiations of
-$\trainsym$ and $\evalsym$ (\autoref{sec:models}).
+$\trainsym$ and $\evalsym$
+(\autoref{sec:models}).
 
@@
 This means that we are given a training data set
-$S = \{ \langle v_i, b_i \rangle \}_{i=1}^n$ of labeled examples, and
+%$S = \{ \langle v_i, b_i \rangle \}_{i=1}^n$
+$S : \List{\V \times \B}$
+of labeled examples, and
 our goal is to use it to build a \emph{classifier}, \ie a rule

### 2017-04-25 study/study.tex
add sample questions for user-study

\documentclass{article}

\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  keywordstyle=\ttfamily,
  escapeinside={(*@}{@*)},
}

\usepackage{xcolor}
\newcommand\hlBlame[1]{\colorbox{red!25}{#1}}

\begin{document}

\section{Debugging and Functional Programming}

You will be shown OCaml programs that \emph{do not type-check}, the code
implicated by the type checker will be \hlBlame{highlighted}. You will
also be given English comments explaining what the program \emph{should}
do, as well as assertions that \emph{should} type check \emph{and} succeed.

\subsection{Concatenating Strings}

\subsubsection{Control}
\begin{lstlisting}
  (* "sepConcat sep [s1;s2;s3]" should insert "sep"
     between "s1", "s2", and "s3", and concatentate
     the result. *)

  let rec sepConcat sep sl =
    match sl with
    | [] -> ""
    | h::t ->
        let f a x = a ^ (sep ^ x) in
        let base = [] in
        (*@\hlBlame{List.fold\_left f base sl}@*)

  assert( sepConcat "," ["foo"; "bar"; "baz"] = "foo,bar,baz" )
\end{lstlisting}

\begin{enumerate}
\item Why is \verb!sepConcat! not well-typed?
\item Fix the \verb!sepConcat! program.
\end{enumerate}

\subsubsection{Treatment}
\begin{lstlisting}
  (* "sepConcat sep [s1;s2;s3]" should insert "sep"
     between "s1", "s2", and "s3", and concatentate
     the result. *)

  let rec sepConcat sep sl =
    match sl with
    | [] -> ""
    | h::t ->
        let f a x = a ^ (sep ^ x) in
        let base = (*@\hlBlame{[]}@*) in
        List.fold_left f base sl

  assert( sepConcat "," ["foo"; "bar"; "baz"] = "foo,bar,baz" )
\end{lstlisting}

\begin{enumerate}
\item Why is \verb!sepConcat! not well-typed?
\item Fix the \verb!sepConcat! program.
\end{enumerate}

\subsection{Padding Lists}

\subsubsection{Control}
\begin{lstlisting}
  (* "padZero xs ys" returns a pair "(xs', ys')"
     where the shorter of "xs" and "ys" has been
     left-padded by zeros until both lists have
     equal length. *)

  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)

  let padZero l1 l2 =
    let n = List.length l1 - List.length l2 in
    if n < 0 then
      (clone 0 ((-1) * n) @ l2, l2)
    else
      (l1, (*@\hlBlame{clone 0 n}@*) :: l2)

  assert( padZero [1;2] [1] = ([1;2], [0;1]) )
\end{lstlisting}

\subsubsection{Treatment}
\begin{lstlisting}
  (* "padZero xs ys" returns a pair "(xs', ys')"
     where the shorter of "xs" and "ys" has been
     left-padded by zeros until both lists have
     equal length. *)

  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)

  let padZero l1 l2 =
    let n = List.length l1 - List.length l2 in
    if n < 0 then
      (clone 0 ((-1) * n) @ l2, l2)
    else
      (l1, clone 0 n (*@\hlBlame{::}@*) l2)

  assert( padZero [1;2] [1] = ([1;2], [0;1]) )
\end{lstlisting}

\begin{enumerate}
\item Why is \verb!padZero! not well-typed?
\item Fix the \verb!padZero! program.
\end{enumerate}

\subsection{Big-Integer Arithmetic}

\subsubsection{Control}
\begin{lstlisting}
  (* "mulByDigit d [n1;n2;n3]" should multiply the
     "big integer" "[n1;n2;n3]"
     by the single digit "d". *)

  let rec mulByDigit d n =
    match List.rev n with
    | []   -> []
    | h::t -> [(*@\hlBlame{mulByDigit d t}@*); (h * d) mod 10]

  assert( mulByDigit 4 [2;5] = [1;0;0] )
\end{lstlisting}

\begin{enumerate}
\item Why is \verb!mulByDigit! not well-typed?
\item Fix the \verb!mulByDigit! program.
\end{enumerate}

\subsubsection{Treatment}
\begin{lstlisting}
  (* "mulByDigit d [n1;n2;n3]" should multiply the
     "big integer" "[n1;n2;n3]"
     by the single digit "d". *)

  let rec mulByDigit d n =
    match List.rev n with
    | []   -> []
    | h::t -> (*@\hlBlame{[mulByDigit d t; (h * d) mod 10]}@*)

  assert( mulByDigit 4 [2;5] = [1;0;0] )
\end{lstlisting}

\begin{enumerate}
\item Why is \verb!mulByDigit! not well-typed?
\item Fix the \verb!mulByDigit! program.
\end{enumerate}

\end{document}

### 2017-04-25 paper/oopsla17-submission/evaluation.tex
typo

@@
 sub-term \HiddenFhTopOne\% of the time
-which, a good \ToolnameWinSherrloc points
+which is a good \ToolnameWinSherrloc points
 higher than the state of

### 2017-05-17 paper/oopsla17-submission/qualitative.tex
tweak
update dataset
update qualitative.tex

@@
-\mysection{Qualitative Evaluation}
+%\mysubsection{Qualitative Evaluation}
+\mysubsection{Interpreting Specific Predictions}
 \label{sec:qualitative}
@@
 
-\mysubsection{Failed Predictions}
+\mysubsubsection{Failed Predictions}
 \label{sec:failed-predictions}
@@
 \mypara{Constructing a List of Duplicates}
-% data/sp14/0219.ml
+% data/sp14/0969.ml
 Our first program is a simple recursive function |clone| that takes an
@@
       else
-        (*@\hlFix{clone} \hlSherrloc{([\hlTree{x}] @\ acc)}@*) (n - 1) in
+        (*@\hlFix{clone} \hlSherrloc{\hlTree{([x] @\ acc)}}@*) (n - 1) in
     loop [] n
@@
 
+\ES{FIX: our first/second predictions are now swapped}
 Our classifier incorrectly predicts that the use of |x| in the recursive
@@
 \mypara{Currying Considered Harmful?}
-% data/sp14/1204.ml
+% data/sp14/1082.ml
 Our next example is another ill-fated attempt at |clone|.
@@
 % \mypara{Composing Functions}
-% % data/sp14/3041.ml
+% % data/sp14/3203.ml
 % Next, let us consider the |pipe| function that composes a list of
@@
 
-\mysubsection{Correct Predictions}
+\mysubsubsection{Correct Predictions}
 \label{sec:correct-predictions}
@@
 \mypara{Extracting the Digits of an Integer}
-% data/sp14/2176.ml
+% data/sp14/1737.ml
 Consider first a simple recursive function |digitsOfInt| that extracts
@@
 \mypara{Padding a list}
-% data/sp14/0442.ml
+% data/sp14/1394.ml
 Our next program, |padZero|, is given two |int list|s as input, and must

### 2017-06-08 paper/oopsla17-submission/rebuttal.md
first draft of the rebuttal

We thank the reviewers for their kind and thoughtful remarks, and will
respond briefly to some of the larger concerns.

1. Size / Diversity of Training Corpus [A,D]

It may have been unclear in the text, but our corpus was drawn from
student solutions to *3 homework sets*, not 3 programs. Each homework
set consisted of many individual programs. In total our corpus contains
solutions to *23 distinct programs* that cover a broad range of FP
idioms, e.g. (tail) recursive functions, higher-order functions,
expression evaluators, etc.

We will expand Section 4.1 (Methodology) to clearly list the total
number of programs students wrote in addition to the number of
assignments they completed.

2. Generalizing NATE Models to Other Datasets [A,D]

We may not have adequately clarified our claims regarding the
generalizability of our models. As reviewers A and D point out, it is
well known that ML models suffer from dataset bias; we cannot claim (nor
do we) that our models generalize to all OCaml type errors.

Rather, we claim that our models generalize with respect to type errors
that occur in our undergraduate course. We substantiate this claim by
training a model on the student programs from one instance of the course
and testing it on programs from another instance (see Section 4.2).

Second, we claim that our models are *easy to train*, i.e. they can be
trained on programs from a single instance of a course with around 50
participants. Thus, even if our models do not generalize to other OCaml
courses (which we have not tested and do not claim), the barrier to
producing a course-specific model is actually quite low (see the final
paragraph of the conclusion).

We will update the text in the Introduction and Evaluation to clarify
the scope of our claim to generality.

3. Supported Fragment of OCaml [B]

Our restriction to OCaml's functional core is simply due to the nature
of the programs in our corpus. We do not teach references, classes,
modules, etc., and thus have no data to train or evaluate on. We do not
anticipate that these additional language features would be difficult to
support, as our features are simple queries over the typed syntax tree.

As we mention in Section 5, however, supporting the unbounded set of
user-defined types in our approach may be difficult. Briefly, the issue
is that we create a new feature for each data- and type-constructor,
so models will be at a disadvantage when confronted with unseen types.

4. Scaling to Larger Programs [C]

The primary concern with respect to scaling to larger programs is
probably the size of the type error slice. The slice greatly prunes the
set of possible blame locations (by a factor of 7 in our dataset). As
the slice increases there will be more room for the classifier to
misrank expressions.

This concern applies to the state-of-the-art SherrLoc and Mycroft as
well, as they also rely implicitly or explicitly on type error slices.
It may also be mitigated by the fact that larger codebases tend to
provide explicit type signatures for top-level functions, thus confining
error slices to single definitions.

5. Determination of "Fixed" Programs [B]

Reviewer B is correct that our determination of the *first* type-correct
program as the fix is a threat to validity, we discuss this and other
threats in Section 4.5. We could also train a model using the final
type-correct program as the fix, or perhaps using a (set of)
known-correct solution(s) as the fix, both of these choices could lead
to different results.

We have made the full corpus available at

https://github.com/ucsd-progsys/yunounderstand-data

and will advertise this in the text. The corpus contains the full series
of student interactions with OCaml and would admit Reviewer B's proposed
analysis of how different the initial and final solutions are, though we
have not attempted it.

### 2017-06-09 paper/oopsla17-submission/rebuttal.md
rebuttal is still blinded
rebuttal: 498 words

@@
 set consisted of many individual programs. In total our corpus contains
-solutions to *23 distinct programs* that cover a broad range of FP
-idioms, e.g. (tail) recursive functions, higher-order functions,
-expression evaluators, etc.
+solutions to *23 distinct programs* that cover a broad range of FP idioms.
 
@@
 well known that ML models suffer from dataset bias; we cannot claim (nor
-do we) that our models generalize to all OCaml type errors.
+do we) that our models generalize to all OCaml type errors. We discuss
+this, and other threats to validity, in Section 4.5.
 
-Rather, we claim that our models generalize with respect to type errors
-that occur in our undergraduate course. We substantiate this claim by
+We claim that our models generalize with respect to type errors that
+occur in our undergraduate course. We substantiate this claim by
 training a model on the student programs from one instance of the course
@@
 
-Second, we claim that our models are *easy to train*, i.e. they can be
+We also claim that our models are *easy to train*, i.e. they can be
 trained on programs from a single instance of a course with around 50
@@
 
-3. Supported Fragment of OCaml [B]
+3. Determination of "Fixed" Programs [B]
 
-Our restriction to OCaml's functional core is simply due to the nature
-of the programs in our corpus. We do not teach references, classes,
-modules, etc., and thus have no data to train or evaluate on. We do not
-anticipate that these additional language features would be difficult to
-support, as our features are simple queries over the typed syntax tree.
+Reviewer B is correct that our determination of the *first* type-correct
+program as the fix is a threat to validity (see Section 4.5 for
+others). We could also train a model using the final type-correct
+program as the fix, or perhaps using a (set of) known-correct
+solution(s) as the fix, both of these choices could lead to different
+results.
 
-As we mention in Section 5, however, supporting the unbounded set of
-user-defined types in our approach may be difficult. Briefly, the issue
-is that we create a new feature for each data- and type-constructor,
-so models will be at a disadvantage when confronted with unseen types.
+The corpus is available at http://tinyurl.com/y9pzvcp7 (NOT ANONYMOUS);
+we will advertise this in the text. It contains the full series of
+student interactions and would admit Reviewer B's proposed analysis of
+how different the initial and final solutions are.
 
@@
 probably the size of the type error slice. The slice greatly prunes the
-set of possible blame locations (by a factor of 7 in our dataset). As
+set of potential blame locations (by a factor of 7 in our dataset). As
 the slice increases there will be more room for the classifier to
-misrank expressions.
-
-This concern applies to the state-of-the-art SherrLoc and Mycroft as
-well, as they also rely implicitly or explicitly on type error slices.
-It may also be mitigated by the fact that larger codebases tend to
-provide explicit type signatures for top-level functions, thus confining
-error slices to single definitions.
-
-5. Determination of "Fixed" Programs [B]
-
-Reviewer B is correct that our determination of the *first* type-correct
-program as the fix is a threat to validity, we discuss this and other
-threats in Section 4.5. We could also train a model using the final
-type-correct program as the fix, or perhaps using a (set of)
-known-correct solution(s) as the fix, both of these choices could lead
-to different results.
-
-We have made the full corpus available at
-
-https://github.com/ucsd-progsys/yunounderstand-data
-
-and will advertise this in the text. The corpus contains the full series
-of student interactions with OCaml and would admit Reviewer B's proposed
-analysis of how different the initial and final solutions are, though we
-have not attempted it.
+misrank expressions. This concern also applies to SherrLoc and Mycroft
+as they rely (implicitly or explicitly) on error slices, and may be
+similarly affected.

### 2017-06-10 paper/oopsla17-submission/rebuttal.md
markdown formattign

@@
 
-1. Size / Diversity of Training Corpus [A,D]
+## Size / Diversity of Training Corpus [A,D]
 
 It may have been unclear in the text, but our corpus was drawn from
-student solutions to *3 homework sets*, not 3 programs. Each homework
+student solutions to **3 homework sets**, not 3 programs. Each homework
 set consisted of many individual programs. In total our corpus contains
-solutions to *23 distinct programs* that cover a broad range of FP idioms.
+solutions to **23 distinct programs** that cover a broad range of FP idioms.
 
@@
 
-2. Generalizing NATE Models to Other Datasets [A,D]
+## Generalizing NATE Models to Other Datasets [A,D]
 
@@
 
-We also claim that our models are *easy to train*, i.e. they can be
+We also claim that our models are **easy to train**, i.e. they can be
 trained on programs from a single instance of a course with around 50
@@
 
-3. Determination of "Fixed" Programs [B]
+## Determination of "Fixed" Programs [B]
 
-Reviewer B is correct that our determination of the *first* type-correct
+Reviewer B is correct that our determination of the **first** type-correct
 program as the fix is a threat to validity (see Section 4.5 for
@@
 
-The corpus is available at http://tinyurl.com/y9pzvcp7 (NOT ANONYMOUS);
+The corpus is available at http://tinyurl.com/y9pzvcp7 (**NOT ANONYMOUS**);
 we will advertise this in the text. It contains the full series of
@@
 
-4. Scaling to Larger Programs [C]
+## Scaling to Larger Programs [C]
 

### 2017-06-12 paper/oopsla17-submission/abstract.tex
clarify generality claims and dataset a bit

@@
 on a set of over 4,500 ill-typed \ocaml
-programs drawn from introductory
-programming classes.
+programs drawn from two instances of an
+introductory programming course.
 %

### 2017-06-12 paper/oopsla17-submission/evaluation.tex
clarify generality claims and dataset a bit

--- 2017-04-24 paper/oopsla17-submission/evaluation.tex
@@
 %
+When trained on student errors from one
+instance of an undergraduate course and
+tested on another instance,
 \toolname's most sophisticated
@@
 sub-term \HiddenFhTopOne\% of the time
-which, a good \ToolnameWinSherrloc points
-higher than the state of
-the art \sherrloc's \SherrlocTopOne\%.
+--- a good \ToolnameWinSherrloc points
+higher than the state-of-the-art
+\sherrloc's \SherrlocTopOne\%.
 %
@@
 That is, once we focus our classifiers on
-slices, purely syntactic features
+slices, purely local syntactic features
 perform as well as the
@@
 %
-Furthermore, we find contextual syntax
+Furthermore, we find contextual syntactic
 features to be mostly (but not entirely)
@@
 models, we found that the models
-do indeed capture some simple and intuitive
+appear to capture some simple and intuitive
 rules for predicting well-typedness.
@@
 %
-We recorded each interaction with the \ocaml top-level system over the
-course of the first three assignments, capturing
-ill-typed programs and, crucially, their subsequent fixes.
+We recorded each interaction with the \ocaml top-level system while the
+students worked on 23 programs from the first three homework
+assignments, capturing ill-typed programs and, crucially, their
+subsequent fixes.
 %

### 2017-06-12 paper/oopsla17-submission/intro2.tex
rebuttal: 498 words
clarify generality claims and dataset a bit

@@
 %
-%Types are awesome.
-When viewed fully, types are awe-inspiring.
+Types are awesome.
+%When viewed fully, types are awe-inspiring.
 %
@@
 of an undergraduate Programming Languages
-course.\footnote{
-% \begin{anonsuppress}
-% UC San Diego (IRB \#140608).
-% \end{anonsuppress}
-% \begin{noanonsuppress}
+course
+\begin{anonsuppress}
+at UC San Diego (IRB \#140608).
+\end{anonsuppress}
+\begin{noanonsuppress}
 at AUTHOR's INSTITUTION (IRB HIDDEN).
-% \end{noanonsuppress}
-}
+\end{noanonsuppress}
 %
@@
 solves this problem with a simple, yet expressive, representation called
-a \emph{Bag-of-Abstracted-Terms (BOAT)} wherein
+a \emph{Bag-of-Abstracted-Terms} (BOAT) wherein
 each program is represented by the \emph{bag}
@@
 %
-Our evaluation lets us identify the best features and
-classifier and empirically characterize the importance
-of different features for error localization.
+We find that \toolname's models \emph{generalize} well
+between instances of the same undergraduate course, outperforming
+the state of the art by at least \ToolnameWinSherrloc
+percentage points at predicting the source of a type error.
+%
+We also investigate which features and classifiers
+are most effective at localizing type errors, and
+empirically characterize the importance of different
+feature sets.
+% Our evaluation lets us identify the best features and
+% classifier and empirically characterize the importance
+% of different features for error localization.
 %
@@
 the \emph{type error slice} \citep{Tip2001-qp} brings our
-classifiers up to par with the state of the art,
+classifiers up to par with the state-of-the-art,
 and further augmenting the data with \emph{contextual}
 features allows our classifiers to outperform
-the state of the art by \ToolnameWinSherrloc percentage points.
+the state-of-the-art by \ToolnameWinSherrloc percentage points.
 % In particular, we find that while machine learning

### 2017-06-19 paper/oopsla17-submission/evaluation-accuracy.tex
update numbers

@@
 programs where they predicted locations that our oracle could not match
-with a program expression: 6--8\% of programs for \mycroft and 3--4\%
+with a program expression: around 6\% of programs for \mycroft and 4\%
 for \sherrloc.
@@
 out (after one minute) or where they encountered an unsupported language
-feature: another 5\% for \mycroft and 12--13\% for \sherrloc.
+feature: another 5\% for \mycroft and 12\% for \sherrloc.
 %

### 2017-06-19 paper/oopsla17-submission/evaluation-utility.tex
update numbers

@@
 %
-As expected, the baseline performs the worst, with a mere 28\% \linear
+As expected, the baseline performs the worst, with a mere 25\% \linear
 Top-1 accuracy.
@@
 However, filtering out expressions that are not part of the slice
-\emph{further} improves the results to 56\% \linear Top-1 accuracy.
+\emph{further} improves the results to 53\% \linear Top-1 accuracy.
 %
@@
 %
-Top-1 accuracy jumps from 33\% to 55\% when we add \InSlice, and only
-improves by 3\% when we filter out expressions that are not part of the
+Top-1 accuracy jumps from 30\% to 54\% when we add \InSlice, and only
+improves by 1\% when we filter out expressions that are not part of the
 error slice.
@@
 \ExprSize is the weakest feature, improving \linear Top-1
-accuracy by less than 1\% and \hiddenFH by only 3\%.
+accuracy by only 1\% and \hiddenFH by only 4\%.
 %
 In contrast, the contextual syntactic features improve \linear Top-1
-accuracy by 4\% (\resp 14\%), and the typing features improve
-Top-1 accuracy by 6\% (\resp 16\%).
+accuracy by 6\% (\resp 16\%), and the typing features improve
+Top-1 accuracy by 7\% (\resp 19\%).
 %
@@
 %
-They improve Top-1 accuracy by an additional 2\%, and are much more
+They improve Top-1 accuracy by an additional 1\% (\resp 3\%), and are much more
 compact --- requiring only 55 typing features compared to 180

### 2017-06-19 paper/oopsla17-submission/evaluation.tex
update numbers

@@
 more than one standard deviation above the mean, establishing a diff
-threshold of 45\%.
+threshold of 40\%.
 %
 This accounts for roughly 14\% of each dataset, leaving us with
-2,425 program pairs for \SPRING and 2,325 pairs for \FALL.
+2,710 program pairs for \SPRING and 2,393 pairs for \FALL.
 

### 2017-06-19 paper/oopsla17-submission/intro2.tex
update numbers

@@
 precision against the state-of-the-art
-on a set of over 4,500 ill-typed \ocaml
+on a set of over 5,000 ill-typed \ocaml
 programs drawn from two instances of an
@@
 Our \emph{first contribution} (\autoref{sec:overview})
-is a set of more than 4,500 labeled programs,
+is a set of more than 5,000 labeled programs \citep{Seidel2017-ko},
 giving us an accurate ground truth of

### 2017-06-19 paper/oopsla17-submission/threats.tex
tweak
update numbers

@@
-\mysection{Threats to Validity}
+\mysubsection{Threats to Validity}
 \label{sec:validity}
@@
 (2) our Top-1 accuracy margin is large enough that even if we assumed
-that \sherrloc were perfect on all excluded programs, it would
-only tie our \hiddenFH. % in Top-1 accuracy.
+that \sherrloc were perfect on all excluded programs,
+% it would only tie our \hiddenFH. % in Top-1 accuracy.
+we would still lead by 10 points.
 %

### 2017-06-19 study/study.R

@@ -44,7 +44,9 @@ fixes0 = c(a0$X2..sepConcat.fix..1.0.pts.,
 
 significance <- function(x,y) {
   print(x)
+  print(mean(x))
   print(y)
+  print(mean(y))
   wilcox.test(x, y, alternative="g")
 }
 

### 2017-06-19 study/study.R
add other user-study data

@@ -2,8 +2,10 @@ library(irr)
 
 a0 = read.csv(file="study-data/Types_Study_A_scores.csv")
 b0 = read.csv(file="study-data/Types_Study_B_scores.csv")
-
-reasons0 = c(a0)
+a1 = read.csv(file="study-data/Types_Study_1_A_scores.csv")
+b1 = read.csv(file="study-data/Types_Study_1_B_scores.csv")
+a2 = read.csv(file="study-data/Types_Study_2_A_scores.csv")
+b2 = read.csv(file="study-data/Types_Study_2_B_scores.csv")
 
 reasons0 = c(a0$X1..sepConcat.explain..1.0.pts.,
              a0$X3..padZero.explain..1.0.pts.,
@@ -11,10 +13,22 @@ reasons0 = c(a0$X1..sepConcat.explain..1.0.pts.,
              b0$X1..sepConcat.explain..1.0.pts.,
              b0$X3..padZero.explain..1.0.pts.,
              b0$X5..mulByDigit.explain..1.0.pts.)
-#reasons = data.frame(reasons0)
-## filter out answers that any annotator marked as INCOMPLETE
-# reasons = reasons[reasons$reasons0 != -1 & reasons$reasons1 != -1 & reasons$reasons2 != -1 & reasons$reasons3 != -1,]
-#reasons = reasons[reasons$reasons0 != -1]
+reasons1 = c(a1$X1..sepConcat.explain..1.0.pts.,
+             a1$X3..padZero.explain..1.0.pts.,
+             a1$X5..mulByDigit.explain..1.0.pts.,
+             b1$X1..sepConcat.explain..1.0.pts.,
+             b1$X3..padZero.explain..1.0.pts.,
+             b1$X5..mulByDigit.explain..1.0.pts.)
+reasons2 = c(a2$X1..sepConcat.explain..1.0.pts.,
+             a2$X3..padZero.explain..1.0.pts.,
+             a2$X5..mulByDigit.explain..1.0.pts.,
+             b2$X1..sepConcat.explain..1.0.pts.,
+             b2$X3..padZero.explain..1.0.pts.,
+             b2$X5..mulByDigit.explain..1.0.pts.)
+reasons = data.frame(reasons0, reasons1, reasons2)
+# filter out answers that any annotator marked as INCOMPLETE
[…]

### 2017-06-21 paper/oopsla17-submission/intro2.tex
widows

@@
 of what kinds of errors are more
-or less likely, rather than
+%or less
+likely, rather than
 adapting to the kinds of mistakes
@@
 \emph{tree-diff} between the two terms to find
-the exact sub-terms that were changed in the fix.
+the exact sub-terms that changed in the fix.
 

### 2017-06-21 paper/oopsla17-submission/overview.tex
acmsmall and basic reformatting
widows

@@
 \begin{figure}[t!]
+\small
 \begin{minipage}{0.45\linewidth}
@@
 to |sumList|, complaining that |sumList|
-returns a |list| where an |int| was
+returns a |list| while an |int| was
 expected by the |+| operator.
@@
 with the programmer's intention and may
-not be helpful debugging information to
-a novice.
+not help the novice understand the error.
 
@@
   \item How can we \emph{represent} blame-labeled programs
-        as numeric vectors amenable to machine learning?
+        in a format amenable to machine learning?
 
@@
 representations, from raw
-sequences of tokens, to
-richly-structured
+token streams to
+rich % richly-structured
 abstract syntax trees (AST) or
@@
 %
-Our choice of working with ASTs is natural
-as that is the representation over which the
-type-checker operates.
+Working with ASTs is a natural choice
+as type-checkers operate on the same representation.
 
@@
 models. Our BOAT representation
-lets us iteratively solve this
-problem by just starting with
+enables an iterative solution
+by starting with
 a simple set of features, and
@@
 %
-This use of ranking to report the
-results of a program analysis is
+The use of ranking to report the
+results of an analysis is
 popular in other problem domains

### 2017-06-21 paper/oopsla17-submission/evaluation-accuracy.tex

@@
 
-In this experiment we compare the accuracy of our predictions to the
+First, we compare the accuracy of our predictions to the
 state of the art in type error localization.
@@
 constraint graph for constraints that participate in many unsatisfiable
-paths and few satisfiable paths.
+paths and comparatively few satisfiable paths.
 %
@@
 \mypara{Our Classifiers}
-We evaluate five classifiers, each trained on the full set of features.
+We evaluate five classifiers, each trained on the full feature set.
 % features: 44 local syntactic features, 176 contextual syntactic
@@
 %
-Our classifiers are:
+These include:
+%Our classifiers are:
 %
@@
 \item[\forest] A random forest \citep{Breiman2001-wo} of 30
-  estimators, trained with an impurity threshold of $10^{-7}$.
+  estimators, with an impurity threshold of $10^{-7}$.
 \item[\hiddenT and \hiddenFH] Two multi-layer perceptron neural

### 2017-06-21 paper/oopsla17-submission/qualitative.tex
update qualitative.tex

@@
 \mypara{Constructing a List of Duplicates}
-% data/sp14/0969.ml
+% data/sp14/0967.ml
 Our first program is a simple recursive function |clone| that takes an
@@
 
-\ES{FIX: our first/second predictions are now swapped}
-Our classifier incorrectly predicts that the use of |x| in the recursive
-call is the most likely source of the error.
+Our top prediction coincides with \sherrloc (and \ocaml), blaming the
+the first argument to |clone| rather than the occurrence of |clone| itself.
 %
-Our second prediction coincides with \sherrloc (and \ocaml), blaming the
-the first argument to |clone|.
+% This is also incorrect, but may be more helpful than our first
+% prediction --- if our student decides that she has certainly provided
+% the correct \emph{argument}, an alternative explanation is that
+% perhaps she has called the wrong \emph{function}.
 %
-This is also incorrect, but may be more helpful than our first
-prediction --- if our student decides that she has certainly provided
-the correct \emph{argument}, an alternative explanation is that
-perhaps she has called the wrong \emph{function}.
-
-We confess that both of these predictions are difficult to explain by
+We confess that this prediction is difficult to explain by
 examining the induced paths.
 %
-In particular, both predictions only reference the expression's context,
+In particular, it only references the expression's context,
 which is surprising.
@@
 \mypara{Currying Considered Harmful?}
-% data/sp14/1082.ml
+% data/sp14/1080.ml
 Our next example is another ill-fated attempt at |clone|.
@@
 \mypara{Extracting the Digits of an Integer}
-% data/sp14/1737.ml
+% data/sp14/1732.ml
 Consider first a simple recursive function |digitsOfInt| that extracts
@@
 \mypara{Padding a list}
-% data/sp14/1394.ml
+% data/sp14/1389.ml
 Our next program, |padZero|, is given two |int list|s as input, and must

### 2017-06-27 paper/oopsla17-submission/qualitative.tex

@@
       else
-        (*@\hlSherrloc{loop} \hlTree{(x, (n - 1), (x :: acc))}@*) in
+        (*@\hlTree{\hlSherrloc{loop} (x, (n - 1), (x :: acc))}@*) in
     loop (x, n, [])
@@
 \sherrloc blames the recursive call to |loop| while our classifier
-blames the tuple of arguments --- a reasonable suggestion, but not
-the answer the student expected.
+blames the entire application.
+% tuple of arguments --- a reasonable suggestion, but not
+% the answer the student expected.
 

### 2017-06-28 paper/oopsla17-submission/qualitative.tex

--- 2017-06-21 paper/oopsla17-submission/qualitative.tex
@@
 %
-Much clearer is why we fail to blame the occurrence of |clone|, the two
-enabled features on the path are:
+More clear is why we fail to blame the occurrence of |clone|, two of
+the enabled features on the path are:
 %
@@
 %
-The model seems to have learned that programmers typically call the
+The model appears to have learned that programmers typically call the
 correct function.
@@
 \sherrloc blames the recursive call to |loop| while our classifier
-blames the tuple of arguments --- a reasonable suggestion, but not
+blames the
+% third element of the tuple.
+tuple of arguments --- a reasonable suggestion, but not
 the answer the student expected.
@@
 %
-(3) the parent expression's type mentions |list|.
+(3) the parent expression is also an application.
 %
@@
 constructor application.
-%
 

### 2017-07-24 paper/oopsla17-cameraready/abstract.tex
copy into oopsla17-cameraready

--- 2017-04-17 paper/oopsla17-submission/abstract.tex
@@
 precision to the state of the art
-on a set of over 4,500 ill-typed \ocaml
-programs drawn from introductory
-programming classes.
+on a set of over 5,000 ill-typed \ocaml
+programs drawn from two instances of an
+introductory programming course.
 %

### 2017-07-24 paper/oopsla17-cameraready/conclusion.tex
copy into oopsla17-cameraready

--- 2017-04-24 paper/oopsla17-submission/conclusion.tex
@@
 state of the art on a new benchmark
-suite comprising 4,500 student programs.
+suite comprising 5,000 student programs.
 %

### 2017-07-24 paper/oopsla17-cameraready/data.tex
copy into oopsla17-cameraready

\pgfplotstableset{col sep=comma}

% \pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/linear/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/decision-tree/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/random-forest/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}

\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/linear/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/decision-tree/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/random-forest/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-10/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-500/results.csv}

\pgfplotstableread{../../models/linear-op+slice-no-slice.cross.csv}{\SliceLinearBench}
\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice-only-slice.cross.csv}
\pgfplotstableread{../../models/hidden-500-op+slice-no-slice.cross.csv}{\SliceHiddenBench}
\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice-only-slice.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\SliceBench}

\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureLinearBench}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type+size.cross.csv}
\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureHiddenBench}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type+size.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\FeatureBench}

### 2017-07-24 paper/oopsla17-cameraready/evaluation-utility-graph.tex
copy into oopsla17-cameraready

--- 2017-04-17 paper/oopsla17-submission/evaluation-utility-graph.tex
@@
   xticklabels={\textsc{Local Syntax}, +\InSlice, \textsc{Filter \InSlice}},
-  %x tick label style={rotate=45, anchor=north east},
+  x tick label style={font=\small},
+  y tick label style={font=\small},
   reverse legend,
@@
   xticklabels={
-    \textsc{Local Syntax}\\(45),
+    \textsc{Local Syn}\\(45),
     \textsc{+Size}\\(46), \textsc{+Context}\\(225), \textsc{+Type}\\(100),
@@
   },
+  x tick label style={font=\small},
+  y tick label style={font=\small},
   %x tick label style={rotate=45, anchor=north east},
@@
   %
-  The total number of features for each set is given in parentheses.
+  The total number of features is given in parentheses.
 }

### 2017-07-24 paper/oopsla17-cameraready/evaluation-utility.tex
copy into oopsla17-cameraready

--- 2017-06-19 paper/oopsla17-submission/evaluation-utility.tex
@@
 %
-Adding \InSlice improves the results substantially with a 46\% \linear Top-1
+Adding \InSlice improves the results substantially with a 45\% \linear Top-1
 accuracy, demonstrating the importance of a minimal error slice.
@@
 However, filtering out expressions that are not part of the slice
-\emph{further} improves the results to 53\% \linear Top-1 accuracy.
+\emph{further} improves the results to 54\% \linear Top-1 accuracy.
 %
@@
 %
-Top-1 accuracy jumps from 30\% to 54\% when we add \InSlice, and only
+Top-1 accuracy jumps from 29\% to 53\% when we add \InSlice, and only
 improves by 1\% when we filter out expressions that are not part of the
@@
 \ExprSize is the weakest feature, improving \linear Top-1
-accuracy by only 1\% and \hiddenFH by only 4\%.
+accuracy by less than 1\% and \hiddenFH by only 4\%.
 %
 In contrast, the contextual syntactic features improve \linear Top-1
-accuracy by 6\% (\resp 16\%), and the typing features improve
-Top-1 accuracy by 7\% (\resp 19\%).
+accuracy by 5\% (\resp 16\%), and the typing features improve
+Top-1 accuracy by 6\% (\resp 18\%).
 %

### 2017-07-24 paper/oopsla17-cameraready/evaluation.tex
copy into oopsla17-cameraready

--- 2017-06-19 paper/oopsla17-submission/evaluation.tex
@@
 \mypara{Feature Selection}
-We extract a set of 282 features from each sub-expression in a student
+We extract 282 features from each sub-expression in a
 program, including:
@@
 This accounts for roughly 14\% of each dataset, leaving us with
-2,710 program pairs for \SPRING and 2,393 pairs for \FALL.
+2,712 program pairs for \SPRING and 2,365 pairs for \FALL.
 

### 2017-07-24 paper/oopsla17-cameraready/feature-table.tex
copy into oopsla17-cameraready

\begin{table}[ht]
\caption{Example Feature Vectors}\label{tab:sumList}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Expression}
  & \IsNil & \IsCaseListP & \ExprSize
  & \HasTypeIntCOne & \HasTypeList & \InSlice \\
\midrule
|[]|
  & 1 & 1 & 1 & 0 & 1 & 1 \\
|hd + sumList tl|
  & 0 & 1 & 5 & 1 & 0 & 1 \\
|sumList tl|
  & 0 & 0 & 3 & 0 & 1 & 1 \\
|tl|
  & 0 & 0 & 1 & 0 & 1 & 0 \\
\bottomrule
\end{tabular}
\bigskip
\caption*{A selection of the features we would extract from the
\lstinline!sumList! program in \autoref{fig:sumList}. A feature is
considered \emph{enabled} if it has a non-zero value, and
\emph{disabled} otherwise. A ``-P'' suffix indicates that the feature
describes the parent of the current expression, a ``-C$n$'' suffix
indicates that the feature describes the $n$-th (left-to-right) child of
the current expression.  Note that, since we rely on a partial typing
derivation, we are subject to the well-known traversal bias and label
the expression \lstinline!sumList tl! as having type
$\tlist{\cdot}$. The model will have to learn to correct for this bias.}
\end{table}

### 2017-07-24 paper/oopsla17-cameraready/discussion.tex
minor tweaks to appease reviewers + cover-letter

--- 2017-04-15 paper/oopsla17-submission/discussion.tex
@@
 %
-However, there is certainly room for improvement.
+% However, there is certainly room for improvement.
 %
@@
 We can mitigate this, to some extent, by adding generic
-syntactic features for data constructors and match expressions, but it
+syntactic features for data constructors and |match| expressions, but it
 remains to be seen how much these help. % \ES{TODO: UW data!}.
@@
 %
-Anything we learn about |::| and |2| can now be transferred directly to
+Anything we learn about |::| and |2| can now be transferred to
 yet unseen types, but we run the risk generalizing \emph{too much} ---
-\ie perhaps programmers make different mistakes with |list|s than they
+\ie perhaps programmers make different errors with |list|s than they
 do with other types, and are thus likely to choose different fixes.

### 2017-07-24 paper/oopsla17-cameraready/evaluation.tex
minor tweaks to appease reviewers + cover-letter

@@
   Are the models \emph{interpretable} using
-  our intuition about the causes of type errors?
+  our intuition about type errors?
   (\autoref{sec:qualitative})
@@
 \mypara{1. Data Beats Algorithms}
-Our main result is that for type error
+Our main result is that for (novice) type error
 localization, data is indeed unreasonably
@@
 However, even \toolname's simple
-\emph{logistic regression} based
+\emph{logistic regression}-based
 classifier is correct \LinearTopOne\% of the time,

### 2017-07-24 paper/oopsla17-cameraready/intro2.tex
minor tweaks to appease reviewers + cover-letter

--- 2017-06-21 paper/oopsla17-submission/intro2.tex
@@
 using different classes of features like the above, and
-with three different classification algorithms: logistic
-regression, decision trees, and neural networks.
+with four different classification algorithms: logistic
+regression, decision trees, random forests, and neural networks.
 %
@@
 the kinds of errors that programmers
-(specifically, novices who are in greatest
+(in particular novices, who are in greatest
 need of precise feedback) actually make

### 2017-07-24 paper/oopsla17-cameraready/learning.tex
minor tweaks to appease reviewers + cover-letter

--- 2017-04-24 paper/oopsla17-submission/learning.tex
@@
 %
-Our models will have to learn good localizations in spite this bias (see
+Our models will have to learn good localizations in spite of this bias (see
 \autoref{sec:evaluation}).
@@
 %
-Each leaf is labeled with a prediction and the fraction of training
-samples that would reach it; the latter quantity can be interpreted as
-the decision tree's confidence in its prediction.
+Each leaf is labeled with a prediction and the fraction of
+correctly-labeled training samples that would reach it; the latter
+quantity can be interpreted as the decision tree's confidence in its
+prediction.
 %
@@
 The last (and most complex) model we use is a type of neural network
-called a \emph{multi-layer perceptron} (see \citealt{Nielsen2015-pu} for
+called a \emph{multi-layer perceptron} (see \citet{Nielsen2015-pu} for
 an introduction to neural networks).

### 2017-07-24 paper/oopsla17-cameraready/overview.tex
minor tweaks to appease reviewers + cover-letter

--- 2017-06-21 paper/oopsla17-submission/overview.tex
@@
 yield a dataset large enough
-(\eg 1000s of programs) needed
+(\eg 1000s of programs)
 to facilitate machine learning.
@@
 %
-While the PL literature is full
+While the Programming Languages literature is full
 of different program
@@
 token streams to
-rich % richly-structured
+richly-structured
 abstract syntax trees (AST) or

### 2017-07-24 paper/oopsla17-cameraready/related.tex
minor tweaks to appease reviewers + cover-letter

--- 2017-04-17 paper/oopsla17-submission/related.tex
@@
 track choices made by the type checker and enumerate potential
-% type (and expression)
+type (and expression)
 changes that would fix the error.

### 2017-07-24 paper/oopsla17-cameraready/threats.tex
copy into oopsla17-cameraready
minor tweaks to appease reviewers + cover-letter

--- 2017-06-19 paper/oopsla17-submission/threats.tex
@@
 We mitigate this threat by including the largest empirical evaluation of
-type error localization that we are aware of: over 4,500 pairs of
+type error localization that we are aware of: over 5,000 pairs of
 ill-typed programs and fixes from two instances of the course, with
@@
 tools timed out or encountered unsupported language features is another
-threat to validity.
+threat.
 %
@@
 % it would only tie our \hiddenFH. % in Top-1 accuracy.
-we would still lead by 10 points.
+we would still lead by 9 points.
 %

### 2017-07-26 paper/oopsla17-cameraready/evaluation-user-study.tex
add user study

\subsection{Blame Utility}
\label{sec:user-study}

%\ES{TODO: explain that we selected programs where nate got the right answer and sherrloc did NOT}

% Finally, to test the explanatory power of our blame assigments, we
% ran a user study (IRB \#2014009900) at the University of Virginia (UVA).
% %
% We included three problems in an exam in the Spring 2017 session of UVA's
% undergraduate Programming Languages course (CS 4501).
% %
% We presented the 31 students in the course with ill-typed \ocaml\
% programs and asked them to
% %
% (1) \emph{explain} the type error, and
% %
% (2) \emph{fix} the type error.
% %
% For each problem the student was given the ill-typed program and
% either \sherrloc or \toolname's blame assignment, with no error message.

We assigned three problems to the students: the |padZero| and
|mulByDigit| programs from \autoref{sec:qualitative}, as well as
the following |sepConcat| program
%
\begin{ecode}
  let rec sepConcat sep sl =
    match sl with
    | [] -> ""
    | h::t ->
        let f a x = a ^ (sep ^ x) in
        let base = (*@\hlTree{[]}@*) in
        (*@\hlSherrloc{List.fold\_left f base sl}@*)
\end{ecode}
%
where the student has erroneously returned the empty list, rather than
the empty string, in the base case of the fold.
% which triggers an error on line 7.
%
% For each problem the students were given the ill-typed program and
% either \ocaml's error or \toolname's jump-compressed trace;
%
The full user study is available in \autoref{sec:user-study-exams}.
%
Due to the nature of an in-class exam not every student answered every
question, but we always received at least 12 (out of a possible 15 or
16) responses for each problem-tool pair.
%
A further complication for this study is that this session of the course
was taught in \textsc{Reason}\footnote{\url{https://reasonml.github.io}},
a dialect of \ocaml with a more C-like syntax, and thus for the study
we transcribed the programs to \textsc{Reason} syntax.

We then instructed three annotators (one of whom is an author, the others
are graduate students at UCSD) to classify the answers as
correct or incorrect.
%
We performed an inter-rater reliability (IRR) analysis to determine the
degree to which the annotators consistently graded the exams.
%
As we had more than two annotators assigning nominal (``correct'' or
``incorrect'') ratings we used Fleiss' kappa~\cite{Fleiss1971-du} to
measure IRR.\@
%
Fleiss' kappa is measured on a scale from $1$, indicating total
agreement, to $-1$, indicating total disagreement, with $0$ indicating
random agreement.

Finally, we used a one-sided Mann-Whitney $U$ test~\cite{Mann1947-fd} to
determine the significance of our results.
%
The null hypothesis was that the responses from students given
\toolname's blame were drawn from the same distribution as those
given \sherrloc's, \ie \toolname had no effect.
%
Since we used a one-sided test, the alternative to the null hypothesis
is that \toolname had a \emph{positive} effect on the responses.
%
We reject the null hypothesis in favor of the alternative if the test
produces a significance level $p < 0.05$, a standard threshold for
determining statistical significance.

\mypara{Threats to Validity}
%
Measuring understanding is difficult, and comes with its own
set of threats. % to validity.

\paragraph{Construct.}
%
We used the correctness of the student's explanation of, and fix for,
the type error as a proxy for her understanding, but it is possible
that other metrics would produce different results.
%
A further threat arises from our decision to use \textsc{Reason} syntax
rather than \ocaml.
%
\textsc{Reason} and \ocaml differ only in syntax, the type system is the
same; however, the difference in syntax may affect students'
understanding of the programs.
%
For example, \textsc{Reason} uses the notation |[h, ...t]| for the list
``cons'' constructor, in contrast to \ocaml's |h::t|.
%
It is quite possible that \textsc{Reason}'s syntax could help students
remember that |h| is a single element while |t| is a list.

\paragraph{Internal.}
%
We assigned students randomly to two groups. The first was given
\sherrloc's blame assignment for |sepConcat| and |mulByDigit|, and
\toolname's blame for |padZero|; the second was given the opposite
assignment. This ensured that each student was given \sherrloc and
\toolname problems. Students without sufficient knowledge of
\textsc{Reason} could affect the results, as could the time-constrained
nature of an exam. Thus, we excluded any answers left blank
from our analysis.

\paragraph{External.}
%
Our experiment used students in the process of learning \textsc{Reason},
and thus may not generalize to all developers. The three programs were
chosen manually, via a random selection and filtering of the programs
from the \SPRING dataset, where \toolname's top prediction was correct
but \sherrloc's was incorrect. A different selection of programs may
lead to different results.

\paragraph{Conclusion.}
%
We collected exams from 31 students, though due to the nature of the
study not every student completed every problem.
%
The number of complete submissions was always at least 12 out of
a maximum of 15 or 16 per program-tool pair.
%
While our results are not statistically significant,
collecting more responses per test pair was not possible as it
would require having students answer the same problem twice (once with
\sherrloc and once with \toolname).

\begin{figure}[t]
\centering
% \centerline{
% \begin{minipage}{1.2\textwidth}
\includegraphics[width=0.7\linewidth]{user-study-reason.png}
\includegraphics[width=0.7\linewidth]{user-study-fix.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption[A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.]
  {A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.
  %
  The students given \toolname's location generally scored
  better than those given \sherrloc's.
  %
  We report the result of a one-sided Mann-Whitney $U$ test for
  statistical significance in parentheses.}
\label{fig:results-user-study}
\end{figure}

\mypara{Results}
%
The measured kappa values were $\kappa = 0.68$ for the explanations and
$\kappa = 0.77$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's blame assignment
were generally more likely to correctly explain and fix the type error
than those given \sherrloc's.
%
While there was no discernible difference between \toolname and
\sherrloc for |sepConcat|, \toolname responses for |padZero| and
|mulByDigit| were marked correct 5--30\% more often than the \sherrloc
responses.
%
The results appear to show a trend in favor of \toolname;
%
however, none rise to the level of statistical significance.
%

### 2017-07-26 paper/oopsla17-cameraready/evaluation.tex
add user study

@@
 %
-We seek to answer three questions in our evaluation:
+We seek to answer four questions in our evaluation:
 %
@@
   (\autoref{sec:qualitative})
+
+\item \textbf{Blame Utility}
+  Do \toolname's blame assignments help
+  users diagnose type errors?
+  (\autoref{sec:user-study})
 \end{itemize}
@@
 
+\mypara{Blame Utility}
+Finally, to test the explanatory power of our blame assigments, we
+ran a user study at the University of Virginia (UVA IRB \#2014009900).
+%
+We included three problems in an exam in the Spring 2017 session of UVA's
+undergraduate Programming Languages course (CS 4501).
+%
+We presented the 31 students in the course with ill-typed \ocaml\
+programs and asked them to
+%
+(1) \emph{explain} the type error, and
+%
+(2) \emph{fix} the type error.
+%
+For each problem the student was given the ill-typed program and
+either \sherrloc or \toolname's blame assignment, with no error message.
+
+\input{threats}
 \input{evaluation-accuracy}
 \input{evaluation-utility}
+\input{qualitative}
+\input{evaluation-user-study}
 

### 2017-07-26 paper/oopsla17-cameraready/plots.py

@@ -1,0 +1,115 @@
+import csv
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+
+UCSD = 'UCSD'
+
+BUCKETS = [0.1, 0.2, 1.0, 10.0, 60.0 ] # range(500, 3001, 500)
+#COLORS=['#90B0D4', '#90D492', '#D4B490', '#D490D2']
+COLORS=['#8dd3c7','#bebada','#ffffb3','#fb8072','#80b1d3','#fdb462']
+COLORS_E=['#8dd3c7','#bebada','#80b1d3','#ffffb3','#fdb462','#fb8072']
+
+def read_csv(f):
+    with open(f) as f:
+        return list(csv.reader(f))
+def read_csv_dict(f):
+    with open(f) as f:
+        return list(csv.DictReader(f))
+
+def plot_user_study():
+    a = read_csv_dict('study/study-data/Types_Study_A_scores.csv')
+    b = read_csv_dict('study/study-data/Types_Study_B_scores.csv')
+
+    def f(xs):
+        return [float(x) for x in xs if float(x) >= 0]
+
+    def err(xs):
+        #p = np.average(xs)
+        #return 100 * np.sqrt(p * (1-p) / len(xs))
+        s = np.std(xs)
+        n = len(xs)
+        return 100 * (s / np.sqrt(n))
+
+    ## REASON
+    sepconcat_a = f([r['1: sepConcat explain (1.0 pts)'] for r in a])
+    padzero_a = f([r['3: padZero explain (1.0 pts)'] for r in a])
+    mulbydigit_a = f([r['5: mulByDigit explain (1.0 pts)'] for r in a])
+
+    sepconcat_b = f([r['1: sepConcat explain (1.0 pts)'] for r in b])
+    padzero_b = f([r['3: padZero explain (1.0 pts)'] for r in b])
+    mulbydigit_b = f([r['5: mulByDigit explain (1.0 pts)'] for r in b])
+
+    ind = np.arange(3)
+    width = 0.35
+
+    fig = plt.figure()
+    p_o = plt.bar(ind,
+                  [100*np.average(sepconcat_a), 100*np.average(padzero_b), 100*np.average(mulbydigit_a)],
+                  width,
+                  color=COLORS[0],
+                  yerr=map(err, [sepconcat_a, padzero_b, mulbydigit_a]),
[…]

### 2017-07-26 paper/oopsla17-cameraready/evaluation-user-study-graph.tex
fiddle with graphs

\begin{figure}[t]
\centering
%% EXPLAIN
\begin{tikzpicture}
\begin{axis}[
  ybar,
  % width=6cm,
  height=6cm,
  title={\large Explanation},
  ylabel={\% Correct},
  bar width=0.5cm,
  ymin=0,
  ymax=1,
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  enlarge x limits=0.25,
  symbolic x coords={{sepConcat}, {padZero}, {mulByDigit}},
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    {\texttt{sepConcat}\\($p=0.48$)},
    {\texttt{padZero}\\($p=0.097$)},
    {\texttt{mulByDigit}\\($p=0.083$)}
  },
  % x tick label style={rotate=45, anchor=north east},
  %x tick label style={font=\small},
  y tick label style={font=\small},
  reverse legend,
  transpose legend,
  legend style={legend pos = south west, font=\footnotesize},
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green2, bar shift=.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.8125)  +- (0, 0.0976)
  (padZero, 1.00)      +- (0, 0)
  (mulByDigit, 0.8125) +- (0, 0.0976)
};
\addlegendentry{\toolname}

\addplot[draw=black, fill=blue2, bar shift=-.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.80)    +- (0, 0.1033)
  (padZero, 0.875)     +- (0, 0.0827)
  (mulByDigit, 0.5714) +- (0, 0.1323)
};
\addlegendentry{\sherrloc}
\end{axis}
\end{tikzpicture}
%% FIX
\begin{tikzpicture}
\begin{axis}[
  ybar,
  % width=6cm,
  height=6cm,
  title={\large Fix},
  ylabel={\% Correct},
  bar width=0.5cm,
  ymin=0,
  ymax=1,
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  enlarge x limits=0.25,
  symbolic x coords={{sepConcat}, {padZero}, {mulByDigit}},
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    {\texttt{sepConcat}\\($p=0.57$)},
    {\texttt{padZero}\\($p=0.33$)},
    {\texttt{mulByDigit}\\($p=0.31$)}
  },
  % x tick label style={rotate=45, anchor=north east},
  %x tick label style={font=\small},
  y tick label style={font=\small},
  reverse legend,
  transpose legend,
  legend style={legend pos = south west, font=\footnotesize},
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green2, bar shift=.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.8125)  +- (0, 0.0976)
  (padZero, 0.9286)    +- (0, 0.0688)
  (mulByDigit, 0.7143) +- (0, 0.1207)
};
\addlegendentry{\toolname}

\addplot[draw=black, fill=blue2, bar shift=-.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.83)    +- (0, 0.1076)
  (padZero, 0.875)     +- (0, 0.0827)
  (mulByDigit, 0.6154) +- (0, 0.1349)
};
\addlegendentry{\sherrloc}
\end{axis}
\end{tikzpicture}
\caption[A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.]
  {A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.
  %
  The students given \toolname's location generally scored
  better than those given \sherrloc's.
  %
  We report the result of a one-sided Mann-Whitney $U$ test for
  statistical significance in parentheses.}
\label{fig:user-study-results}
\end{figure}

### 2017-07-26 paper/oopsla17-cameraready/evaluation-user-study.tex
fiddle with graphs

@@
 
-\begin{figure}[t]
-\centering
-% \centerline{
-% \begin{minipage}{1.2\textwidth}
-\includegraphics[width=0.7\linewidth]{user-study-reason.png}
-\includegraphics[width=0.7\linewidth]{user-study-fix.png}
-% \end{minipage}
-% }
-% \vspace{3ex}
-\caption[A classification of students' explanations and fixes for type
-  errors, given either \sherrloc or \toolname's blame assignment.]
-  {A classification of students' explanations and fixes for type
-  errors, given either \sherrloc or \toolname's blame assignment.
-  %
-  The students given \toolname's location generally scored
-  better than those given \sherrloc's.
-  %
-  We report the result of a one-sided Mann-Whitney $U$ test for
-  statistical significance in parentheses.}
-\label{fig:results-user-study}
-\end{figure}
+\input{evaluation-user-study-graph}
+
+% \begin{figure}[t]
+% \centering
+% % \centerline{
+% % \begin{minipage}{1.2\textwidth}
+% % \centering
+% \includegraphics[width=0.49\linewidth]{user-study-reason.png}
+% \includegraphics[width=0.49\linewidth]{user-study-fix.png}
+% % \end{minipage}
+% % }
+% % \vspace{3ex}
+% \caption[A classification of students' explanations and fixes for type
+%   errors, given either \sherrloc or \toolname's blame assignment.]
+%   {A classification of students' explanations and fixes for type
+%   errors, given either \sherrloc or \toolname's blame assignment.
+%   %
+%   The students given \toolname's location generally scored
+%   better than those given \sherrloc's.
+%   %
+%   We report the result of a one-sided Mann-Whitney $U$ test for
+%   statistical significance in parentheses.}
+% \label{fig:results-user-study}
+% \end{figure}
 

### 2017-07-26 paper/oopsla17-cameraready/evaluation-utility-graph.tex
fiddle with graphs

@@
   width=0.5\linewidth,
-  height=4cm,
+  height=5cm,
   %title={Impact of Type Error Slice},
@@
   bar width=0.5cm,
-  ymin=0.2,
+  ymin=0,
   ymax=1,
-  ytick={0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  ytick={0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
@@
   width=0.5\linewidth,
-  height=4cm,
+  height=5cm,
   ylabel={Recall},
   axis y line*=right,
-  ymin=0.2,
+  ymin=0,
   ymax=1,
-  ytick={0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+  ytick={0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
   yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
@@
   width=0.9\linewidth,
-  height=5cm,
+  height=6cm,
   %title={Impact of Contextual Feature Classes},
@@
   width=0.9\linewidth,
-  height=5cm,
+  height=6cm,
   ylabel={Recall},

### 2017-07-26 paper/oopsla17-cameraready/plots.py

@@ -43,6 +43,10 @@ def plot_user_study():
     ind = np.arange(3)
     width = 0.35
 
+    print 'EXPLAIN'
+    print 'sherrloc'
+    print [100*np.average(sepconcat_a), 100*np.average(padzero_b), 100*np.average(mulbydigit_a)]
+    print map(err, [sepconcat_a, padzero_b, mulbydigit_a])
     fig = plt.figure()
     p_o = plt.bar(ind,
                   [100*np.average(sepconcat_a), 100*np.average(padzero_b), 100*np.average(mulbydigit_a)],
@@ -51,6 +55,9 @@ def plot_user_study():
                   yerr=map(err, [sepconcat_a, padzero_b, mulbydigit_a]),
                   error_kw={'linewidth': 3, 'ecolor': 'gray', 'capsize': 6, 'capthick': 3}
     )
+    print 'nate'
+    print [100*np.average(sepconcat_b), 100*np.average(padzero_a), 100*np.average(mulbydigit_b)]
+    print map(err, [sepconcat_b, padzero_a, mulbydigit_b])
     p_n = plt.bar(ind + width,
                   [100*np.average(sepconcat_b), 100*np.average(padzero_a), 100*np.average(mulbydigit_b)],
                   width,
@@ -59,11 +66,11 @@ def plot_user_study():
                   error_kw={'linewidth': 3, 'ecolor': 'gray', 'capsize': 6, 'capthick': 3}
     )
 
-    plt.title('Explanation',fontsize=24)
+    plt.title('Explanation',fontsize=30)
     # plt.xlabel('Problem', fontsize=20)
-    plt.ylabel('% Correct', fontsize=20)
-    plt.xticks(ind + width, ['sepconcat\n(p = 0.48)', 'padzero\n(p = 0.097)', 'mulbydigit\n(p = 0.083)'], fontsize='large')
-    plt.legend(('SHErrLoc', 'Nate'), loc='lower right', fontsize=16)
+    plt.ylabel('% Correct', fontsize=24)
+    plt.xticks(ind + width, ['sepConcat\n(p = 0.48)', 'padZero\n(p = 0.097)', 'mulByDigit\n(p = 0.083)'], fontsize=20)
+    plt.legend(('SHErrLoc', 'Nate'), loc='lower right', fontsize=20)
     # autolabel(plt, p_o)
     # autolabel(plt, p_n)
 
[…]

### 2017-07-26 paper/oopsla17-cameraready/evaluation-user-study.tex
tweak

@@
 \sherrloc for |sepConcat|, \toolname responses for |padZero| and
-|mulByDigit| were marked correct 5--30\% more often than the \sherrloc
+|mulByDigit| were marked correct 5--25\% more often than the \sherrloc
 responses.
 %
-The results appear to show a trend in favor of \toolname;
+While the results appear to show a trend in favor of \toolname,
+they do not rise to the level of statistical significance.
+% The results appear to show a trend in favor of \toolname;
+% %
+% however, none rise to the level of statistical significance.
 %
-however, none rise to the level of statistical significance.
-%

### 2017-07-27 paper/oopsla17-cameraready/evaluation-user-study.tex
some Wes comments

@@
 %
-Due to the nature of an in-class exam not every student answered every
+Due to the nature of an in-class exam, not every student answered every
 question, but we always received at least 12 (out of a possible 15 or
@@
 %
-A further complication for this study is that this session of the course
-was taught in \textsc{Reason}\footnote{\url{https://reasonml.github.io}},
+%A further complication for this study is that
+This session of the course
+was taught in \textsc{Reason},\footnote{\url{https://reasonml.github.io}}
 a dialect of \ocaml with a more C-like syntax, and thus for the study
@@
 
-\paragraph{Conclusion.}
+\paragraph{Subjects.}
 %
@@
 %
-While our results are not statistically significant,
-collecting more responses per test pair was not possible as it
-would require having students answer the same problem twice (once with
-\sherrloc and once with \toolname).
+% While our results are not statistically significant,
+% collecting more responses per test pair was not possible as it
+% would require having students answer the same problem twice (once with
+% \sherrloc and once with \toolname).
 
@@
 %
-While there was no discernible difference between \toolname and
-\sherrloc for |sepConcat|, \toolname responses for |padZero| and
+There was no discernible difference between \toolname and
+\sherrloc for |sepConcat|; however, \toolname responses for |padZero| and
 |mulByDigit| were marked correct 5--25\% more often than the \sherrloc
@@
 While the results appear to show a trend in favor of \toolname,
-they do not rise to the level of statistical significance.
+they do not rise to the level of statistical significance;
+further investigation is merited.
 % The results appear to show a trend in favor of \toolname;

### 2017-07-27 paper/oopsla17-cameraready/evaluation-user-study.tex
rearrange threats per wes+kamalika

@@
 
-We assigned three problems to the students: the |padZero| and
-|mulByDigit| programs from \autoref{sec:qualitative}, as well as
-the following |sepConcat| program
+We have demonstrated in the preceding sections that we can produce
+more \emph{accurate} blame assignments by learning from the collective
+mistakes of prior students; however, users are the final judge of
+the \emph{utility} of an error message.
+%
+Thus, in this final experiment we ask whether \toolname's correct blame
+assignments aid users in \emph{understanding} type errors
+more than incorrect assignments.
+
+We assigned three problems to the students in our user study: the
+|padZero| and |mulByDigit| programs from \autoref{sec:qualitative}, as
+well as the following |sepConcat| program
 %
@@
 %
+For each problem the students were additionally given either \toolname's
+correct blame assignment or \sherrloc's incorrect blame assignment,
+with no error message.
 % For each problem the students were given the ill-typed program and
@@
 determining statistical significance.
+
+\input{evaluation-user-study-graph}
+
+% \begin{figure}[t]
+% \centering
+% % \centerline{
+% % \begin{minipage}{1.2\textwidth}
+% % \centering
+% \includegraphics[width=0.49\linewidth]{user-study-reason.png}
+% \includegraphics[width=0.49\linewidth]{user-study-fix.png}
+% % \end{minipage}
+% % }
+% % \vspace{3ex}
+% \caption[A classification of students' explanations and fixes for type
+%   errors, given either \sherrloc or \toolname's blame assignment.]
+%   {A classification of students' explanations and fixes for type
+%   errors, given either \sherrloc or \toolname's blame assignment.
+%   %
+%   The students given \toolname's location generally scored
+%   better than those given \sherrloc's.
+%   %
+%   We report the result of a one-sided Mann-Whitney $U$ test for
+%   statistical significance in parentheses.}
+% \label{fig:results-user-study}
+% \end{figure}
+
+\mypara{Results}
+%
+The measured kappa values were $\kappa = 0.68$ for the explanations and
+$\kappa = 0.77$ for the fixes; while there is no formal notion for what
+consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
+above $0.60$ are often called ``substantial''
+agreement~\cite{Landis1977-ey}.
+%
+Figure~\ref{fig:results-user-study} summarizes a single annotator's
+results, which show that students given \toolname's blame assignment
+were generally more likely to correctly explain and fix the type error
+than those given \sherrloc's.
+%
+There was no discernible difference between \toolname and
+\sherrloc for |sepConcat|; however, \toolname responses for |padZero| and
+|mulByDigit| were marked correct 5--25\% more often than the \sherrloc
+responses.
+%
+While the results appear to show a trend in favor of \toolname,
+they do not rise to the level of statistical significance
+in this experiment; further investigation is merited.
+% The results appear to show a trend in favor of \toolname;
+% %
+% however, none rise to the level of statistical significance.
+%
 
@@
 % \sherrloc and once with \toolname).
-
-\input{evaluation-user-study-graph}
-
-% \begin{figure}[t]
-% \centering
-% % \centerline{
-% % \begin{minipage}{1.2\textwidth}
-% % \centering
-% \includegraphics[width=0.49\linewidth]{user-study-reason.png}
-% \includegraphics[width=0.49\linewidth]{user-study-fix.png}
-% % \end{minipage}
-% % }
-% % \vspace{3ex}
-% \caption[A classification of students' explanations and fixes for type
-%   errors, given either \sherrloc or \toolname's blame assignment.]
-%   {A classification of students' explanations and fixes for type
-%   errors, given either \sherrloc or \toolname's blame assignment.
-%   %
-%   The students given \toolname's location generally scored
-%   better than those given \sherrloc's.
-%   %
-%   We report the result of a one-sided Mann-Whitney $U$ test for
-%   statistical significance in parentheses.}
-% \label{fig:results-user-study}
-% \end{figure}
-
-\mypara{Results}
-%
-The measured kappa values were $\kappa = 0.68$ for the explanations and
-$\kappa = 0.77$ for the fixes; while there is no formal notion for what
-consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
-above $0.60$ are often called ``substantial''
-agreement~\cite{Landis1977-ey}.
-%
-Figure~\ref{fig:results-user-study} summarizes a single annotator's
-results, which show that students given \toolname's blame assignment
-were generally more likely to correctly explain and fix the type error
-than those given \sherrloc's.
-%
-There was no discernible difference between \toolname and
-\sherrloc for |sepConcat|; however, \toolname responses for |padZero| and
-|mulByDigit| were marked correct 5--25\% more often than the \sherrloc
-responses.
-%
-While the results appear to show a trend in favor of \toolname,
-they do not rise to the level of statistical significance;
-further investigation is merited.
-% The results appear to show a trend in favor of \toolname;
-% %
-% however, none rise to the level of statistical significance.
-%

### 2017-07-27 paper/oopsla17-cameraready/threats.tex
rearrange threats per wes+kamalika

@@
 errors more accurately than the state of the art and that our features are
-relevant to blame assignment, our results may not generalize.
+relevant to blame assignment, our results may not generalize
+to other problem domains or program sets.
 

### 2017-08-24 paper/oopsla17-cameraready/data.tex
checkpoint

@@
 
-\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/linear/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/random-forest/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
+\pgfplotstableread{data/sp14-baseline.csv}{\SpringBench}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-ocaml-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-mycroft-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-sherrloc-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-linear-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-decision-tree-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-random-forest-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-hidden-10-results.csv}
+\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-hidden-500-results.csv}
 
-\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/linear/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/decision-tree/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/random-forest/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-10/results.csv}
-\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-500/results.csv}
+\pgfplotstableread{data/fa15-baseline.csv}{\FallBench}
+\pgfplotstablevertcat{\FallBench}{data/fa15-ocaml-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-mycroft-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-sherrloc-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-linear-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-decision-tree-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-random-forest-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-hidden-10-results.csv}
+\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-hidden-500-results.csv}
 
-\pgfplotstableread{../../models/linear-op+slice-no-slice.cross.csv}{\SliceLinearBench}
-\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice.cross.csv}
-\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice-only-slice.cross.csv}
-\pgfplotstableread{../../models/hidden-500-op+slice-no-slice.cross.csv}{\SliceHiddenBench}
-\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice.cross.csv}
-\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice-only-slice.cross.csv}
+\pgfplotstableread{models/linear-op+slice-no-slice.cross.csv}{\SliceLinearBench}
+\pgfplotstablevertcat{\SliceLinearBench}{models/linear-op+slice.cross.csv}
+\pgfplotstablevertcat{\SliceLinearBench}{models/linear-op+slice-only-slice.cross.csv}
+\pgfplotstableread{models/hidden-500-op+slice-no-slice.cross.csv}{\SliceHiddenBench}
+\pgfplotstablevertcat{\SliceHiddenBench}{models/hidden-500-op+slice.cross.csv}
+\pgfplotstablevertcat{\SliceHiddenBench}{models/hidden-500-op+slice-only-slice.cross.csv}
 % \pgfplotstablecreatecol[create col/assign/.code={%
@@
 
-\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureLinearBench}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+size.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+size.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type+size.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type.cross.csv}
-\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type+size.cross.csv}
-\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureHiddenBench}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+size.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+size.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type+size.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type.cross.csv}
-\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type+size.cross.csv}
+\pgfplotstableread{models/linear-op.cross.csv}{\FeatureLinearBench}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+size.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+type.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+size.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+type+size.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+type.cross.csv}
+\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+type+size.cross.csv}
+\pgfplotstableread{models/linear-op.cross.csv}{\FeatureHiddenBench}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+size.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+type.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+size.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+type+size.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+type.cross.csv}
+\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+type+size.cross.csv}
 % \pgfplotstablecreatecol[create col/assign/.code={%

### 2017-08-25 paper/oopsla17-cameraready/evaluation-utility-graph.tex
arghhhh

@@
   width=0.9\linewidth,
-  height=6cm,
+  height=5cm,
   %title={Impact of Contextual Feature Classes},
@@
   width=0.9\linewidth,
-  height=6cm,
+  height=5cm,
   ylabel={Recall},

### 2017-09-19 paper/oopsla17-cameraready/intro2.tex
sigh.. update acmart.cls again..

@@
 of an undergraduate Programming Languages
-course
-\begin{anonsuppress}
-at UC San Diego (IRB \#140608).
-\end{anonsuppress}
-\begin{noanonsuppress}
-at AUTHOR's INSTITUTION (IRB HIDDEN).
-\end{noanonsuppress}
+course at UC San Diego (IRB \#140608).
 %
